### 2022-05-24 Overview.Rmd
initial commit

---
title: "overview"
author: "Justin Pomeranz"
date: "5/24/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ABT
### attempt 1
Size spectra are fundamental descriptions of communities and are being used increasingly to assess ecological health. Numerous methods exist for estimating size spectra, each of which introduces its own unique systematic biases making direct comparisons of results difficult. But no studies have been peformed to examine if different methodologies are able to detect differences in size spectra parameters, or if the *relative* differences within a method are consistent. Therefore, we set out to answer this question to guide future research, as well as to leverage previously published results to synthesize knowledge of how size spectra relationships respond across studies.

### attempt 2
Community body size distributions are a fundamental characteristic of natural communities related to energy availability and trophic transfer efficiency. Numerous methodologies for estimating size spectra relationships have been proposed and compared, with maximum likelihood estimates generally being the most accurate. However, to date, no study has been conducted to assess which methodologies are able to detect known differences in size spectra parameters (slope or exponent). The present study addresses this shortfall by analyzing data using common methods on data simulated from know distributions.

## Motivation

Size spectra relationships describe the negative pattern of abundance with increasing body size. Communities across habitats, biomes, and scales have shown remarkably similar patterns, leading to body size distributions being proposed as a "universal" relationship (Pethcey and Belgrano 2010). Furthermore, deviations from this expected pattern have been demonstrated in communities which have experienced extreme anthropogenic stress, including from overfishing marine communities in the North Sea (Blanchard/Jennings (always forget which one)), and stream macroinvertebrate communities from acid mine drainage in New Zealand (Pomeranz et al. 2018). The generally consistent pattern in undisturbed communities and response to stressors of size spectra is increasing interest in the use of size spectra as a universal indicator in ecology.

Numerous methodologies have been proposed for estimating size spectra relationships. Traditionally, body sizes (M) have been grouped together in bins, and an ordinary least squares regression on the abundance (N) within each bin has been conducted in log-log space to estimate the slope (exponent) of decreasing N in response to increasing M. Many binning methods have been proposed and used previously in the literature. However, numerous studies have shown that binning methods introduce systematic error, and are generally do not provide accurate estimates of the slope in log-log space, or the exponent in real space. Maximum likelihood estimates of the exponent describing the power law have generally been shown to provide the most accurate estimate.

Despite the increasing consensus of MLE analyses, no study has yet been conducted to assess which method, if any, can accurately detect differences in parameters describing size spectra relationships. Knowing which methods are able to detect differences is an important goal for ecologists moving forward, particularly with the increasing interest in using size spectra relationships to characterize communities. Additionally, having an understanding of how previous estimates of size spectra parameters vary based on methodologies will allow future studies to leverage previously published results. Because of the variation in estimates based on methodologies, it is currently not possible to conduct meta-analyses synthesizing previously published data.

## methods

1. Simulate data
  + sizeSpectra::rPLB(n = 1000, b, xmin, xmax)
  + b = c(-1.5, -2, -2.5); b = c(-1.9, -2, -2.1)
  + xmin and xmax based on macroinvertebrate sizes
  + Also use Edwards values to test sensitivity
  + Simulate from other distributions? rlnorm(), rweibull() etc.?
2. Analyze simulated data
  + MLE methods from Edwards et al. 2018
  + log2 bins - Pomeranz, Blanchard, etc
  + ELB - 6, 8, 10 bins. Woodward's group generally uses 6 equal bins
  + Others from Edwards' 8 methods?
3. assessment
  + which methods provide best estimate of *real* slopes?
  + what is *relative* difference in estimates? i.e. do all methods result in a relative change of 0.5 between data "groups"?

### 2022-05-24 MLE_tidy.R

@@ -1,0 +1,49 @@
+MLE_tidy <- function(df, rsp_var){
+  # define variables
+  x <- df[[rsp_var]]
+  xmin = min(x)
+  xmax = max(x)
+  log.x = log(x)
+  sum.log.x = sum(log.x)
+  
+  # initial starting point for parameter estimate
+  PL.bMLE = 1/(log(min(x)) - sum.log.x/length(x)) - 1
+  
+  # non-linear minimization  
+  PLB.minLL = nlm(negLL.PLB, 
+                  p = PL.bMLE, x = x, n = length(x), 
+                  xmin = xmin, xmax = xmax,
+                  sumlogx = sum.log.x)
+  
+  # estimate for b
+  PLB.bMLE = PLB.minLL$estimate
+  # minimum estimate of b
+  PLB.minNegLL = PLB.minLL$minimum
+  
+  ## 95% CI calculation
+  bvec = seq(PLB.bMLE - 0.5, PLB.bMLE + 0.5, 1e-05)
+  PLB.LLvals = vector(length = length(bvec))
+  for (i in 1:length(bvec)) {
+    PLB.LLvals[i] = negLL.PLB(bvec[i],
+                              x = x,
+                              n = length(x), 
+                              xmin = xmin,
+                              xmax = xmax,
+                              sumlogx = sum.log.x)
+  }
+  critVal = PLB.minNegLL + qchisq(0.95, 1)/2
+  bIn95 = bvec[PLB.LLvals < critVal]
+  # confidence interval
+  PLB.MLE.bConf = c(min(bIn95), max(bIn95))
+  if (PLB.MLE.bConf[1] == min(bvec) | 
+      PLB.MLE.bConf[2] == max(bvec)) {
+    dev.new()
+    plot(bvec, PLB.LLvals)
+    abline(h = critVal, col = "red")
+    stop("Need to make bvec larger - see R window")
+  }
+  # return b estimate and min/max 95% CI
+  return(data.frame(b = PLB.bMLE,
+                    minCI = min(bIn95),
+                    maxCI = max(bIn95)))
+}

### 2022-05-24 custom_functions.R

@@ -1,0 +1,187 @@
+# custom functions
+# these are largely copied from Pomeranz et al. 2018 Freshwater Biology and Pomeranz et al. 2022 Global Change Biology
+# copied 5/24/22 --> go through these and make sure doing what I think I remember them doing. 
+
+# MLE in tidy format
+MLE_tidy <- function(df, rsp_var){
+  # define variables
+  x <- df[[rsp_var]]
+  xmin = min(x)
+  xmax = max(x)
+  log.x = log(x)
+  sum.log.x = sum(log.x)
+  
+  # initial starting point for parameter estimate
+  PL.bMLE = 1/(log(min(x)) - sum.log.x/length(x)) - 1
+  
+  # non-linear minimization  
+  PLB.minLL = nlm(negLL.PLB, 
+                  p = PL.bMLE, x = x, n = length(x), 
+                  xmin = xmin, xmax = xmax,
+                  sumlogx = sum.log.x)
+  
+  # estimate for b
+  PLB.bMLE = PLB.minLL$estimate
+  # minimum estimate of b
+  PLB.minNegLL = PLB.minLL$minimum
+  
+  ## 95% CI calculation
+  bvec = seq(PLB.bMLE - 0.5, PLB.bMLE + 0.5, 1e-05)
+  PLB.LLvals = vector(length = length(bvec))
+  for (i in 1:length(bvec)) {
+    PLB.LLvals[i] = negLL.PLB(bvec[i],
+                              x = x,
+                              n = length(x), 
+                              xmin = xmin,
+                              xmax = xmax,
+                              sumlogx = sum.log.x)
+  }
+  critVal = PLB.minNegLL + qchisq(0.95, 1)/2
+  bIn95 = bvec[PLB.LLvals < critVal]
+  # confidence interval
+  PLB.MLE.bConf = c(min(bIn95), max(bIn95))
+  if (PLB.MLE.bConf[1] == min(bvec) | 
+      PLB.MLE.bConf[2] == max(bvec)) {
+    dev.new()
+    plot(bvec, PLB.LLvals)
+    abline(h = critVal, col = "red")
+    stop("Need to make bvec larger - see R window")
+  }
+  # return b estimate and min/max 95% CI
+  return(data.frame(b = PLB.bMLE,
+                    minCI = min(bIn95),
+                    maxCI = max(bIn95)))
+}
+# binning method
+bin_and_center <- function(data, var, breaks, ...){
+  # data is a data frame
[…]

### 2022-05-24 data_simulation.R

@@ -1,0 +1,239 @@
+# script 1
+# simulate data sets
+# analyze using different methods
+# Compare results
+
+
+# Preliminaries -----------------------------------------------------------
+
+library(tidyverse)
+library(sizeSpectra)
+source("custom_functions.R")
+
+
+# Simulate ----------------------------------------------------------------
+
+
+n <-  1000
+b <- c(-1.5, -2, -2.5)
+x_15 <- rPLB(n = n, b = b[1])
+x_20 <- rPLB(n = n, b = b[2])
+x_25 <- rPLB(n = n, b = b[3])
+x_sample <- c(x_15, x_20, x_25)
+
+df <- tibble(known_b = 
+               rep(c(-1.5, -2, -2.5),
+                   each = n),
+             n = n,
+             name = 
+               rep(c("m_invert_1.5",
+                     "m_invert_2.0",
+                     "m_invert_2.5"),
+                   each = n))
+df$m <- x_sample
+
+# just MLE estimate
+# mle_lambda <- df %>%
+#   group_by(known_b) %>%
+#   nest() %>%
+#   mutate(lambda = map(data,
+#                       MLE_tidy,
+#                       "m")) %>%
+#   unnest(cols = lambda) %>%
+#   select(-data) %>%
+#   ungroup()
+
+# compare slopes method from Pomeranz et al. 2022
+dw_range = range(df$m, na.rm = TRUE)
+
+(test_method <- df %>%
+  # mutate(date = as.Date(collectDate)) %>%
+  group_by(known_b) %>%
+  #filter(siteID == "CARI") %>%
+  # create list-column
+  nest() %>% 
+  mutate(method_compare = 
+           map(data,
+               compare_slopes,
+               rsp_var = "m",
+               dw_range = dw_range)) %>%
+  ungroup() %>%
+  select(-data) %>%
+  unnest(cols = method_compare))
+
+# Scale up simulation ####
+
+n <-  1000
+b <- c(-1.5, -2, -2.5)
+rep <- 100
+sim_out <- list()
+#df <- tibble()
+
+for(i in 1:rep){
+  df <- tibble(known_b = 
+                 rep(b,
+                     each = n),
+               n = n)
+  x_15 <- rPLB(n = n, b = b[1])
+  x_20 <- rPLB(n = n, b = b[2])
+  x_25 <- rPLB(n = n, b = b[3])
+  x_sample <- c(x_15, x_20, x_25)
+  df$m <- x_sample
+  df$rep <- i
[…]

### 2022-06-13 Manuscript.Rmd
update simulation and analysis scripts

---
title: "Manuscript"
author: "Justin Pomeranz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
output:
  pdf_document: 
    toc: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE)
```

**NOTE** I need to integrate reference manager with R markdown. Any favorites or recommendations?? I generally use Zotero (free), but think I have access to Mendeley through CMU...

**NOTE** I re-ran all of these after setting the seed for reproducibility. Figures should be up to date, but the numbers may change slightly i.e., 94.8% might become 95.1%, mean $\pm$ SD values will be different.

# Abstract

1.  Size spectra represent a fundamental attribute of community organization and are increasingly being used in assessments of fresh waters.
2.  Many methods have been proposed for constructing size spectra relationships, but recent work has shown that varied methods return biased estimates of relationship parameters, and different methods in fact are not estimating the same parameter. Despite this variability in estimates, it is unclear if the relative change across environmental gradients is consistent across methodologies. Here, we simulate data sets across an hypothetical environmental gradient and estimate the size spectra parameter (slope or exponent, depending on method) at each site, and, importantly, estimate the relationship of parameters across the hypothetical gradient. We also use two previously published body size datasets across an anthropogenic stress gradient and an environmental temperature gradient and assess how the conclusions of these studies would vary based on methods used.
3.  We find that maximum likelihood methods always perform as well or better than common binning methods. Additionally, the variance in estimates using MLE methods is markedly reduced when compared to binning methods.
4.  The uncertainty and variation in estimates when using binning methods is often greater than or equal to the variation previously published in experimental and observational studies, bringing into question the effect size of previously published results. However, re-analysis of two previously published datasets does not markedly alter the conclusions reached. Further study is needed to identify how and when these estimates may be biased, and when the general pattern of results can be considered to be "true".

# Introduction

Body size distributions are a fundamental characteristic of communities. The remarkable consistency of these relationships across spatiotemporal scales and habitat types has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010). Size metrics have commonly been used in marine systems, and are increasingly being applied to assess the condition of freshwater ecosystems (Martinez et al. 2016, Pomeranz 2018).

Individual size distributions (ISD, also referred to as abundance size spectra) are commonly used. Generally, there is a negative relationship between body size (M) and abundance (N). Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$. Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated from OLS regressions of log transformed data of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$. Myriad binning methods have been proposed, including linear and logarithmic bin widths. Likewise, methods can rely on the absolute count in the bins as well as normalization techniques where the count is divided by the bin width (especially common with logarithmic binning). Alternatively, $\lambda$ can be estimated directly using maximum likelihood techniques.

Previous work has shown that the estimates of $\lambda$ differ between MLE and OLS techniques. OLS methods are particularly sensitive to decisions made in the binning process. Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ (White et al. 2007, Edwards et al. 2017), and reanalysis of empirical datasets also indicates that the conclusions are dependent on the methodology used (White et al. 2007, Edwards et al. 2020). However, recent empirical analysis of stream macroinvertebrate communities across the NAtional Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used (Pomeranz et al. 2022). While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time. In other words, if the data within a study are all treated the same, does a relative change of OLS slope parameters of 0.1 coincide with a relative change of MLE estimates of 0.1? In order to answer this question, we simulate body size observations from bounded power law distributions with varied $\lambda$ exponents across a hypothetical environmental gradient and compare the results obtained using three different methodologies common in the literature. In addition, we re-analyze two previously published datasets of stream community body sizes across a stress and environmental gradient. We find that the MLE method more accurately estimates the site-specific $\lambda$ exponents as well as the relative change in exponents across the hypothetical gradient, as well as having smaller variation. The logarthmic binning methods generally perform well, but have larger confidence intervals around the estimates. However, the overall conclusions of previously published empirical data sets are not dependent on the method used.

# Methods

## Data Simulation

In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in Edwards et al. (2017). Let $M$ be a random variable of body sizes described by the probability density function:

$$f(M) = CM^\lambda, M_{min} \le M \le M_{max} $$

$M_{min} = 0.0026$ and $M_{max} = 1.2 *10^3$. These values are based on empirical body sizes of stream benthic communities reported in Pomeranz et al. (2020). Our results are not dependent on the range of body sizes, and we show the results of other ranges in the supplemental information.

For the main analysis, presented here, we sampled $n = 1000$ body sizes from distributions described by five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The supplemental information contains results when the value of n is varied (or maybe include in main?). Each value of $\lambda$ was assumed to come from a community across a hypothetical environmental gradient $X$. Values of $X$ for each community are uniformly distributed from -1 to 1. We repeated the data simulation 1000 times (reps). The main results presented here were not dependent on the range of *x*-values or the number of sites (supplemental information).

### Sample size, n

The number of observations in our simulations may bias the results. Therefore, we repeated the simulations described above, but varied the sample size $n$. We tested values of $n = 200, 500, 1000, 5000, 10 000$. (originally had n=100, but thre errors based on random seed. Bumped it to 200 to try and get consistent error-free runs)

### "No" relationship

It is possible that our sampling or simulation framework could impact the results obtained, possibly by inflating the type I error rate. In order to assess whether or not our simulations are robust to this, we performed independent random samples from a bounded power law for five communities across a hypothetical gradient as described above, but we set $\lambda = -2$ for all five communities. Hence, we are simulating a scenario where the size spectra relationship is invariant to our hypothetical environmental gradient, and we should expect the estimated relationship coefficient $\beta_1 = 0$.

## Finer resolution of $\lambda$ values

For the main analysis, we were interested in determining whether or not the methods were able to detect relatively coarse changes to size spectra parameters, specifically, a change of 0.5 units across each value of $X$, or an absolute change or 1.0 units. However, many seminal works on variation in size spectra relationships in freshwater are on a much finer scale (on the order 0f 0.1 to 0.25 units). Therefore, we performed simulations as described above, but varied $\lambda$ from -1.9 to -2.1 across the hypothetical environmental gradient. This equates to a know relationship coefficient of $\beta_1 = 0.1$ across the hypothetical gradient, and an absolute change of 0.2 units.

## Estimation of Size spectra parameter $\lambda$

For each sample of $M$ and each of the 1000 replicates, we estimated the exponent $\lambda$ using MLE methods modified from the the `sizeSpectra` package (Edwards et al. 2020). In addition, we use two common binning methods to estimate the OLS slope parameter in log-log space. For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes. The count in each bin was normalized by dividing by the bin width. This method has been used by (refs here, I think a lot of Woodward's group uses this...) Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn. The second method was similar to ELBn, but bins of Log~2~ widths are used. The count in each bin is normalized in the same way. This method has been used by (Pomeranz et al. 2018, I **think** McGarvey \~2018, Blanchard (log 2, ?normalized?)) and is referred to as the Normalized Abundance Spectrum (NAS).

## Estimation of relationship across gradient, $\beta_1$

For each simulation replicate, we estimated how the size spectra parameters varied across the hypothetical gradient. Simple OLS regression were conducted in the form $\lambda_{estimate} ~ \beta_0 + \beta_1 * X$. The distribution of the relationship coefficient, $\beta_1$, were plotted compared to the known relationship. Likewise, the confidence interval of $\beta_1$ was assessed to determine the proportion of estimates which contained the true value of the known relationship.

## Empirical Data

We re-analyze two datasets of benthic macroinvertebrate communities from stream habitats across two different gradients. In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage stress gradient. Details of the sample collection and processing can be found in Pomeranz et al. (2018). Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope. Body mass was estimated using published length weight regressions.

The second data set was from the wadeable stream sites of the National Ecological Observatory Network (NEON; **data product XXX**). These sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska. Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat. All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions. This data has been analyzed previously using size spectra methods as described in Pomeranz et al. (2022). Detailed methods of the sampling collection and data processing methods can be found on the NEON website (URL, or macroinvertebrate DPI pubs).

# Results

## Relationship across the hypothetical environmental gradient

![Relationship estimates across the hypothetical gradient for each replicate. Each panel is a different method for estimating the size spectra parameter.](figures/PLB_sim_main.png)

All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient. The regressions using the MLE method found a significant relationship 100% of the time, while the ELBn and NAS method found significant relatoinships 95.8% and 96.9% of the time, respectively.

The confidence intervals for the relationship coefficients ($\beta_1$) for both the MLE and ELBn method contained the true value of the known relationship across the gradient $\ge$ 95% of the time, whereas the confidence interval for the NAS method only had the true value 83.1% of the time. Despite having similar accuracy, the width of the CIs for the ELBn method were more than 3 times that of the MLE method (ELBn_CI_ = 0.386 $\pm$ 0.20; MLE_CI_ = 0.121 $\pm$ 0.056). The CI's for the NAS were slightly smaller than the ELBn method, but still $>$ 2.5 times as wide as the MLE method (NAS_CI = 0.339 $\pm$ 0.172).

On average, the relationship across the gradient was under estimated: MLE -0.002 ($\pm$ 0.022); ELBn -0.039 ($\pm$ 0.062); and NA -0.081 ($\pm$ 0.061).

![Distribution of relationship coefficient estimates. Vertical line is the known relationship. All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.](figures/PLB_sim_relationship_density.png)

Interestingly, there was an interactive effect of the estimate accuracy between sample size and the value of $\lambda$. All methods were more accurate with larger sample sizes, and smaller values of $\lambda$.

![Distribution of size spectra parameter estimates. Vertical line is the known parameter wich describes the bounded power law distribution from which the body size estimates were sampled. As n increases (top to bottom) and $\lambda$ increases (left to right), the accuracy of the estimate improves across all methods.](figures/n_vary_est_b.png)

![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).](figures/n_vary_main.png)

## No relationship

All of the methods performed similarly when there was no relationship across the hypothetical gradient. The Type I error rate for MLE, ELBn and NAS were 5%, 5.7% and 4.6%, respectively. However, the confidence intervals for the binning methods were \~3 times as wide as the MLE method (mean CI widths: MLE = 0.116, ELBn = 0.372; NAS = 0.332).

Of the relationships which were estimated to be significant, the MLE, ELBn, and NAS method had a 58%, 49.1% and 54.3% probability of indicating a negative relationship, respectively.

![Individual regression estimates when no relationship exists across the hypothetical environmental gradient. The methods are separated by rows, and the left and right column show relationships which were non-significant and significant, respectively. The individual regressions coefficients are colored-coded to indicate positive and negative relationship estimates. ](figures/static_b_lambda_spaghetti.png)

## Small variation in lambda

The performance of all methods declined when trying to detect variation in the $\lambda$ parameter between -1.9 to -2.1 across the environmental gradient. The mean $\beta_1$ coefficient estimates for all methods were closely associated with the known relationship. Once again the variation in the estimates using the MLE method were much smaller that the estimates using both of the binning methods (Figure 4).

![Mean difference between relationship estimate and known relationship coefficient for each method. Here, $\lambda$ varied from -1.9 to -2.1, with a kown relationaship of -0.1 across the gradient.](figures/PLB_sim_small_relationship_density.png)

Of the 1000 simulation replicates, the MLE only detected a significant relationship 90% of the time, whereas the ELBn and NAS methods detected significant relationships 19 and 23% of the time, respectively. Of the replicates which were significant, the confidence interval contained the true value of the relationship 94.6, 88.4, and 89.1% of the time for the MLE, ELBn, and NAS methods, respectively.

![Linear regressions across the hypothetical gradient when lambda varies from -1.9 to -2.1. Each method is displayed in the rows, and the left colum shows the non-significant estimates, while the right column shows the relationships which were signficant ($\beta_1$ coefficient p-value \< 0.05). The colors indicate when the beta coefficient estimate was negative (pink) and when it was positive (purple).](figures/small_lambda_spaghetti.png)

Once again, the width of the confidence intervals for both of the binning methods were $\ge$ 2.8 times that of the MLE method (mean $\pm$ SD MLE CI width: 0.117 $\pm$ 0.051; ELBn CI width: 0.378 $\pm$ 0.169; NAS CI width: 0.332 $\pm$ 0.150).

On average, the relationship estimates across the gradient were similar to the know value (mean deviation: MLE = 0.00001, ELBn = -0.009, NAS = -0.012). However, when only looking at the relationships which were significant, this variation increased considerably(MLE = -0.002, ELBN = -0.052, NAS = 0.034).

## Empirical data

For both empirical datasets, the direction and magnitude of chnage (i.e. $\beta_1$ coefficients) are generally in agreement. Size spectra parameters consistently increase (become flatter) in the AMD data (Fig. XX A). Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig XX D)

**NOTE** combine all the figures below into one with 4 panels (AMD = A, B; NEON = C, D)

![A](figures/AMD_plot.png)

![B.Relationship estimates across AMD gradient](figures/amd_relationship.png)

![C.](figures/NEON_plot.png)

![D.](figures/neon_relationship.png)

Because the $\beta_1$ coefficient estimates are similar, and the range of the gradient in the AMD data is relatively small (9.5), the absolute change in the size spectra parameter across the AMD gradient are similar regardless of method used (range: 0.59 to 0.72). Likewise, the absolute change in size spectra parameters across the NEON temperature gradient ranges from 0.06 to 0.165, depending on the method used.

*alternative* absolute difference of \~0.1 units in $\lambda$ across gradients is similar to differences reported in other studies (seasonally: 0.16, Mcgarvey and Kirk; landuse: 0.15 Martinez; O'Gorman 2017 \~0.25, Yvon and Dossena papers \~0.1? [need to redownload some of these and compare])

## Simulating shallower lambdas

Estimates of the relationship between lambda and a hypothetical gradient varied depending on the method used. However, interpretations of empirical data were broadly consistent across methodologies. Upon closer inspection, the values of size spectra parameters in the empirical data were considerably shallower than -2. Given that we found the performance of all methods increased with shallower $\lambda$'s we wanted to investigate how the methods performed with simulated values more similar to the empirical estimates of parameters describing size spectra relationships. Therefore, we repeated the simulation process as in the main analysis, but used $\lambda$ values ranging from -1.1 to -1.5.

![Relationship estimates across the hypothetical gradient for each replicate. Each panel is a different method for estimating the size spectra parameter.](figures/shallow_lambda_main.png)

We found generally less variability in the relationship estimates across a gradient of distributions described with shallower $\lambda$ parameters. All of the relationship coefficient estimates ($\beta_1$)'s across all replicates were significant regardless of method used. The confidence interval for $\beta_1$ estimate for the MLE method contained the true value 94.7% of the time. However, the true value was in the confidence interval for the ELBN and NAS method only 90.9% and 76.9% of the time, respectively. Once again, the mean width of the CI for the binning methods were ~2 times as large as for the MLE method. Mean $\pm$ SD CI width: MLE = 0.0442 $\pm$ 0.0196; ELBn = 0.0920 $\pm$ 0.0482, NAS = 0.0798 $\pm$ 0.0324.

On average, the relationship across the gradient was under estimated, but by nearly an order of magnitude less than when simulating across steeper values of $\lambda$. Differences between known and estimated relationship coefficients: MLE -0.0007 ($\pm$ 0.0079); ELBn -0.0058 ($\pm$ 0.0177); NAS -0.0251 ($\pm$ 0.0124).

![Distribution of relationship coefficient estimates. Vertical line is the known relationship. All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.](figures/shallow_lambda_relationship_density.png)

# Discussion

Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology. Previous work has investigated the accuracy and inherent biases associated with different estimation methods. However, how these innaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detecting variation in size spectra parameters across environmental gradients with confidence. Here, we sampled body sizes from known distributions and estimated their parameters across a hypothetical gradient in order to assess the relationship coefficients obtained.

Generally, there was large agreement in the direction and magnitude of average relationship coefficient estimates across methods. The estimate from the MLE method was always closer to the known relationship, and always had the smallest variation around the estimate.

All of the methods performed equally well when the $\lambda$ value did not vary across the hypothetical gradient, with a Type I error rate of \~5%, exactly as would be expected under the assumptions of frequentist statistics. This indicates that previous studies which have detected no difference in size spectra parameters across sites can be interpreted with confidence, regardless of the method used.

The performance of all methods improved as sample size increases, and as the exponent of the size spectra relationship gets larger (more shallow). As either or both of these variables change, the difference in estimated relationship coefficients declines across the methods. This is particularly interesting given the fact that both empirical data sets of stream communities examined here are shallower than the expected relationship of $\lambda = -2$, and the conclusions of the change in size spectra relationships across the gradients are not dependent on method used. If communities under study are in fact described by shallower exponents, the method used may not be critical in the conclusions reached. However, the MLE method performs as well or better than the two binning methods examined here under all contexts and should be the preferred method moving forward. At a minimum, future studies should report MLE estimates to ensure that the results are not dependent on the methodology used.

Alarmingly, the performance of all methods declined dramatically when trying to detect smaller variation in $\lambda$ values. We simulated observations from distributions varying by an absolute value of 0.2 units. This range of variation is similar to that which has been reported in many seminal works on variation in size spectra parameters. When using the MLE method, significant relationships were only detected \~88% of the time, whereas both of the binning methods only detected significant trends \~20% of the time. Of the simulation replicates which were significant, nearly all of them were negative across all methods except for one single replicate using the NAS method which was positive (Figure. 5).

Despite the drop in performance with reduced variation in $\lambda$ values, when a significant relationship was observed, it was generally in the correct direction and of a similar magnitude. This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible, and the magnitude of the relative change is a reasonable estimate. Given that all of the data within a study is treated identically, the the over all chnage in size spectra parameters is likely reasonable. However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to compare the relative changes across different published studies which use different methods. Publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental stressors and perturbations.

## Concluding Remarks

We reiterate the recommendations of White et al. (2007), Sprules and Barth (2016) and Edwards et al. (2017) to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context. Furthermore, we strongly encourage authors to publish individual size data whenever possible. This will allow for the consistent re-analysis of existing data sets as methodologies develop and improve. This will aid in the ability for size spectra work to be synthesized between research groups and across scales.

\newpage

# Supplementary material {.unnumbered}

\beginsupplement

## Range of body sizes, $M$

![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100. ](figures/PLB_small_m_main.png)

![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_est_b_density.png)

![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_relationship_density.png)

## Large environmental gradient

![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_main.png)

![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values. Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_est_b_density.png)

![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5. Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_relationship_density.png)

## Varying number of sites

### 10 sites

![Individual regressions for ten sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis](figures/PLB_10_sites_main.png)

![Distribution of estimated $\lambda$ coefficient for ten sites across a hypothetical gradient with known values. All other parameters are the same as in the main analysis](figures/PLB_10_sites_est_b_density.png)

![Distribution of estimated relationship ($\beta_1$) coefficient's for ten sites across a hypothetical gradient with known value of 0.5. All other parameters are the same as in the main analysis](figures/PLB_10_sites_relationship_density.png)

### Three sites 
![Individual regressions for three sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis.](figures/PLB_3_sites_main.png)

![Distribution of estimated $\lambda$ coefficient for three sites across a hypothetical gradient with known values. All other parameters are the same as in the main analysis.](figures/PLB_3_sites_est_b_density.png)

![Distribution of estimated relationship ($\beta_1$) coefficient's for three sites across a hypothetical gradient with known value of 0.5. All other parameters are the same as in the main analysis](figures/PLB_3_sites_relationship_density.png)

## Sample size

**note** These plots still have n = 100. Need to re run and update if we end up going with n = 200

![Plot showing the relationships between sample size and the estimated $\lambda$ parameters. the dashed line indicates the known value of $\lambda$.](figures/lambda_n_5_sites.png)

### 2022-06-13 1_simulations.R

@@ -1,0 +1,352 @@
+# Main simulation
+
+
+## this takes about 4 hours to run...
+# I'm sure the time could be improved with the use of parallel computing
+# But I have a 4 month old and let the sims run over night instead of going down an optimizing/troubleshooting rabbit hole. 
+# probably less time efficient in the long run but {shruggy guy emoji}
+
+
+## added "possibly()" command to custom functions
+# load PLB_sim, save as something else and rerun simulation
+# Does old (pre-possibly()) data match new data with possibly()?
+# Need to re-run all sims with modified functions??
+
+## Possibly didn't seem to work, chnaged n=100 to n=200 and it seems to be good. Need to go through and delete comments / clean it all up. 
+
+# for reproducibility
+set.seed(2806)
+
+# Preliminaries -----------------------------------------------------------
+
+library(tidyverse)
+library(sizeSpectra) #bounded power law and MLE functions
+source("custom_functions.R")
+
+# Simulation and summary --------------------------------------------------
+
+# PLB ####
+# lambda parameter varies between -1.5 and -2.5 across a hypothetical gradient from -1 to 1. i.e. relationship of beta = 0.5
+# setting main analysis to 5 sites
+lambda = seq(from = -1.5, to = -2.5, length.out = 5)
+env_gradient = seq(from = -1, to = 1, length.out = 5)
+
+lambda
+env_gradient
+
+rep = 1000
+# body sizes in NEON data set varies from 0.0026 to 1.2e4
+# setting sample range based on that minimum, but reducing max by order of magnitude
+m_lower = 0.0026
+m_upper = 1.2e3
+
+PLB_sim <- sim_result(n = 1000,
+                      b = lambda,
+                      env_gradient = env_gradient,
+                      rep = rep, 
+                      m_lower = m_lower,
+                      m_upper = m_upper,
+                      distribution = "PLB")
+
+saveRDS(PLB_sim, file = paste0("data_sim/",
+                               substitute(PLB_sim),
+                               "_dat.rds"))
+plot_sim(PLB_sim)
+#rm(PLB_sim)
+
+# PLB no relationship
+# for reproducibility
+set.seed(2806)
+
+PLB_static_b <- sim_result(n = 1000,
+                           b = rep(-2, length(env_gradient)),
+                           env_gradient = env_gradient,
+                           rep = rep, 
[…]

### 2022-06-13 analysis.R

@@ -1,89 +1,0 @@
-# Cleaned up simulation
-
-# Preliminaries -----------------------------------------------------------
-
-library(tidyverse)
-library(sizeSpectra) #bounded power law and MLE functions
-source("custom_functions.R")
-
-
-# Simulation and summary --------------------------------------------------
-
-# PLB ####
-rep = 1000
-m_lower = 0.0026
-m_upper = 1.2e4
-
-start_reg = proc.time()
-PLB_sim <- sim_result(b = c(-1.5, -2, -2.5),
-                      env_gradient = c(-1, 0, 1),
-                      rep = rep, 
-                      m_lower = m_lower,
-                      m_upper = m_upper,
-                      distribution = "PLB")
-plot_sim(PLB_sim)
-rm(PLB_sim)
-
-# PLB no relationship
-PLB_static_b <- sim_result(b = c(-2, -2, -2),
-                      env_gradient = c(-1, 0, 1),
-                      rep = rep, 
-                      m_lower = m_lower,
-                      m_upper = m_upper,
-                      distribution = "PLB")
-plot_sim(PLB_static_b)
-rm(PLB_static_b)
-
-proc.time() - start_reg
-
-# PLB "small" relationship
-PLB_sim_small <- sim_result(b = c(-1.9, -2, -2.1),
-                      env_gradient = c(-1, 0, 1),
-                      rep = rep, 
-                      m_lower = m_lower,
-                      m_upper = m_upper,
-                      distribution = "PLB")
-plot_sim(PLB_sim_small)
-rm(PLB_sim_small)
-
-# sensitivity, supplementary information ----------------------------------
-
-# different scale of env_gradient (predictor, x)
-PLB_large_x <- sim_result(b = c(-1.5, -2, -2.5),
-                          env_gradient = c(-100, 0, 100),
-                          rep = rep, 
-                          m_lower = m_lower,
-                          m_upper = m_upper,
-                          distribution = "PLB")
-
-plot_sim(PLB_large_x)
-rm(PLB_large_x)
-
-# change range of m sampled in rPLB
-PLB_small_m <- sim_result(b = c(-1.5, -2, -2.5),
-                          env_gradient = c(-1, 0, 1),
-                          rep = rep, 
-                          m_lower = 1,
-                          m_upper = 100,
-                          distribution = "PLB")
-
-plot_sim(PLB_small_m)
-rm(PLB_small_m)
-
[…]

### 2022-06-13 custom_functions.R

@@ -1,8 +1,14 @@
 # custom functions
 # these are largely copied from Pomeranz et al. 2018 Freshwater Biology and Pomeranz et al. 2022 Global Change Biology
-# copied 5/24/22 --> go through these and make sure doing what I think I remember them doing. 
+# copied 5/24/22 --> go through these and make sure doing what I think I remember them doing.
+
+# These functions require the tidyverse (Wickham et al.) and sizeSpectra (Edwards et al. 207, 2020)
 
 # MLE in tidy format
+# this is a wrapper function I wrote to be able to use the MLE functions from size Spectra, but used on data in the tidy format. 
+# it was designed to specifically work with list-columns of the data you want to analyze. 
+# I make no guarantees on it's functionality 
+
 MLE_tidy <- function(df, rsp_var){
   # define variables
   x <- df[[rsp_var]]
@@ -16,8 +22,11 @@ MLE_tidy <- function(df, rsp_var){
   
   # non-linear minimization  
   PLB.minLL = nlm(negLL.PLB, 
-                  p = PL.bMLE, x = x, n = length(x), 
-                  xmin = xmin, xmax = xmax,
+                  p = PL.bMLE,
+                  x = x,
+                  n = length(x), 
+                  xmin = xmin,
+                  xmax = xmax,
                   sumlogx = sum.log.x)
   
   # estimate for b
@@ -203,11 +212,14 @@ sim_result <- function(n = 1000,
                        m_upper,
                        distribution){
   stopifnot(length(b) == length(env_gradient))
+  known_relationship <- -(max(b) - min(b)) / 
+    (max(env_gradient) - min(env_gradient)) 
   # simulate values from distribution
   sim_out <- list()
   for(i in 1:rep){
     df <- tibble(
       known_b = rep(b, each = n),
+      known_relationship = known_relationship,
       env_gradient = rep(env_gradient, each = n),
       n = n,
       distribution = distribution,
@@ -246,13 +258,18 @@ sim_result <- function(n = 1000,
   out <- sim_df %>%
     group_by(rep,
              known_b,
+             known_relationship,
              env_gradient,
              distribution,
              m_lower,
              m_upper) %>%
     nest() %>% 
     mutate(method_compare = 
-             map(data, # change map() to furrr::future_map()???
+             map(data, 
+                 # added the possibly() function on 6/12/2022
[…]

### 2022-06-13 lambda_pairwise.R

@@ -1,0 +1,105 @@
+# pairwise estimates of lambda
+
+dat <- readRDS("data_sim/PLB_shallow_lambda_dat.rds")
+
+names(dat)
+
+# estimate the relationship (beta_1_ estimate) across the gradient
+relationship_estimate <- dat %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
+               ~lm(estimate ~ env_gradient, data = .x))) %>%
+  mutate(tidied = map(lm_mod, broom::tidy)) %>%
+  unnest(tidied) %>%
+  filter(term == "env_gradient") %>%
+  select(-data, -lm_mod, -statistic)
+
+# what is the pairwise difference in spectra parameter estimates?
+lambda_diff <- dat %>%
+  select(env_gradient, rep, name, estimate) %>%
+  pivot_wider(names_from = name, values_from = estimate) %>%
+  mutate(MLE_NAS = MLE - NAS,
+         MLE_ELB = MLE - ELBn, 
+         NAS_ELB = NAS - ELBn)
+
+lambda_diff %>%
+  pivot_longer(cols = MLE_NAS:NAS_ELB) %>%
+  ggplot(aes(y = ..scaled..,
+             x = value,
+             fill = name)) +
+  geom_density(alpha = 0.55,
+               adjust = 2) +
+  geom_vline(aes(xintercept = 0),
+             linetype = "dashed",
+             color = "black",
+             size = 1) +
+  theme_bw() +
+  scale_fill_viridis_d(option = "A", direction = -1)
+
+
+# proportion of relationships which are significant
+relationship_estimate %>%
+  filter(p.value < 0.05) %>%
+  group_by(name) %>%
+  summarize(n() / 1000)
+
+# how far does the estimated relationship deviate from the known relationship?
+relationship_estimate %>%
+  mutate(diff = known_relationship - estimate) %>%
+  group_by(name) %>%
+  summarize(mean(diff), 
+            sd(diff))
+
+
+dat_conf <- dat %>%
+  #filter(rep <= 10) %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
+               ~lm(estimate ~ env_gradient, data = .x))) %>%
+  mutate(conf_low = map(lm_mod, 
[…]

### 2022-06-13 main_analysis_CI_plot.R

@@ -1,0 +1,91 @@
+# main analysis CI plot
+
+library(tidyverse)
+library(sizeSpectra) #bounded power law and MLE functions
+source("custom_functions.R")
+
+# main simulation data
+dat <- readRDS("data_sim/PLB_sim_dat.rds")
+
+names(dat)
+
+# estimate the relationship (beta_1_ estimate) across the gradient
+relationship_estimate <- dat %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
+               ~lm(estimate ~ env_gradient, data = .x))) %>%
+  mutate(tidied = map(lm_mod, broom::tidy)) %>%
+  unnest(tidied) %>%
+  filter(term == "env_gradient") %>%
+  select(-data, -lm_mod, -statistic)
+
+# what is the pairwise difference in spectra parameter estimates?
+lambda_diff <- dat %>%
+  select(env_gradient, rep, name, estimate) %>%
+  pivot_wider(names_from = name, values_from = estimate) %>%
+  mutate(MLE_NAS = MLE - NAS,
+         MLE_ELB = MLE - ELBn, 
+         NAS_ELB = NAS - ELBn)
+
+lambda_diff %>%
+  pivot_longer(cols = MLE_NAS:NAS_ELB) %>%
+  ggplot(aes(y = ..scaled..,
+             x = value,
+             fill = name)) +
+  geom_density(alpha = 0.55,
+               adjust = 2) +
+  geom_vline(aes(xintercept = 0),
+             linetype = "dashed",
+             color = "black",
+             size = 1) +
+  theme_bw() +
+  scale_fill_viridis_d(option = "A", direction = -1)
+
+
+# proportion of relationships which are significant
+relationship_estimate %>%
+  filter(p.value < 0.05) %>%
+  group_by(name) %>%
+  summarize(n())
+
+# how far does the estimated relationship deviate from the known relationship?
+relationship_estimate %>%
+  mutate(diff = known_relationship - estimate) %>%
+  group_by(name) %>%
+  summarize(mean(diff), 
+            sd(diff))
+
+
+dat_conf <- dat %>%
+  #filter(rep <= 10) %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
[…]

### 2022-06-13 shallow_lambda.R

@@ -1,0 +1,10 @@
+#
+# ISD exponents vary from 1.2 to 1.4 in NEON data
+
+
+# Preliminaries -----------------------------------------------------------
+
+library(tidyverse)
+library(sizeSpectra) #bounded power law and MLE functions
+source("custom_functions.R")
+

### 2022-06-13 sim_small_CI_plots.R

@@ -1,0 +1,204 @@
+# small variation in lambda CI plot
+
+library(tidyverse)
+library(sizeSpectra) #bounded power law and MLE functions
+source("custom_functions.R")
+
+# simulation data
+dat <- readRDS("data_sim/PLB_sim_small_dat.rds")
+
+names(dat)
+
+# estimate the relationship (beta_1_ estimate) across the gradient
+relationship_estimate <- dat %>%
+  #filter(rep<10) %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
+               ~lm(estimate ~ env_gradient, data = .x))) %>%
+  mutate(tidied = map(lm_mod, broom::tidy)) %>%
+  unnest(tidied) %>%
+  filter(term == "env_gradient") %>%
+  select(-data, -lm_mod, -statistic)
+
+# what is the pairwise difference in spectra parameter estimates?
+lambda_diff <- dat %>%
+  select(env_gradient, rep, name, estimate) %>%
+  pivot_wider(names_from = name, values_from = estimate) %>%
+  mutate(MLE_NAS = MLE - NAS,
+         MLE_ELB = MLE - ELBn, 
+         NAS_ELB = NAS - ELBn)
+
+lambda_diff
+
+lambda_diff %>%
+  pivot_longer(cols = MLE_NAS:NAS_ELB) %>%
+  ggplot(aes(y = ..scaled..,
+             x = value,
+             fill = name)) +
+  geom_density(alpha = 0.55,
+               adjust = 2) +
+  geom_vline(aes(xintercept = 0),
+             linetype = "dashed",
+             color = "black",
+             size = 1) +
+  theme_bw() +
+  scale_fill_viridis_d(option = "A", direction = -1)
+
+
+lambda_diff %>%
+  pivot_longer(cols = MLE_NAS:NAS_ELB) %>%
+  ggplot(aes(y = value,
+             x = rep,
+             fill = name,
+             color = name)) +
+  geom_pointrange(aes(ymin = 0, ymax = 0),
+                  alpha = 0.3) +
+  theme_bw()
+
+dat %>%
+  filter(rep %in% seq(0, 1000, by = 10)) %>%
+  ggplot(aes(y = estimate - known_b,
+             x = rep,
+             fill = name,
+             color = name)) +
+  geom_point() +
+  theme_bw() +
+  facet_wrap(~known_b)
+
+# proportion of relationships which are significant
+relationship_estimate %>%
[…]

### 2022-06-13 static_b_CI_plot.R

@@ -1,0 +1,141 @@
+# no relationship CI plot
+
+library(tidyverse)
+library(sizeSpectra) #bounded power law and MLE functions
+source("custom_functions.R")
+
+# no relationship data
+dat <- readRDS("data_sim/PLB_static_b_dat.rds")
+
+names(dat)
+
+# calculate beta coefficients for all reps
+relationship_estimate <- dat %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
+               ~lm(estimate ~ env_gradient, data = .x))) %>%
+  mutate(tidied = map(lm_mod, broom::tidy)) %>%
+  unnest(tidied) %>%
+  filter(term == "env_gradient") %>%
+  select(-data, -lm_mod, -statistic)
+
+# how many show a "significant" result?
+relationship_estimate %>%
+  filter(p.value < 0.05) %>%
+  group_by(name) %>%
+  summarize(n()/1000)
+
+spaghetti_dat <- dat %>%
+  #filter(rep < 600) %>%
+  group_by(rep, name, known_relationship) %>%
+  nest() %>%
+  mutate(lm_mod =
+           map(data,
+               ~lm(estimate ~ env_gradient, data = .x))) %>%
+  mutate(fit = map(lm_mod,
+                   ~fitted(object = .))) %>%
+  unnest(c(fit, data)) %>%
+  select(rep, name, env_gradient, fit, lm_mod) %>%
+  mutate(tidied = map(lm_mod, broom::tidy)) %>%
+  unnest(tidied) %>%
+  filter(term == "env_gradient") %>%
+  mutate(significant = p.value < 0.05,
+         negative = estimate < 0) %>%
+  select(-lm_mod, -term, -std.error, -statistic) %>%
+  ungroup() %>%
+  select(rep, name, env_gradient,fit, p.value, significant, negative) 
+
+spaghetti_dat %>%
+  ggplot(aes(y = fit,
+             x = env_gradient,
+             group = rep, 
+             color = negative)) +
+  geom_point() +
+  stat_smooth(geom = "line",
+              method = "lm",
+              alpha = 0.25,
+              se = FALSE,
+              size = 1) +
+  facet_wrap(name~significant,
+             labeller = label_both,
+             ncol = 2) +
+  scale_color_viridis_d(option = "plasma",
[…]

### 2022-06-13 vary_n.R

@@ -1,0 +1,130 @@
+# sample size analysis
+
+library(tidyverse)
+library(sizeSpectra) #bounded power law and MLE functions
+source("custom_functions.R")
+
+# read in datasets with variable sample size, n
+# All other params are as in main analysis
+n100 <- readRDS("data_sim/PLB_sim_n100_dat.rds")
+n500 <- readRDS("data_sim/PLB_sim_n500_dat.rds")
+n5000 <- readRDS("data_sim/PLB_sim_n5000_dat.rds")
+n10000 <- readRDS("data_sim/PLB_sim_n10000_dat.rds")
+n1000 <- readRDS("data_sim/PLB_sim_dat.rds")
+n1000$n <- 1000
+
+
+### ELBn struggles to estimate b when sampled from -2 and -2.5 PLB ###
+# Filter out all reps which have an estimate == NA
+
+# make vector of reps which have an NA value for the estimate
+rep_na <- n100$rep[which(is.na(n100$estimate))]
+
+# filter out all the reps with NA values
+# the !rep means (reps which do not match values in rep_na)
+PLB_sim_n100 <- n100 %>%
+  filter(!rep %in% rep_na)
+
+# combine all data sets
+dat <- bind_rows(n100, n500, n5000, n10000, n1000)
+
+# How does the lambda estimate vary by sample size?
+ggplot(dat, aes(x = n, y = estimate, color = name)) +
+  geom_point(alpha = 0.1) +
+  #stat_smooth(method = "loess") +
+  stat_summary(geom = "line",  fun = mean, size = 1.5) +
+  geom_hline(aes(yintercept = known_b), linetype = "dashed") +
+  facet_wrap(~known_b,
+             scales = "free") +
+  theme_bw() +
+  labs(title = "Lambda ~ n",
+       caption = "dashed line is known lambda value") +
+  NULL
+
+ggsave("figures/lambda_n_5_sites.png", width = 8, height = 8)
+
+
+# how does estimate of lambda vary with sample size?
+ggplot(dat,
+       aes(x = estimate,
+           y = ..scaled..,
+           fill = name)) +
+  geom_density(alpha = 0.5,
+               adjust = 2) +
+  facet_wrap(n ~ known_b,
+             ncol = 5,
+             labeller = label_both) +
+  geom_vline(aes(xintercept = known_b),
+             size = 1,
+             alpha = 0.75,
+             color = "black")+
+  scale_fill_viridis_d(option = "plasma") +
[…]

### 2022-06-30 Manuscript.Rmd
updated rmd yaml for word doc export. added initial .bib file for references.

@@
 title: "Manuscript"
-author: "Justin Pomeranz"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-header-includes:
-   - \usepackage{lineno}
-   - \linenumbers
-   - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+author: 
+  - Justin Pomeranz:
+      email: jfpomeranz@gmail.com
+      institute: cmu
+      correspondence: true
+  - name: James R. Junker
+    institute: [glrc, lumcon]
+  - name: Jeff S. Wesner
+    institute: usd
+institute:
+  - cmu: Colorado Mesa University
+  - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
+  - lumcon: Louisiana Universities Marine Consortium, Chauvin, LA USA
+  - usd: Dept. of Biology, University of South Dakota, Vermillion, SD, USA
 output:
@@
      toc: yes
+     keep_tex: true
+     pandoc_args:
+     - '--lua-filter=scholarly-metadata.lua'
+     - '--lua-filter=author-info-blocks.lua'
+   word_document:
+      reference_docx: working_docx_template.docx
+      toc: no
+      pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
    html_document: default
+      pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
+header-includes:
+  - \usepackage{lineno}
+  - \usepackage{amsmath}
+  - \usepackage{indentfirst}
+  - \linenumbers
+  - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+indent: true
+linestretch: 1
+bibliography: refs.bib
+link-citations: no
+link-color: grey
+csl: ecology.csl
 ---
@@
 **NOTE** I need to integrate reference manager with R markdown. Any favorites or recommendations?? I generally use Zotero (free), but think I have access to Mendeley through CMU...
+
+**NOTE from Jim** I use Zotero, mainly for free and open-source reasons, but I also can have a standalone program that doesn't require internet access, can backup my reference library on the cloud and sync among multiple computers, can spit out bibtex to connect to markdown, and can have group citation lists if neccessary.
+I use it with a betterBibtex add on that allows for modifying bibtex keys for calling refs in Rmarkdown.
 
@@
 Body size distributions are a fundamental characteristic of communities.
-The remarkable consistency of these relationships across spatiotemporal scales and habitat types has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010).
+The remarkable consistency of these relationships across spatiotemporal scales and habitat types has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010)[@petchey2010].
 Size metrics have commonly been used in marine systems, and are increasingly being applied to assess the condition of freshwater ecosystems (Martinez et al.
-2016, Pomeranz 2018).
+2016, Pomeranz 2018)[@martinez2016; @pomeranz2018].
 
@@
 2007, Edwards et al.
-2017), and reanalysis of empirical datasets also indicates that the conclusions are dependent on the methodology used (White et al.
+2017)[@white2007; @edwards2017], and reanalysis of empirical datasets also indicates that the conclusions are dependent on the methodology used (White et al.
 2007, Edwards et al.
-2020).
+2020)[@white2007; @edwards2020].
 However, recent empirical analysis of stream macroinvertebrate communities across the NAtional Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used (Pomeranz et al.
-2022).
+2022)[@pomeranz2022].
 While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time.
@@
 In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in Edwards et al.
-(2017).
+(2017)[-@edwards2017].
 Let $M$ be a random variable of body sizes described by the probability density function:
@@
 These values are based on empirical body sizes of stream benthic communities reported in Pomeranz et al.
-(2020).
+(2020)[@pomeranz2020].
 Our results are not dependent on the range of body sizes, and we show the results of other ranges in the supplemental information.
@@
 This method has been used by (Pomeranz et al.
-2018, I **think** McGarvey \~2018, Blanchard (log 2, ?normalized?)) and is referred to as the Normalized Abundance Spectrum (NAS).
+2018, I **think** McGarvey \~2018, Blanchard (log 2, ?normalized?))[@pomeranz2018; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 Details of the sample collection and processing can be found in Pomeranz et al.
-(2018).
+(2018)[@pomeranz2018].
 Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
@@
 
-The second data set was from the wadeable stream sites of the National Ecological Observatory Network (NEON; **data product XXX**).
+The second data set was from the wadeable stream sites of the National Ecological Observatory Network (NEON; **data product XXX**)[@NEON_Inverts2022].
 These sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska.
@@
 This data has been analyzed previously using size spectra methods as described in Pomeranz et al.
-(2022).
-Detailed methods of the sampling collection and data processing methods can be found on the NEON website (URL, or macroinvertebrate DPI pubs).
+(2022)[@pomeranz2022].
+Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([https://www.neonscience.org/data-collection/macroinvertebrates], or macroinvertebrate DPI pubs).
 
@@
 *alternative* absolute difference of \~0.1 units in $\lambda$ across gradients is similar to differences reported in other studies (seasonally: 0.16, Mcgarvey and Kirk; landuse: 0.15 Martinez; O'Gorman 2017 \~0.25, Yvon and Dossena papers \~0.1?
-[need to redownload some of these and compare])
+[need to redownload some of these and compare])[@mcgarvey2018; @dossena2012; @ogorman2017]
 
@@
 We reiterate the recommendations of White et al.
-(2007), Sprules and Barth (2016) and Edwards et al.
-(2017) to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
+(2007)[-@white2007], Sprules and Barth (2016)[-@sprules2016] and Edwards et al.
+(2017)[-@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
 Furthermore, we strongly encourage authors to publish individual size data whenever possible.
@@
 This will aid in the ability for size spectra work to be synthesized between research groups and across scales.
+
+# References
+
+<div id="refs"></div>
 

### 2022-06-30 Manuscript.tex
updated rmd yaml for word doc export. added initial .bib file for references.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Manuscript},
  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Jeff S. Wesner4},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{indentfirst}
\linenumbers
\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Manuscript}
\author{Justin Pomeranz\textsuperscript{1,*} \and James R.
Junker\textsuperscript{2,3} \and Jeff S. Wesner\textsuperscript{4}}
\date{30 June, 2022}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\setstretch{1}
\textsuperscript{1} Colorado Mesa University\\
\textsuperscript{2} Great Lakes Research Center, Michigan Technological
University, Houghton, MI USA\\
\textsuperscript{3} Louisiana Universities Marine Consortium, Chauvin,
LA USA\\
\textsuperscript{4} Dept. of Biology, University of South Dakota,
Vermillion, SD, USA

\textsuperscript{*} Correspondence:
\href{mailto:jfpomeranz@gmail.com}{Justin Pomeranz
\textless{}\href{mailto:jfpomeranz@gmail.com}{\nolinkurl{jfpomeranz@gmail.com}}\textgreater{}}

\textbf{NOTE} I need to integrate reference manager with R markdown. Any
favorites or recommendations?? I generally use Zotero (free), but think
I have access to Mendeley through CMU\ldots{}

\textbf{NOTE from Jim} I use Zotero, mainly for free and open-source
reasons, but I also can have a standalone program that doesn't require
internet access, can backup my reference library on the cloud and sync
among multiple computers, can spit out bibtex to connect to markdown,
and can have group citation lists if neccessary. I use it with a
betterBibtex add on that allows for modifying bibtex keys for calling
refs in Rmarkdown.

\textbf{NOTE} I re-ran all of these after setting the seed for
reproducibility. Figures should be up to date, but the numbers may
change slightly i.e., 94.8\% might become 95.1\%, mean \(\pm\) SD values
will be different.

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Size spectra represent a fundamental attribute of community
  organization and are increasingly being used in assessments of fresh
  waters.
\item
  Many methods have been proposed for constructing size spectra
  relationships, but recent work has shown that varied methods return
  biased estimates of relationship parameters, and different methods in
  fact are not estimating the same parameter. Despite this variability
  in estimates, it is unclear if the relative change across
  environmental gradients is consistent across methodologies. Here, we
  simulate data sets across an hypothetical environmental gradient and
  estimate the size spectra parameter (slope or exponent, depending on
  method) at each site, and, importantly, estimate the relationship of
  parameters across the hypothetical gradient. We also use two
  previously published body size datasets across an anthropogenic stress
  gradient and an environmental temperature gradient and assess how the
  conclusions of these studies would vary based on methods used.
\item
  We find that maximum likelihood methods always perform as well or
  better than common binning methods. Additionally, the variance in
  estimates using MLE methods is markedly reduced when compared to
  binning methods.
\item
  The uncertainty and variation in estimates when using binning methods
  is often greater than or equal to the variation previously published
  in experimental and observational studies, bringing into question the
  effect size of previously published results. However, re-analysis of
  two previously published datasets does not markedly alter the
  conclusions reached. Further study is needed to identify how and when
  these estimates may be biased, and when the general pattern of results
  can be considered to be ``true''.
\end{enumerate}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Body size distributions are a fundamental characteristic of communities.
The remarkable consistency of these relationships across spatiotemporal
scales and habitat types has led them to be recommended as a
``universal'' indicator of ecological status (Petchey and Belgrano
2010)(Petchey and Belgrano 2010). Size metrics have commonly been used
in marine systems, and are increasingly being applied to assess the
condition of freshwater ecosystems (Martinez et al.~2016, Pomeranz
2018)(\textbf{martinez2016?}, \textbf{pomeranz2018?}).

Individual size distributions (ISD, also referred to as abundance size
spectra) are commonly used. Generally, there is a negative relationship
between body size (M) and abundance (N). Theoretical and empirical data
support this relationship being described as a simple power law with
exponent \(\lambda\) in the form of \(N \sim M^{\lambda}\). Commonly,
\(N\) is the count of body sizes grouped into bins, and \(\lambda\) is
estimated from OLS regressions of log transformed data of \(N\) and the
mid point of the body size bin \(M_{bin}\):
\(\log(N_{count}) = \lambda \log(M_{bin})\). Myriad binning methods have
been proposed, including linear and logarithmic bin widths. Likewise,
methods can rely on the absolute count in the bins as well as
normalization techniques where the count is divided by the bin width
(especially common with logarithmic binning). Alternatively, \(\lambda\)
can be estimated directly using maximum likelihood techniques.

Previous work has shown that the estimates of \(\lambda\) differ between
MLE and OLS techniques. OLS methods are particularly sensitive to
decisions made in the binning process. Simulation studies have shown
that MLE offers consistently more accurate estimates of \(\lambda\)
(White et al.~2007, Edwards et al.~2017)(White et al. 2007, Edwards et
al. 2017), and reanalysis of empirical datasets also indicates that the
conclusions are dependent on the methodology used (White et al.~2007,
Edwards et al.~2020)(White et al. 2007, Edwards et al. 2020). However,
recent empirical analysis of stream macroinvertebrate communities across
the NAtional Ecological Observatory Network (NEON, USA) showed that
while the estimates of \(\lambda\) varied, the relative change across
the environmental gradient was consistent regardless of method used
(Pomeranz et al.~2022)(Pomeranz et al. 2022). While there is a growing
consensus that MLE methods offer more reliable estimates of \(\lambda\),
and binning methods result in biased estimates, it remains unclear if
these biases are consistent and systematic or stochastic, and whether or
not the relative change in ISD parameters is consistent across space and
time. In other words, if the data within a study are all treated the
same, does a relative change of OLS slope parameters of 0.1 coincide
with a relative change of MLE estimates of 0.1? In order to answer this
question, we simulate body size observations from bounded power law
distributions with varied \(\lambda\) exponents across a hypothetical
environmental gradient and compare the results obtained using three
different methodologies common in the literature. In addition, we
re-analyze two previously published datasets of stream community body
sizes across a stress and environmental gradient. We find that the MLE
method more accurately estimates the site-specific \(\lambda\) exponents
as well as the relative change in exponents across the hypothetical
gradient, as well as having smaller variation. The logarthmic binning
methods generally perform well, but have larger confidence intervals
around the estimates. However, the overall conclusions of previously
published empirical data sets are not dependent on the method used.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-simulation}{%
\subsection{Data Simulation}\label{data-simulation}}

In order to investigate the performance of commonly used methods, we
simulate body size observations from a bounded power law distribution
using the inverse method, as described in Edwards et al.~(2017)(2017).
Let \(M\) be a random variable of body sizes described by the
probability density function:

\[f(M) = CM^\lambda, M_{min} \le M \le M_{max} \]

\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
on empirical body sizes of stream benthic communities reported in
Pomeranz et al.~(2020)(\textbf{pomeranz2020?}). Our results are not
dependent on the range of body sizes, and we show the results of other
ranges in the supplemental information.

For the main analysis, presented here, we sampled \(n = 1000\) body
sizes from distributions described by five different \(\lambda\)'s:
(-1.5, -1.75, -2.00, -2.25, -2.5). The supplemental information contains
results when the value of n is varied (or maybe include in main?). Each
value of \(\lambda\) was assumed to come from a community across a
hypothetical environmental gradient \(X\). Values of \(X\) for each
community are uniformly distributed from -1 to 1. We repeated the data
simulation 1000 times (reps). The main results presented here were not
dependent on the range of \emph{x}-values or the number of sites
(supplemental information).

\hypertarget{sample-size-n}{%
\subsubsection{Sample size, n}\label{sample-size-n}}

The number of observations in our simulations may bias the results.
Therefore, we repeated the simulations described above, but varied the
sample size \(n\). We tested values of
\(n = 200, 500, 1000, 5000, 10 000\). (originally had n=100, but thre
errors based on random seed. Bumped it to 200 to try and get consistent
error-free runs)

\hypertarget{no-relationship}{%
\subsubsection{``No'' relationship}\label{no-relationship}}

It is possible that our sampling or simulation framework could impact
the results obtained, possibly by inflating the type I error rate. In
order to assess whether or not our simulations are robust to this, we
performed independent random samples from a bounded power law for five
communities across a hypothetical gradient as described above, but we
set \(\lambda = -2\) for all five communities. Hence, we are simulating
a scenario where the size spectra relationship is invariant to our
hypothetical environmental gradient, and we should expect the estimated
relationship coefficient \(\beta_1 = 0\).

\hypertarget{finer-resolution-of-lambda-values}{%
\subsection{\texorpdfstring{Finer resolution of \(\lambda\)
values}{Finer resolution of \textbackslash lambda values}}\label{finer-resolution-of-lambda-values}}

For the main analysis, we were interested in determining whether or not
the methods were able to detect relatively coarse changes to size
spectra parameters, specifically, a change of 0.5 units across each
value of \(X\), or an absolute change or 1.0 units. However, many
seminal works on variation in size spectra relationships in freshwater
are on a much finer scale (on the order 0f 0.1 to 0.25 units).
Therefore, we performed simulations as described above, but varied
\(\lambda\) from -1.9 to -2.1 across the hypothetical environmental
gradient. This equates to a know relationship coefficient of
\(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
change of 0.2 units.

\hypertarget{estimation-of-size-spectra-parameter-lambda}{%
\subsection{\texorpdfstring{Estimation of Size spectra parameter
\(\lambda\)}{Estimation of Size spectra parameter \textbackslash lambda}}\label{estimation-of-size-spectra-parameter-lambda}}

For each sample of \(M\) and each of the 1000 replicates, we estimated
the exponent \(\lambda\) using MLE methods modified from the the
\texttt{sizeSpectra} package (Edwards et al.~2020). In addition, we use
two common binning methods to estimate the OLS slope parameter in
log-log space. For the first binning method, we created 6 equal
logarithmic bins covering the range of body sizes. The count in each bin
was normalized by dividing by the bin width. This method has been used
by (refs here, I think a lot of Woodward's group uses this\ldots)
Throughout the manuscript, the normalized equal logarithmic binning
method will be referred to as ELBn. The second method was similar to
ELBn, but bins of Log\textsubscript{2} widths are used. The count in
each bin is normalized in the same way. This method has been used by
(Pomeranz et al.~2018, I \textbf{think} McGarvey \textasciitilde2018,
Blanchard (log 2, ?normalized?))(McGarvey et al. 2019,
\textbf{pomeranz2018?}) and is referred to as the Normalized Abundance
Spectrum (NAS).

\hypertarget{estimation-of-relationship-across-gradient-beta_1}{%
\subsection{\texorpdfstring{Estimation of relationship across gradient,
\(\beta_1\)}{Estimation of relationship across gradient, \textbackslash beta\_1}}\label{estimation-of-relationship-across-gradient-beta_1}}

For each simulation replicate, we estimated how the size spectra
parameters varied across the hypothetical gradient. Simple OLS
regression were conducted in the form
\(\lambda_{estimate} ~ \beta_0 + \beta_1 * X\). The distribution of the
relationship coefficient, \(\beta_1\), were plotted compared to the
known relationship. Likewise, the confidence interval of \(\beta_1\) was
assessed to determine the proportion of estimates which contained the
true value of the known relationship.

\hypertarget{empirical-data}{%
\subsection{Empirical Data}\label{empirical-data}}

We re-analyze two datasets of benthic macroinvertebrate communities from
stream habitats across two different gradients. In the first,
quantitative macroinvertebrate samples were collected from streams
across an acid mine drainage stress gradient. Details of the sample
collection and processing can be found in Pomeranz et
al.~(2018)(\textbf{pomeranz2018?}). Briefly, all individuals from each
sample were identified to the lowest practical taxonomic unit and body
lengths were measured using image processing software from photos taken
with a camera mounted to a dissecting microscope. Body mass was
estimated using published length weight regressions.

The second data set was from the wadeable stream sites of the National
Ecological Observatory Network (NEON; \textbf{data product
XXX})(National Ecological Observatory Network (NEON) 2022). These sites
are located across a wide temperature gradient in the United States,
from Puerto Rico to Alaska. Quantitative macroinvertebrate samples were
collected using the most appropriate method based on the local habitat.
All individuals were identified and had their body lengths measured, and
body mass was estimated using published length weight regressions. This
data has been analyzed previously using size spectra methods as
described in Pomeranz et al.~(2022)(Pomeranz et al. 2022). Detailed
methods of the sampling collection and data processing methods can be
found on the NEON website
({[}\url{https://www.neonscience.org/data-collection/macroinvertebrates}{]},
or macroinvertebrate DPI pubs).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{relationship-across-the-hypothetical-environmental-gradient}{%
\subsection{Relationship across the hypothetical environmental
gradient}\label{relationship-across-the-hypothetical-environmental-gradient}}

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_main.png}
\caption{Relationship estimates across the hypothetical gradient for
each replicate. Each panel is a different method for estimating the size
spectra parameter.}
\end{figure}

All methods performed reasonably well in detecting the known
relationship across the hypothetical environmental gradient. The
regressions using the MLE method found a significant relationship 100\%
of the time, while the ELBn and NAS method found significant
relatoinships 95.8\% and 96.9\% of the time, respectively.

The confidence intervals for the relationship coefficients (\(\beta_1\))
for both the MLE and ELBn method contained the true value of the known
relationship across the gradient \(\ge\) 95\% of the time, whereas the
confidence interval for the NAS method only had the true value 83.1\% of
the time. Despite having similar accuracy, the width of the CIs for the
ELBn method were more than 3 times that of the MLE method (ELBn\_CI\_ =
0.386 \(\pm\) 0.20; MLE\_CI\_ = 0.121 \(\pm\) 0.056). The CI's for the
NAS were slightly smaller than the ELBn method, but still \(>\) 2.5
times as wide as the MLE method (NAS\_CI = 0.339 \(\pm\) 0.172).

On average, the relationship across the gradient was under estimated:
MLE -0.002 (\(\pm\) 0.022); ELBn -0.039 (\(\pm\) 0.062); and NA -0.081
(\(\pm\) 0.061).

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods.}
\end{figure}

Interestingly, there was an interactive effect of the estimate accuracy
between sample size and the value of \(\lambda\). All methods were more
accurate with larger sample sizes, and smaller values of \(\lambda\).

\begin{figure}
\centering
\includegraphics{figures/n_vary_est_b.png}
\caption{Distribution of size spectra parameter estimates. Vertical line
is the known parameter wich describes the bounded power law distribution
from which the body size estimates were sampled. As n increases (top to
bottom) and \(\lambda\) increases (left to right), the accuracy of the
estimate improves across all methods.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/n_vary_main.png}
\caption{Individual regression estimates across the hypothetical
gradient based on sample size (rows) and methodology used (columns).}
\end{figure}

\hypertarget{no-relationship-1}{%
\subsection{No relationship}\label{no-relationship-1}}

All of the methods performed similarly when there was no relationship
across the hypothetical gradient. The Type I error rate for MLE, ELBn
and NAS were 5\%, 5.7\% and 4.6\%, respectively. However, the confidence
intervals for the binning methods were \textasciitilde3 times as wide as
the MLE method (mean CI widths: MLE = 0.116, ELBn = 0.372; NAS = 0.332).

Of the relationships which were estimated to be significant, the MLE,
ELBn, and NAS method had a 58\%, 49.1\% and 54.3\% probability of
indicating a negative relationship, respectively.

\begin{figure}
\centering
\includegraphics{figures/static_b_lambda_spaghetti.png}
\caption{Individual regression estimates when no relationship exists
across the hypothetical environmental gradient. The methods are
separated by rows, and the left and right column show relationships
which were non-significant and significant, respectively. The individual
regressions coefficients are colored-coded to indicate positive and
negative relationship estimates.}
\end{figure}

\hypertarget{small-variation-in-lambda}{%
\subsection{Small variation in lambda}\label{small-variation-in-lambda}}

The performance of all methods declined when trying to detect variation
in the \(\lambda\) parameter between -1.9 to -2.1 across the
environmental gradient. The mean \(\beta_1\) coefficient estimates for
all methods were closely associated with the known relationship. Once
again the variation in the estimates using the MLE method were much
smaller that the estimates using both of the binning methods (Figure 4).

\begin{figure}
\centering
\includegraphics{figures/PLB_sim_small_relationship_density.png}
\caption{Mean difference between relationship estimate and known
relationship coefficient for each method. Here, \(\lambda\) varied from
-1.9 to -2.1, with a kown relationaship of -0.1 across the gradient.}
\end{figure}

Of the 1000 simulation replicates, the MLE only detected a significant
relationship 90\% of the time, whereas the ELBn and NAS methods detected
significant relationships 19 and 23\% of the time, respectively. Of the
replicates which were significant, the confidence interval contained the
true value of the relationship 94.6, 88.4, and 89.1\% of the time for
the MLE, ELBn, and NAS methods, respectively.

\begin{figure}
\centering
\includegraphics{figures/small_lambda_spaghetti.png}
\caption{Linear regressions across the hypothetical gradient when lambda
varies from -1.9 to -2.1. Each method is displayed in the rows, and the
left colum shows the non-significant estimates, while the right column
shows the relationships which were signficant (\(\beta_1\) coefficient
p-value \textless{} 0.05). The colors indicate when the beta coefficient
estimate was negative (pink) and when it was positive (purple).}
\end{figure}

Once again, the width of the confidence intervals for both of the
binning methods were \(\ge\) 2.8 times that of the MLE method (mean
\(\pm\) SD MLE CI width: 0.117 \(\pm\) 0.051; ELBn CI width: 0.378
\(\pm\) 0.169; NAS CI width: 0.332 \(\pm\) 0.150).

On average, the relationship estimates across the gradient were similar
to the know value (mean deviation: MLE = 0.00001, ELBn = -0.009, NAS =
-0.012). However, when only looking at the relationships which were
significant, this variation increased considerably(MLE = -0.002, ELBN =
-0.052, NAS = 0.034).

\hypertarget{empirical-data-1}{%
\subsection{Empirical data}\label{empirical-data-1}}

For both empirical datasets, the direction and magnitude of chnage
(i.e.~\(\beta_1\) coefficients) are generally in agreement. Size spectra
parameters consistently increase (become flatter) in the AMD data (Fig.
XX A). Likewise, the size spectra parameters consistently increase
(become steeper) with increasing temperature across the NEON sites (Fig
XX D)

\textbf{NOTE} combine all the figures below into one with 4 panels (AMD
= A, B; NEON = C, D)

\begin{figure}
\centering
\includegraphics{figures/AMD_plot.png}
\caption{A}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/amd_relationship.png}
\caption{B.Relationship estimates across AMD gradient}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/NEON_plot.png}
\caption{C.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/neon_relationship.png}
\caption{D.}
\end{figure}

Because the \(\beta_1\) coefficient estimates are similar, and the range
of the gradient in the AMD data is relatively small (9.5), the absolute
change in the size spectra parameter across the AMD gradient are similar
regardless of method used (range: 0.59 to 0.72). Likewise, the absolute
change in size spectra parameters across the NEON temperature gradient
ranges from 0.06 to 0.165, depending on the method used.

\emph{alternative} absolute difference of \textasciitilde0.1 units in
\(\lambda\) across gradients is similar to differences reported in other
studies (seasonally: 0.16, Mcgarvey and Kirk; landuse: 0.15 Martinez;
O'Gorman 2017 \textasciitilde0.25, Yvon and Dossena papers
\textasciitilde0.1? {[}need to redownload some of these and
compare{]})(Dossena et al. 2012, O'Gorman et al. 2017, McGarvey and Kirk
2018)

\hypertarget{simulating-shallower-lambdas}{%
\subsection{Simulating shallower
lambdas}\label{simulating-shallower-lambdas}}

Estimates of the relationship between lambda and a hypothetical gradient
varied depending on the method used. However, interpretations of
empirical data were broadly consistent across methodologies. Upon closer
inspection, the values of size spectra parameters in the empirical data
were considerably shallower than -2. Given that we found the performance
of all methods increased with shallower \(\lambda\)'s we wanted to
investigate how the methods performed with simulated values more similar
to the empirical estimates of parameters describing size spectra
relationships. Therefore, we repeated the simulation process as in the
main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5.

\begin{figure}
\centering
\includegraphics{figures/shallow_lambda_main.png}
\caption{Relationship estimates across the hypothetical gradient for
each replicate. Each panel is a different method for estimating the size
spectra parameter.}
\end{figure}

We found generally less variability in the relationship estimates across
a gradient of distributions described with shallower \(\lambda\)
parameters. All of the relationship coefficient estimates
(\(\beta_1\))'s across all replicates were significant regardless of
method used. The confidence interval for \(\beta_1\) estimate for the
MLE method contained the true value 94.7\% of the time. However, the
true value was in the confidence interval for the ELBN and NAS method
only 90.9\% and 76.9\% of the time, respectively. Once again, the mean
width of the CI for the binning methods were \textasciitilde2 times as
large as for the MLE method. Mean \(\pm\) SD CI width: MLE = 0.0442
\(\pm\) 0.0196; ELBn = 0.0920 \(\pm\) 0.0482, NAS = 0.0798 \(\pm\)
0.0324.

On average, the relationship across the gradient was under estimated,
but by nearly an order of magnitude less than when simulating across
steeper values of \(\lambda\). Differences between known and estimated
relationship coefficients: MLE -0.0007 (\(\pm\) 0.0079); ELBn -0.0058
(\(\pm\) 0.0177); NAS -0.0251 (\(\pm\) 0.0124).

\begin{figure}
\centering
\includegraphics{figures/shallow_lambda_relationship_density.png}
\caption{Distribution of relationship coefficient estimates. Vertical
line is the known relationship. All methods under estimate the value,
but the mean magnitude and distribution of values is greater for the
ELBn and NAS methods.}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Measuring parameters describing the decline in abundance with increasing
body size in communities is being done with increasing frequency across
ecology. Previous work has investigated the accuracy and inherent biases
associated with different estimation methods. However, how these
innaccuracies and biases compound across environmental gradients remains
uncertain, making it difficult to detecting variation in size spectra
parameters across environmental gradients with confidence. Here, we
sampled body sizes from known distributions and estimated their
parameters across a hypothetical gradient in order to assess the
relationship coefficients obtained.

Generally, there was large agreement in the direction and magnitude of
average relationship coefficient estimates across methods. The estimate
from the MLE method was always closer to the known relationship, and
always had the smallest variation around the estimate.

All of the methods performed equally well when the \(\lambda\) value did
not vary across the hypothetical gradient, with a Type I error rate of
\textasciitilde5\%, exactly as would be expected under the assumptions
of frequentist statistics. This indicates that previous studies which
have detected no difference in size spectra parameters across sites can
be interpreted with confidence, regardless of the method used.

The performance of all methods improved as sample size increases, and as
the exponent of the size spectra relationship gets larger (more
shallow). As either or both of these variables change, the difference in
estimated relationship coefficients declines across the methods. This is
particularly interesting given the fact that both empirical data sets of
stream communities examined here are shallower than the expected
relationship of \(\lambda = -2\), and the conclusions of the change in
size spectra relationships across the gradients are not dependent on
method used. If communities under study are in fact described by
shallower exponents, the method used may not be critical in the
conclusions reached. However, the MLE method performs as well or better
than the two binning methods examined here under all contexts and should
be the preferred method moving forward. At a minimum, future studies
should report MLE estimates to ensure that the results are not dependent
on the methodology used.

Alarmingly, the performance of all methods declined dramatically when
trying to detect smaller variation in \(\lambda\) values. We simulated
observations from distributions varying by an absolute value of 0.2
units. This range of variation is similar to that which has been
reported in many seminal works on variation in size spectra parameters.
When using the MLE method, significant relationships were only detected
\textasciitilde88\% of the time, whereas both of the binning methods
only detected significant trends \textasciitilde20\% of the time. Of the
simulation replicates which were significant, nearly all of them were
negative across all methods except for one single replicate using the
NAS method which was positive (Figure. 5).

Despite the drop in performance with reduced variation in \(\lambda\)
values, when a significant relationship was observed, it was generally
in the correct direction and of a similar magnitude. This suggests that
previously reported significant changes in size spectra parameters
across environmental gradients and in experimental manipulations are
plausible, and the magnitude of the relative change is a reasonable
estimate. Given that all of the data within a study is treated
identically, the the over all chnage in size spectra parameters is
likely reasonable. However, the biases and inconsistencies in
relationship estimates presented here suggest that it would be difficult
if not impossible to compare the relative changes across different
published studies which use different methods. Publication of individual
body size data with future studies of size spectra relationships would
greatly aid in our ability to generalize changes to this fundamental
aspect of community organization across spatiotemporal scales and in
response to environmental stressors and perturbations.

\hypertarget{concluding-remarks}{%
\subsection{Concluding Remarks}\label{concluding-remarks}}

We reiterate the recommendations of White et al.~(2007)(2007), Sprules
and Barth (2016)(2016) and Edwards et al.~(2017)(2017) to estimate size
spectra relationships using MLE methods due to their superior
performance in nearly every context. Furthermore, we strongly encourage
authors to publish individual size data whenever possible. This will
allow for the consistent re-analysis of existing data sets as
methodologies develop and improve. This will aid in the ability for size
spectra work to be synthesized between research groups and across
scales.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-dossena2012}{}}%
Dossena, M., G. Yvon-Durocher, J. Grey, J. M. Montoya, D. M. Perkins, M.
Trimmer, and G. Woodward. 2012.
\href{https://doi.org/10.1098/rspb.2012.0394}{Warming alters community
size structure and ecosystem functioning}. Proceedings of the Royal
Society B: Biological Sciences 279:3011--3019.

\leavevmode\vadjust pre{\hypertarget{ref-edwards2020}{}}%
Edwards, A. M., J. P. W. Robinson, J. L. Blanchard, J. K. Baum, and M.
J. Plank. 2020. \href{https://doi.org/10.3354/meps13230}{Accounting for
the bin structure of data removes bias when fitting size spectra}.
Marine Ecology Progress Series 636:19--33.

\leavevmode\vadjust pre{\hypertarget{ref-edwards2017}{}}%
Edwards, A. M., J. P. W. Robinson, M. J. Plank, J. K. Baum, and J. L.
Blanchard. 2017. \href{https://doi.org/10.1111/2041-210X.12641}{Testing
and recommending methods for fitting size spectra to data}. Methods in
Ecology and Evolution 8:57--67.

\leavevmode\vadjust pre{\hypertarget{ref-mcgarvey2018}{}}%
McGarvey, D. J., and A. J. Kirk. 2018.
\href{https://doi.org/10.1007/s10750-017-3448-0}{Seasonal comparison of
community-level size-spectra in southern coalfield streams of {West
Virginia} ({USA})}. Hydrobiologia 809:65--77.

\leavevmode\vadjust pre{\hypertarget{ref-mcgarvey2019}{}}%
McGarvey, D. J., T. E. Woods, and A. J. Kirk. 2019.
\href{https://doi.org/10.3791/59945}{Modeling the {Size Spectrum} for
{Macroinvertebrates} and {Fishes} in {Stream Ecosystems}}. Journal of
Visualized Experiments: JoVE.

\leavevmode\vadjust pre{\hypertarget{ref-NEON_Inverts2022}{}}%
National Ecological Observatory Network (NEON). 2022.
\href{https://doi.org/10.48443/GN8X-K322}{Macroinvertebrate collection
({DP1}.20120.001)}. {National Ecological Observatory Network (NEON)}.

\leavevmode\vadjust pre{\hypertarget{ref-ogorman2017}{}}%
O'Gorman, E. J., L. Zhao, D. E. Pichler, G. Adams, N. Friberg, B. C.
Rall, A. Seeney, H. Zhang, D. C. Reuman, and G. Woodward. 2017.
\href{https://doi.org/10.1038/nclimate3368}{Unexpected changes in
community size structure in a natural warming experiment}. Nature
Climate Change 7:659--663.

\leavevmode\vadjust pre{\hypertarget{ref-petchey2010}{}}%
Petchey, O. L., and A. Belgrano. 2010.
\href{https://doi.org/10.1098/rsbl.2010.0240}{Body-size distributions
and size-spectra: Universal indicators of ecological status?} Biology
Letters 6:434--437.

\leavevmode\vadjust pre{\hypertarget{ref-pomeranz2022}{}}%
Pomeranz, J. P. F., J. R. Junker, and J. S. Wesner. 2022.
\href{https://doi.org/10.1111/gcb.15862}{Individual size distributions
across {North American} streams vary with local temperature}. Global
Change Biology 28:848--858.

\leavevmode\vadjust pre{\hypertarget{ref-sprules2016}{}}%
Sprules, W. G., and L. E. Barth. 2016.
\href{https://doi.org/10.1139/cjfas-2015-0115}{Surfing the biomass size
spectrum: Some remarks on history, theory, and application}. Canadian
Journal of Fisheries and Aquatic Sciences 73:477--495.

\leavevmode\vadjust pre{\hypertarget{ref-white2007}{}}%
White, E. P., S. K. M. Ernest, A. J. Kerkhoff, and B. J. Enquist. 2007.
\href{https://doi.org/10.1016/j.tree.2007.03.007}{Relationships between
body size and abundance in ecology}. Trends in Ecology \& Evolution
22:323--330.

\end{CSLReferences}

\newpage

\hypertarget{supplementary-material}{%
\section*{Supplementary material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary material}

\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\hypertarget{range-of-body-sizes-m}{%
\subsection{\texorpdfstring{Range of body sizes,
\(M\)}{Range of body sizes, M}}\label{range-of-body-sizes-m}}

\begin{figure}
\centering
\includegraphics{figures/PLB_small_m_main.png}
\caption{Individual regressions for five sites across a hypothetical
gradient with a known relationship of 0.5. Range of body sizes is
reduced and is from 1, to 100.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_small_m_est_b_density.png}
\caption{Distribution of estimated \(\lambda\) coefficient for five
sites across a hypothetical gradient with known values. Range of body
sizes is reduced and is from 1, to 100.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_small_m_relationship_density.png}
\caption{Distribution of estimated relationship (\(\beta_1\))
coefficient's for five sites across a hypothetical gradient with known
value of 0.5. Range of body sizes is reduced and is from 1, to 100.}
\end{figure}

\hypertarget{large-environmental-gradient}{%
\subsection{Large environmental
gradient}\label{large-environmental-gradient}}

\begin{figure}
\centering
\includegraphics{figures/PLB_large_x_main.png}
\caption{Individual regressions for five sites across a hypothetical
gradient with a known relationship of 0.5. Range of environmental values
(\emph{x}-axis) increased to be -1000, to 1000.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_large_x_est_b_density.png}
\caption{Distribution of estimated \(\lambda\) coefficient for five
sites across a hypothetical gradient with known values. Range of
environmental values (\emph{x}-axis) increased to be -1000, to 1000.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_large_x_relationship_density.png}
\caption{Distribution of estimated relationship (\(\beta_1\))
coefficient's for five sites across a hypothetical gradient with known
value of 0.5. Range of environmental values (\emph{x}-axis) increased to
be -1000, to 1000.}
\end{figure}

\hypertarget{varying-number-of-sites}{%
\subsection{Varying number of sites}\label{varying-number-of-sites}}

\hypertarget{sites}{%
\subsubsection{10 sites}\label{sites}}

\begin{figure}
\centering
\includegraphics{figures/PLB_10_sites_main.png}
\caption{Individual regressions for ten sites across a hypothetical
gradient with a known relationship of 0.5. All other parameters are the
same as in the main analysis}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_10_sites_est_b_density.png}
\caption{Distribution of estimated \(\lambda\) coefficient for ten sites
across a hypothetical gradient with known values. All other parameters
are the same as in the main analysis}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_10_sites_relationship_density.png}
\caption{Distribution of estimated relationship (\(\beta_1\))
coefficient's for ten sites across a hypothetical gradient with known
value of 0.5. All other parameters are the same as in the main analysis}
\end{figure}

\hypertarget{three-sites}{%
\subsubsection{Three sites}\label{three-sites}}

\begin{figure}
\centering
\includegraphics{figures/PLB_3_sites_main.png}
\caption{Individual regressions for three sites across a hypothetical
gradient with a known relationship of 0.5. All other parameters are the
same as in the main analysis.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_3_sites_est_b_density.png}
\caption{Distribution of estimated \(\lambda\) coefficient for three
sites across a hypothetical gradient with known values. All other
parameters are the same as in the main analysis.}
\end{figure}

\begin{figure}
\centering
\includegraphics{figures/PLB_3_sites_relationship_density.png}
\caption{Distribution of estimated relationship (\(\beta_1\))
coefficient's for three sites across a hypothetical gradient with known
value of 0.5. All other parameters are the same as in the main analysis}
\end{figure}

\hypertarget{sample-size}{%
\subsection{Sample size}\label{sample-size}}

\textbf{note} These plots still have n = 100. Need to re run and update
if we end up going with n = 200

\begin{figure}
\centering
\includegraphics{figures/lambda_n_5_sites.png}
\caption{Plot showing the relationships between sample size and the
estimated \(\lambda\) parameters. the dashed line indicates the known
value of \(\lambda\).}
\end{figure}

\end{document}

### 2022-07-19 Manuscript.Rmd
big stage to save everything after resetting commit history

@@
 
-For the main analysis, presented here, we sampled $n = 1000$ body sizes from distributions described by five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5).
+For the main analysis, presented here, we sampled $n = 1000$ body sizes from distributions described by five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5), such that our model for $\lambda$ is:
+
+$$\lambda = Norm(\alpha -0.5X, \sigma)$$
+
 The supplemental information contains results when the value of n is varied (or maybe include in main?).

### 2022-07-19 AMD_estimate.R

@@ -10,6 +10,7 @@ amd_dat <- read_csv("site_dw_pca.csv")
 
 dw_range = range(amd_dat$dw, na.rm = TRUE)
 
+# estimate spectra paramaters across AMD gradient using three methods
 (amd_result <- amd_dat %>%
   # mutate(date = as.Date(collectDate)) %>%
   group_by(site, pca1) %>%
@@ -25,6 +26,7 @@ dw_range = range(amd_dat$dw, na.rm = TRUE)
   select(-data) %>%
   unnest(cols = method_compare))
 
+# plot parameter estimates (slope or exponent) across AMD gradient
 ggplot(amd_result,
        aes(x = pca1,
            y = estimate,
@@ -36,9 +38,11 @@ ggplot(amd_result,
   labs(title = "Change in exponent across an AMD gradient", 
        x = "AMD gradient (PCA)",
        y = "Slope/exponent estimate")
+# save the plot
 ggsave("figures/AMD_plot.png")
 
-amd_result %>%
+# estimate the beta_1_ relationship across the gradient
+amd_relationship <- amd_result %>%
   group_by(name) %>%
   nest() %>%
   mutate(lm_mod = 
@@ -47,19 +51,16 @@ amd_result %>%
   mutate(tidied = map(lm_mod, broom::tidy)) %>%
   unnest(tidied) %>%
   filter(term == "pca1") %>%
-  select(-data, -lm_mod) %>%
+  select(-data, -lm_mod) 
+
+# absolute variation across gradient
+amd_relationship %>%
+  mutate(abs_change = estimate * 
+           (max(amd_dat$pca1) - min(amd_dat$pca1))) %>%
   write_csv("results/AMD_estimates.csv")
 
-amd_result %>%
-  group_by(name) %>%
-  nest() %>%
-  mutate(lm_mod = 
-           map(data,
-               ~lm(estimate ~ pca1, data = .x))) %>%
-  mutate(tidied = map(lm_mod, broom::tidy)) %>%
-  unnest(tidied) %>%
-  filter(term == "pca1") %>%
-  select(-data, -lm_mod) %>%
+# plot the mean \pm STD.E of the estimated relationship across the gradient based on method used. 
+amd_relationship %>%
   ggplot(aes(y = estimate, 
              ymin = estimate - std.error, 
              ymax = estimate + std.error,
@@ -68,5 +69,30 @@ amd_result %>%
   geom_pointrange()+
   theme_bw() +
   labs(y = "Relationship estimate",
[…]

### 2022-07-19 NEON_estimate.R

@@ -6,9 +6,13 @@ source("custom_functions.R")
 
 # compare estimates with empirical data
 # data is from Pomeranz et al. 2018 Freshwater Biology
-neon_dat <- read_csv("neon_data.csv") 
+neon_dat <- read_csv("neon_data.csv")
 
 dw_range = range(neon_dat$dw, na.rm = TRUE)
+
+# not using the MLEbin method, as in Pomeranz et al. 2022
+# just using the "normal" MLE method to compare
+# NEON "Bins" individuals into 1 mm widths, so doesn't drastically alter body size estimates
 
 (neon_result <- neon_dat %>%
     # mutate(date = as.Date(collectDate)) %>%
@@ -34,10 +38,11 @@ ggplot(neon_result,
   labs(title = "Change in exponent across NEON", 
        x = "Mean annual air temp",
        y = "Slope/exponent estimate")
+
 ggsave("figures/NEON_plot.png")
 
 
-neon_result %>%
+neon_relationship <- neon_result %>%
   group_by(name) %>%
   nest() %>%
   mutate(lm_mod = 
@@ -46,19 +51,14 @@ neon_result %>%
   mutate(tidied = map(lm_mod, broom::tidy)) %>%
   unnest(tidied) %>%
   filter(term == "mat.c") %>%
-  select(-data, -lm_mod)%>%
+  select(-data, -lm_mod)
+
+neon_relationship %>%
+  mutate(abs_change = estimate * 
+           (max(neon_dat$mat.c) - min(neon_dat$mat.c))) %>%
   write_csv("results/NEON_estimates.csv")
 
-neon_result %>%
-  group_by(name) %>%
-  nest() %>%
-  mutate(lm_mod = 
-           map(data,
-               ~lm(estimate ~ mat.c, data = .x))) %>%
-  mutate(tidied = map(lm_mod, broom::tidy)) %>%
-  unnest(tidied) %>%
-  filter(term == "mat.c") %>%
-  select(-data, -lm_mod) %>%
+neon_relationship %>%
   ggplot(aes(y = estimate, 
              ymin = estimate - std.error, 
              ymax = estimate + std.error,
@@ -67,5 +67,102 @@ neon_result %>%
   geom_pointrange()+
   theme_bw() +
   labs(y = "Relationship estimate",
-       title = "Change in CSS parameter across Neon")
+       title = "Mean +/- Std. Error Beta across Neon")
+
 ggsave("figures/neon_relationship.png")
+
+
+
+# estimate by sample ------------------------------------------------------
+
[…]

### 2022-07-19 replot_sims_adjust2.R

@@ -1,0 +1,29 @@
+# Fix distribution plots with adjust = 2
+
+source("custom_functions.R")
+PLB_sim <- readRDS("data_sim/PLB_sim_dat.rds")
+PLB_static_b <- readRDS("data_sim/PLB_static_b_dat.rds")
+PLB_sim_small <- readRDS("data_sim/PLB_sim_small_dat.rds")
+PLB_large_x <- readRDS("data_sim/PLB_large_x_dat.rds")
+PLB_small_m <- readRDS("data_sim/PLB_small_m_dat.rds")
+PLB_3_sites <- readRDS("data_sim/PLB_3_sites_dat.rds")
+PLB_10_sites <- readRDS("data_sim/PLB_10_sites_dat.rds")
+PLB_sim_n10000 <- readRDS("data_sim/PLB_sim_n10000_dat.rds")
+PLB_sim_n5000 <- readRDS("data_sim/PLB_sim_n5000_dat.rds")
+PLB_sim_n500 <- readRDS("data_sim/PLB_sim_n500_dat.rds")
+PLB_sim_n100 <- readRDS("data_sim/PLB_sim_n100_dat.rds")
+shallow_lambda<- readRDS("data_sim/PLB_shallow_lambda_dat.rds")
+
+
+plot_sim(PLB_sim)
+plot_sim(PLB_static_b)
+plot_sim(PLB_sim_small)
+plot_sim(PLB_large_x)
+plot_sim(PLB_small_m)
+plot_sim(PLB_3_sites)
+plot_sim(PLB_10_sites)
+plot_sim(PLB_sim_n10000)
+plot_sim(PLB_sim_n5000)
+plot_sim(PLB_sim_n500)
+plot_sim(PLB_sim_n100)
+plot_sim(shallow_lambda)

### 2022-08-27 Manuscript.Rmd
new discussion approach HP

@@
 institute:
-  - cmu: Colorado Mesa University
+  - cmu: Colorado Mesa University, Grand Junction, CO, USA
   - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
@@
 The count in each bin is normalized in the same way.
-This method has been used by (Pomeranz et al.
-2018, I **think** McGarvey \~2018, Blanchard (log 2, ?normalized?))[@pomeranz2018; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).
+This method has been used by [@pomeranz2018; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 The distribution of the relationship coefficient, $\beta_1$, were plotted compared to the known relationship.
-Likewise, the confidence interval of $\beta_1$ was assessed to determine the proportion of estimates which contained the true value of the known relationship.
+Likewise, the mean $\beta_1$ estimate ($\pm$ 1 SD) for all simulations were also plotted.
 
@@
 
-We re-analyze two datasets of benthic macroinvertebrate communities from stream habitats across two different gradients.
+We re-analyze two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients.
 In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage stress gradient.
-Details of the sample collection and processing can be found in Pomeranz et al.
-(2018)[@pomeranz2018].
+Details of the sample collection and processing can be found in [@pomeranz2018].
 Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
-Body mass was estimated using published length weight regressions.
+Body mass was estimated using taxon-specific published length weight regressions.
 
-The second data set was from the wadeable stream sites of the National Ecological Observatory Network (NEON; **data product XXX**)[@NEON_Inverts2022].
-These sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska.
+The second data set was from the wadeable stream sites of the National Ecological Observatory Network [@NEON_Inverts2022].
+NEON stream sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska.
 Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat.
 All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions.
-This data has been analyzed previously using size spectra methods as described in Pomeranz et al.
-(2022)[@pomeranz2022].
+This data has been analyzed previously using size spectra methods as described in [@pomeranz2022].
 Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([https://www.neonscience.org/data-collection/macroinvertebrates], or macroinvertebrate DPI pubs).
@@
 
-All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient.
-The regressions using the MLE method found a significant relationship 100% of the time, while the ELBn and NAS method found significant relatoinships 95.8% and 96.9% of the time, respectively.
+All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient (mean $\beta_1$ estimates ranged from -0.419 to -0.498).
+However, the SD for the NAS and ELBn methods were 2-3x larger that that of the MLE method (Figure ~XX~).
+This result was consistent across simulation parameters, including sample size, number of sites, and $\lambda$ values.
 
-The confidence intervals for the relationship coefficients ($\beta_1$) for both the MLE and ELBn method contained the true value of the known relationship across the gradient $\ge$ 95% of the time, whereas the confidence interval for the NAS method only had the true value 83.1% of the time.
-Despite having similar accuracy, the width of the CIs for the ELBn method were more than 3 times that of the MLE method (ELBn_CI_ = 0.386 $\pm$ 0.20; MLE_CI_ = 0.121 $\pm$ 0.056).
-The CI's for the NAS were slightly smaller than the ELBn method, but still $>$ 2.5 times as wide as the MLE method (NAS_CI = 0.339 $\pm$ 0.172).
-
-On average, the relationship across the gradient was under estimated: MLE -0.002 ($\pm$ 0.022); ELBn -0.039 ($\pm$ 0.062); and NA -0.081 ($\pm$ 0.061).
+![Mean estimate of the relationship coefficient for main simulation (A), with ten sites (B), with small variation in lambda (C), and with no relationship.Error bars indicate 1 standard deviation.](figures/slopes_combined.png)
 
@@
 All of the methods performed similarly when there was no relationship across the hypothetical gradient.
-The Type I error rate for MLE, ELBn and NAS were 5%, 5.7% and 4.6%, respectively.
-However, the confidence intervals for the binning methods were \~3 times as wide as the MLE method (mean CI widths: MLE = 0.116, ELBn = 0.372; NAS = 0.332).
-
-Of the relationships which were estimated to be significant, the MLE, ELBn, and NAS method had a 58%, 49.1% and 54.3% probability of indicating a negative relationship, respectively.
+The mean $\beta_1$ estimates for all three methods ranged from -0.0005 to 0.0002.
+However, once again, the SD for the NAS and ELBn methods were ~3x larger than the SD for the MLE method.
 
@@
 
-The performance of all methods declined when trying to detect variation in the $\lambda$ parameter between -1.9 to -2.1 across the environmental gradient.
-The mean $\beta_1$ coefficient estimates for all methods were closely associated with the known relationship.
-Once again the variation in the estimates using the MLE method were much smaller that the estimates using both of the binning methods (Figure 4).
+The performance of all methods to detect small variation in the $\lambda$ parameter (between -1.9 to -2.1) across the environmental gradient was also investigated.
+The mean $\beta_1$ coefficient estimates ranged from -0.10 to -0.09.
+The mean estimate for the MLE method was -0.1, which exaclty matched the value for the hypothetical relationship.
+Likewise, the upper 97.5 percentile of the distribution was below 0, meaning that most of the estimates had the correct sign.
+In contrast, the upper 97.5 percentile for both the NAS and ELBn methods were greater than 0 at 0.027 and 0.052, respectively.
+(Figure 4).
 
-![Mean difference between relationship estimate and known relationship coefficient for each method.
-Here, $\lambda$ varied from -1.9 to -2.1, with a kown relationaship of -0.1 across the gradient.](figures/PLB_sim_small_relationship_density.png)
-
-Of the 1000 simulation replicates, the MLE only detected a significant relationship 90% of the time, whereas the ELBn and NAS methods detected significant relationships 19 and 23% of the time, respectively.
-Of the replicates which were significant, the confidence interval contained the true value of the relationship 94.6, 88.4, and 89.1% of the time for the MLE, ELBn, and NAS methods, respectively.
+![Distribution of the estimated relationship for each method.
+Here, $\lambda$ varied from -1.9 to -2.1, with a known relationaship of -0.1 across the gradient indicated by the dashed vertical line.](figures/PLB_sim_small_relationship_density.png)
 
@@
 
-Once again, the width of the confidence intervals for both of the binning methods were $\ge$ 2.8 times that of the MLE method (mean $\pm$ SD MLE CI width: 0.117 $\pm$ 0.051; ELBn CI width: 0.378 $\pm$ 0.169; NAS CI width: 0.332 $\pm$ 0.150).
-
-On average, the relationship estimates across the gradient were similar to the know value (mean deviation: MLE = 0.00001, ELBn = -0.009, NAS = -0.012).
-However, when only looking at the relationships which were significant, this variation increased considerably(MLE = -0.002, ELBN = -0.052, NAS = 0.034).
+Once again, the SD for both of the binning methods were $\ge$ 2.8 times that of the MLE method.
 
@@
 
-![D.](figures/neon_relationship.png)
+![Need to load data on githib and rerun to make figure match new format above.](figures/neon_relationship.png)
 
@@
 Previous work has investigated the accuracy and inherent biases associated with different estimation methods.
-However, how these innaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detecting variation in size spectra parameters across environmental gradients with confidence.
+However, how these innaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
 Here, we sampled body sizes from known distributions and estimated their parameters across a hypothetical gradient in order to assess the relationship coefficients obtained.
+Likewise, we compared how the interpretation of previously published results could change depending on the methodology used to estimate the relationship.
 
-Generally, there was large agreement in the direction and magnitude of average relationship coefficient estimates across methods.
-The estimate from the MLE method was always closer to the known relationship, and always had the smallest variation around the estimate.
+Generally, the estimate from the MLE method was always closer to the known relationship, and always had the smallest variation around the estimate.
+One of the benefits of using MLE is that all data points are retained for use within the model.
 
-All of the methods performed equally well when the $\lambda$ value did not vary across the hypothetical gradient, with a Type I error rate of \~5%, exactly as would be expected under the assumptions of frequentist statistics.
-This indicates that previous studies which have detected no difference in size spectra parameters across sites can be interpreted with confidence, regardless of the method used.
+Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies.
+However, aggregating individuals into logarthmic bins removes a large source of the variation within the data by collapsing body size variation into a single values within bins.
+For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
+Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh ~2 grams or ~4 grams.
+By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use.
 
@@
 
-Alarmingly, the performance of all methods declined dramatically when trying to detect smaller variation in $\lambda$ values.
+Alarmingly, the performance of the ELBn and NAS all methods declined dramatically when trying to detect smaller variation in $\lambda$ values.
 We simulated observations from distributions varying by an absolute value of 0.2 units.
 This range of variation is similar to that which has been reported in many seminal works on variation in size spectra parameters.
-When using the MLE method, significant relationships were only detected \~88% of the time, whereas both of the binning methods only detected significant trends \~20% of the time.
-Of the simulation replicates which were significant, nearly all of them were negative across all methods except for one single replicate using the NAS method which was positive (Figure.
-5).
 
@@
 This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible, and the magnitude of the relative change is a reasonable estimate.
-Given that all of the data within a study is treated identically, the the over all chnage in size spectra parameters is likely reasonable.
+Given that all of the data within a study is treated identically, the the over all change in size spectra parameters is likely reasonable.
 However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to compare the relative changes across different published studies which use different methods.
@@
 ## Concluding Remarks
+Both of the binning methods can perform as well as MLE under certain circumstances (large sample size, shallow $\lambda$).
+However, it would be extremely difficult to know if this was the case beforehand.
+Furthermore, in order to confirm this result, MLE methods would likely need to be employed.
+With the publication of the sizespectra package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
+Therefore, we recommend using this in all future studies of size spectra relationships.
 

### 2022-08-27 2_mean_SD_slope_plots.R

@@ -1,0 +1,111 @@
+# 2 mean and SD estimate plots
+
+library(tidyverse)
+source("custom_functions.R")
+library(ggpubr)
+plb <- read_csv("results/PLB_sim.csv")
+plb
+ggplot(plb, 
+       aes(x = name,
+           y = mu_slope, 
+           ymin = mu_slope - sd_slope, 
+           ymax = mu_slope + sd_slope))+
+  geom_pointrange()
+
+ggplot(plb, 
+       aes(y = name,
+           color = name,
+           x = mu_slope, 
+           xmin = mu_slope - sd_slope, 
+           xmax = mu_slope + sd_slope))+
+  geom_pointrange(size = 2) +
+  geom_errorbar(inherit.aes = FALSE,
+                aes(y = name, 
+                    color = name, 
+                    xmin = p25, 
+                    xmax = p975)) +
+  scale_color_viridis_d(option = "plasma") +
+  labs(x = "Slope estimate",
+       y = "Method") +
+  theme_bw() +
+  theme(legend.position="none")
+
+plot_mean_slope_estimate <- function(dat){
+  ggplot(dat, 
+         aes(y = name,
+             color = name,
+             x = p50, 
+             xmin = p25, 
+             xmax = p975))+
+    geom_pointrange() +
+    scale_color_viridis_d(option = "plasma") +
+    labs(x = "Slope estimate",
+         y = "Method") +
+    theme_bw() +
+    theme(legend.position="none")
+}
+
+(pa <- plot_mean_slope_estimate(plb)+
+  geom_vline(xintercept =  -0.5, linetype = "dashed"))
+ggsave(file = "figures/plb_slope_estimate.png",
+       height = 3,
+       width = 5)
+
+pb <- plot_mean_slope_estimate(plb_10)+
+  geom_vline(xintercept =  -0.5, linetype = "dashed")
+ggsave(file = "figures/plb_10_slope_estimate.png",
+       height = 3,
+       width = 5)
+
+pc <- plot_mean_slope_estimate(plb_small)+
+  geom_vline(xintercept =  -0.1, linetype = "dashed")
+ggsave(file = "figures/plb_small_slope_estimate.png",
+       height = 3,
+       width = 5)
+
+pd <- plot_mean_slope_estimate(plb_static_b) +
+  geom_vline(xintercept =  0, linetype = "dashed")
[…]

### 2022-08-27 AMD_estimate.R

@@ -61,17 +61,21 @@ amd_relationship %>%
 
 # plot the mean \pm STD.E of the estimated relationship across the gradient based on method used. 
 amd_relationship %>%
-  ggplot(aes(y = estimate, 
-             ymin = estimate - std.error, 
-             ymax = estimate + std.error,
-             x = name,
+  ggplot(aes(x = estimate, 
+             xmin = estimate - std.error, 
+             xmax = estimate + std.error,
+             y = name,
              color = name)) +
   geom_pointrange()+
   theme_bw() +
-  labs(y = "Relationship estimate",
-       title = "Mean +/- Std. Error Beta across AMD gradient")
+  labs(x = "Relationship estimate",
+       y = "Method",
+       title = "Mean +/- Std. Error Beta across AMD gradient") +
+  theme(legend.position="none")
 
-ggsave("figures/amd_relationship.png")
+ggsave("figures/amd_relationship.png",
+       height = 3,
+       width = 5)
 
 
 

### 2022-08-27 custom_functions.R

@@ -282,7 +282,7 @@ sim_result <- function(n = 1000,
 
 plot_sim <- function(sim_data,
                      display = FALSE,
-                     adjust = 2){
+                     adjust = 1){
   # plot each regression "rep"
   main_plot <- ggplot(sim_data,
          aes(x = env_gradient,

### 2022-08-30 1_simulations.R
updated custom functions and simulation scripts. Halfeye plots

@@ -21,6 +21,7 @@ set.seed(2806)
 
 library(tidyverse)
 library(sizeSpectra) #bounded power law and MLE functions
+library(tidybayes) # for nicer looking plots
 source("custom_functions.R")
 
 # Simulation and summary --------------------------------------------------

### 2022-08-30 custom_functions.R

@@ -315,22 +315,44 @@ plot_sim <- function(sim_data,
                   "_main.png"))
   
 if(sim_data$distribution[1] == "PLB"){
+  #### this is the old plot, keeping for now ####
+  # estimate_density <- sim_data %>%
+  #   ggplot(aes(y = ..scaled..,
+  #              x = estimate,
+  #              fill = name)) +
+  #   geom_density(alpha = 0.5,
+  #                adjust = adjust) +
+  #   theme_bw() +
+  #   facet_wrap(~known_b) +
+  #   geom_vline(aes(xintercept = known_b),
+  #              size = 1,
+  #              linetype = "dashed",
+  #              color = "black")+
+  #   scale_fill_viridis_d(option = "plasma") +
+  #   labs(title = sim_data$distribution,
+  #        x = "slope estimate") +
+  #   NULL
+  #### New 8/30/22 "halfeye" plot
   estimate_density <- sim_data %>%
-    ggplot(aes(y = ..scaled..,
-               x = estimate,
-               fill = name)) +
-    geom_density(alpha = 0.5,
-                 adjust = adjust) +
+    mutate(Model = factor(name,
+                          levels = 
+                            c("MLE",
+                              "ELBn", 
+                              "NAS"))) %>%
+    ggplot(
+      aes(x = estimate, 
+          y = Model,
+          fill = Model)) +
+    stat_halfeye(.width = c(0.66, 0.95)) +
+    scale_fill_manual(
+      values = c("#019AFF", "#FF914A", "#FF1984" )) +
     theme_bw() +
-    facet_wrap(~known_b) +
     geom_vline(aes(xintercept = known_b),
-               size = 1,
-               linetype = "dashed",
-               color = "black")+
-    scale_fill_viridis_d(option = "plasma") +
-    labs(title = sim_data$distribution,
-         x = "slope estimate") +
-    NULL}
+               linetype = "dashed") +
+    labs(
+      x = "Lambda estimate") +
+    facet_wrap(~known_b)
+  }
   
   if(sim_data$distribution[1] == "tpareto"){
     estimate_density <- sim_data %>%
@@ -383,20 +405,42 @@ if(sim_data$distribution[1] == "PLB"){
                           substitute(sim_data),
                           ".csv"))
   # plot density of relationship estimate
-    relationship_density <- relationship_estimate %>%
-    ggplot(aes(y = ..scaled..,
-               x = estimate, 
-               fill = name)) + 
-    geom_density(alpha = 0.5,
[…]

### 2022-08-30 AMD_estimate.R
update empirical figs to match new style

@@ -27,13 +27,20 @@ dw_range = range(amd_dat$dw, na.rm = TRUE)
   unnest(cols = method_compare))
 
 # plot parameter estimates (slope or exponent) across AMD gradient
-ggplot(amd_result,
-       aes(x = pca1,
+amd_result %>%
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
+  ggplot(aes(x = pca1,
            y = estimate,
-           color = name))+
+           color = Model))+
   geom_point() +
   stat_smooth(method = "lm",
               se = FALSE)+
+  scale_color_manual(
+    values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
   labs(title = "Change in exponent across an AMD gradient", 
        x = "AMD gradient (PCA)",
@@ -61,12 +68,19 @@ amd_relationship %>%
 
 # plot the mean \pm STD.E of the estimated relationship across the gradient based on method used. 
 amd_relationship %>%
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
   ggplot(aes(x = estimate, 
              xmin = estimate - std.error, 
              xmax = estimate + std.error,
-             y = name,
-             color = name)) +
+             y = Model,
+             color = Model)) +
   geom_pointrange()+
+  scale_color_manual(
+    values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
   labs(x = "Relationship estimate",
        y = "Method",

### 2022-08-30 NEON_estimate.R

@@ -59,12 +59,19 @@ neon_relationship %>%
   write_csv("results/NEON_estimates.csv")
 
 neon_relationship %>%
-  ggplot(aes(y = estimate, 
-             ymin = estimate - std.error, 
-             ymax = estimate + std.error,
-             x = name,
-             color = name)) +
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
+  ggplot(aes(x = estimate, 
+             xmin = estimate - std.error, 
+             xmax = estimate + std.error,
+             y = Model,
+             color = Model)) +
   geom_pointrange()+
+  scale_fill_manual(
+    values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
   labs(y = "Relationship estimate",
        title = "Mean +/- Std. Error Beta across Neon")

### 2022-08-30 2_mean_SD_slope_plots.R
delete mean_sd_slope script, changing to halfeye plots

@@ -1,111 +1,0 @@
-# 2 mean and SD estimate plots
-
-library(tidyverse)
-source("custom_functions.R")
-library(ggpubr)
-plb <- read_csv("results/PLB_sim.csv")
-plb
-ggplot(plb, 
-       aes(x = name,
-           y = mu_slope, 
-           ymin = mu_slope - sd_slope, 
-           ymax = mu_slope + sd_slope))+
-  geom_pointrange()
-
-ggplot(plb, 
-       aes(y = name,
-           color = name,
-           x = mu_slope, 
-           xmin = mu_slope - sd_slope, 
-           xmax = mu_slope + sd_slope))+
-  geom_pointrange(size = 2) +
-  geom_errorbar(inherit.aes = FALSE,
-                aes(y = name, 
-                    color = name, 
-                    xmin = p25, 
-                    xmax = p975)) +
-  scale_color_viridis_d(option = "plasma") +
-  labs(x = "Slope estimate",
-       y = "Method") +
-  theme_bw() +
-  theme(legend.position="none")
-
-plot_mean_slope_estimate <- function(dat){
-  ggplot(dat, 
-         aes(y = name,
-             color = name,
-             x = p50, 
-             xmin = p25, 
-             xmax = p975))+
-    geom_pointrange() +
-    scale_color_viridis_d(option = "plasma") +
-    labs(x = "Slope estimate",
-         y = "Method") +
-    theme_bw() +
-    theme(legend.position="none")
-}
-
-(pa <- plot_mean_slope_estimate(plb)+
-  geom_vline(xintercept =  -0.5, linetype = "dashed"))
-ggsave(file = "figures/plb_slope_estimate.png",
-       height = 3,
-       width = 5)
-
-pb <- plot_mean_slope_estimate(plb_10)+
-  geom_vline(xintercept =  -0.5, linetype = "dashed")
-ggsave(file = "figures/plb_10_slope_estimate.png",
-       height = 3,
-       width = 5)
-
-pc <- plot_mean_slope_estimate(plb_small)+
-  geom_vline(xintercept =  -0.1, linetype = "dashed")
-ggsave(file = "figures/plb_small_slope_estimate.png",
-       height = 3,
-       width = 5)
-
-pd <- plot_mean_slope_estimate(plb_static_b) +
-  geom_vline(xintercept =  0, linetype = "dashed")
[…]

### 2022-08-30 NEON_estimate.R
neon plots updated to color scheme

@@ -27,14 +27,23 @@ dw_range = range(neon_dat$dw, na.rm = TRUE)
     select(-data) %>%
     unnest(cols = method_compare))
 
-ggplot(neon_result,
-       aes(x = mat.c,
+neon_result %>%
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
+ggplot(aes(x = mat.c,
            y = estimate,
-           color = name))+
+           color = Model))+
   geom_point() +
   stat_smooth(method = "lm",
               se = FALSE)+
   theme_bw() +
+  scale_color_manual(
+    values = c("#019AFF",
+               "#FF914A",
+               "#FF1984" )) +
   labs(title = "Change in exponent across NEON", 
        x = "Mean annual air temp",
        y = "Slope/exponent estimate")
@@ -70,7 +79,7 @@ neon_relationship %>%
              y = Model,
              color = Model)) +
   geom_pointrange()+
-  scale_fill_manual(
+  scale_color_manual(
     values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
   labs(y = "Relationship estimate",

### 2022-08-30 delete_this_script.R
test push to save PAT

@@ -1,0 +1,2 @@
+# delete this script
+# saving and pushing this to get my personal access token to save on this machine

### 2022-08-30 delete_this_script.R
delete test scripts

@@ -1,2 +1,0 @@
-# delete this script
-# saving and pushing this to get my personal access token to save on this machine

### 2022-08-31 Manuscript.Rmd
cleaning up MS

@@
 
+![Distribution of all lambda estimates across sites when there is no relationaship.
+The coeficient estimates for the binning methods tend to be shallower and have wider distributions than the MLE method.
+](figures/PLB_static_b_est_b_density.png)
+
 ![Individual regression estimates when no relationship exists across the hypothetical environmental gradient.
-The methods are separated by rows, and the left and right column show relationships which were non-significant and significant, respectively.
-The individual regressions coefficients are colored-coded to indicate positive and negative relationship estimates.
-](figures/static_b_lambda_spaghetti.png)
+The methods are separated by rows.
+The median for all methocs is approximately zero, but the distribution for the MLE method is narrower than the two binning methods.
+](figures/PLB_static_b_relationship_density.png)
 
@@
 ![Distribution of the estimated relationship for each method.
-Here, $\lambda$ varied from -1.9 to -2.1, with a known relationaship of -0.1 across the gradient indicated by the dashed vertical line.](figures/PLB_sim_small_relationship_density.png)
-
-![Linear regressions across the hypothetical gradient when lambda varies from -1.9 to -2.1.
-Each method is displayed in the rows, and the left colum shows the non-significant estimates, while the right column shows the relationships which were signficant ($\beta_1$ coefficient p-value \< 0.05).
-The colors indicate when the beta coefficient estimate was negative (pink) and when it was positive (purple).](figures/small_lambda_spaghetti.png)
+Here, $\lambda$ varied from -1.9 to -2.1, with a known relationaship of -0.1 across the gradient indicated by the dashed vertical line.
+The distribution of relationship estimates was approximately centered at the known relationship for all methods, with a narrower distribution observed for the MLE estimates.
+](figures/PLB_sim_small_relationship_density.png)
 
@@
 Vertical line is the known relationship.
-All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.](figures/shallow_lambda_relationship_density.png)
+All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.](figures/PLB_shallow_lambda_relationship_density.png)
 

### 2022-08-31 custom_functions.R

@@ -351,7 +351,8 @@ if(sim_data$distribution[1] == "PLB"){
                linetype = "dashed") +
     labs(
       x = "Lambda estimate") +
-    facet_wrap(~known_b)
+    facet_wrap(~known_b,
+               scales = "free_x")
   }
   
   if(sim_data$distribution[1] == "tpareto"){

### 2022-08-31 vary_n.R

@@ -6,7 +6,7 @@ source("custom_functions.R")
 
 # read in datasets with variable sample size, n
 # All other params are as in main analysis
-n100 <- readRDS("data_sim/PLB_sim_n100_dat.rds")
+n200 <- readRDS("data_sim/PLB_sim_n200_dat.rds")
 n500 <- readRDS("data_sim/PLB_sim_n500_dat.rds")
 n5000 <- readRDS("data_sim/PLB_sim_n5000_dat.rds")
 n10000 <- readRDS("data_sim/PLB_sim_n10000_dat.rds")
@@ -18,15 +18,15 @@ n1000$n <- 1000
 # Filter out all reps which have an estimate == NA
 
 # make vector of reps which have an NA value for the estimate
-rep_na <- n100$rep[which(is.na(n100$estimate))]
+#rep_na <- n100$rep[which(is.na(n100$estimate))]
 
 # filter out all the reps with NA values
 # the !rep means (reps which do not match values in rep_na)
-PLB_sim_n100 <- n100 %>%
-  filter(!rep %in% rep_na)
+#PLB_sim_n100 <- n100 %>%
+#  filter(!rep %in% rep_na)
 
 # combine all data sets
-dat <- bind_rows(n100, n500, n5000, n10000, n1000)
+dat <- bind_rows(n200, n500, n5000, n10000, n1000)
 
 # How does the lambda estimate vary by sample size?
 ggplot(dat, aes(x = n, y = estimate, color = name)) +
@@ -45,23 +45,46 @@ ggsave("figures/lambda_n_5_sites.png", width = 8, height = 8)
 
 
 # how does estimate of lambda vary with sample size?
-ggplot(dat,
-       aes(x = estimate,
-           y = ..scaled..,
-           fill = name)) +
-  geom_density(alpha = 0.5,
-               adjust = 2) +
-  facet_wrap(n ~ known_b,
-             ncol = 5,
-             labeller = label_both) +
-  geom_vline(aes(xintercept = known_b),
-             size = 1,
-             alpha = 0.75,
-             color = "black")+
-  scale_fill_viridis_d(option = "plasma") +
-  labs(title = "Sample size and known b",
-       x = "slope estimate") +
+# ggplot(dat,
+#        aes(x = estimate,
+#            y = ..scaled..,
+#            fill = name)) +
+#   geom_density(alpha = 0.5,
+#                adjust = 2) +
+#   facet_wrap(n ~ known_b,
[…]

### 2022-09-04 AMD_estimate.R
update figures, combine empirical, alt version main plot

@@ -44,9 +44,13 @@ amd_result %>%
   theme_bw() +
   labs(title = "Change in exponent across an AMD gradient", 
        x = "AMD gradient (PCA)",
-       y = "Slope/exponent estimate")
+       y = "Slope/exponent estimate") +
+  theme(legend.position = "none")
 # save the plot
-ggsave("figures/AMD_plot.png")
+ggsave("figures/AMD_plot.png",
+       width = 1600,
+       height = 753, 
+       units = "px")
 
 # estimate the beta_1_ relationship across the gradient
 amd_relationship <- amd_result %>%

### 2022-09-04 empirical plot.R

@@ -1,0 +1,23 @@
+# empirical data plot
+
+# combine the AMD and NEON data into a single multipanel plot
+
+library(ggplot2)
+library(png)
+library(grid)
+# library(gridExtra)
+library(ggpubr)
+amd_a <- readPNG("figures/AMD_plot.png")
+amd_b <- readPNG("figures/amd_relationship.png")
+neon_a <- readPNG("figures/NEON_plot.png")
+neon_b <- readPNG("figures/neon_relationship.png")
+ggarrange(rasterGrob(amd_a),
+             rasterGrob(amd_b),
+             rasterGrob(neon_a),
+             rasterGrob(neon_b),
+             labels = c("A", "B", "C", "D"))
+
+ggsave("figures/empirical_combined.png",
+       height = 6,
+       width = 4, 
+       units = "in")

### 2022-09-04 main_plot_v2.R

@@ -1,0 +1,40 @@
+# main plot version 2
+
+# if we like this replace in custom functions script and re run main simulation. 
+
+library(tidyverse)
+
+sim_data <- readRDS("data_sim/PLB_sim_dat.rds")
+sim_data %>%
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
+  ggplot(aes(x = env_gradient,
+             y = estimate, 
+             group = rep,
+             color = Model)) +
+  scale_color_manual(
+    values = c("#019AFF", "#FF914A", "#FF1984" )) +
+  stat_smooth(geom = "line",
+              method = "lm",
+              alpha = 0.15,
+              se = FALSE, 
+              color = "black") +
+  geom_point(alpha = 0.5) +
+  facet_wrap(~Model)+
+  labs(title = sim_data$distribution,
+       subtitle = 
+         paste0("m_range =(",
+                sim_data$m_lower,
+                ",",
+                sim_data$m_upper,
+                ") \nb = (",
+                min(sim_data$known_b),
+                ",",
+                max(sim_data$known_b),
+                ")")) +
+  theme_bw() +
+  theme(legend.position="none")
+ggsave("figures/PLB_main_v2.png", width = 2100, height = 2100, units = "px")

### 2022-09-04 vary_n.R

@@ -3,6 +3,7 @@
 library(tidyverse)
 library(sizeSpectra) #bounded power law and MLE functions
 source("custom_functions.R")
+library(tidybayes)
 
 # read in datasets with variable sample size, n
 # All other params are as in main analysis
@@ -28,22 +29,25 @@ n1000$n <- 1000
 # combine all data sets
 dat <- bind_rows(n200, n500, n5000, n10000, n1000)
 
-# How does the lambda estimate vary by sample size?
-ggplot(dat, aes(x = n, y = estimate, color = name)) +
-  geom_point(alpha = 0.1) +
-  #stat_smooth(method = "loess") +
-  stat_summary(geom = "line",  fun = mean, size = 1.5) +
-  geom_hline(aes(yintercept = known_b), linetype = "dashed") +
-  facet_wrap(~known_b,
-             scales = "free") +
-  theme_bw() +
-  labs(title = "Lambda ~ n",
-       caption = "dashed line is known lambda value") +
-  NULL
+# labda ~n line graph old? ####
+# # How does the lambda estimate vary by sample size?
+# ggplot(dat, aes(x = n, y = estimate, color = name)) +
+#   geom_point(alpha = 0.1) +
+#   #stat_smooth(method = "loess") +
+#   stat_summary(geom = "line",  fun = mean, size = 1.5) +
+#   geom_hline(aes(yintercept = known_b), linetype = "dashed") +
+#   facet_wrap(~known_b,
+#              scales = "free") +
+#   theme_bw() +
+#   labs(title = "Lambda ~ n",
+#        caption = "dashed line is known lambda value") +
+#   NULL
+# 
+# ggsave("figures/lambda_n_5_sites.png",
+#        width = 8,
+#        height = 8)
 
-ggsave("figures/lambda_n_5_sites.png", width = 8, height = 8)
-
-
+# lambda ~ n density old? ####
 # how does estimate of lambda vary with sample size?
 # ggplot(dat,
 #        aes(x = estimate,
@@ -63,6 +67,8 @@ ggsave("figures/lambda_n_5_sites.png", width = 8, height = 8)
 #        x = "slope estimate") +
 #   theme_bw() +
 #   NULL
+
+# lambda estimate halfeye ####
 dat %>%
   mutate(Model = factor(name,
                         levels = 
@@ -85,6 +91,7 @@ dat %>%
              ncol = 5,
              labeller = label_both,
              scales = "free_x") +
+  theme(legend.position = "none") +
   NULL
 
[…]

### 2022-09-04 Manuscript.Rmd
update manuscript

@@
       - '--lua-filter=author-info-blocks.lua'
-   html_document: default
-      pandoc_args:
-      - '--lua-filter=scholarly-metadata.lua'
-      - '--lua-filter=author-info-blocks.lua'
 header-includes:
@@
   echo = TRUE)
+library(tidyverse)
 ```
@@
 Size metrics have commonly been used in marine systems, and are increasingly being applied to assess the condition of freshwater ecosystems (Martinez et al.
-2016, Pomeranz 2018)[@martinez2016; @pomeranz2018].
+2016, Pomeranz 2018)[@martinez2016; @pomeranz2019].
 
@@
 These values are based on empirical body sizes of stream benthic communities reported in Pomeranz et al.
-(2020)[@pomeranz2020].
-Our results are not dependent on the range of body sizes, and we show the results of other ranges in the supplemental information.
+(2020)[@pomeranz2022].
+Our results are not dependent on the range of body sizes, and we show the results of other body size ranges in the supplemental information.
 
@@
 
-The supplemental information contains results when the value of n is varied (or maybe include in main?).
 Each value of $\lambda$ was assumed to come from a community across a hypothetical environmental gradient $X$.
@@
 
-### Sample size, n
+### Sample size, $n$
 
@@
 We tested values of $n = 200, 500, 1000, 5000, 10 000$.
-(originally had n=100, but thre errors based on random seed.
-Bumped it to 200 to try and get consistent error-free runs)
 
@@
 
-It is possible that our sampling or simulation framework could impact the results obtained, possibly by inflating the type I error rate.
+It is possible that our sampling or simulation framework could bias the results obtained.
 In order to assess whether or not our simulations are robust to this, we performed independent random samples from a bounded power law for five communities across a hypothetical gradient as described above, but we set $\lambda = -2$ for all five communities.
@@
 
-For the main analysis, we were interested in determining whether or not the methods were able to detect relatively coarse changes to size spectra parameters, specifically, a change of 0.5 units across each value of $X$, or an absolute change or 1.0 units.
+For the main analysis, we were interested in determining whether or not the methods were able to detect relatively coarse changes to size spectra parameters, specifically, a change of 0.5 units across each value of $X$, or an absolute change in $\lambda$ values of 1.0 units.
 However, many seminal works on variation in size spectra relationships in freshwater are on a much finer scale (on the order 0f 0.1 to 0.25 units).
@@
 The count in each bin is normalized in the same way.
-This method has been used by [@pomeranz2018; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).
+This method has been used by [@pomeranz2019; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).
+Previous work has investigated the biases in $\lambda$ estimates when using different methods (Edwards et al.
+(2017)[-@edwards2017], Edwards et al.
+2020[-@edwards2020], White et al.
+2008[-@white2007]).
+The focus of the present work is to investigate biases when estimating the relationship or relative change in parameters across an environmental gradient.
+Hence, we do not discuss the $\lambda$ estimates in the main text, but the results are presented in the Supplmental Information.
 
@@
 The distribution of the relationship coefficient, $\beta_1$, were plotted compared to the known relationship.
-Likewise, the mean $\beta_1$ estimate ($\pm$ 1 SD) for all simulations were also plotted.
 
@@
 We re-analyze two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients.
-In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage stress gradient.
-Details of the sample collection and processing can be found in [@pomeranz2018].
+In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage (AMD) stress gradient.
+Details of the sample collection and processing can be found in [@pomeranz2019].
 Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
@@
 This data has been analyzed previously using size spectra methods as described in [@pomeranz2022].
-Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([https://www.neonscience.org/data-collection/macroinvertebrates], or macroinvertebrate DPI pubs).
+Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([<https://www.neonscience.org/data-collection/macroinvertebrates>], or macroinvertebrate DPI pubs).
+Coefficient estimates ($\pm$ 1 standard deviation) are compared across methods.
+This allows us to determine whether or not the main results would differ depending on the method used.
 
@@
 
+Alternate plot version
+
+![alternate plot version, thoughts on which you prefer??](figures/PLB_main_v2.png)
+
 All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient (mean $\beta_1$ estimates ranged from -0.419 to -0.498).
@@
 
-![Mean estimate of the relationship coefficient for main simulation (A), with ten sites (B), with small variation in lambda (C), and with no relationship.Error bars indicate 1 standard deviation.](figures/slopes_combined.png)
+\newpage
 
-![Distribution of relationship coefficient estimates.
+![Halfeye plots for the estimated relationship coefficient estimates across 1000 replicates and three different methods.
+The point below each distribution plot is the median estimate, the thick and thin bars represent the 66th and 95th percentile, respectively.
 Vertical line is the known relationship.
-All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.](figures/PLB_sim_relationship_density.png)
+The median estimate for the MLE method was closest to the known relationship, and had the narrowest distribution across replicates.](figures/PLB_sim_relationship_density.png)
 
-Interestingly, there was an interactive effect of the estimate accuracy between sample size and the value of $\lambda$.
-All methods were more accurate with larger sample sizes, and smaller values of $\lambda$.
+\newpage
 
-![Distribution of size spectra parameter estimates.
-Vertical line is the known parameter wich describes the bounded power law distribution from which the body size estimates were sampled.
-As n increases (top to bottom) and $\lambda$ increases (left to right), the accuracy of the estimate improves across all methods.](figures/n_vary_est_b.png)
+```{r, echo=FALSE, warning=FALSE}
+read.csv("results/PLB_sim.csv") %>%
+  mutate(Model = name) %>%
+  select(Model, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  arrange(p50, p975)
+```
 
 ![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).](figures/n_vary_main.png)
+
+\newpage
 
@@
 The mean $\beta_1$ estimates for all three methods ranged from -0.0005 to 0.0002.
-However, once again, the SD for the NAS and ELBn methods were ~3x larger than the SD for the MLE method.
-
-![Distribution of all lambda estimates across sites when there is no relationaship.
-The coeficient estimates for the binning methods tend to be shallower and have wider distributions than the MLE method.
-](figures/PLB_static_b_est_b_density.png)
+However, the distribution of coefficient estimates was much wider for the two binning methods; the 95th percentile of the MLE distributions is narrower thatn the 66th percentiles for the NAS and ELBn methods.
 
@@
 The methods are separated by rows.
-The median for all methocs is approximately zero, but the distribution for the MLE method is narrower than the two binning methods.
-](figures/PLB_static_b_relationship_density.png)
+The median for all methocs is approximately zero, but the distribution for the MLE method is narrower than the two binning methods.](figures/PLB_static_b_relationship_density.png)
+
+```{r, echo=FALSE, warning=FALSE}
+read.csv("results/PLB_static_b.csv") %>%
+  mutate(Model = name) %>%
+  select(Model, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  arrange(p50, p975)
+```
 
@@
 The performance of all methods to detect small variation in the $\lambda$ parameter (between -1.9 to -2.1) across the environmental gradient was also investigated.
-The mean $\beta_1$ coefficient estimates ranged from -0.10 to -0.09.
-The mean estimate for the MLE method was -0.1, which exaclty matched the value for the hypothetical relationship.
-Likewise, the upper 97.5 percentile of the distribution was below 0, meaning that most of the estimates had the correct sign.
+The median $\beta_1$ coefficient estimates for the MLE method was -0.1, which exactly matched the value for the hypothetical relationship, while the other two methods both had median estimates of 0.9.
+The upper 97.5 percentile of the distribution was below 0 for the MLE method, meaning that most of the estimates had the correct sign.
 In contrast, the upper 97.5 percentile for both the NAS and ELBn methods were greater than 0 at 0.027 and 0.052, respectively.
@@
 Here, $\lambda$ varied from -1.9 to -2.1, with a known relationaship of -0.1 across the gradient indicated by the dashed vertical line.
-The distribution of relationship estimates was approximately centered at the known relationship for all methods, with a narrower distribution observed for the MLE estimates.
-](figures/PLB_sim_small_relationship_density.png)
+The distribution of relationship estimates was approximately centered at the known relationship for all methods, with a narrower distribution observed for the MLE estimates.](figures/PLB_sim_small_relationship_density.png)
 
-Once again, the SD for both of the binning methods were $\ge$ 2.8 times that of the MLE method.
+```{r, echo=FALSE, warning=FALSE}
+read.csv("results/PLB_sim_small.csv") %>%
+  mutate(Model = name) %>%
+  select(Model, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  arrange(p50, p975)
+```
 
@@
 
-For both empirical datasets, the direction and magnitude of chnage (i.e.
+For both empirical data sets, the direction and magnitude of change (i.e.
 $\beta_1$ coefficients) are generally in agreement.
@@
 
-**NOTE** combine all the figures below into one with 4 panels (AMD = A, B; NEON = C, D)
+**NOTE** I need to re-plot these in ggplot and and stitich together. Currently stitching pre-made png files, I think that's why it looks so funky.
 
-![A](figures/AMD_plot.png)
+\newpage
 
-![B.Relationship estimates across AMD gradient](figures/amd_relationship.png)
+![Figure of empirical data estimates.
+All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.](figures/empirical_combined.png)
 
-![C.](figures/NEON_plot.png)
+\newpage
 
-![Need to load data on githib and rerun to make figure match new format above.](figures/neon_relationship.png)
+The $\beta_1$ coefficient estimates for the AMD data are all positive and of relatively similar magnitudes across methods.
+The MLE estimate has the smallest magnitude and SD (0.063 $\pm$ 0.011 SD), The ELBn method is slightly larger and has a wider SD : (0.072 $\pm$ 0.018).
+Finally, the NAS method has the largest estimate and an intermediate standard deviation (0.076 $\pm$ 0.013).
+The range of the gradient in the AMD data is relatively small (9.5), meaning that the absolute change in the size spectra parameter ($\lambda$) across the AMD gradient are similar regardless of method used (range of absolute: 0.59 to 0.72).
 
-Because the $\beta_1$ coefficient estimates are similar, and the range of the gradient in the AMD data is relatively small (9.5), the absolute change in the size spectra parameter across the AMD gradient are similar regardless of method used (range: 0.59 to 0.72).
-Likewise, the absolute change in size spectra parameters across the NEON temperature gradient ranges from 0.06 to 0.165, depending on the method used.
-
-*alternative* absolute difference of \~0.1 units in $\lambda$ across gradients is similar to differences reported in other studies (seasonally: 0.16, Mcgarvey and Kirk; landuse: 0.15 Martinez; O'Gorman 2017 \~0.25, Yvon and Dossena papers \~0.1?
-[need to redownload some of these and compare])[@mcgarvey2018; @dossena2012; @ogorman2017]
+Similar to the AMD results, the $\beta_1$ coefficient estimates for the NEON data are all negative and have similar magnitudes across methods.
+The MLE estimate is once again the smallest (-0.006 $\pm$ 0.001 SD).
+The NAS and ELBn methods are larger at -0.004 $\pm$ 0.002, and -0.002 $\pm$ 0.002, respectively.
+Although the estimates are similar and have a small magnitude, the range of the gradient in this data is larger (\~30$\circ$C), meaning that the absolute change in $\lambda$ estimates ranges from 0.06 to 0.17, depending on the method used.
 
@@
 Upon closer inspection, the values of size spectra parameters in the empirical data were considerably shallower than -2.
-Given that we found the performance of all methods increased with shallower $\lambda$'s we wanted to investigate how the methods performed with simulated values more similar to the empirical estimates of parameters describing size spectra relationships.
-Therefore, we repeated the simulation process as in the main analysis, but used $\lambda$ values ranging from -1.1 to -1.5.
+Given that we found the performance of all methods increased with shallower $\lambda$'s we wanted to investigate how the methods performed with simulated values closer to the empirical estimates of parameters describing size spectra relationships.
+Therefore, we repeated the simulation process as in the main analysis, but used $\lambda$ values ranging from -1.1 to -1.5, with a known relationship of $\beta_1 = -0.2$
+
+\newpage
 
@@
 
-We found generally less variability in the relationship estimates across a gradient of distributions described with shallower $\lambda$ parameters.
-All of the relationship coefficient estimates ($\beta_1$)'s across all replicates were significant regardless of method used.
-The confidence interval for $\beta_1$ estimate for the MLE method contained the true value 94.7% of the time.
-However, the true value was in the confidence interval for the ELBN and NAS method only 90.9% and 76.9% of the time, respectively.
-Once again, the mean width of the CI for the binning methods were ~2 times as large as for the MLE method.
-Mean $\pm$ SD CI width: MLE = 0.0442 $\pm$ 0.0196; ELBn = 0.0920 $\pm$ 0.0482, NAS = 0.0798 $\pm$ 0.0324.
-
-On average, the relationship across the gradient was under estimated, but by nearly an order of magnitude less than when simulating across steeper values of $\lambda$.
-Differences between known and estimated relationship coefficients: MLE -0.0007 ($\pm$ 0.0079); ELBn -0.0058 ($\pm$ 0.0177); NAS -0.0251 ($\pm$ 0.0124).
+We found generally less variability in the relationship estimates ($\beta_1$) across a gradient of distributions described with shallower $\lambda$ parameters.
+However, the median MLE estimate was once again closest to the hypothetical relationship and had the narrowest distribution.
+The median ELBn estimate was close to the hypothetical relationship but had a slightly wider distribution.
+Finally, 95% of the NAS estimates were greater than the hypothetical relationship.
+In general, the distributions from the different methodologies followed the same overall pattern, with the MLE estimates being the smallest, and the binning methods being larger (i.e., more positive).
 
@@
 
+```{r, echo=FALSE, warning=FALSE}
+read.csv("results/shallow_lambda.csv") %>%
+  mutate(Model = name) %>%
+  select(Model, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  arrange(p50, p975)
+```
+
+\newpage
+
 # Discussion
@@
 Previous work has investigated the accuracy and inherent biases associated with different estimation methods.
-However, how these innaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
-Here, we sampled body sizes from known distributions and estimated their parameters across a hypothetical gradient in order to assess the relationship coefficients obtained.
+However, how these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
+Here, we sampled body sizes from known distributions with varying parameters ($\lambda$) and estimated the coefficient of the relationship ($\beta_1$) across a hypothetical gradient in order to assess how the results would vary depnding on the method used.
 Likewise, we compared how the interpretation of previously published results could change depending on the methodology used to estimate the relationship.
 
-Generally, the estimate from the MLE method was always closer to the known relationship, and always had the smallest variation around the estimate.
-One of the benefits of using MLE is that all data points are retained for use within the model.
+Generally, the estimate from the MLE method was always closer to the known relationship in simulated data, and always had the smallest variation around the estimate.
 
 Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies.
-However, aggregating individuals into logarthmic bins removes a large source of the variation within the data by collapsing body size variation into a single values within bins.
+However, aggregating individuals into logarithmic bins removes a large source of the variation within the data by collapsing body size variation into a single values within bins.
 For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
-Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh ~2 grams or ~4 grams.
+Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2 grams or \~4 grams.
 By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use.
+
+One of the benefits of using MLE is that all data points are retained for use within the model.
 
@@
 If communities under study are in fact described by shallower exponents, the method used may not be critical in the conclusions reached.
-However, the MLE method performs as well or better than the two binning methods examined here under all contexts and should be the preferred method moving forward.
+However, the MLE method performs as well or better than the two binning methods examined here under all contexts and should be the preferred method in future studies.
 At a minimum, future studies should report MLE estimates to ensure that the results are not dependent on the methodology used.
-
-Alarmingly, the performance of the ELBn and NAS all methods declined dramatically when trying to detect smaller variation in $\lambda$ values.
-We simulated observations from distributions varying by an absolute value of 0.2 units.
-This range of variation is similar to that which has been reported in many seminal works on variation in size spectra parameters.
 
@@
 ## Concluding Remarks
+
 Both of the binning methods can perform as well as MLE under certain circumstances (large sample size, shallow $\lambda$).
@@
 
+\newpage
+
 # References
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
+```{=tex}
 \beginsupplement
-
+```
 ## Range of body sizes, $M$
 
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100. ](figures/PLB_small_m_main.png)
+![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_main.png)
 
-![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values.
-Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_est_b_density.png)
+\newpage
 
@@
 Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_relationship_density.png)
+
+\newpage
 
@@
 
+\newpage
+
 ![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values.
@@
 
+\newpage
+
 ![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
 Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_relationship_density.png)
+
+\newpage
 
@@
 
+\newpage
+
 ![Distribution of estimated $\lambda$ coefficient for ten sites across a hypothetical gradient with known values.
 All other parameters are the same as in the main analysis](figures/PLB_10_sites_est_b_density.png)
+
+\newpage
 
@@
 
+\newpage
+
 ### Three sites
+
 ![Individual regressions for three sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis.](figures/PLB_3_sites_main.png)
+
+\newpage
 
@@
 
+\newpage
+
 ![Distribution of estimated relationship ($\beta_1$) coefficient's for three sites across a hypothetical gradient with known value of 0.5.
@@
 
+\newpage
+
 ## Sample size
 
-**note** These plots still have n = 100. Need to re run and update if we end up going with n = 200
+![Distribution of size spectra parameter estimates.
+Vertical line is the known parameter (dashed line) wich describes the bounded power law distribution from which the body size estimates were sampled.
+As n increases (top to bottom) and $\lambda$ increases (left to right), the accuracy of the estimate improves across all methods.](figures/n_vary_est_b.png)
 
-![Plot showing the relationships between sample size and the estimated $\lambda$ parameters. the dashed line indicates the known value of $\lambda$.](figures/lambda_n_5_sites.png)
+\newpage
+
+![Individual regressions when varying sample size](figures/n_vary_main.png)
+
+\newpage
+
+![Distribtuion of the relationship coefficients with varying sample size](figures/n_vary_relationship_density.png)
+
+\newpage
+
+## $\lambda$ and relationship estimates
+
+In the main analysis, we simulate body size data from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.
+For the results presented in the main text, we held the number of sites at 5, scaled the environmental gradient from -1 to 1, and set the minimum and maximum body sizes to $0.0026$ and $1.2 *10^3$ respectively.
+Here, we plot the results of varying the number of sites (3, 10), increasing the scale of the environmental gradient (-100 to 100) and decreasing the range of body sizes (min = 1, max = 100).
+Generally, the results reported in the main manuscript are robust to changing these parameters: the MLE estimate is nearly always closer to the known parameters, and the variation in these estimates is usually smaller than the binning methods.
+
+\newpage
+
+### Number of sites = 3
+
+![Distribution of estimated $\lambda$ coefficient across a hypothetical gradient with known values (dashed line).
+The number of sites is reduced to 3, compared with 5 sites presented in the main anaysis.](figures/PLB_3_sites_est_b_density.png)
+
+\newpage
+
+![Relationship estimate](figures/PLB_3_sites_relationship_density.png)
+
+\newpage
+
+### Number of sites = 10
+
+![Distribution of estimated $\lambda$ coefficient across a hypothetical gradient with known values(dashed line).
+The number of sites is increased to 10, compared with 5 sites presented in the main anaysis.](figures/PLB_10_sites_est_b_density.png)
+
+\newpage
+
+![Relationship estimate](figures/PLB_10_sites_relationship_density.png)
+
+\newpage
+
+### Environmental Gradient
+
+![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values (dashed line).
+The range of the environmental gradient was increased (-100 to 100).](figures/PLB_large_x_est_b_density.png)
+
+\newpage
+
+![Relationship estimate](figures/PLB_large_x_relationship_density.png)
+
+\newpage
+
+### Range of body sizes = 1 to 100
+
+![Distribution of estimated \$\\lambda\$ coefficient for five sites across a hypothetical gradient with known values(dashed line).
+Range of body sizes is smaller than main anaysis and ranges from 1, to 100.](figures/PLB_small_m_est_b_density.png)
+
+\newpage
+
+![Relationship estimate](figures/PLB_small_m_relationship_density.png)

### 2022-09-04 Manuscript.tex
update manuscript

@@
 Junker\textsuperscript{2,3} \and Jeff S. Wesner\textsuperscript{4}}
-\date{30 June, 2022}
+\date{04 September, 2022}
 
@@
 \setstretch{1}
-\textsuperscript{1} Colorado Mesa University\\
+\textsuperscript{1} Colorado Mesa University, Grand Junction, CO, USA\\
 \textsuperscript{2} Great Lakes Research Center, Michigan Technological
@@
 condition of freshwater ecosystems (Martinez et al.~2016, Pomeranz
-2018)(\textbf{martinez2016?}, \textbf{pomeranz2018?}).
+2018)(Pomeranz et al. 2019, \textbf{martinez2016?}).
 
@@
 on empirical body sizes of stream benthic communities reported in
-Pomeranz et al.~(2020)(\textbf{pomeranz2020?}). Our results are not
+Pomeranz et al.~(2020)(Pomeranz et al. 2022). Our results are not
 dependent on the range of body sizes, and we show the results of other
-ranges in the supplemental information.
+body size ranges in the supplemental information.
 
@@
 sizes from distributions described by five different \(\lambda\)'s:
-(-1.5, -1.75, -2.00, -2.25, -2.5). The supplemental information contains
-results when the value of n is varied (or maybe include in main?). Each
-value of \(\lambda\) was assumed to come from a community across a
+(-1.5, -1.75, -2.00, -2.25, -2.5), such that our model for \(\lambda\)
+is:
+
+\[\lambda = Norm(\alpha -0.5X, \sigma)\]
+
+Each value of \(\lambda\) was assumed to come from a community across a
 hypothetical environmental gradient \(X\). Values of \(X\) for each
@@
 \hypertarget{sample-size-n}{%
-\subsubsection{Sample size, n}\label{sample-size-n}}
+\subsubsection{\texorpdfstring{Sample size,
+\(n\)}{Sample size, n}}\label{sample-size-n}}
 
@@
 sample size \(n\). We tested values of
-\(n = 200, 500, 1000, 5000, 10 000\). (originally had n=100, but thre
-errors based on random seed. Bumped it to 200 to try and get consistent
-error-free runs)
+\(n = 200, 500, 1000, 5000, 10 000\).
 
@@
 
-It is possible that our sampling or simulation framework could impact
-the results obtained, possibly by inflating the type I error rate. In
-order to assess whether or not our simulations are robust to this, we
-performed independent random samples from a bounded power law for five
-communities across a hypothetical gradient as described above, but we
-set \(\lambda = -2\) for all five communities. Hence, we are simulating
-a scenario where the size spectra relationship is invariant to our
-hypothetical environmental gradient, and we should expect the estimated
-relationship coefficient \(\beta_1 = 0\).
+It is possible that our sampling or simulation framework could bias the
+results obtained. In order to assess whether or not our simulations are
+robust to this, we performed independent random samples from a bounded
+power law for five communities across a hypothetical gradient as
+described above, but we set \(\lambda = -2\) for all five communities.
+Hence, we are simulating a scenario where the size spectra relationship
+is invariant to our hypothetical environmental gradient, and we should
+expect the estimated relationship coefficient \(\beta_1 = 0\).
 
@@
 spectra parameters, specifically, a change of 0.5 units across each
-value of \(X\), or an absolute change or 1.0 units. However, many
-seminal works on variation in size spectra relationships in freshwater
-are on a much finer scale (on the order 0f 0.1 to 0.25 units).
-Therefore, we performed simulations as described above, but varied
-\(\lambda\) from -1.9 to -2.1 across the hypothetical environmental
-gradient. This equates to a know relationship coefficient of
-\(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
+value of \(X\), or an absolute change in \(\lambda\) values of 1.0
+units. However, many seminal works on variation in size spectra
+relationships in freshwater are on a much finer scale (on the order 0f
+0.1 to 0.25 units). Therefore, we performed simulations as described
+above, but varied \(\lambda\) from -1.9 to -2.1 across the hypothetical
+environmental gradient. This equates to a know relationship coefficient
+of \(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
 change of 0.2 units.
@@
 each bin is normalized in the same way. This method has been used by
-(Pomeranz et al.~2018, I \textbf{think} McGarvey \textasciitilde2018,
-Blanchard (log 2, ?normalized?))(McGarvey et al. 2019,
-\textbf{pomeranz2018?}) and is referred to as the Normalized Abundance
-Spectrum (NAS).
+(Pomeranz et al. 2019, McGarvey et al. 2019) and is referred to as the
+Normalized Abundance Spectrum (NAS). Previous work has investigated the
+biases in \(\lambda\) estimates when using different methods (Edwards et
+al.~(2017)(2017), Edwards et al.~2020(2020), White et al.~2008(2007)).
+The focus of the present work is to investigate biases when estimating
+the relationship or relative change in parameters across an
+environmental gradient. Hence, we do not discuss the \(\lambda\)
+estimates in the main text, but the results are presented in the
+Supplmental Information.
 
@@
 relationship coefficient, \(\beta_1\), were plotted compared to the
-known relationship. Likewise, the confidence interval of \(\beta_1\) was
-assessed to determine the proportion of estimates which contained the
-true value of the known relationship.
+known relationship.
 
@@
 
-We re-analyze two datasets of benthic macroinvertebrate communities from
-stream habitats across two different gradients. In the first,
+We re-analyze two data sets of benthic macroinvertebrate communities
+from stream habitats across two different gradients. In the first,
 quantitative macroinvertebrate samples were collected from streams
-across an acid mine drainage stress gradient. Details of the sample
-collection and processing can be found in Pomeranz et
-al.~(2018)(\textbf{pomeranz2018?}). Briefly, all individuals from each
-sample were identified to the lowest practical taxonomic unit and body
-lengths were measured using image processing software from photos taken
-with a camera mounted to a dissecting microscope. Body mass was
-estimated using published length weight regressions.
+across an acid mine drainage (AMD) stress gradient. Details of the
+sample collection and processing can be found in (Pomeranz et al. 2019).
+Briefly, all individuals from each sample were identified to the lowest
+practical taxonomic unit and body lengths were measured using image
+processing software from photos taken with a camera mounted to a
+dissecting microscope. Body mass was estimated using taxon-specific
+published length weight regressions.
 
 The second data set was from the wadeable stream sites of the National
-Ecological Observatory Network (NEON; \textbf{data product
-XXX})(National Ecological Observatory Network (NEON) 2022). These sites
-are located across a wide temperature gradient in the United States,
-from Puerto Rico to Alaska. Quantitative macroinvertebrate samples were
-collected using the most appropriate method based on the local habitat.
-All individuals were identified and had their body lengths measured, and
-body mass was estimated using published length weight regressions. This
-data has been analyzed previously using size spectra methods as
-described in Pomeranz et al.~(2022)(Pomeranz et al. 2022). Detailed
-methods of the sampling collection and data processing methods can be
-found on the NEON website
+Ecological Observatory Network (National Ecological Observatory Network
+(NEON) 2022). NEON stream sites are located across a wide temperature
+gradient in the United States, from Puerto Rico to Alaska. Quantitative
+macroinvertebrate samples were collected using the most appropriate
+method based on the local habitat. All individuals were identified and
+had their body lengths measured, and body mass was estimated using
+published length weight regressions. This data has been analyzed
+previously using size spectra methods as described in (Pomeranz et al.
+2022). Detailed methods of the sampling collection and data processing
+methods can be found on the NEON website
 ({[}\url{https://www.neonscience.org/data-collection/macroinvertebrates}{]},
-or macroinvertebrate DPI pubs).
+or macroinvertebrate DPI pubs). Coefficient estimates (\(\pm\) 1
+standard deviation) are compared across methods. This allows us to
+determine whether or not the main results would differ depending on the
+method used.
 
@@
 
+Alternate plot version
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_main_v2.png}
+\caption{alternate plot version, thoughts on which you prefer??}
+\end{figure}
+
 All methods performed reasonably well in detecting the known
-relationship across the hypothetical environmental gradient. The
-regressions using the MLE method found a significant relationship 100\%
-of the time, while the ELBn and NAS method found significant
-relatoinships 95.8\% and 96.9\% of the time, respectively.
+relationship across the hypothetical environmental gradient (mean
+\(\beta_1\) estimates ranged from -0.419 to -0.498). However, the SD for
+the NAS and ELBn methods were 2-3x larger that that of the MLE method
+(Figure \textsubscript{XX}). This result was consistent across
+simulation parameters, including sample size, number of sites, and
+\(\lambda\) values.
 
-The confidence intervals for the relationship coefficients (\(\beta_1\))
-for both the MLE and ELBn method contained the true value of the known
-relationship across the gradient \(\ge\) 95\% of the time, whereas the
-confidence interval for the NAS method only had the true value 83.1\% of
-the time. Despite having similar accuracy, the width of the CIs for the
-ELBn method were more than 3 times that of the MLE method (ELBn\_CI\_ =
-0.386 \(\pm\) 0.20; MLE\_CI\_ = 0.121 \(\pm\) 0.056). The CI's for the
-NAS were slightly smaller than the ELBn method, but still \(>\) 2.5
-times as wide as the MLE method (NAS\_CI = 0.339 \(\pm\) 0.172).
-
-On average, the relationship across the gradient was under estimated:
-MLE -0.002 (\(\pm\) 0.022); ELBn -0.039 (\(\pm\) 0.062); and NA -0.081
-(\(\pm\) 0.061).
+\newpage
 
@@
 \includegraphics{figures/PLB_sim_relationship_density.png}
-\caption{Distribution of relationship coefficient estimates. Vertical
-line is the known relationship. All methods under estimate the value,
-but the mean magnitude and distribution of values is greater for the
-ELBn and NAS methods.}
+\caption{Halfeye plots for the estimated relationship coefficient
+estimates across 1000 replicates and three different methods. The point
+below each distribution plot is the median estimate, the thick and thin
+bars represent the 66th and 95th percentile, respectively. Vertical line
+is the known relationship. The median estimate for the MLE method was
+closest to the known relationship, and had the narrowest distribution
+across replicates.}
 \end{figure}
 
-Interestingly, there was an interactive effect of the estimate accuracy
-between sample size and the value of \(\lambda\). All methods were more
-accurate with larger sample sizes, and smaller values of \(\lambda\).
+\newpage
 
-\begin{figure}
-\centering
-\includegraphics{figures/n_vary_est_b.png}
-\caption{Distribution of size spectra parameter estimates. Vertical line
-is the known parameter wich describes the bounded power law distribution
-from which the body size estimates were sampled. As n increases (top to
-bottom) and \(\lambda\) increases (left to right), the accuracy of the
-estimate improves across all methods.}
-\end{figure}
+\begin{verbatim}
+##   Model   p50   p25  p975
+## 1   MLE -0.49 -0.50 -0.48
+## 2  ELBn -0.49 -0.52 -0.29
+## 3   NAS -0.44 -0.46 -0.30
+\end{verbatim}
 
@@
 
+\newpage
+
 \hypertarget{no-relationship-1}{%
@@
 All of the methods performed similarly when there was no relationship
-across the hypothetical gradient. The Type I error rate for MLE, ELBn
-and NAS were 5\%, 5.7\% and 4.6\%, respectively. However, the confidence
-intervals for the binning methods were \textasciitilde3 times as wide as
-the MLE method (mean CI widths: MLE = 0.116, ELBn = 0.372; NAS = 0.332).
-
-Of the relationships which were estimated to be significant, the MLE,
-ELBn, and NAS method had a 58\%, 49.1\% and 54.3\% probability of
-indicating a negative relationship, respectively.
+across the hypothetical gradient. The mean \(\beta_1\) estimates for all
+three methods ranged from -0.0005 to 0.0002. However, the distribution
+of coefficient estimates was much wider for the two binning methods; the
+95th percentile of the MLE distributions is narrower thatn the 66th
+percentiles for the NAS and ELBn methods.
 
@@
 \centering
-\includegraphics{figures/static_b_lambda_spaghetti.png}
+\includegraphics{figures/PLB_static_b_relationship_density.png}
 \caption{Individual regression estimates when no relationship exists
 across the hypothetical environmental gradient. The methods are
-separated by rows, and the left and right column show relationships
-which were non-significant and significant, respectively. The individual
-regressions coefficients are colored-coded to indicate positive and
-negative relationship estimates.}
+separated by rows. The median for all methocs is approximately zero, but
+the distribution for the MLE method is narrower than the two binning
+methods.}
 \end{figure}
+
+\begin{verbatim}
+##   Model p50   p25 p975
+## 1   MLE   0 -0.01 0.04
+## 2   NAS   0 -0.04 0.12
+## 3  ELBn   0 -0.05 0.13
+\end{verbatim}
 
@@
 
-The performance of all methods declined when trying to detect variation
-in the \(\lambda\) parameter between -1.9 to -2.1 across the
-environmental gradient. The mean \(\beta_1\) coefficient estimates for
-all methods were closely associated with the known relationship. Once
-again the variation in the estimates using the MLE method were much
-smaller that the estimates using both of the binning methods (Figure 4).
+The performance of all methods to detect small variation in the
+\(\lambda\) parameter (between -1.9 to -2.1) across the environmental
+gradient was also investigated. The median \(\beta_1\) coefficient
+estimates for the MLE method was -0.1, which exactly matched the value
+for the hypothetical relationship, while the other two methods both had
+median estimates of 0.9. The upper 97.5 percentile of the distribution
+was below 0 for the MLE method, meaning that most of the estimates had
+the correct sign. In contrast, the upper 97.5 percentile for both the
+NAS and ELBn methods were greater than 0 at 0.027 and 0.052,
+respectively. (Figure 4).
 
@@
 \includegraphics{figures/PLB_sim_small_relationship_density.png}
-\caption{Mean difference between relationship estimate and known
-relationship coefficient for each method. Here, \(\lambda\) varied from
--1.9 to -2.1, with a kown relationaship of -0.1 across the gradient.}
+\caption{Distribution of the estimated relationship for each method.
+Here, \(\lambda\) varied from -1.9 to -2.1, with a known relationaship
+of -0.1 across the gradient indicated by the dashed vertical line. The
+distribution of relationship estimates was approximately centered at the
+known relationship for all methods, with a narrower distribution
+observed for the MLE estimates.}
 \end{figure}
 
-Of the 1000 simulation replicates, the MLE only detected a significant
-relationship 90\% of the time, whereas the ELBn and NAS methods detected
-significant relationships 19 and 23\% of the time, respectively. Of the
-replicates which were significant, the confidence interval contained the
-true value of the relationship 94.6, 88.4, and 89.1\% of the time for
-the MLE, ELBn, and NAS methods, respectively.
-
-\begin{figure}
-\centering
-\includegraphics{figures/small_lambda_spaghetti.png}
-\caption{Linear regressions across the hypothetical gradient when lambda
-varies from -1.9 to -2.1. Each method is displayed in the rows, and the
-left colum shows the non-significant estimates, while the right column
-shows the relationships which were signficant (\(\beta_1\) coefficient
-p-value \textless{} 0.05). The colors indicate when the beta coefficient
-estimate was negative (pink) and when it was positive (purple).}
-\end{figure}
-
-Once again, the width of the confidence intervals for both of the
-binning methods were \(\ge\) 2.8 times that of the MLE method (mean
-\(\pm\) SD MLE CI width: 0.117 \(\pm\) 0.051; ELBn CI width: 0.378
-\(\pm\) 0.169; NAS CI width: 0.332 \(\pm\) 0.150).
-
-On average, the relationship estimates across the gradient were similar
-to the know value (mean deviation: MLE = 0.00001, ELBn = -0.009, NAS =
--0.012). However, when only looking at the relationships which were
-significant, this variation increased considerably(MLE = -0.002, ELBN =
--0.052, NAS = 0.034).
+\begin{verbatim}
+##   Model   p50   p25  p975
+## 1   MLE -0.10 -0.11 -0.06
+## 2   NAS -0.09 -0.13  0.03
+## 3  ELBn -0.09 -0.13  0.05
+\end{verbatim}
 
@@
 
-For both empirical datasets, the direction and magnitude of chnage
+For both empirical data sets, the direction and magnitude of change
 (i.e.~\(\beta_1\) coefficients) are generally in agreement. Size spectra
@@
 
-\textbf{NOTE} combine all the figures below into one with 4 panels (AMD
-= A, B; NEON = C, D)
+\textbf{NOTE} I need to re-plot these in ggplot and and stitich
+together. Currently stitching pre-made png files, I think that's why it
+looks so funky.
+
+\newpage
 
@@
 \centering
-\includegraphics{figures/AMD_plot.png}
-\caption{A}
+\includegraphics{figures/empirical_combined.png}
+\caption{Figure of empirical data estimates. All of the methods estimate
+the same sign of the relationship, but the estimates from the binning
+methods are generally greater than the MLE estimates.}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics{figures/amd_relationship.png}
-\caption{B.Relationship estimates across AMD gradient}
-\end{figure}
+\newpage
 
-\begin{figure}
-\centering
-\includegraphics{figures/NEON_plot.png}
-\caption{C.}
-\end{figure}
+The \(\beta_1\) coefficient estimates for the AMD data are all positive
+and of relatively similar magnitudes across methods. The MLE estimate
+has the smallest magnitude and SD (0.063 \(\pm\) 0.011 SD), The ELBn
+method is slightly larger and has a wider SD : (0.072 \(\pm\) 0.018).
+Finally, the NAS method has the largest estimate and an intermediate
+standard deviation (0.076 \(\pm\) 0.013). The range of the gradient in
+the AMD data is relatively small (9.5), meaning that the absolute change
+in the size spectra parameter (\(\lambda\)) across the AMD gradient are
+similar regardless of method used (range of absolute: 0.59 to 0.72).
 
-\begin{figure}
-\centering
-\includegraphics{figures/neon_relationship.png}
-\caption{D.}
-\end{figure}
-
-Because the \(\beta_1\) coefficient estimates are similar, and the range
-of the gradient in the AMD data is relatively small (9.5), the absolute
-change in the size spectra parameter across the AMD gradient are similar
-regardless of method used (range: 0.59 to 0.72). Likewise, the absolute
-change in size spectra parameters across the NEON temperature gradient
-ranges from 0.06 to 0.165, depending on the method used.
-
-\emph{alternative} absolute difference of \textasciitilde0.1 units in
-\(\lambda\) across gradients is similar to differences reported in other
-studies (seasonally: 0.16, Mcgarvey and Kirk; landuse: 0.15 Martinez;
-O'Gorman 2017 \textasciitilde0.25, Yvon and Dossena papers
-\textasciitilde0.1? {[}need to redownload some of these and
-compare{]})(Dossena et al. 2012, O'Gorman et al. 2017, McGarvey and Kirk
-2018)
+Similar to the AMD results, the \(\beta_1\) coefficient estimates for
+the NEON data are all negative and have similar magnitudes across
+methods. The MLE estimate is once again the smallest (-0.006 \(\pm\)
+0.001 SD). The NAS and ELBn methods are larger at -0.004 \(\pm\) 0.002,
+and -0.002 \(\pm\) 0.002, respectively. Although the estimates are
+similar and have a small magnitude, the range of the gradient in this
+data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
+change in \(\lambda\) estimates ranges from 0.06 to 0.17, depending on
+the method used.
 
@@
 of all methods increased with shallower \(\lambda\)'s we wanted to
-investigate how the methods performed with simulated values more similar
-to the empirical estimates of parameters describing size spectra
+investigate how the methods performed with simulated values closer to
+the empirical estimates of parameters describing size spectra
 relationships. Therefore, we repeated the simulation process as in the
-main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5.
+main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5,
+with a known relationship of \(\beta_1 = -0.2\)
+
+\newpage
 
@@
 
-We found generally less variability in the relationship estimates across
-a gradient of distributions described with shallower \(\lambda\)
-parameters. All of the relationship coefficient estimates
-(\(\beta_1\))'s across all replicates were significant regardless of
-method used. The confidence interval for \(\beta_1\) estimate for the
-MLE method contained the true value 94.7\% of the time. However, the
-true value was in the confidence interval for the ELBN and NAS method
-only 90.9\% and 76.9\% of the time, respectively. Once again, the mean
-width of the CI for the binning methods were \textasciitilde2 times as
-large as for the MLE method. Mean \(\pm\) SD CI width: MLE = 0.0442
-\(\pm\) 0.0196; ELBn = 0.0920 \(\pm\) 0.0482, NAS = 0.0798 \(\pm\)
-0.0324.
-
-On average, the relationship across the gradient was under estimated,
-but by nearly an order of magnitude less than when simulating across
-steeper values of \(\lambda\). Differences between known and estimated
-relationship coefficients: MLE -0.0007 (\(\pm\) 0.0079); ELBn -0.0058
-(\(\pm\) 0.0177); NAS -0.0251 (\(\pm\) 0.0124).
+We found generally less variability in the relationship estimates
+(\(\beta_1\)) across a gradient of distributions described with
+shallower \(\lambda\) parameters. However, the median MLE estimate was
+once again closest to the hypothetical relationship and had the
+narrowest distribution. The median ELBn estimate was close to the
+hypothetical relationship but had a slightly wider distribution.
+Finally, 95\% of the NAS estimates were greater than the hypothetical
+relationship. In general, the distributions from the different
+methodologies followed the same overall pattern, with the MLE estimates
+being the smallest, and the binning methods being larger (i.e., more
+positive).
 
@@
 \centering
-\includegraphics{figures/shallow_lambda_relationship_density.png}
+\includegraphics{figures/PLB_shallow_lambda_relationship_density.png}
 \caption{Distribution of relationship coefficient estimates. Vertical
@@
 \end{figure}
+
+\begin{verbatim}
+##   Model   p50   p25  p975
+## 1   MLE -0.20 -0.21 -0.18
+## 2  ELBn -0.19 -0.21 -0.16
+## 3   NAS -0.17 -0.18 -0.15
+\end{verbatim}
+
+\newpage
 
@@
 associated with different estimation methods. However, how these
-innaccuracies and biases compound across environmental gradients remains
-uncertain, making it difficult to detecting variation in size spectra
+inaccuracies and biases compound across environmental gradients remains
+uncertain, making it difficult to detect variation in size spectra
 parameters across environmental gradients with confidence. Here, we
-sampled body sizes from known distributions and estimated their
-parameters across a hypothetical gradient in order to assess the
-relationship coefficients obtained.
+sampled body sizes from known distributions with varying parameters
+(\(\lambda\)) and estimated the coefficient of the relationship
+(\(\beta_1\)) across a hypothetical gradient in order to assess how the
+results would vary depnding on the method used. Likewise, we compared
+how the interpretation of previously published results could change
+depending on the methodology used to estimate the relationship.
 
-Generally, there was large agreement in the direction and magnitude of
-average relationship coefficient estimates across methods. The estimate
-from the MLE method was always closer to the known relationship, and
-always had the smallest variation around the estimate.
+Generally, the estimate from the MLE method was always closer to the
+known relationship in simulated data, and always had the smallest
+variation around the estimate.
 
-All of the methods performed equally well when the \(\lambda\) value did
-not vary across the hypothetical gradient, with a Type I error rate of
-\textasciitilde5\%, exactly as would be expected under the assumptions
-of frequentist statistics. This indicates that previous studies which
-have detected no difference in size spectra parameters across sites can
-be interpreted with confidence, regardless of the method used.
+Binning methods are easy to use and interpret, which most likely
+accounts for their wide use in ecological studies. However, aggregating
+individuals into logarithmic bins removes a large source of the
+variation within the data by collapsing body size variation into a
+single values within bins. For example, all individuals placed into a
+bin that ranges from 2-4 grams of mass are all treated as having a mass
+of 3 grams, the midpoint of that bin. Likewise, a single abundance value
+is taken for this bin, despite that fact that there is almost certainly
+variation in the abundance of individuals that weigh \textasciitilde2
+grams or \textasciitilde4 grams. By homogenizing the data in this way,
+it seems that binning methods will always produce noisier results than
+MLE from the binning process alone, since binning is akin to deleting
+information that the model could otherwise use.
+
+One of the benefits of using MLE is that all data points are retained
+for use within the model.
 
@@
 than the two binning methods examined here under all contexts and should
-be the preferred method moving forward. At a minimum, future studies
+be the preferred method in future studies. At a minimum, future studies
 should report MLE estimates to ensure that the results are not dependent
 on the methodology used.
-
-Alarmingly, the performance of all methods declined dramatically when
-trying to detect smaller variation in \(\lambda\) values. We simulated
-observations from distributions varying by an absolute value of 0.2
-units. This range of variation is similar to that which has been
-reported in many seminal works on variation in size spectra parameters.
-When using the MLE method, significant relationships were only detected
-\textasciitilde88\% of the time, whereas both of the binning methods
-only detected significant trends \textasciitilde20\% of the time. Of the
-simulation replicates which were significant, nearly all of them were
-negative across all methods except for one single replicate using the
-NAS method which was positive (Figure. 5).
 
@@
 estimate. Given that all of the data within a study is treated
-identically, the the over all chnage in size spectra parameters is
+identically, the the over all change in size spectra parameters is
 likely reasonable. However, the biases and inconsistencies in
@@
 
+Both of the binning methods can perform as well as MLE under certain
+circumstances (large sample size, shallow \(\lambda\)). However, it
+would be extremely difficult to know if this was the case beforehand.
+Furthermore, in order to confirm this result, MLE methods would likely
+need to be employed. With the publication of the sizespectra package
+(2017), producing MLE estimates of size spectra parameters is a
+relatively easy task. Therefore, we recommend using this in all future
+studies of size spectra relationships.
+
 We reiterate the recommendations of White et al.~(2007)(2007), Sprules
@@
 
+\newpage
+
 \hypertarget{references}{%
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-dossena2012}{}}%
-Dossena, M., G. Yvon-Durocher, J. Grey, J. M. Montoya, D. M. Perkins, M.
-Trimmer, and G. Woodward. 2012.
-\href{https://doi.org/10.1098/rspb.2012.0394}{Warming alters community
-size structure and ecosystem functioning}. Proceedings of the Royal
-Society B: Biological Sciences 279:3011--3019.
-
 \leavevmode\vadjust pre{\hypertarget{ref-edwards2020}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-mcgarvey2018}{}}%
-McGarvey, D. J., and A. J. Kirk. 2018.
-\href{https://doi.org/10.1007/s10750-017-3448-0}{Seasonal comparison of
-community-level size-spectra in southern coalfield streams of {West
-Virginia} ({USA})}. Hydrobiologia 809:65--77.
-
 \leavevmode\vadjust pre{\hypertarget{ref-mcgarvey2019}{}}%
@@
 ({DP1}.20120.001)}. {National Ecological Observatory Network (NEON)}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ogorman2017}{}}%
-O'Gorman, E. J., L. Zhao, D. E. Pichler, G. Adams, N. Friberg, B. C.
-Rall, A. Seeney, H. Zhang, D. C. Reuman, and G. Woodward. 2017.
-\href{https://doi.org/10.1038/nclimate3368}{Unexpected changes in
-community size structure in a natural warming experiment}. Nature
-Climate Change 7:659--663.
 
@@
 Change Biology 28:848--858.
+
+\leavevmode\vadjust pre{\hypertarget{ref-pomeranz2019}{}}%
+Pomeranz, J. P. F., H. J. Warburton, and J. S. Harding. 2019.
+\href{https://doi.org/10.1111/fwb.13196}{Anthropogenic mining alters
+macroinvertebrate size spectra in streams}. Freshwater Biology
+64:81--92.
 
@@
 
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
+\beginsupplement
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_small_m_est_b_density.png}
-\caption{Distribution of estimated \(\lambda\) coefficient for five
-sites across a hypothetical gradient with known values. Range of body
-sizes is reduced and is from 1, to 100.}
-\end{figure}
+\newpage
 
@@
 \end{figure}
+
+\newpage
 
@@
 
+\newpage
+
 \begin{figure}
@@
 
+\newpage
+
 \hypertarget{sample-size}{%
@@
 
-\textbf{note} These plots still have n = 100. Need to re run and update
-if we end up going with n = 200
+\begin{figure}
+\centering
+\includegraphics{figures/n_vary_est_b.png}
+\caption{Distribution of size spectra parameter estimates. Vertical line
+is the known parameter (dashed line) wich describes the bounded power
+law distribution from which the body size estimates were sampled. As n
+increases (top to bottom) and \(\lambda\) increases (left to right), the
+accuracy of the estimate improves across all methods.}
+\end{figure}
+
+\newpage
 
@@
 \centering
-\includegraphics{figures/lambda_n_5_sites.png}
-\caption{Plot showing the relationships between sample size and the
-estimated \(\lambda\) parameters. the dashed line indicates the known
-value of \(\lambda\).}
+\includegraphics{figures/n_vary_main.png}
+\caption{Individual regressions when varying sample size}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/n_vary_relationship_density.png}
+\caption{Distribtuion of the relationship coefficients with varying
+sample size}
+\end{figure}
+
+\newpage
+
+\hypertarget{lambda-and-relationship-estimates}{%
+\subsection{\texorpdfstring{\(\lambda\) and relationship
+estimates}{\textbackslash lambda and relationship estimates}}\label{lambda-and-relationship-estimates}}
+
+In the main analysis, we simulate body size data from bounded power law
+distributions while varying the \(\lambda\) exponent which describes the
+distribution. For the results presented in the main text, we held the
+number of sites at 5, scaled the environmental gradient from -1 to 1,
+and set the minimum and maximum body sizes to \(0.0026\) and
+\(1.2 *10^3\) respectively. Here, we plot the results of varying the
+number of sites (3, 10), increasing the scale of the environmental
+gradient (-100 to 100) and decreasing the range of body sizes (min = 1,
+max = 100). Generally, the results reported in the main manuscript are
+robust to changing these parameters: the MLE estimate is nearly always
+closer to the known parameters, and the variation in these estimates is
+usually smaller than the binning methods.
+
+\newpage
+
+\hypertarget{number-of-sites-3}{%
+\subsubsection{Number of sites = 3}\label{number-of-sites-3}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_3_sites_est_b_density.png}
+\caption{Distribution of estimated \(\lambda\) coefficient across a
+hypothetical gradient with known values (dashed line). The number of
+sites is reduced to 3, compared with 5 sites presented in the main
+anaysis.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_3_sites_relationship_density.png}
+\caption{Relationship estimate}
+\end{figure}
+
+\newpage
+
+\hypertarget{number-of-sites-10}{%
+\subsubsection{Number of sites = 10}\label{number-of-sites-10}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_10_sites_est_b_density.png}
+\caption{Distribution of estimated \(\lambda\) coefficient across a
+hypothetical gradient with known values(dashed line). The number of
+sites is increased to 10, compared with 5 sites presented in the main
+anaysis.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_10_sites_relationship_density.png}
+\caption{Relationship estimate}
+\end{figure}
+
+\newpage
+
+\hypertarget{environmental-gradient}{%
+\subsubsection{Environmental Gradient}\label{environmental-gradient}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_large_x_est_b_density.png}
+\caption{Distribution of estimated \(\lambda\) coefficient for five
+sites across a hypothetical gradient with known values (dashed line).
+The range of the environmental gradient was increased (-100 to 100).}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_large_x_relationship_density.png}
+\caption{Relationship estimate}
+\end{figure}
+
+\newpage
+
+\hypertarget{range-of-body-sizes-1-to-100}{%
+\subsubsection{Range of body sizes = 1 to
+100}\label{range-of-body-sizes-1-to-100}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_small_m_est_b_density.png}
+\caption{Distribution of estimated \$\textbackslash lambda\$ coefficient
+for five sites across a hypothetical gradient with known values(dashed
+line). Range of body sizes is smaller than main anaysis and ranges from
+1, to 100.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_small_m_relationship_density.png}
+\caption{Relationship estimate}
 \end{figure}

### 2022-09-04 custom_functions.R

@@ -332,7 +332,7 @@ if(sim_data$distribution[1] == "PLB"){
   #   labs(title = sim_data$distribution,
   #        x = "slope estimate") +
   #   NULL
-  #### New 8/30/22 "halfeye" plot
+  #### New 8/30/22 "halfeye" plot ####
   estimate_density <- sim_data %>%
     mutate(Model = factor(name,
                           levels = 

### 2022-09-04 Manuscript.Rmd
re-knit updated manuscript sept 4

@@
     institute: [glrc, lumcon]
+  - name: Vojsava Gjoni
+    institute: usd
   - name: Jeff S. Wesner
@@
 output:
+   word_document:
+      reference_docx: working_docx_template.docx
+      toc: no
+      pandoc_args:
+      - '--lua-filter=scholarly-metadata.lua'
+      - '--lua-filter=author-info-blocks.lua'
    pdf_document:
@@
      keep_tex: true
-     pandoc_args:
-     - '--lua-filter=scholarly-metadata.lua'
-     - '--lua-filter=author-info-blocks.lua'
-   word_document:
-      reference_docx: working_docx_template.docx
-      toc: no
      pandoc_args:

### 2022-09-05 Manuscript.Rmd
test knit with office machine

--- 2022-09-04 Manuscript.Rmd
@@
 
-**NOTE** I need to integrate reference manager with R markdown. Any favorites or recommendations?? I generally use Zotero (free), but think I have access to Mendeley through CMU...
+**NOTE** I'm still working on integrating with Zotero and importing all the necessary references (thanks Jim for the help and instructions!).
+Don't worry about the **??** references in the rendered version for now.
 
-**NOTE from Jim** I use Zotero, mainly for free and open-source reasons, but I also can have a standalone program that doesn't require internet access, can backup my reference library on the cloud and sync among multiple computers, can spit out bibtex to connect to markdown, and can have group citation lists if neccessary.
-I use it with a betterBibtex add on that allows for modifying bibtex keys for calling refs in Rmarkdown.
-
-**NOTE** I re-ran all of these after setting the seed for reproducibility.
-Figures should be up to date, but the numbers may change slightly i.e., 94.8% might become 95.1%, mean $\pm$ SD values will be different.
+**Savina** I'm particularly keen to get your feedback on this.
+I think you mentioned including some other references and discussion in Jeff's Bayesian model paper.
+If you think they fit in here please let me know or feel free to include them.
+Likewise, please include any other comments you have, or if you have any insight into how the interpretation of results from different methods would vary.
 
@@
 The focus of the present work is to investigate biases when estimating the relationship or relative change in parameters across an environmental gradient.
-Hence, we do not discuss the $\lambda$ estimates in the main text, but the results are presented in the Supplmental Information.
+Hence, we do not discuss the $\lambda$ estimates in the main text, but the results are presented in the Supplemental Information (**NOTE** I'm thinking about dropping this from SI.
+Might be too much info, and it's already been published on by Edwards.
+Does this add anything to the story??).
 
@@
 
+**NOTE** I'm considering dropping the $\lambda$ estimate plots from the manuscript. I think it's too much info, but if it adds to the story I'm happy to leave it in. **END**
+
 In the main analysis, we simulate body size data from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.

### 2022-09-05 Manuscript.tex
re-knit updated manuscript sept 4
test knit with office machine

@@
   pdftitle={Manuscript},
-  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Jeff S. Wesner4},
+  pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Vojsava Gjoni4; Jeff S. Wesner4},
   hidelinks,
@@
 \author{Justin Pomeranz\textsuperscript{1,*} \and James R.
-Junker\textsuperscript{2,3} \and Jeff S. Wesner\textsuperscript{4}}
-\date{04 September, 2022}
+Junker\textsuperscript{2,3} \and Vojsava
+Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
+\date{05 September, 2022}
 
@@
 
-\textbf{NOTE} I need to integrate reference manager with R markdown. Any
-favorites or recommendations?? I generally use Zotero (free), but think
-I have access to Mendeley through CMU\ldots{}
-
-\textbf{NOTE from Jim} I use Zotero, mainly for free and open-source
-reasons, but I also can have a standalone program that doesn't require
-internet access, can backup my reference library on the cloud and sync
-among multiple computers, can spit out bibtex to connect to markdown,
-and can have group citation lists if neccessary. I use it with a
-betterBibtex add on that allows for modifying bibtex keys for calling
-refs in Rmarkdown.
-
-\textbf{NOTE} I re-ran all of these after setting the seed for
-reproducibility. Figures should be up to date, but the numbers may
-change slightly i.e., 94.8\% might become 95.1\%, mean \(\pm\) SD values
-will be different.
+\textbf{NOTE} I'm still working on integrating with Zotero and importing
+all the necessary references (thanks Jim for the help and
+instructions!). Don't worry about the \textbf{??} references in the
+rendered version for now.
+
+\textbf{Savina} I'm particularly keen to get your feedback on this. I
+think you mentioned including some other references and discussion in
+Jeff's Bayesian model paper. If you think they fit in here please let me
+know or feel free to include them. Likewise, please include any other
+comments you have, or if you have any insight into how the
+interpretation of results from different methods would vary.
 
@@
 estimates in the main text, but the results are presented in the
-Supplmental Information.
+Supplemental Information (\textbf{NOTE} I'm thinking about dropping this
+from SI. Might be too much info, and it's already been published on by
+Edwards. Does this add anything to the story??).
 
@@
 
+\textbf{NOTE} I'm considering dropping the \(\lambda\) estimate plots
+from the manuscript. I think it's too much info, but if it adds to the
+story I'm happy to leave it in. \textbf{END}
+
 In the main analysis, we simulate body size data from bounded power law

### 2022-09-24 Manuscript.Rmd
new ms knit. Added pub estimates to SI

@@
 ---
-title: "Manuscript"
+title: "Manuscript: Detecting differences in Size Spectra"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
+## Data Availabity
+
+The data used in this manuscript is already publicly available (Pomeranz et al.
+2019 data dryad DOI **here**, NEON data citation **here**.
+R Scripts to reproduce the full simulation and analysis are available at: ([<https://github.com/Jpomz/detecting-spectra-differences>]).
+
 # Results
@@
 
-Alternate plot version
+![alternate plot version; I think I like this one better. What are your thoughts on which you prefer??](figures/PLB_main_v2.png)
 
-![alternate plot version, thoughts on which you prefer??](figures/PLB_main_v2.png)
-
-All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient (mean $\beta_1$ estimates ranged from -0.419 to -0.498).
-However, the SD for the NAS and ELBn methods were 2-3x larger that that of the MLE method (Figure ~XX~).
-This result was consistent across simulation parameters, including sample size, number of sites, and $\lambda$ values.
+The distribution of $\beta_1$ estimates from MLE method were the most accurate and precise, and the NAS method was the least accurate and precise (Table *xx*).
+The median estimates for all methods were closer to the known relationship.
+All of the methods produced estimate distributions with long right tails, but this pattern was greater in the ELBn and NAS methods (e.g., SD for ELBn and NAS methods were 2-3x larger, Figure *xx* Halfeye plots).This result was consistent across simulation parameters, including sample size, number of sites, and $\lambda$ values.
 
@@
 ```{r, echo=FALSE, warning=FALSE}
+# will turn these into "real" formatted tables if we decide they're useful
 read.csv("results/PLB_sim.csv") %>%
-  mutate(Model = name) %>%
-  select(Model, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  mutate(Model = name,
+         mean = mu_slope) %>%
+  select(Model, mean, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 3))) %>%
   arrange(p50, p975)
@@
 
-![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).](figures/n_vary_main.png)
+![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).
+(match this figure to "new" style if we like that better)](figures/n_vary_main.png)
 
@@
 The mean $\beta_1$ estimates for all three methods ranged from -0.0005 to 0.0002.
-However, the distribution of coefficient estimates was much wider for the two binning methods; the 95th percentile of the MLE distributions is narrower thatn the 66th percentiles for the NAS and ELBn methods.
+Although the distributions were closely centered on 0, the distribution of coefficient estimates was much wider for the two binning methods; the 95th percentile of the MLE distributions is narrower than the 66th percentiles for the NAS and ELBn methods.
 
@@
 read.csv("results/PLB_static_b.csv") %>%
-  mutate(Model = name) %>%
-  select(Model, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  mutate(Model = name,
+         mean = mu_slope) %>%
+  select(Model, mean, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 3))) %>%
   arrange(p50, p975)
@@
 The performance of all methods to detect small variation in the $\lambda$ parameter (between -1.9 to -2.1) across the environmental gradient was also investigated.
-The median $\beta_1$ coefficient estimates for the MLE method was -0.1, which exactly matched the value for the hypothetical relationship, while the other two methods both had median estimates of 0.9.
+The median $\beta_1$ coefficient estimates for the MLE method was -0.1, which exactly matched the value for the hypothetical relationship, while the other two methods both had median estimates of -0.09.
 The upper 97.5 percentile of the distribution was below 0 for the MLE method, meaning that most of the estimates had the correct sign.
@@
 read.csv("results/PLB_sim_small.csv") %>%
-  mutate(Model = name) %>%
-  select(Model, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  mutate(Model = name,
+         mean = mu_slope) %>%
+  select(Model, mean, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 3))) %>%
   arrange(p50, p975)
@@
 
-**NOTE** I need to re-plot these in ggplot and and stitich together. Currently stitching pre-made png files, I think that's why it looks so funky.
-
 \newpage
@@
 ![Figure of empirical data estimates.
-All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.](figures/empirical_combined.png)
+All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.
+**NOTE** I need to re-plot these in ggplot and and stitch together.
+Currently stitching pre-made png files, I think that's why it looks so funky.](figures/empirical_combined.png)
 
@@
 Finally, the NAS method has the largest estimate and an intermediate standard deviation (0.076 $\pm$ 0.013).
-The range of the gradient in the AMD data is relatively small (9.5), meaning that the absolute change in the size spectra parameter ($\lambda$) across the AMD gradient are similar regardless of method used (range of absolute: 0.59 to 0.72).
+The range of the gradient in the AMD data is relatively small (9.5), but the absolute change in the size spectra parameter ($\lambda$) across the AMD gradient vary by ~0.13 units depending on the method used (range of absolute change: 0.59 to 0.72).
 
@@
 The NAS and ELBn methods are larger at -0.004 $\pm$ 0.002, and -0.002 $\pm$ 0.002, respectively.
-Although the estimates are similar and have a small magnitude, the range of the gradient in this data is larger (\~30$\circ$C), meaning that the absolute change in $\lambda$ estimates ranges from 0.06 to 0.17, depending on the method used.
+Although the estimates are similar and have a small magnitude, the range of the gradient in this data is larger (\~30$\circ$C), meaning that the absolute change in $\lambda$ estimates varies by 0.11 units depending on the methods used (range of absolute change across gradient: 0.06 to 0.17).
 
@@
 Upon closer inspection, the values of size spectra parameters in the empirical data were considerably shallower than -2.
-Given that we found the performance of all methods increased with shallower $\lambda$'s we wanted to investigate how the methods performed with simulated values closer to the empirical estimates of parameters describing size spectra relationships.
+Given that we found the performance of all methods increased with shallower $\lambda$'s we wanted to investigate how the methods performed with simulated values closer to the empirical estimates of parameters describing size spectra relationships in streams.
 Therefore, we repeated the simulation process as in the main analysis, but used $\lambda$ values ranging from -1.1 to -1.5, with a known relationship of $\beta_1 = -0.2$
@@
 
-![Relationship estimates across the hypothetical gradient for each replicate. Each panel is a different method for estimating the size spectra parameter.](figures/shallow_lambda_main.png)
+![Relationship estimates across the hypothetical gradient for each replicate.
+Each panel is a different method for estimating the size spectra parameter.
+(match to "new" style if needed)](figures/shallow_lambda_main.png)
 
@@
 Vertical line is the known relationship.
-All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.](figures/PLB_shallow_lambda_relationship_density.png)
+All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.
+(not sure why this is such different proportions, need to re plot with right dimensions)](figures/PLB_shallow_lambda_relationship_density.png)
 
@@
 read.csv("results/shallow_lambda.csv") %>%
-  mutate(Model = name) %>%
-  select(Model, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 2))) %>%
+  mutate(Model = name,
+         mean = mu_slope) %>%
+  select(Model, mean, p50, p25, p975) %>%
+  mutate_at(vars(-Model), funs(round(., 3))) %>%
   arrange(p50, p975)
 ```
+**NOTE** Not sure if these tables are usful or not. Shows same information as figures, but might be easier to see the quantitative values instead of estimating from plot
 
@@
 
-Generally, the estimate from the MLE method was always closer to the known relationship in simulated data, and always had the smallest variation around the estimate.
+Generally, the estimate from the MLE method was always closer to the known $\lambda$ value, and resulted in the most accurate $\beta_1$ coefficient estimate for the hypothetical relationship in simulated data, and always had the smallest variation around the estimate.
 
@@
 By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use.
+One of the benefits of using MLE is that all data points are retained for use within the model.
 
-One of the benefits of using MLE is that all data points are retained for use within the model.
+The variation in estimated relationships across the empirical datasets varied by ~0.1 units depending on the method used.
+At first glance this may not appear to be a concerning magnitude of variation.
+However, when one considers the range of variation reported in many seminal works on size spectra relationships in streams is ~0.1-0.2 (SI Table *xx*), it is apparent that many of these results may be dependent on the method used.
 
@@
 
-Despite the drop in performance with reduced variation in $\lambda$ values, when a significant relationship was observed, it was generally in the correct direction and of a similar magnitude.
-This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible, and the magnitude of the relative change is a reasonable estimate.
+Despite the drop in performance with reduced variation in $\lambda$ values, the estimated relationship coefficient was generally in the correct direction and of a similar magnitude.
+This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
 Given that all of the data within a study is treated identically, the the over all change in size spectra parameters is likely reasonable.
 However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to compare the relative changes across different published studies which use different methods.
-Publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental stressors and perturbations.
+Publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
+
+**NOTE** not sure if the idea below belongs in this paper or not, but it's something I've been thinking about for a while.
+I'm using this as an excuse to write out my thoughts and to see what y'all think about it.
+Maybe it's another paper to write in the future?
+
+An interesting finding is that the empirical data sets used here were quite a bit shallower than is expected based on theory, and that what has been observed empirically in marine systems.
+There are two, non-mutually exclusive, possible explanations for this: 1) freshwater streams are more dependent on detrital resources, 2) streams are effectively 2D systems, whereas marine systems are 3D.
+Blanchard et al.
+(2009) observed that benthic marine communities had shallower slopes than pelagic communities, and that this was likely due to a stronger reliance on a shared detrital basal resource in the benthic community, whereas the pelagic community was largely driven by energy transfer through predator-prey interactions.
+The importance of allochthonous resources in stream ecosystems has long been recognized (Vannote et al.
+1980).
+The stream data here largely comes from low order, headwater streams which may be reasonably considered to be driven by allochthonous resources.
+It is possible that the communities sampled here are structured based on a shared basal resource, and that the importance of predator-prey interactions is reduced.
+An alternate hypothesis is the relative dimensionality of search space in benthic and pelagic systems.
+Consumption rates and interaction strenghts have been shown to vary markedly based on search dimensionality.
+Consumers generally encounter resources more frequently in 3D environments (i.e., aerial, pelagic) than in 2D environments (i.e., benthic, terrestrial), and lead to stronger consumption rates and interaction strengths.
+Likewise, it has been hypothesized that the scaling exponent for mass-specific search rates may vary based on search space dimensionality instead of being set at approximately 0.75 (Pawar, Dell, Savage, 2012).
 
@@
 Both of the binning methods can perform as well as MLE under certain circumstances (large sample size, shallow $\lambda$).
-However, it would be extremely difficult to know if this was the case beforehand.
+However, it would be unlikely for a researcher to know if this was the case beforehand.
 Furthermore, in order to confirm this result, MLE methods would likely need to be employed.
-With the publication of the sizespectra package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
+With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
 Therefore, we recommend using this in all future studies of size spectra relationships.
@@
 ## Range of body sizes, $M$
+
+**NOTE** udate all plots to be consistent with main text
 
@@
 ![Relationship estimate](figures/PLB_small_m_relationship_density.png)
+
+# Comparison with other published estimates
+
+SI Table.
+This table shows published estimates of the variation in size spectra slopes (or exponents) in empirical studies.
+It is unclear how to directly compare estimates of the slope with different methods.However, the published estimates here range from ~0.1 to 0.2 across the gradients studied.
+For comparison, the 2.5-95% quantiles around the relationship estimate for the MLE method were ~0.1, whereas for the ELBn and NAS method they were ~0.25 and ~0.2, respectively.
+b_diff is the change in estimate (b-low - b-min).
+System refers to stream communities or mesocosm experiments.
+Method: MLE = maximum likelihod estimate, ELB = equal logarthmic binning, the number before indicates the number of bins used.
+The normalization process shifts the estimates by an absolute value of 1.0.
+Hence, direct comparison the relative change in normalized and non-normalized studies should not introduce any bias.
+The O'Gorman et al.
+2017 study used average species size and abundance (Local Size Density Relationship, *sensu* White et al.
+2007) as opposed to individual size distribution.
+These methods are related, but it is unclear how to directly compare estimates from each method.
+
+```{r}
+read.csv("lit comparison.csv")
+```

### 2022-09-24 Manuscript.tex
new ms knit. Added pub estimates to SI

@@
 \hypersetup{
-  pdftitle={Manuscript},
+  pdftitle={Manuscript: Detecting differences in Size Spectra},
   pdfauthor={Justin Pomeranz1,; James R. Junker2,3; Vojsava Gjoni4; Jeff S. Wesner4},
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx}
@@
 
-\title{Manuscript}
+\title{Manuscript: Detecting differences in Size Spectra}
 \author{Justin Pomeranz\textsuperscript{1,*} \and James R.
@@
 Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
-\date{05 September, 2022}
+\date{24 September, 2022}
 
@@
 
+\hypertarget{data-availabity}{%
+\subsection{Data Availabity}\label{data-availabity}}
+
+The data used in this manuscript is already publicly available (Pomeranz
+et al.~2019 data dryad DOI \textbf{here}, NEON data citation
+\textbf{here}. R Scripts to reproduce the full simulation and analysis
+are available at:
+({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).
+
 \hypertarget{results}{%
@@
 
-Alternate plot version
-
 \begin{figure}
@@
 \includegraphics{figures/PLB_main_v2.png}
-\caption{alternate plot version, thoughts on which you prefer??}
-\end{figure}
-
-All methods performed reasonably well in detecting the known
-relationship across the hypothetical environmental gradient (mean
-\(\beta_1\) estimates ranged from -0.419 to -0.498). However, the SD for
-the NAS and ELBn methods were 2-3x larger that that of the MLE method
-(Figure \textsubscript{XX}). This result was consistent across
-simulation parameters, including sample size, number of sites, and
-\(\lambda\) values.
+\caption{alternate plot version; I think I like this one better. What
+are your thoughts on which you prefer??}
+\end{figure}
+
+The distribution of \(\beta_1\) estimates from MLE method were the most
+accurate and precise, and the NAS method was the least accurate and
+precise (Table \emph{xx}). The median estimates for all methods were
+closer to the known relationship. All of the methods produced estimate
+distributions with long right tails, but this pattern was greater in the
+ELBn and NAS methods (e.g., SD for ELBn and NAS methods were 2-3x
+larger, Figure \emph{xx} Halfeye plots).This result was consistent
+across simulation parameters, including sample size, number of sites,
+and \(\lambda\) values.
 
@@
 \begin{verbatim}
-##   Model   p50   p25  p975
-## 1   MLE -0.49 -0.50 -0.48
-## 2  ELBn -0.49 -0.52 -0.29
-## 3   NAS -0.44 -0.46 -0.30
+##   Model   mean    p50    p25   p975
+## 1   MLE -0.495 -0.494 -0.499 -0.483
+## 2  ELBn -0.460 -0.489 -0.524 -0.291
+## 3   NAS -0.419 -0.442 -0.458 -0.302
 \end{verbatim}
@@
 \caption{Individual regression estimates across the hypothetical
-gradient based on sample size (rows) and methodology used (columns).}
+gradient based on sample size (rows) and methodology used (columns).
+(match this figure to ``new'' style if we like that better)}
 \end{figure}
@@
 across the hypothetical gradient. The mean \(\beta_1\) estimates for all
-three methods ranged from -0.0005 to 0.0002. However, the distribution
-of coefficient estimates was much wider for the two binning methods; the
-95th percentile of the MLE distributions is narrower thatn the 66th
-percentiles for the NAS and ELBn methods.
+three methods ranged from -0.0005 to 0.0002. Although the distributions
+were closely centered on 0, the distribution of coefficient estimates
+was much wider for the two binning methods; the 95th percentile of the
+MLE distributions is narrower than the 66th percentiles for the NAS and
+ELBn methods.
 
@@
 \begin{verbatim}
-##   Model p50   p25 p975
-## 1   MLE   0 -0.01 0.04
-## 2   NAS   0 -0.04 0.12
-## 3  ELBn   0 -0.05 0.13
+##   Model   mean    p50    p25  p975
+## 1   MLE -0.001 -0.001 -0.012 0.038
+## 2   NAS  0.000 -0.001 -0.039 0.119
+## 3  ELBn  0.000 -0.001 -0.046 0.134
 \end{verbatim}
@@
 for the hypothetical relationship, while the other two methods both had
-median estimates of 0.9. The upper 97.5 percentile of the distribution
+median estimates of -0.09. The upper 97.5 percentile of the distribution
 was below 0 for the MLE method, meaning that most of the estimates had
@@
 \begin{verbatim}
-##   Model   p50   p25  p975
-## 1   MLE -0.10 -0.11 -0.06
-## 2   NAS -0.09 -0.13  0.03
-## 3  ELBn -0.09 -0.13  0.05
+##   Model   mean    p50    p25   p975
+## 1   MLE -0.100 -0.101 -0.112 -0.062
+## 2  ELBn -0.091 -0.095 -0.134  0.052
+## 3   NAS -0.088 -0.091 -0.126  0.027
 \end{verbatim}
@@
 
-\textbf{NOTE} I need to re-plot these in ggplot and and stitich
-together. Currently stitching pre-made png files, I think that's why it
-looks so funky.
-
 \newpage
@@
 the same sign of the relationship, but the estimates from the binning
-methods are generally greater than the MLE estimates.}
+methods are generally greater than the MLE estimates. \textbf{NOTE} I
+need to re-plot these in ggplot and and stitch together. Currently
+stitching pre-made png files, I think that's why it looks so funky.}
 \end{figure}
@@
 standard deviation (0.076 \(\pm\) 0.013). The range of the gradient in
-the AMD data is relatively small (9.5), meaning that the absolute change
-in the size spectra parameter (\(\lambda\)) across the AMD gradient are
-similar regardless of method used (range of absolute: 0.59 to 0.72).
+the AMD data is relatively small (9.5), but the absolute change in the
+size spectra parameter (\(\lambda\)) across the AMD gradient vary by
+\textasciitilde0.13 units depending on the method used (range of
+absolute change: 0.59 to 0.72).
 
@@
 data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
-change in \(\lambda\) estimates ranges from 0.06 to 0.17, depending on
-the method used.
+change in \(\lambda\) estimates varies by 0.11 units depending on the
+methods used (range of absolute change across gradient: 0.06 to 0.17).
 
@@
 the empirical estimates of parameters describing size spectra
-relationships. Therefore, we repeated the simulation process as in the
-main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5,
-with a known relationship of \(\beta_1 = -0.2\)
+relationships in streams. Therefore, we repeated the simulation process
+as in the main analysis, but used \(\lambda\) values ranging from -1.1
+to -1.5, with a known relationship of \(\beta_1 = -0.2\)
 
@@
 each replicate. Each panel is a different method for estimating the size
-spectra parameter.}
+spectra parameter. (match to ``new'' style if needed)}
 \end{figure}
@@
 but the mean magnitude and distribution of values is greater for the
-ELBn and NAS methods.}
+ELBn and NAS methods. (not sure why this is such different proportions,
+need to re plot with right dimensions)}
 \end{figure}
@@
 \begin{verbatim}
-##   Model   p50   p25  p975
-## 1   MLE -0.20 -0.21 -0.18
-## 2  ELBn -0.19 -0.21 -0.16
-## 3   NAS -0.17 -0.18 -0.15
+##   Model   mean    p50    p25   p975
+## 1   MLE -0.199 -0.199 -0.205 -0.183
+## 2  ELBn -0.195 -0.195 -0.207 -0.161
+## 3   NAS -0.174 -0.174 -0.183 -0.151
 \end{verbatim}
+
+\textbf{NOTE} Not sure if these tables are usful or not. Shows same
+information as figures, but might be easier to see the quantitative
+values instead of estimating from plot
 
@@
 Generally, the estimate from the MLE method was always closer to the
-known relationship in simulated data, and always had the smallest
-variation around the estimate.
+known \(\lambda\) value, and resulted in the most accurate \(\beta_1\)
+coefficient estimate for the hypothetical relationship in simulated
+data, and always had the smallest variation around the estimate.
 
@@
 MLE from the binning process alone, since binning is akin to deleting
-information that the model could otherwise use.
-
-One of the benefits of using MLE is that all data points are retained
-for use within the model.
+information that the model could otherwise use. One of the benefits of
+using MLE is that all data points are retained for use within the model.
+
+The variation in estimated relationships across the empirical datasets
+varied by \textasciitilde0.1 units depending on the method used. At
+first glance this may not appear to be a concerning magnitude of
+variation. However, when one considers the range of variation reported
+in many seminal works on size spectra relationships in streams is
+\textasciitilde0.1-0.2 (SI Table \emph{xx}), it is apparent that many of
+these results may be dependent on the method used.
 
@@
 Despite the drop in performance with reduced variation in \(\lambda\)
-values, when a significant relationship was observed, it was generally
-in the correct direction and of a similar magnitude. This suggests that
+values, the estimated relationship coefficient was generally in the
+correct direction and of a similar magnitude. This suggests that
 previously reported significant changes in size spectra parameters
 across environmental gradients and in experimental manipulations are
-plausible, and the magnitude of the relative change is a reasonable
-estimate. Given that all of the data within a study is treated
+plausible. Given that all of the data within a study is treated
 identically, the the over all change in size spectra parameters is
@@
 aspect of community organization across spatiotemporal scales and in
-response to environmental stressors and perturbations.
+response to environmental conditions.
+
+\textbf{NOTE} not sure if the idea below belongs in this paper or not,
+but it's something I've been thinking about for a while. I'm using this
+as an excuse to write out my thoughts and to see what y'all think about
+it. Maybe it's another paper to write in the future?
+
+An interesting finding is that the empirical data sets used here were
+quite a bit shallower than is expected based on theory, and that what
+has been observed empirically in marine systems. There are two,
+non-mutually exclusive, possible explanations for this: 1) freshwater
+streams are more dependent on detrital resources, 2) streams are
+effectively 2D systems, whereas marine systems are 3D. Blanchard et
+al.~(2009) observed that benthic marine communities had shallower slopes
+than pelagic communities, and that this was likely due to a stronger
+reliance on a shared detrital basal resource in the benthic community,
+whereas the pelagic community was largely driven by energy transfer
+through predator-prey interactions. The importance of allochthonous
+resources in stream ecosystems has long been recognized (Vannote et
+al.~1980). The stream data here largely comes from low order, headwater
+streams which may be reasonably considered to be driven by allochthonous
+resources. It is possible that the communities sampled here are
+structured based on a shared basal resource, and that the importance of
+predator-prey interactions is reduced. An alternate hypothesis is the
+relative dimensionality of search space in benthic and pelagic systems.
+Consumption rates and interaction strenghts have been shown to vary
+markedly based on search dimensionality. Consumers generally encounter
+resources more frequently in 3D environments (i.e., aerial, pelagic)
+than in 2D environments (i.e., benthic, terrestrial), and lead to
+stronger consumption rates and interaction strengths. Likewise, it has
+been hypothesized that the scaling exponent for mass-specific search
+rates may vary based on search space dimensionality instead of being set
+at approximately 0.75 (Pawar, Dell, Savage, 2012).
 
@@
 circumstances (large sample size, shallow \(\lambda\)). However, it
-would be extremely difficult to know if this was the case beforehand.
-Furthermore, in order to confirm this result, MLE methods would likely
-need to be employed. With the publication of the sizespectra package
-(2017), producing MLE estimates of size spectra parameters is a
-relatively easy task. Therefore, we recommend using this in all future
-studies of size spectra relationships.
+would be unlikely for a researcher to know if this was the case
+beforehand. Furthermore, in order to confirm this result, MLE methods
+would likely need to be employed. With the publication of the
+\emph{sizespectra} package (2017), producing MLE estimates of size
+spectra parameters is a relatively easy task. Therefore, we recommend
+using this in all future studies of size spectra relationships.
 
@@
 \(M\)}{Range of body sizes, M}}\label{range-of-body-sizes-m}}
+
+\textbf{NOTE} udate all plots to be consistent with main text
 
@@
 
+\hypertarget{comparison-with-other-published-estimates}{%
+\section{Comparison with other published
+estimates}\label{comparison-with-other-published-estimates}}
+
+SI Table. This table shows published estimates of the variation in size
+spectra slopes (or exponents) in empirical studies. It is unclear how to
+directly compare estimates of the slope with different methods.However,
+the published estimates here range from \textasciitilde0.1 to 0.2 across
+the gradients studied. For comparison, the 2.5-95\% quantiles around the
+relationship estimate for the MLE method were \textasciitilde0.1,
+whereas for the ELBn and NAS method they were \textasciitilde0.25 and
+\textasciitilde0.2, respectively. b\_diff is the change in estimate
+(b-low - b-min). System refers to stream communities or mesocosm
+experiments. Method: MLE = maximum likelihod estimate, ELB = equal
+logarthmic binning, the number before indicates the number of bins used.
+The normalization process shifts the estimates by an absolute value of
+1.0. Hence, direct comparison the relative change in normalized and
+non-normalized studies should not introduce any bias. The O'Gorman et
+al.~2017 study used average species size and abundance (Local Size
+Density Relationship, \emph{sensu} White et al.~2007) as opposed to
+individual size distribution. These methods are related, but it is
+unclear how to directly compare estimates from each method.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{read.csv}\NormalTok{(}\StringTok{"lit comparison.csv"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##                                      Author b_diff error error_type b_low
+## 1                      Pomeranz et al. 2021   0.12    NA            -1.37
+## 2     Yvon Durocher et al. 2011 (Community)   0.13    NA            -0.79
+## 3             Dossena et al. 2012 (October)   0.14 0.070         se    NA
+## 4                      O'Gorman et al. 2017   0.15    NA            -0.85
+## 5                      Martinez et al. 2016   0.15    NA            -1.11
+## 6 Yvon Durocher et al. 2011 (Phytoplankton)   0.19    NA            -0.46
+## 7                    McGarvey and Kirk 2018   0.19    NA            -1.81
+## 8               Dossena et al. 2012 (April)   0.21 0.064         se    NA
+##   b_high       System      Driver
+## 1  -1.25      Streams Temperature
+## 2  -0.92 FW mesocosms Temperature
+## 3     NA FW mesocosms Temperature
+## 4  -0.70      Streams Temperature
+## 5  -0.96      Streams    Land Use
+## 6  -0.65 FW mesocosms Temperature
+## 7  -1.62      Streams Seasonality
+## 8     NA FW mesocosms Temperature
+##                                                Method
+## 1                                                 MLE
+## 2                              10 ELB, not normalized
+## 3                               6 ELB, not normalized
+## 4 ln (mean species abundance) ~ ln(mean species mass)
+## 5                               6 ELB, not normalized
+## 6                              10 ELB, not normalized
+## 7                                      NAS, Log2 bins
+## 8                               6 ELB, not normalized
+\end{verbatim}
+
 \end{document}

### 2022-10-24 update_oct_6.Rmd
rmd oct 6 update

---
title: "Manuscript: Detecting differences in Size Spectra"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "jpz"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Quick update post discussion on 10/6

It seemed to me that there were three main points to cover based on our discussion.
1. How do $\lambda$ estimates change across known "shallow" to "steep" values?
2. How do our estimates of the relationships of $\lambda$ differ across "lambda windows"?
3. how do our estimates perform when we have a strong or weak relationship (i.e. $\beta_1$) across a gradient?

# 1 $\lambda$ estimates

![lambda estimates by method from sampling from random distributions of steep (lambda = -2.5) to shallow (lambda = -0.5) exponents](figures/est_lambda_est_b_density.png)

**NOTE** the difference in x-axis scale. 
The MLE (blue) distribution is nearly always centered on the known value of lambda exponent (dashed line, and facet title).

For reference, here is the plot without allown the x-axis to vary. I think it's really hard to see anyhting in this, but the varying x-axis above is also hard to intuit, so not sure which to use. 
![Same plot but with "fixed" x-scale](figures/est_lambda_est_b_density_fixed_x.png)

![Plot of the known lambda (wrong x-axis label) and the lambda estimate from different methods. The dashed line is the 1:1 line, and almost perfectly matches the MLE (Blue) estimate. The two binning method cross the 1:1 line, meaning they systematically underestimate more positive (shallow) lambdas, and over estimate more negative (steeper) lambdas.](figures/known_est_b_line.png)

This figure shows that methods are comparble when lambda = ~-1, but are systematically biased at extreme lambda values. Also note that the theoretical exponent for abundance spectra are $\lambda = 2$, and there is considerable variation at this point in the figure.

# 2 lambda "windows"

Because of the markedly different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of beta = -0.5 across a hypothetical environmental gradient. The lambda "windows" are steep (-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).

![Distribution of relationship estimates (beta_1) in three different "windows" of lambda values. The dashed vertical line is the known relationship value of -0.5.  ](figures/lambda_angle_plot.png)

The distribution of relationship estimates from the MLE method (blue) is approximately centered at the known relationship value. The distribution is widest in the "steep" window, but is still markedly narrower than the binning methods. In addition to having much wider distributions, the binning methods systematically overestimate the known relationship.

# 3 varying the known relationship

Finally, we were interested in finding if the three methods performed differently with varying values of the known relaitonship. Here, we present estimates

![regressions for three different known relationship values: beta = 0, 0.25, and 0.5](figures/vary_beta_plot.png)

Finally, we can also plot the distribution of relationship estimates

![Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
All the methods perform well when no relationship is present. The performance of the binning methods (median values) declines, with relationships being systematically over estimated. Likewise, the width of the distributions increases with stronger relationships across a hypothetical gradient. However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has markedly narrower distributions.

### 2022-12-30 Manuscript.Rmd
finally finishing rewrite from oct 06

@@
 
-1.  Size spectra represent a fundamental attribute of community organization and are increasingly being used in assessments of fresh waters.
+1.  The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra.
 2.
@@
 Body size distributions are a fundamental characteristic of communities.
-The remarkable consistency of these relationships across spatiotemporal scales and habitat types has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010)[@petchey2010].
-Size metrics have commonly been used in marine systems, and are increasingly being applied to assess the condition of freshwater ecosystems (Martinez et al.
-2016, Pomeranz 2018)[@martinez2016; @pomeranz2019].
+In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities (Damuth 1981, 1991, 1998).
+This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology (Nee 1991, Brown et al.
+2004).
+In a similar way $\lambda$ of size spectra is expected to be -2.
+The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010)[@petchey2010].
+Deviations from the putative universal $\lambda$ have been reported, but little attention has been paid to whether these are systematic.
 
-Individual size distributions (ISD, also referred to as abundance size spectra) are commonly used.
-Generally, there is a negative relationship between body size (M) and abundance (N).
+Systematic variation in size spectra would provide evidence against the universality of the MTE (Brown et al. 2004), with important consequences for our understanding of nature.
+
+Individual size distributions (ISD (*sensu* White et al.
+2007[@white2007]), also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
+ISDs show the relationship between the abundances of individual body size regardless of taxonomy.
+Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
 Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$.
-Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated from OLS regressions of log transformed data of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
+Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions of log transformed data of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
 Myriad binning methods have been proposed, including linear and logarithmic bin widths.
@@
 Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ (White et al.
-2007, Edwards et al.
-2017)[@white2007; @edwards2017], and reanalysis of empirical datasets also indicates that the conclusions are dependent on the methodology used (White et al.
-2007, Edwards et al.
-2020)[@white2007; @edwards2020].
-However, recent empirical analysis of stream macroinvertebrate communities across the NAtional Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used (Pomeranz et al.
+2008, Edwards et al.
+2017), and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used (White et al.
+2008, Edwards et al.
+2020)[ @edwards2020].
+However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used (Pomeranz et al.
 2022)[@pomeranz2022].
@@
 In other words, if the data within a study are all treated the same, does a relative change of OLS slope parameters of 0.1 coincide with a relative change of MLE estimates of 0.1?
-In order to answer this question, we simulate body size observations from bounded power law distributions with varied $\lambda$ exponents across a hypothetical environmental gradient and compare the results obtained using three different methodologies common in the literature.
-In addition, we re-analyze two previously published datasets of stream community body sizes across a stress and environmental gradient.
-We find that the MLE method more accurately estimates the site-specific $\lambda$ exponents as well as the relative change in exponents across the hypothetical gradient, as well as having smaller variation.
-The logarthmic binning methods generally perform well, but have larger confidence intervals around the estimates.
-However, the overall conclusions of previously published empirical data sets are not dependent on the method used.
+
+We had three primary objectives in this study: 1) how well do different methods estimate site-specific $\lambda$s?
+2) Do the estimates of the hypothetical relationship obtained from different methods agree?
+and 3) Are the conclusions reached from previously published data sets dependent on the method used?
+
+In order to answer the first two questions, we simulate body size observations from bounded power law distributions with varied $\lambda$ exponents across a hypothetical environmental gradient and compare the results obtained using three different methodologies common in the literature.
+To answert the third question, we re-analyze two previously published data sets of stream community body sizes across a stress and environmental gradient, respectively.
+We find that the MLE method more accurately estimates the site-specific $\lambda$ exponents as well as the relative change in exponents across the hypothetical gradient, as well as having smaller variation around the estimates.
+The logarithmic binning methods generally perform well, but have larger confidence intervals around the site-specific estimates.
+In addition, the variation of the estimates of the hypothetical relationship obtained from the binning methods is dependent on the value of $\lambda$ used (under and over estimates steep and shallow $\lambda$, respectively).
 
@@
 
-For the main analysis, presented here, we sampled $n = 1000$ body sizes from distributions described by five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5), such that our model for $\lambda$ is:
+## Site-specific $\lambda$ estimates
 
-$$\lambda = Norm(\alpha -0.5X, \sigma)$$
-
-Each value of $\lambda$ was assumed to come from a community across a hypothetical environmental gradient $X$.
-Values of $X$ for each community are uniformly distributed from -1 to 1.
-We repeated the data simulation 1000 times (reps).
-The main results presented here were not dependent on the range of *x*-values or the number of sites (supplemental information).
-
-### Sample size, $n$
-
-The number of observations in our simulations may bias the results.
-Therefore, we repeated the simulations described above, but varied the sample size $n$.
-We tested values of $n = 200, 500, 1000, 5000, 10 000$.
-
-### "No" relationship
-
-It is possible that our sampling or simulation framework could bias the results obtained.
-In order to assess whether or not our simulations are robust to this, we performed independent random samples from a bounded power law for five communities across a hypothetical gradient as described above, but we set $\lambda = -2$ for all five communities.
-Hence, we are simulating a scenario where the size spectra relationship is invariant to our hypothetical environmental gradient, and we should expect the estimated relationship coefficient $\beta_1 = 0$.
-
-## Finer resolution of $\lambda$ values
-
-For the main analysis, we were interested in determining whether or not the methods were able to detect relatively coarse changes to size spectra parameters, specifically, a change of 0.5 units across each value of $X$, or an absolute change in $\lambda$ values of 1.0 units.
-However, many seminal works on variation in size spectra relationships in freshwater are on a much finer scale (on the order 0f 0.1 to 0.25 units).
-Therefore, we performed simulations as described above, but varied $\lambda$ from -1.9 to -2.1 across the hypothetical environmental gradient.
-This equates to a know relationship coefficient of $\beta_1 = 0.1$ across the hypothetical gradient, and an absolute change of 0.2 units.
+We sampled $n = 1000$ body sizes from distributions described by five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5).
+The values of $\lambda$ describe how quickly the abundance of large body sizes decline within a community, e.g., a value of -1.5 means there would be a relatively high number of large body sizes (shallow) whereas a value of -2.5 means there would be relatively very few large body sizes (steep).
+We used three different methods (described below) to estimate the value of $\lambda$.
+We repeated this process 1000 times (reps) and plot the distribution of values obtained for each method.
 
@@
 
-For each sample of $M$ and each of the 1000 replicates, we estimated the exponent $\lambda$ using MLE methods modified from the the `sizeSpectra` package (Edwards et al.
+For each sample of $M$ and each of the 1000 replicates, we estimated the exponent $\lambda$ directly using MLE methods modified from the the `sizeSpectra` package (Edwards et al.
 2020).
+Throughout the manuscript, these estimates are referred to as MLE.
+
+### Equal Logarithmic Bins: ELBn
+
 In addition, we use two common binning methods to estimate the OLS slope parameter in log-log space.
+It is important to note that the binning methods are estimating a slope ($\beta_{1}$) from an OLS regression analysis, and not $\lambda$ directly.
+However, in the interest of clarity, we will refer to all estimates as $\lambda$ regardless of the method used, and reserve the term $\beta_{env}$ to refer to the estimate of the size spectra relationships across the hypothetical environmental gradient.
+
 For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes.
-The count in each bin was normalized by dividing by the bin width.
-This method has been used by (refs here, I think a lot of Woodward's group uses this...) Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
+This method has been used extensively in previous studies (for example: Perkins et al.
+(2018), Martinez et al.
+(2016), Dossena et al.
+(2012)).
+In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal width of the bin sizes.
+The process of normalizing shifts the estimate by 1.
+In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data (Sprules and Barth 2016, Edwards et al.
+2018, Pomeranz et al.
+2022).
+Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
+
+### Log~2~ Bins: NAS
 The second method was similar to ELBn, but bins of Log~2~ widths are used.
+The number of bins is dependent on the size ranges present in the data.
+When working with empirical data with different size ranges this can alter the number of bins per site.
+However, since the data here were simulated from a known size range, the number of bins for each site is identical.
 The count in each bin is normalized in the same way.
-This method has been used by [@pomeranz2019; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).
-Previous work has investigated the biases in $\lambda$ estimates when using different methods (Edwards et al.
-(2017)[-@edwards2017], Edwards et al.
-2020[-@edwards2020], White et al.
-2008[-@white2007]).
-The focus of the present work is to investigate biases when estimating the relationship or relative change in parameters across an environmental gradient.
-Hence, we do not discuss the $\lambda$ estimates in the main text, but the results are presented in the Supplemental Information (**NOTE** I'm thinking about dropping this from SI.
-Might be too much info, and it's already been published on by Edwards.
-Does this add anything to the story??).
+This method has been  used (Jennings & Blanchard, 2004; Jennings, Warr, & Mackinson, 2002; Sprules and Barth 2016, @pomeranz2019; @mcgarvey2019) and is referred to as the Normalized Abundance Spectrum (NAS).
 
-## Estimation of relationship across gradient, $\beta_1$
+For the two binning methods, simple OLS regression were conducted in the form $Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)$. Where $\beta_{1}$ is considered to be the $\lambda_{estimate}$.
 
-For each simulation replicate, we estimated how the size spectra parameters varied across the hypothetical gradient.
-Simple OLS regression were conducted in the form $\lambda_{estimate} ~ \beta_0 + \beta_1 * X$.
-The distribution of the relationship coefficient, $\beta_1$, were plotted compared to the known relationship.
+## Variation in $\lambda$ Across a Gradient
+
+### Hypothetical relationshiop of -0.5
+
+Previous work has investigated the biases in $\lambda$ estimates when using different methods [-@edwards2017, @edwards2020, @white2007].
+However, the focus of the present work is to investigate biases when estimating the relationship parameters across an environmental gradient.
+
+To do this, we again sampled body sizes from distributions described by five different $\lambda$'s.
+However, because the estimates from the binning methods varied depending on the value of $\lambda$ (see results), we chose three different "windows" of lambda values.
+The three windows were steep ($\lambda = [-1.50,~ -1.75,~ -2.00,~ -2.25, -2.50]$), medium ($\lambda = [-1.00,~ -1.25,~ -1.50,~ -1.75,~ -2.00]$), and shallow (($\lambda = [-0.50,~ -0.75,~ -1.00,~ -1.25,~ -1.50]$).
+
+**NOTE** This MS seems to be getting long. Possibly present "medium" window in main results, and include other windows in SI
+
+Each sample of body sizes from different $\lambda$ values was assumed to come from a community across a hypothetical environmental gradient $X$.
+Values of $X$ for each community are uniformly distributed from -1 to 1.
+For each $\lambda$ "window", the relative change in $\lambda$ values ($\beta_{env}$) across the gradient is equal to -0.5, such that our model is:
+
+**JEFF** Is this correct? for some reason this seems wrong to me.
+
+$$\beta_{env} = Norm(\alpha -0.5X,~\sigma)$$
+
+We repeated the data simulation 1000 times (reps). The main results presented here were not dependent on the range of *x*-values or the number of sites (supplemental information).
+
+### Varying Hypothetical Relationship
+
+Finally, we also varied the relationship ($\beta_{env}$) across the hypothetical gradient.
+We tested $\beta_{env}$ values of 0, -0.25, and -0.5 by sampling 5 body sizes communities uniformly distributed across the following sets of $\lambda$ values: [-1.5], [-1.25, -1.75], [-1, -2].
+
+## Estimation of relationship across gradient, $\beta_{env}$
+
+For each simulation replicate and size spectra method, we estimated how the size spectra parameters varied across the hypothetical gradient.
+We did this by performing simple OLS regression of the $\lambda_{estimate}$ across the hypothetical gradient ($X$):
+
+$$\lambda_{estimate} = \beta_0 + \beta_{env} * X$$
+
+The distribution of the relationship coefficient, $\beta_{env}$, were plotted compared to the known relationship.
 
@@
 
+# $\lambda$ estimates
+
+There was considerable variation in the $\lambda$ estimate across methodologies (Figure 1).
+The distribution of estimates from the MLE method was almost always symmetrical and centered at the known value of $\lambda$ (Figure 1, blue distribution and dashed line).
+The distribution of estimates from the binning methods were generally wider, and sometimes showed assymetrical distributions.
+
+![Figure 1.
+Distribution of Lambda estimates by method (color) from random samples of body sizes from bounded power law distributions with varying exponents (-2.5, -0.5).
+The figure is facetted by the known lambda parameter (facet title) and is also shown as the dashed line in each facet.](figures/est_lambda_est_b_density.png)
+
+Furthermore, the binning methods systematically over estimated steep size spectra relationships ($\lambda = ~\sim-2.5~to-1.5$, Figure 2), and slightly underestimated shallow size spectra relationships ($\lambda = ~\sim-0.5$).
+This finding was more pronounced in the NAS method compared with the ELBn method.
+
+![Plot of the known and estimated lambda from different methods (color).
+The dashed line is the 1:1 line, and almost perfectly matches the MLE (Blue) estimate.
+The two binning method cross the 1:1 line, meaning they systematically underestimate shallow lambdas, and over estimate steeper lambdas.](figures/known_est_b_line.png)
+
 ## Relationship across the hypothetical environmental gradient
 
-![Relationship estimates across the hypothetical gradient for each replicate. Each panel is a different method for estimating the size spectra parameter.](figures/PLB_sim_main.png)
+### $\lambda$ "windows"
 
-![alternate plot version; I think I like this one better. What are your thoughts on which you prefer??](figures/PLB_main_v2.png)
+Because of the markedly different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of $\beta_{env} = -0.5$ across a hypothetical environmental gradient.
+The lambda "windows" are steep (-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).
 
-The distribution of $\beta_1$ estimates from MLE method were the most accurate and precise, and the NAS method was the least accurate and precise (Table *xx*).
-The median estimates for all methods were closer to the known relationship.
-All of the methods produced estimate distributions with long right tails, but this pattern was greater in the ELBn and NAS methods (e.g., SD for ELBn and NAS methods were 2-3x larger, Figure *xx* Halfeye plots).This result was consistent across simulation parameters, including sample size, number of sites, and $\lambda$ values.
+![Distribution of relationship estimates ($beta_{env}$) in three different "windows" of lambda values. The dashed vertical line is the known relationship value of -0.5.](figures/lambda_angle_plot.png)
 
-\newpage
+The distribution of relationship estimates from the MLE method (Figure 3, blue) is approximately centered at the known relationship value in each of the "windows".
+The distribution is widest in the "steep" window, but is still markedly narrower than the binning methods.
+In addition to having much wider distributions, the binning methods systematically overestimate the known relationship.
 
-![Halfeye plots for the estimated relationship coefficient estimates across 1000 replicates and three different methods.
-The point below each distribution plot is the median estimate, the thick and thin bars represent the 66th and 95th percentile, respectively.
-Vertical line is the known relationship.
-The median estimate for the MLE method was closest to the known relationship, and had the narrowest distribution across replicates.](figures/PLB_sim_relationship_density.png)
+### Varying the Known Relationship
 
-\newpage
+Finally, we were interested in finding if the three methods performed differently with varying values of the known relationship.
+Here, we present estimates from each method when the known relationship ($\beta_{env}$) across the hypothetical gradient was -0.5, -0.25, and 0, respectively (Figure 4).
 
-```{r, echo=FALSE, warning=FALSE}
-# will turn these into "real" formatted tables if we decide they're useful
-read.csv("results/PLB_sim.csv") %>%
-  mutate(Model = name,
-         mean = mu_slope) %>%
-  select(Model, mean, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 3))) %>%
-  arrange(p50, p975)
-```
+![regressions for three different known relationship values: beta = 0, 0.25, and 0.5](figures/vary_beta_plot.png)
 
-![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).
-(match this figure to "new" style if we like that better)](figures/n_vary_main.png)
+Likewise, we also present the distribution of the $\beta_{env}$ relationship estimates (Figure 5).
+All the methods perform reasonably well when no relationship is present.
+The performance of the binning methods (median values) declines, with relationships being systematically over estimated.
+Likewise, the width of the distributions increases with stronger relationships across a hypothetical gradient.
+However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has markedly narrower distributions.
 
-\newpage
-
-## No relationship
-
-All of the methods performed similarly when there was no relationship across the hypothetical gradient.
-The mean $\beta_1$ estimates for all three methods ranged from -0.0005 to 0.0002.
-Although the distributions were closely centered on 0, the distribution of coefficient estimates was much wider for the two binning methods; the 95th percentile of the MLE distributions is narrower than the 66th percentiles for the NAS and ELBn methods.
-
-![Individual regression estimates when no relationship exists across the hypothetical environmental gradient.
-The methods are separated by rows.
-The median for all methocs is approximately zero, but the distribution for the MLE method is narrower than the two binning methods.](figures/PLB_static_b_relationship_density.png)
-
-```{r, echo=FALSE, warning=FALSE}
-read.csv("results/PLB_static_b.csv") %>%
-  mutate(Model = name,
-         mean = mu_slope) %>%
-  select(Model, mean, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 3))) %>%
-  arrange(p50, p975)
-```
-
-## Small variation in lambda
-
-The performance of all methods to detect small variation in the $\lambda$ parameter (between -1.9 to -2.1) across the environmental gradient was also investigated.
-The median $\beta_1$ coefficient estimates for the MLE method was -0.1, which exactly matched the value for the hypothetical relationship, while the other two methods both had median estimates of -0.09.
-The upper 97.5 percentile of the distribution was below 0 for the MLE method, meaning that most of the estimates had the correct sign.
-In contrast, the upper 97.5 percentile for both the NAS and ELBn methods were greater than 0 at 0.027 and 0.052, respectively.
-(Figure 4).
-
-![Distribution of the estimated relationship for each method.
-Here, $\lambda$ varied from -1.9 to -2.1, with a known relationaship of -0.1 across the gradient indicated by the dashed vertical line.
-The distribution of relationship estimates was approximately centered at the known relationship for all methods, with a narrower distribution observed for the MLE estimates.](figures/PLB_sim_small_relationship_density.png)
-
-```{r, echo=FALSE, warning=FALSE}
-read.csv("results/PLB_sim_small.csv") %>%
-  mutate(Model = name,
-         mean = mu_slope) %>%
-  select(Model, mean, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 3))) %>%
-  arrange(p50, p975)
-```
+![Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
 
@@
 
-## Simulating shallower lambdas
-
-Estimates of the relationship between lambda and a hypothetical gradient varied depending on the method used.
-However, interpretations of empirical data were broadly consistent across methodologies.
-Upon closer inspection, the values of size spectra parameters in the empirical data were considerably shallower than -2.
-Given that we found the performance of all methods increased with shallower $\lambda$'s we wanted to investigate how the methods performed with simulated values closer to the empirical estimates of parameters describing size spectra relationships in streams.
-Therefore, we repeated the simulation process as in the main analysis, but used $\lambda$ values ranging from -1.1 to -1.5, with a known relationship of $\beta_1 = -0.2$
-
-\newpage
-
-![Relationship estimates across the hypothetical gradient for each replicate.
-Each panel is a different method for estimating the size spectra parameter.
-(match to "new" style if needed)](figures/shallow_lambda_main.png)
-
-We found generally less variability in the relationship estimates ($\beta_1$) across a gradient of distributions described with shallower $\lambda$ parameters.
-However, the median MLE estimate was once again closest to the hypothetical relationship and had the narrowest distribution.
-The median ELBn estimate was close to the hypothetical relationship but had a slightly wider distribution.
-Finally, 95% of the NAS estimates were greater than the hypothetical relationship.
-In general, the distributions from the different methodologies followed the same overall pattern, with the MLE estimates being the smallest, and the binning methods being larger (i.e., more positive).
-
-![Distribution of relationship coefficient estimates.
-Vertical line is the known relationship.
-All methods under estimate the value, but the mean magnitude and distribution of values is greater for the ELBn and NAS methods.
-(not sure why this is such different proportions, need to re plot with right dimensions)](figures/PLB_shallow_lambda_relationship_density.png)
-
-```{r, echo=FALSE, warning=FALSE}
-read.csv("results/shallow_lambda.csv") %>%
-  mutate(Model = name,
-         mean = mu_slope) %>%
-  select(Model, mean, p50, p25, p975) %>%
-  mutate_at(vars(-Model), funs(round(., 3))) %>%
-  arrange(p50, p975)
-```
-**NOTE** Not sure if these tables are usful or not. Shows same information as figures, but might be easier to see the quantitative values instead of estimating from plot
-
-\newpage
-
 # Discussion
 
+The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by Brown 1995, Gaston and Blackburn 2000, White et al.
+2007).
+Empirical patterns are remarkably consistent, and can be explained by the metabolic theory of Ecology (Brown et al.
+2004).
 Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology.
@@
 However, how these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
-Here, we sampled body sizes from known distributions with varying parameters ($\lambda$) and estimated the coefficient of the relationship ($\beta_1$) across a hypothetical gradient in order to assess how the results would vary depnding on the method used.
+Here, we sampled body sizes from known distributions with varying parameters ($\lambda$) and estimated the coefficient of the relationship ($\beta_{env}$) across a hypothetical gradient in order to assess how the results would vary depending on the method used.
 Likewise, we compared how the interpretation of previously published results could change depending on the methodology used to estimate the relationship.
 
-Generally, the estimate from the MLE method was always closer to the known $\lambda$ value, and resulted in the most accurate $\beta_1$ coefficient estimate for the hypothetical relationship in simulated data, and always had the smallest variation around the estimate.
+Generally, the estimate of the size spectra parameter from the MLE method was always closer to the known $\lambda$ value, regardless of how steep or shallow the $\lambda$ exponent was.
+This is in agreement with Edwards et at.
+(2017) which found that the MLE estimate more consistently recovered known $\lambda$ values ranging from -2.5 to -0.5.
+We further this work by discovering that the magnitude of the error associated with the two binning methods increases as the $\lambda$ value moves away from $\sim$ -1.
+
+Likewise, the MLE method also resulted in the most accurate $\beta_{env}$ estimates for the hypothetical relationship in simulated data regardless of the $\lambda$ "window" investigated.
+Furthermore, the variation around the estimate was always smallest with the MLE method.
+However, the variation in the MLE estimate did increase with steeper $\lambda$ "windows".
+This is not surprising, however because steep relationships like this make it very unlikely to sample the largest body sizes, and capturing the "true" relationships without the full range of body sizes present is more difficult (Edwards et al.
+2017).
+Despite this, the MLE method is nearly always the best option to use, likely because all the data are used in the analysis.
 
 Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies.
-However, aggregating individuals into logarithmic bins removes a large source of the variation within the data by collapsing body size variation into a single values within bins.
+However, aggregating individuals into logarithmic bins removes a large source of the variation within the data by collapsing body size variation into a single value within each bin.
 For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
-Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2 grams or \~4 grams.
+Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams.
 By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use.
@@
 At first glance this may not appear to be a concerning magnitude of variation.
-However, when one considers the range of variation reported in many seminal works on size spectra relationships in streams is ~0.1-0.2 (SI Table *xx*), it is apparent that many of these results may be dependent on the method used.
+However, when one considers the range of variation reported in many seminal works on variation in size spectra relationships is ~0.1-0.2 (SI Table *xx*), it becomes clear how important an error in estimates of this magnitude is.
 
-The performance of all methods improved as sample size increases, and as the exponent of the size spectra relationship gets larger (more shallow).
-As either or both of these variables change, the difference in estimated relationship coefficients declines across the methods.
-This is particularly interesting given the fact that both empirical data sets of stream communities examined here are shallower than the expected relationship of $\lambda = -2$, and the conclusions of the change in size spectra relationships across the gradients are not dependent on method used.
-If communities under study are in fact described by shallower exponents, the method used may not be critical in the conclusions reached.
-However, the MLE method performs as well or better than the two binning methods examined here under all contexts and should be the preferred method in future studies.
-At a minimum, future studies should report MLE estimates to ensure that the results are not dependent on the methodology used.
+Despite the differences in magnitude of the empirical relationship coefficients, they were generally in the correct direction and of a similar magnitude.
+This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
+Given that all of the data within a study is treated identically, the over all change in size spectra parameters is likely reasonable.
+However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.
+However, the relatively systematic error of $\lambda$ estimates across known exponents suggests that it may be possible to correct previous estimates onto the same scale (Figure 2).
 
-Despite the drop in performance with reduced variation in $\lambda$ values, the estimated relationship coefficient was generally in the correct direction and of a similar magnitude.
-This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
-Given that all of the data within a study is treated identically, the the over all change in size spectra parameters is likely reasonable.
-However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to compare the relative changes across different published studies which use different methods.
-Publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
-
-**NOTE** not sure if the idea below belongs in this paper or not, but it's something I've been thinking about for a while.
-I'm using this as an excuse to write out my thoughts and to see what y'all think about it.
-Maybe it's another paper to write in the future?
-
-An interesting finding is that the empirical data sets used here were quite a bit shallower than is expected based on theory, and that what has been observed empirically in marine systems.
-There are two, non-mutually exclusive, possible explanations for this: 1) freshwater streams are more dependent on detrital resources, 2) streams are effectively 2D systems, whereas marine systems are 3D.
-Blanchard et al.
-(2009) observed that benthic marine communities had shallower slopes than pelagic communities, and that this was likely due to a stronger reliance on a shared detrital basal resource in the benthic community, whereas the pelagic community was largely driven by energy transfer through predator-prey interactions.
-The importance of allochthonous resources in stream ecosystems has long been recognized (Vannote et al.
-1980).
-The stream data here largely comes from low order, headwater streams which may be reasonably considered to be driven by allochthonous resources.
-It is possible that the communities sampled here are structured based on a shared basal resource, and that the importance of predator-prey interactions is reduced.
-An alternate hypothesis is the relative dimensionality of search space in benthic and pelagic systems.
-Consumption rates and interaction strenghts have been shown to vary markedly based on search dimensionality.
-Consumers generally encounter resources more frequently in 3D environments (i.e., aerial, pelagic) than in 2D environments (i.e., benthic, terrestrial), and lead to stronger consumption rates and interaction strengths.
-Likewise, it has been hypothesized that the scaling exponent for mass-specific search rates may vary based on search space dimensionality instead of being set at approximately 0.75 (Pawar, Dell, Savage, 2012).
+Despite this possibility, publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
 
@@
 
-Both of the binning methods can perform as well as MLE under certain circumstances (large sample size, shallow $\lambda$).
-However, it would be unlikely for a researcher to know if this was the case beforehand.
-Furthermore, in order to confirm this result, MLE methods would likely need to be employed.
+The MLE method outperformed both of the binning methods under nearly any measure.
 With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
@@
 ```
-## Range of body sizes, $M$
+# $\lambda$ and relationship estimates
 
-**NOTE** udate all plots to be consistent with main text
-
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_main.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
-Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_relationship_density.png)
-
-\newpage
-
-## Large environmental gradient
-
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_relationship_density.png)
-
-\newpage
+In the main analysis, we simulate body size data from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.
+For the results presented in the main text, we held the number of sites at 5, scaled the environmental gradient from -1 to 1, and set the minimum and maximum body sizes to $0.0026$ and $1.2 *10^3$ respectively.
+Here, we plot the results of varying the number of sites (3, 10), increasing the scale of the environmental gradient (-100 to 100) and decreasing the range of body sizes (min = 1, max = 100).
+Generally, the results reported in the main manuscript are robust to changing these parameters: the MLE estimate is nearly always closer to the known parameters, and the variation in these estimates is usually smaller than the binning methods.
 
@@
 
-## Sample size
+## Large environmental gradient
+
+![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5.
+Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_main.png)
+
+\newpage
+
+![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values.
+Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_est_b_density.png)
+
+\newpage
+
+![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
+Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_relationship_density.png)
+
+\newpage
+
+## Range of body sizes, $M$
+
+![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_main.png)
+
+\newpage
+
+![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
+Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_relationship_density.png)
+
+\newpage
+
+## Sample size, $n$
+
+The number of observations in our simulations may bias the results.
+Therefore, we repeated the simulations described above, but varied the sample size $n$.
+We tested values of $n = 200, 500, 1000, 5000, 10 000$.
+Results of this analysis are presented in the Supplemental Information.
+
+![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).
+(match this figure to "new" style if we like that better)](figures/n_vary_main.png)
 
@@
 
-![Individual regressions when varying sample size](figures/n_vary_main.png)
-
-\newpage
-
 ![Distribtuion of the relationship coefficients with varying sample size](figures/n_vary_relationship_density.png)
-
-\newpage
-
-## $\lambda$ and relationship estimates
-
-**NOTE** I'm considering dropping the $\lambda$ estimate plots from the manuscript. I think it's too much info, but if it adds to the story I'm happy to leave it in. **END**
-
-In the main analysis, we simulate body size data from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.
-For the results presented in the main text, we held the number of sites at 5, scaled the environmental gradient from -1 to 1, and set the minimum and maximum body sizes to $0.0026$ and $1.2 *10^3$ respectively.
-Here, we plot the results of varying the number of sites (3, 10), increasing the scale of the environmental gradient (-100 to 100) and decreasing the range of body sizes (min = 1, max = 100).
-Generally, the results reported in the main manuscript are robust to changing these parameters: the MLE estimate is nearly always closer to the known parameters, and the variation in these estimates is usually smaller than the binning methods.
-
-\newpage
-
-### Number of sites = 3
-
-![Distribution of estimated $\lambda$ coefficient across a hypothetical gradient with known values (dashed line).
-The number of sites is reduced to 3, compared with 5 sites presented in the main anaysis.](figures/PLB_3_sites_est_b_density.png)
-
-\newpage
-
-![Relationship estimate](figures/PLB_3_sites_relationship_density.png)
-
-\newpage
-
-### Number of sites = 10
-
-![Distribution of estimated $\lambda$ coefficient across a hypothetical gradient with known values(dashed line).
-The number of sites is increased to 10, compared with 5 sites presented in the main anaysis.](figures/PLB_10_sites_est_b_density.png)
-
-\newpage
-
-![Relationship estimate](figures/PLB_10_sites_relationship_density.png)
-
-\newpage
-
-### Environmental Gradient
-
-![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values (dashed line).
-The range of the environmental gradient was increased (-100 to 100).](figures/PLB_large_x_est_b_density.png)
-
-\newpage
-
-![Relationship estimate](figures/PLB_large_x_relationship_density.png)
 

### 2022-12-30 1_simulations.R

@@ -132,6 +132,8 @@ saveRDS(PLB_small_m, file = paste0("data_sim/",
                                substitute(PLB_small_m),
                                "_dat.rds"))
 
+#PLB_small_m <- readRDS("data_sim/PLB_small_m_dat.rds")
+
 plot_sim(PLB_small_m)
 rm(PLB_small_m)
 

### 2022-12-30 AMD_estimate.R

@@ -27,7 +27,7 @@ dw_range = range(amd_dat$dw, na.rm = TRUE)
   unnest(cols = method_compare))
 
 # plot parameter estimates (slope or exponent) across AMD gradient
-amd_result %>%
+amd_plot <- amd_result %>%
   mutate(Model = factor(name,
                         levels = 
                           c("MLE",
@@ -42,15 +42,17 @@ amd_result %>%
   scale_color_manual(
     values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
-  labs(title = "Change in exponent across an AMD gradient", 
-       x = "AMD gradient (PCA)",
-       y = "Slope/exponent estimate") +
+  labs(title = "AMD data", 
+       x = "AMD gradient",
+       y = "Parameter estimate") +
   theme(legend.position = "none")
+
 # save the plot
-ggsave("figures/AMD_plot.png",
-       width = 1600,
-       height = 753, 
-       units = "px")
+saveRDS(amd_plot, "figures/AMD_plot.rds")
+# ggsave("figures/AMD_plot.png",
+#        width = 1600,
+#        height = 753, 
+#        units = "px")
 
 # estimate the beta_1_ relationship across the gradient
 amd_relationship <- amd_result %>%
@@ -71,7 +73,7 @@ amd_relationship %>%
   write_csv("results/AMD_estimates.csv")
 
 # plot the mean \pm STD.E of the estimated relationship across the gradient based on method used. 
-amd_relationship %>%
+amd_relationship_plot <- amd_relationship %>%
   mutate(Model = factor(name,
                         levels = 
                           c("MLE",
@@ -87,34 +89,8 @@ amd_relationship %>%
     values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
   labs(x = "Relationship estimate",
-       y = "Method",
-       title = "Mean +/- Std. Error Beta across AMD gradient") +
+       y = "Method") +
   theme(legend.position="none")
 
-ggsave("figures/amd_relationship.png",
-       height = 3,
-       width = 5)
+saveRDS(amd_relationship_plot, "figures/AMD_relationship.rds")
 
-
-
-
-
-# estimate by sample ------------------------------------------------------
-# don't think I'm doing this anymore...
-# 
-# # need to re-export data from AMD project and include "smaple" number
-# 
-# (amd_result <- amd_dat %>%
[…]

### 2022-12-30 NEON_estimate.R

@@ -27,7 +27,7 @@ dw_range = range(neon_dat$dw, na.rm = TRUE)
     select(-data) %>%
     unnest(cols = method_compare))
 
-neon_result %>%
+neon_plot <- neon_result %>%
   mutate(Model = factor(name,
                         levels = 
                           c("MLE",
@@ -44,11 +44,13 @@ ggplot(aes(x = mat.c,
     values = c("#019AFF",
                "#FF914A",
                "#FF1984" )) +
-  labs(title = "Change in exponent across NEON", 
+  labs(title = "NEON data", 
        x = "Mean annual air temp",
-       y = "Slope/exponent estimate")
+       y = "Parameter estimate") +
+  theme(legend.position = "none")
 
-ggsave("figures/NEON_plot.png")
+saveRDS(neon_plot, "figures/NEON_plot.rds")
+#ggsave("figures/NEON_plot.png")
 
 
 neon_relationship <- neon_result %>%
@@ -67,7 +69,7 @@ neon_relationship %>%
            (max(neon_dat$mat.c) - min(neon_dat$mat.c))) %>%
   write_csv("results/NEON_estimates.csv")
 
-neon_relationship %>%
+neon_relationship_plot <- neon_relationship %>%
   mutate(Model = factor(name,
                         levels = 
                           c("MLE",
@@ -82,103 +84,105 @@ neon_relationship %>%
   scale_color_manual(
     values = c("#019AFF", "#FF914A", "#FF1984" )) +
   theme_bw() +
-  labs(y = "Relationship estimate",
-       title = "Mean +/- Std. Error Beta across Neon")
+  labs(y = "Method",
+       x = "Relationship estimate") 
 
-ggsave("figures/neon_relationship.png")
+saveRDS(neon_relationship_plot, "figures/neon_relationship.rds")
+#ggsave("figures/neon_relationship.png")
 
 
 
 # estimate by sample ------------------------------------------------------
 
+# delete this??? ####
 # not sure where to go with this want to compare what the estimates are for each sample (i.e. surber sample), and see how that compares to the avergae across samples, and the estimate when combining all of the data for one estimate. 
 
 # need to re-export NEON data from project and include sample number
 # estimate CSS parameter for each sample
 # compare "total" estimate with individual sample estimates and mean_sample estimates
 
[…]

### 2022-12-30 empirical plot.R

@@ -3,21 +3,19 @@
 # combine the AMD and NEON data into a single multipanel plot
 
 library(ggplot2)
-library(png)
-library(grid)
-# library(gridExtra)
 library(ggpubr)
-amd_a <- readPNG("figures/AMD_plot.png")
-amd_b <- readPNG("figures/amd_relationship.png")
-neon_a <- readPNG("figures/NEON_plot.png")
-neon_b <- readPNG("figures/neon_relationship.png")
-ggarrange(rasterGrob(amd_a),
-             rasterGrob(amd_b),
-             rasterGrob(neon_a),
-             rasterGrob(neon_b),
+amd_a <- readRDS("figures/AMD_plot.rds")
+amd_b <- readRDS("figures/amd_relationship.rds")
+neon_a <- readRDS("figures/NEON_plot.rds")
+neon_b <- readRDS("figures/neon_relationship.rds")
+
+ggarrange(amd_a,
+             amd_b,
+             neon_a,
+             neon_b,
              labels = c("A", "B", "C", "D"))
 
 ggsave("figures/empirical_combined.png",
-       height = 6,
-       width = 4, 
+       height = 8,
+       width = 8, 
        units = "in")

### 2022-12-30 vary_n.R

@@ -69,7 +69,7 @@ dat <- bind_rows(n200, n500, n5000, n10000, n1000)
 #   NULL
 
 # lambda estimate halfeye ####
-dat %>%
+dat %>% 
   mutate(Model = factor(name,
                         levels = 
                           c("MLE",
@@ -102,19 +102,44 @@ ggsave(paste0("figures/",
 
 # regressions ####
 # what are the estimated relationships across gradient with varying n?
-dat %>%
-  ggplot(aes(x = env_gradient,
-             y = estimate, 
-             group = rep,
-             color = rep)) +
-  stat_smooth(geom = "line",
-              method = "lm",
+# dat %>%
+#   ggplot(aes(x = env_gradient,
+#              y = estimate, 
+#              group = rep,
+#              color = rep)) +
+#   stat_smooth(geom = "line",
+#               method = "lm",
+#               alpha = 0.15,
+#               se = FALSE)+
+#   geom_point() +
+#   facet_wrap(n~name,
+#              ncol = 3,
+#              labeller = label_both)+
+#   theme_bw()
+
+dat %>% 
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
+  ggplot(
+    aes(x = env_gradient, 
+        y = estimate, 
+        color = name,
+        group = rep)) +
+  stat_smooth(method = "lm",
+              se = FALSE, 
+              geom = "line",
+              size = 1.75,
               alpha = 0.15,
-              se = FALSE)+
+              color = "black") +
   geom_point() +
   facet_wrap(n~name,
              ncol = 3,
              labeller = label_both)+
+  scale_color_manual(
+    values = c("#FF914A", "#019AFF",  "#FF1984" )) +
   theme_bw()
 
 ggsave(paste0("figures/", 

### 2023-01-07 Manuscript.Rmd
MS updates. biblio is now working

@@
 linestretch: 1
-bibliography: refs.bib
+bibliography: detecting-size-spectra-difference.bib
 link-citations: no
@@
 
-**Savina** I'm particularly keen to get your feedback on this.
-I think you mentioned including some other references and discussion in Jeff's Bayesian model paper.
-If you think they fit in here please let me know or feel free to include them.
-Likewise, please include any other comments you have, or if you have any insight into how the interpretation of results from different methods would vary.
-
 # Abstract
 
-1.  The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra.
+1.  The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra relationships.
 2.
@@
 3.
-We find that maximum likelihood methods always perform as well or better than common binning methods.
+We find that maximum likelihood methods always perform better than common binning methods.
 Additionally, the variance in estimates using MLE methods is markedly reduced when compared to binning methods.
@@
 2004).
-In a similar way $\lambda$ of size spectra is expected to be -2.
-The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010)[@petchey2010].
-Deviations from the putative universal $\lambda$ have been reported, but little attention has been paid to whether these are systematic.
+The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010)[@Petchey2010].
+Variation in size-abundance relationships have been documented through space (Pomeranz 2021), time (newish Sprules papers?), seasonality (McGarvey and Kirk), to human activities (fish abundance with no fishing paper).
+Likwise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized.
+For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected (Perkins).
+However, recent research has shown that these results may be an artifact of how the data were treated.
+Edwards et al.
+(2020?) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodolugy used.
 
-Systematic variation in size spectra would provide evidence against the universality of the MTE (Brown et al. 2004), with important consequences for our understanding of nature.
-
-Individual size distributions (ISD (*sensu* White et al.
-2007[@white2007]), also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
+Individual size distributions (ISD *sensu* [@White2007]), also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
 ISDs show the relationship between the abundances of individual body size regardless of taxonomy.
 Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
-Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$.
+Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$.
 Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions of log transformed data of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
@@
 2008, Edwards et al.
-2017), and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used (White et al.
-2008, Edwards et al.
-2020)[ @edwards2020].
-However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used (Pomeranz et al.
-2022)[@pomeranz2022].
+2017), and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008 @edwards2020].
+However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used [@Pomeranz2022].
 While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time.
@@
 
-We had three primary objectives in this study: 1) how well do different methods estimate site-specific $\lambda$s?
+We had three primary objectives in this study: 1) how well do different methods estimate site-specific $\lambda$'s?
 2) Do the estimates of the hypothetical relationship obtained from different methods agree?
@@
 In order to answer the first two questions, we simulate body size observations from bounded power law distributions with varied $\lambda$ exponents across a hypothetical environmental gradient and compare the results obtained using three different methodologies common in the literature.
-To answert the third question, we re-analyze two previously published data sets of stream community body sizes across a stress and environmental gradient, respectively.
+To answer the third question, we re-analyze two previously published data sets of stream community body sizes across a stress and environmental gradient, respectively.
 We find that the MLE method more accurately estimates the site-specific $\lambda$ exponents as well as the relative change in exponents across the hypothetical gradient, as well as having smaller variation around the estimates.
-The logarithmic binning methods generally perform well, but have larger confidence intervals around the site-specific estimates.
+The logarithmic binning methods generally perform well, but have larger variation around the site-specific estimates.
 In addition, the variation of the estimates of the hypothetical relationship obtained from the binning methods is dependent on the value of $\lambda$ used (under and over estimates steep and shallow $\lambda$, respectively).
@@
 
-In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in Edwards et al.
-(2017)[-@edwards2017].
+In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@Edwards2017].
 Let $M$ be a random variable of body sizes described by the probability density function:
 
-$$f(M) = CM^\lambda, M_{min} \le M \le M_{max} $$
+$$f(M) = CM^\lambda, M_{min} \le M \le M_{max}$$
 
 $M_{min} = 0.0026$ and $M_{max} = 1.2 *10^3$.
-These values are based on empirical body sizes of stream benthic communities reported in Pomeranz et al.
-(2020)[@pomeranz2022].
+These values are based on empirical body sizes of stream benthic communities reported in [@Pomeranz2022].
 Our results are not dependent on the range of body sizes, and we show the results of other body size ranges in the supplemental information.
@@
 For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes.
-This method has been used extensively in previous studies (for example: Perkins et al.
-(2018), Martinez et al.
-(2016), Dossena et al.
-(2012)).
-In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal width of the bin sizes.
-The process of normalizing shifts the estimate by 1.
-In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data (Sprules and Barth 2016, Edwards et al.
-2018, Pomeranz et al.
-2022).
+This method has been used extensively in previous studies.
+For example see [@Perkins2018, @martinez2016, @dossenaWarmingAltersCommunity2012].
+In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes.
+The process of normalizing shifts the estimate by -1.
+In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprulesSurfingBiomassSize2015, Edwards2018, Pomeranz2022].
 Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
@@
 The count in each bin is normalized in the same way.
-This method has been  used (Jennings & Blanchard, 2004; Jennings, Warr, & Mackinson, 2002; Sprules and Barth 2016, @pomeranz2019; @mcgarvey2019) and is referred to as the Normalized Abundance Spectrum (NAS).
+This method has been  used [@jennings2004, @Jennings2002, @sprulesSurfingBiomassSize2015, @pomeranzAnthropogenicMiningAlters2019; @mcgarveySeasonalComparisonCommunitylevel2018) and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 
-Previous work has investigated the biases in $\lambda$ estimates when using different methods [-@edwards2017, @edwards2020, @white2007].
+Previous work has investigated the biases in $\lambda$ estimates when using different methods [-@Edwards2017, @edwards2020, @White2007].
 However, the focus of the present work is to investigate biases when estimating the relationship parameters across an environmental gradient.
@@
 Each sample of body sizes from different $\lambda$ values was assumed to come from a community across a hypothetical environmental gradient $X$.
-Values of $X$ for each community are uniformly distributed from -1 to 1.
+Values of $X$ for each community are uniformly distributed from -1 to 1 (site-specific x-values: $(-1.0, -0.5, 0, 0.5, 1.0)$.
 For each $\lambda$ "window", the relative change in $\lambda$ values ($\beta_{env}$) across the gradient is equal to -0.5, such that our model is:
@@
 In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage (AMD) stress gradient.
-Details of the sample collection and processing can be found in [@pomeranz2019].
+Details of the sample collection and processing can be found in [@pomeranzAnthropogenicMiningAlters2019].
 Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
@@
 All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions.
-This data has been analyzed previously using size spectra methods as described in [@pomeranz2022].
+This data has been analyzed previously using size spectra methods as described in [@Pomeranz2022].
 Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([<https://www.neonscience.org/data-collection/macroinvertebrates>], or macroinvertebrate DPI pubs).
@@
 
-The data used in this manuscript is already publicly available (Pomeranz et al.
-2019 data dryad DOI **here**, NEON data citation **here**.
+The data used in this manuscript is already publicly available [@Pomeranz2019] data dryad DOI <https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s>, [@NEON_Inverts2022].
 R Scripts to reproduce the full simulation and analysis are available at: ([<https://github.com/Jpomz/detecting-spectra-differences>]).
@@
 Distribution of Lambda estimates by method (color) from random samples of body sizes from bounded power law distributions with varying exponents (-2.5, -0.5).
-The figure is facetted by the known lambda parameter (facet title) and is also shown as the dashed line in each facet.](figures/est_lambda_est_b_density.png)
+The figure is facetted by the known lambda parameter (facet title) and is also shown as the dashed line in each facet.
+Note that the x-axis varies in each facet.](figures/est_lambda_est_b_density.png)
 
@@
 
-![Plot of the known and estimated lambda from different methods (color).
+![Figure 2.
+Plot of the known and estimated lambda from different methods (color).
 The dashed line is the 1:1 line, and almost perfectly matches the MLE (Blue) estimate.
@@
 
-Because of the markedly different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of $\beta_{env} = -0.5$ across a hypothetical environmental gradient.
+Because of the different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of $\beta_{env} = -0.5$ across a hypothetical environmental gradient.
 The lambda "windows" are steep (-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).
 
-![Distribution of relationship estimates ($beta_{env}$) in three different "windows" of lambda values. The dashed vertical line is the known relationship value of -0.5.](figures/lambda_angle_plot.png)
+![Figure 3.
+Distribution of relationship estimates ($beta_{env}$) in three different "windows" of lambda values.
+The dashed vertical line is the known relationship value of -0.5.](figures/lambda_angle_plot.png)
 
 The distribution of relationship estimates from the MLE method (Figure 3, blue) is approximately centered at the known relationship value in each of the "windows".
-The distribution is widest in the "steep" window, but is still markedly narrower than the binning methods.
+The distribution for the MLE estimate is widest in the "steep" window, but is still markedly narrower than the binning methods.
 In addition to having much wider distributions, the binning methods systematically overestimate the known relationship.
@@
 
-![regressions for three different known relationship values: beta = 0, 0.25, and 0.5](figures/vary_beta_plot.png)
+![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows): 0, 0.25, and 0.5](figures/vary_beta_plot.png)
 
 Likewise, we also present the distribution of the $\beta_{env}$ relationship estimates (Figure 5).
-All the methods perform reasonably well when no relationship is present.
-The performance of the binning methods (median values) declines, with relationships being systematically over estimated.
+All the methods perform similarly and reasonably well when no relationship is present.
+The performance of the binning methods declines (center of distributions), with relationships being systematically over estimated.
 Likewise, the width of the distributions increases with stronger relationships across a hypothetical gradient.
-However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has markedly narrower distributions.
+However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has relatively narrow distributions.
 
-![Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
+![Figure 5. Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
 
@@
 Size spectra parameters consistently increase (become flatter) in the AMD data (Fig.
-XX A).
-Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig XX D)
+6A).
+However, the magnitude of the change across the AMD gradient is weakest using the MLE method, and the width of the estimates ($\pm$ 1 SD) is larger in the binning methods, particularly the ELBn method (Fig.
+6B).
+Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig.
+6C).
+The magnitude of the relationships is weakest in the MLE method, although the uncertainty around the estimates are similar (Fig.
+6D).
 
@@
 
-![Figure of empirical data estimates.
+![Figure 6.
+of empirical data estimates.
 All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.
@@
 
-The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by Brown 1995, Gaston and Blackburn 2000, White et al.
-2007).
-Empirical patterns are remarkably consistent, and can be explained by the metabolic theory of Ecology (Brown et al.
-2004).
+The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@Brown1995, GastonBlackburn2000, @White2007]).
+Empirical patterns are remarkably consistent, and can be explained by the metabolic theory of Ecology [@Brown2004].
 Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology.
@@
 We further this work by discovering that the magnitude of the error associated with the two binning methods increases as the $\lambda$ value moves away from $\sim$ -1.
+
+It is unsurprising that all methods performed less well in estimating steep values of $\lambda$ (i.e., -2.5).
+The rapid decline of large body sizes makes it unlikely to sample any individuals with a large body size.
+Furthermore, the binning methods inherently "lose" data by aggregating individual observations into a single bin.
 
@@
 The MLE method outperformed both of the binning methods under nearly any measure.
-With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
+With the publication of the *sizespectra* package [-@Edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
 Therefore, we recommend using this in all future studies of size spectra relationships.
 
-We reiterate the recommendations of White et al.
-(2007)[-@white2007], Sprules and Barth (2016)[-@sprules2016] and Edwards et al.
-(2017)[-@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
+We reiterate the recommendations of [-@White2007, -@sprulesSurfingBiomassSize2015, -@Edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
 Furthermore, we strongly encourage authors to publish individual size data whenever possible.
@@
 
-```{r}
-read.csv("lit comparison.csv")
+```{r, echo=FALSE}
+df <- read.csv("lit comparison.csv")
+knitr::kable(df,
+             col.names = c("Citation",
+                           "$\\beta_1$ Difference",
+                           "Error", 
+                           "Error Type",
+                           "$\\beta_{low}$",
+                           "$\\beta_{high}$",
+                           "System",
+                           "Driver", 
+                           "Method"))
+
 ```

### 2023-01-11 Manuscript.Rmd
new version knit for coauthors jan 11 2023

@@
 2004).
-The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status (Petchey and Belgrano 2010)[@Petchey2010].
-Variation in size-abundance relationships have been documented through space (Pomeranz 2021), time (newish Sprules papers?), seasonality (McGarvey and Kirk), to human activities (fish abundance with no fishing paper).
-Likwise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized.
-For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected (Perkins).
+The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010].
+Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], to human activities [@jennings2004].
+Likewise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized.
+For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected [@perkins2018].
 However, recent research has shown that these results may be an artifact of how the data were treated.
 Edwards et al.
-(2020?) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodolugy used.
+(2020) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodology used.
 
-Individual size distributions (ISD *sensu* [@White2007]), also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
+Individual size distributions (ISD *sensu* [@White2007], also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
 ISDs show the relationship between the abundances of individual body size regardless of taxonomy.
 Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
-Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$.
+Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972, @Andersen2006a].
 Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions of log transformed data of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
@@
 2017), and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008 @edwards2020].
-However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied, the relative change across the environmental gradient was consistent regardless of method used [@Pomeranz2022].
+However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied based on method, the relative change across the environmental gradient was consistent across methods [@pomeranz2022].
 While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time.
-In other words, if the data within a study are all treated the same, does a relative change of OLS slope parameters of 0.1 coincide with a relative change of MLE estimates of 0.1?
+In other words, if the data within a study are all treated the same, does a relative change of OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?
 
@@
 In order to answer the first two questions, we simulate body size observations from bounded power law distributions with varied $\lambda$ exponents across a hypothetical environmental gradient and compare the results obtained using three different methodologies common in the literature.
-To answer the third question, we re-analyze two previously published data sets of stream community body sizes across a stress and environmental gradient, respectively.
+To answer the third question, we re-analyze two previously published data sets of community body sizes across a stress and environmental gradient, respectively [@pomeranz2019, @pomeranz2022].
 We find that the MLE method more accurately estimates the site-specific $\lambda$ exponents as well as the relative change in exponents across the hypothetical gradient, as well as having smaller variation around the estimates.
@@
 
-In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@Edwards2017].
+In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@edwards2017].
 Let $M$ be a random variable of body sizes described by the probability density function:
@@
 $M_{min} = 0.0026$ and $M_{max} = 1.2 *10^3$.
-These values are based on empirical body sizes of stream benthic communities reported in [@Pomeranz2022].
+These values are based on empirical body sizes of stream benthic communities reported in [@pomeranz2022].
 Our results are not dependent on the range of body sizes, and we show the results of other body size ranges in the supplemental information.
@@
 This method has been used extensively in previous studies.
-For example see [@Perkins2018, @martinez2016, @dossenaWarmingAltersCommunity2012].
+For example see [@perkins2018, @martinez2016, @dossena2012].
 In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes.
 The process of normalizing shifts the estimate by -1.
-In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprulesSurfingBiomassSize2015, Edwards2018, Pomeranz2022].
+In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015, @edwards2017, @pomeranz2022].
 Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
@@
 The second method was similar to ELBn, but bins of Log~2~ widths are used.
-The number of bins is dependent on the size ranges present in the data.
-When working with empirical data with different size ranges this can alter the number of bins per site.
+The number of bins is dependent on the range of body sizes present in the data.
+When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global min and max body size present.
 However, since the data here were simulated from a known size range, the number of bins for each site is identical.
-The count in each bin is normalized in the same way.
-This method has been  used [@jennings2004, @Jennings2002, @sprulesSurfingBiomassSize2015, @pomeranzAnthropogenicMiningAlters2019; @mcgarveySeasonalComparisonCommunitylevel2018) and is referred to as the Normalized Abundance Spectrum (NAS).
+The count in each bin is normalized in the same way as described above.
+This method has been  used [@jennings2004, @Jennings2002, @sprules2015, @pomeranz2019; @mcgarvey2018) and is referred to as the Normalized Abundance Spectrum (NAS).
 
-For the two binning methods, simple OLS regression were conducted in the form $Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)$. Where $\beta_{1}$ is considered to be the $\lambda_{estimate}$.
+For the two binning methods, simple OLS regression were conducted in the form 
+$$Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)$$
+
+Where $\beta_{1}$ is considered to be the $\lambda_{estimate}$.
 
@@
 
-Previous work has investigated the biases in $\lambda$ estimates when using different methods [-@Edwards2017, @edwards2020, @White2007].
+Previous work has investigated the biases in $\lambda$ estimates when using different methods [-@edwards2017, @edwards2020, @White2007].
 However, the focus of the present work is to investigate biases when estimating the relationship parameters across an environmental gradient.
@@
 Each sample of body sizes from different $\lambda$ values was assumed to come from a community across a hypothetical environmental gradient $X$.
-Values of $X$ for each community are uniformly distributed from -1 to 1 (site-specific x-values: $(-1.0, -0.5, 0, 0.5, 1.0)$.
+Values of $X$ for each community are uniformly distributed from -1 to 1 (site-specific x-values: $(-1.0,~-0.5,~0,~0.5,~1.0)$.
 For each $\lambda$ "window", the relative change in $\lambda$ values ($\beta_{env}$) across the gradient is equal to -0.5, such that our model is:
@@
 In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage (AMD) stress gradient.
-Details of the sample collection and processing can be found in [@pomeranzAnthropogenicMiningAlters2019].
+Details of the sample collection and processing can be found in [@pomeranz2019].
 Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
@@
 
-The second data set was from the wadeable stream sites of the National Ecological Observatory Network [@NEON_Inverts2022].
+The second data set was from the wadeable stream sites of NEON [@NEON_Inverts2022].
 NEON stream sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska.
@@
 All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions.
-This data has been analyzed previously using size spectra methods as described in [@Pomeranz2022].
-Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([<https://www.neonscience.org/data-collection/macroinvertebrates>], or macroinvertebrate DPI pubs).
+This data has been analyzed previously using size spectra methods as described in [@pomeranz2022].
+Detailed methods of the sampling collection and data processing methods can be in the macroinvertebrate DPI pubs found on the NEON website <https://www.neonscience.org/data-collection/macroinvertebrates>.
 Coefficient estimates ($\pm$ 1 standard deviation) are compared across methods.
@@
 R Scripts to reproduce the full simulation and analysis are available at: ([<https://github.com/Jpomz/detecting-spectra-differences>]).
+(To be archived upon acceptance.)
 
@@
 of empirical data estimates.
-All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.
-**NOTE** I need to re-plot these in ggplot and and stitch together.
-Currently stitching pre-made png files, I think that's why it looks so funky.](figures/empirical_combined.png)
+All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.](figures/empirical_combined.png)
 
@@
 
-The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@Brown1995, GastonBlackburn2000, @White2007]).
+The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@brown1995, GastonBlackburn2000, @White2007]).
 Empirical patterns are remarkably consistent, and can be explained by the metabolic theory of Ecology [@Brown2004].
@@
 The MLE method outperformed both of the binning methods under nearly any measure.
-With the publication of the *sizespectra* package [-@Edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
+With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
 Therefore, we recommend using this in all future studies of size spectra relationships.
 
-We reiterate the recommendations of [-@White2007, -@sprulesSurfingBiomassSize2015, -@Edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
+We reiterate the recommendations of [-@White2007, -@sprules2015, -@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
 Furthermore, we strongly encourage authors to publish individual size data whenever possible.
@@
 ```
+
+```{r, echo=TRUE}
+# kable() above looks awful
+# here is another bad option
+# if we decide to include this I'll make it look nice in excel. 
+read.csv("lit comparison.csv")
+```

### 2023-01-11 Manuscript.tex
new version knit for coauthors jan 11 2023

--- 2022-09-05 Manuscript.tex
@@
 ]{article}
-\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
 \usepackage{setspace}
-\usepackage{iftex}
-\ifPDFTeX
+\usepackage{amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[T1]{fontenc}
@@
   \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
 \else % if luatex or xetex
@@
 \usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{lineno}
+\usepackage{amsmath}
+\usepackage{indentfirst}
+\linenumbers
+\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
 \newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
@@
   % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
   \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
  }%
@@
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\usepackage{lineno}
-\usepackage{amsmath}
-\usepackage{indentfirst}
-\linenumbers
-\newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
-
-\title{Manuscript}
+
+\title{Manuscript: Detecting differences in Size Spectra}
 \author{Justin Pomeranz\textsuperscript{1,*} \and James R.
@@
 Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
-\date{05 September, 2022}
+\date{07 January, 2023}
 
@@
 
-\textbf{Savina} I'm particularly keen to get your feedback on this. I
-think you mentioned including some other references and discussion in
-Jeff's Bayesian model paper. If you think they fit in here please let me
-know or feel free to include them. Likewise, please include any other
-comments you have, or if you have any insight into how the
-interpretation of results from different methods would vary.
-
 \hypertarget{abstract}{%
@@
 \item
-  Size spectra represent a fundamental attribute of community
-  organization and are increasingly being used in assessments of fresh
-  waters.
+  The distribution of body size in communities is remarkably consistent
+  across habitats and taxa and can be represented by size spectra
+  relationships.
 \item
@@
 \item
-  We find that maximum likelihood methods always perform as well or
-  better than common binning methods. Additionally, the variance in
-  estimates using MLE methods is markedly reduced when compared to
-  binning methods.
+  We find that maximum likelihood methods always perform better than
+  common binning methods. Additionally, the variance in estimates using
+  MLE methods is markedly reduced when compared to binning methods.
 \item
@@
   these estimates may be biased, and when the general pattern of results
-  can be considered to be ``true''.
+  can be considered to be ``true.''
 \end{enumerate}
@@
 Body size distributions are a fundamental characteristic of communities.
-The remarkable consistency of these relationships across spatiotemporal
-scales and habitat types has led them to be recommended as a
-``universal'' indicator of ecological status (Petchey and Belgrano
-2010)(Petchey and Belgrano 2010). Size metrics have commonly been used
-in marine systems, and are increasingly being applied to assess the
-condition of freshwater ecosystems (Martinez et al.~2016, Pomeranz
-2018)(Pomeranz et al. 2019, \textbf{martinez2016?}).
-
-Individual size distributions (ISD, also referred to as abundance size
-spectra) are commonly used. Generally, there is a negative relationship
-between body size (M) and abundance (N). Theoretical and empirical data
-support this relationship being described as a simple power law with
-exponent \(\lambda\) in the form of \(N \sim M^{\lambda}\). Commonly,
-\(N\) is the count of body sizes grouped into bins, and \(\lambda\) is
-estimated from OLS regressions of log transformed data of \(N\) and the
+In general, size-abundance relationships are expected to scale with the
+value of -0.75 across all biological communities (Damuth 1981, 1991,
+1998). This is thought to be a consequence of simple size-dependent
+metabolic constraints on organisms' energy use predicted by the
+metabolic theory of ecology (Nee 1991, Brown et al.~2004). The
+remarkable consistency of these relationships across spatiotemporal
+scales and ecosystems has led them to be recommended as a ``universal''
+indicator of ecological status (Petchey and Belgrano 2010)(Petchey and
+Belgrano 2010). Variation in size-abundance relationships have been
+documented through space (Pomeranz 2021), time (newish Sprules papers?),
+seasonality (McGarvey and Kirk), to human activities (fish abundance
+with no fishing paper). Likwise, variation in size spectra relationships
+have been used to explain fundamental differences in how communities are
+organized. For example, external resource subsidies ``bend the rules''
+and allow higher abundances of large body sizes than would be expected
+(Perkins). However, recent research has shown that these results may be
+an artifact of how the data were treated. Edwards et al.~(2020?)
+reanalyzed a time series of marine fisheries data and found that
+previously reported changes through time were actually dependent on the
+methodolugy used.
+
+Individual size distributions (ISD \emph{sensu} (White et al. 2007)),
+also referred to as abundance size spectra) are one of the
+size-abundance relationships commonly used. ISDs show the relationship
+between the abundances of individual body size regardless of taxonomy.
+Generally, there is a negative relationship between individual body size
+(M) on the \emph{x}-axis and abundance (N) on the \emph{y}-axis.
+Theoretical and empirical data support this relationship being described
+as a simple power law with exponent \(\lambda\) in the form of
+\(N \sim M^{\lambda}\), where \(\lambda = 2\). Commonly, \(N\) is the
+count of body sizes grouped into bins, and \(\lambda\) is estimated as
+the slope from OLS regressions of log transformed data of \(N\) and the
 mid point of the body size bin \(M_{bin}\):
@@
 that MLE offers consistently more accurate estimates of \(\lambda\)
-(White et al.~2007, Edwards et al.~2017)(White et al. 2007, Edwards et
-al. 2017), and reanalysis of empirical datasets also indicates that the
-conclusions are dependent on the methodology used (White et al.~2007,
-Edwards et al.~2020)(White et al. 2007, Edwards et al. 2020). However,
-recent empirical analysis of stream macroinvertebrate communities across
-the NAtional Ecological Observatory Network (NEON, USA) showed that
-while the estimates of \(\lambda\) varied, the relative change across
-the environmental gradient was consistent regardless of method used
-(Pomeranz et al.~2022)(Pomeranz et al. 2022). While there is a growing
-consensus that MLE methods offer more reliable estimates of \(\lambda\),
-and binning methods result in biased estimates, it remains unclear if
-these biases are consistent and systematic or stochastic, and whether or
-not the relative change in ISD parameters is consistent across space and
-time. In other words, if the data within a study are all treated the
-same, does a relative change of OLS slope parameters of 0.1 coincide
-with a relative change of MLE estimates of 0.1? In order to answer this
-question, we simulate body size observations from bounded power law
-distributions with varied \(\lambda\) exponents across a hypothetical
-environmental gradient and compare the results obtained using three
-different methodologies common in the literature. In addition, we
-re-analyze two previously published datasets of stream community body
-sizes across a stress and environmental gradient. We find that the MLE
+(White et al.~2008, Edwards et al.~2017), and reanalysis of empirical
+data also indicates that the conclusions are dependent on the
+methodology used Edwards et al. (2020). However, recent empirical
+analysis of stream macroinvertebrate communities across the National
+Ecological Observatory Network (NEON, USA) showed that while the
+estimates of \(\lambda\) varied, the relative change across the
+environmental gradient was consistent regardless of method used
+(Pomeranz et al. 2022). While there is a growing consensus that MLE
+methods offer more reliable estimates of \(\lambda\), and binning
+methods result in biased estimates, it remains unclear if these biases
+are consistent and systematic or stochastic, and whether or not the
+relative change in ISD parameters is consistent across space and time.
+In other words, if the data within a study are all treated the same,
+does a relative change of OLS slope parameters of 0.1 coincide with a
+relative change of MLE estimates of 0.1?
+
+We had three primary objectives in this study: 1) how well do different
+methods estimate site-specific \(\lambda\)'s? 2) Do the estimates of the
+hypothetical relationship obtained from different methods agree? and 3)
+Are the conclusions reached from previously published data sets
+dependent on the method used?
+
+In order to answer the first two questions, we simulate body size
+observations from bounded power law distributions with varied
+\(\lambda\) exponents across a hypothetical environmental gradient and
+compare the results obtained using three different methodologies common
+in the literature. To answer the third question, we re-analyze two
+previously published data sets of stream community body sizes across a
+stress and environmental gradient, respectively. We find that the MLE
 method more accurately estimates the site-specific \(\lambda\) exponents
 as well as the relative change in exponents across the hypothetical
-gradient, as well as having smaller variation. The logarthmic binning
-methods generally perform well, but have larger confidence intervals
-around the estimates. However, the overall conclusions of previously
-published empirical data sets are not dependent on the method used.
+gradient, as well as having smaller variation around the estimates. The
+logarithmic binning methods generally perform well, but have larger
+variation around the site-specific estimates. In addition, the variation
+of the estimates of the hypothetical relationship obtained from the
+binning methods is dependent on the value of \(\lambda\) used (under and
+over estimates steep and shallow \(\lambda\), respectively).
 
@@
 simulate body size observations from a bounded power law distribution
-using the inverse method, as described in Edwards et al.~(2017)(2017).
-Let \(M\) be a random variable of body sizes described by the
-probability density function:
-
-\[f(M) = CM^\lambda, M_{min} \le M \le M_{max} \]
+using the inverse method, as described in (Edwards et al. 2017). Let
+\(M\) be a random variable of body sizes described by the probability
+density function:
+
+\[f(M) = CM^\lambda, M_{min} \le M \le M_{max}\]
 
@@
 on empirical body sizes of stream benthic communities reported in
-Pomeranz et al.~(2020)(Pomeranz et al. 2022). Our results are not
-dependent on the range of body sizes, and we show the results of other
-body size ranges in the supplemental information.
-
-For the main analysis, presented here, we sampled \(n = 1000\) body
-sizes from distributions described by five different \(\lambda\)'s:
-(-1.5, -1.75, -2.00, -2.25, -2.5), such that our model for \(\lambda\)
-is:
-
-\[\lambda = Norm(\alpha -0.5X, \sigma)\]
-
-Each value of \(\lambda\) was assumed to come from a community across a
-hypothetical environmental gradient \(X\). Values of \(X\) for each
-community are uniformly distributed from -1 to 1. We repeated the data
-simulation 1000 times (reps). The main results presented here were not
-dependent on the range of \emph{x}-values or the number of sites
-(supplemental information).
-
-\hypertarget{sample-size-n}{%
-\subsubsection{\texorpdfstring{Sample size,
-\(n\)}{Sample size, n}}\label{sample-size-n}}
-
-The number of observations in our simulations may bias the results.
-Therefore, we repeated the simulations described above, but varied the
-sample size \(n\). We tested values of
-\(n = 200, 500, 1000, 5000, 10 000\).
-
-\hypertarget{no-relationship}{%
-\subsubsection{``No'' relationship}\label{no-relationship}}
-
-It is possible that our sampling or simulation framework could bias the
-results obtained. In order to assess whether or not our simulations are
-robust to this, we performed independent random samples from a bounded
-power law for five communities across a hypothetical gradient as
-described above, but we set \(\lambda = -2\) for all five communities.
-Hence, we are simulating a scenario where the size spectra relationship
-is invariant to our hypothetical environmental gradient, and we should
-expect the estimated relationship coefficient \(\beta_1 = 0\).
-
-\hypertarget{finer-resolution-of-lambda-values}{%
-\subsection{\texorpdfstring{Finer resolution of \(\lambda\)
-values}{Finer resolution of \textbackslash lambda values}}\label{finer-resolution-of-lambda-values}}
-
-For the main analysis, we were interested in determining whether or not
-the methods were able to detect relatively coarse changes to size
-spectra parameters, specifically, a change of 0.5 units across each
-value of \(X\), or an absolute change in \(\lambda\) values of 1.0
-units. However, many seminal works on variation in size spectra
-relationships in freshwater are on a much finer scale (on the order 0f
-0.1 to 0.25 units). Therefore, we performed simulations as described
-above, but varied \(\lambda\) from -1.9 to -2.1 across the hypothetical
-environmental gradient. This equates to a know relationship coefficient
-of \(\beta_1 = 0.1\) across the hypothetical gradient, and an absolute
-change of 0.2 units.
+(Pomeranz et al. 2022). Our results are not dependent on the range of
+body sizes, and we show the results of other body size ranges in the
+supplemental information.
+
+\hypertarget{site-specific-lambda-estimates}{%
+\subsection{\texorpdfstring{Site-specific \(\lambda\)
+estimates}{Site-specific \textbackslash lambda estimates}}\label{site-specific-lambda-estimates}}
+
+We sampled \(n = 1000\) body sizes from distributions described by five
+different \(\lambda\)'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The values
+of \(\lambda\) describe how quickly the abundance of large body sizes
+decline within a community, e.g., a value of -1.5 means there would be a
+relatively high number of large body sizes (shallow) whereas a value of
+-2.5 means there would be relatively very few large body sizes (steep).
+We used three different methods (described below) to estimate the value
+of \(\lambda\). We repeated this process 1000 times (reps) and plot the
+distribution of values obtained for each method.
 
@@
 For each sample of \(M\) and each of the 1000 replicates, we estimated
-the exponent \(\lambda\) using MLE methods modified from the the
-\texttt{sizeSpectra} package (Edwards et al.~2020). In addition, we use
-two common binning methods to estimate the OLS slope parameter in
-log-log space. For the first binning method, we created 6 equal
-logarithmic bins covering the range of body sizes. The count in each bin
-was normalized by dividing by the bin width. This method has been used
-by (refs here, I think a lot of Woodward's group uses this\ldots)
+the exponent \(\lambda\) directly using MLE methods modified from the
+the \texttt{sizeSpectra} package (Edwards et al.~2020). Throughout the
+manuscript, these estimates are referred to as MLE.
+
+\hypertarget{equal-logarithmic-bins-elbn}{%
+\subsubsection{Equal Logarithmic Bins:
+ELBn}\label{equal-logarithmic-bins-elbn}}
+
+In addition, we use two common binning methods to estimate the OLS slope
+parameter in log-log space. It is important to note that the binning
+methods are estimating a slope (\(\beta_{1}\)) from an OLS regression
+analysis, and not \(\lambda\) directly. However, in the interest of
+clarity, we will refer to all estimates as \(\lambda\) regardless of the
+method used, and reserve the term \(\beta_{env}\) to refer to the
+estimate of the size spectra relationships across the hypothetical
+environmental gradient.
+
+For the first binning method, we created 6 equal logarithmic bins
+covering the range of body sizes. This method has been used extensively
+in previous studies. For example see Dossena et al. (2012). In the
+present study, the count in each bin was normalized by dividing by the
+bin width to account for the unequal bin sizes. The process of
+normalizing shifts the estimate by -1. In other words, an un-normalized
+estimate of -0.75 would result in an estimate of -1.75 when normalizing
+the data (Sprules and Barth 2015 pp. Edwards2018, Pomeranz2022).
 Throughout the manuscript, the normalized equal logarithmic binning
-method will be referred to as ELBn. The second method was similar to
-ELBn, but bins of Log\textsubscript{2} widths are used. The count in
-each bin is normalized in the same way. This method has been used by
-(Pomeranz et al. 2019, McGarvey et al. 2019) and is referred to as the
-Normalized Abundance Spectrum (NAS). Previous work has investigated the
-biases in \(\lambda\) estimates when using different methods (Edwards et
-al.~(2017)(2017), Edwards et al.~2020(2020), White et al.~2008(2007)).
-The focus of the present work is to investigate biases when estimating
-the relationship or relative change in parameters across an
-environmental gradient. Hence, we do not discuss the \(\lambda\)
-estimates in the main text, but the results are presented in the
-Supplemental Information (\textbf{NOTE} I'm thinking about dropping this
-from SI. Might be too much info, and it's already been published on by
-Edwards. Does this add anything to the story??).
-
-\hypertarget{estimation-of-relationship-across-gradient-beta_1}{%
+method will be referred to as ELBn.
+
+\hypertarget{log2-bins-nas}{%
+\subsubsection{\texorpdfstring{Log\textsubscript{2} Bins:
+NAS}{Log2 Bins: NAS}}\label{log2-bins-nas}}
+
+The second method was similar to ELBn, but bins of Log\textsubscript{2}
+widths are used. The number of bins is dependent on the size ranges
+present in the data. When working with empirical data with different
+size ranges this can alter the number of bins per site. However, since
+the data here were simulated from a known size range, the number of bins
+for each site is identical. The count in each bin is normalized in the
+same way. This method has been used {[}Jennings and Blanchard (2004),
+Jennings et al. (2002), Sprules and Barth (2015), Pomeranz et al.
+(2019a); McGarvey and Kirk (2018)) and is referred to as the Normalized
+Abundance Spectrum (NAS).
+
+For the two binning methods, simple OLS regression were conducted in the
+form \(Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)\). Where
+\(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).
+
+\hypertarget{variation-in-lambda-across-a-gradient}{%
+\subsection{\texorpdfstring{Variation in \(\lambda\) Across a
+Gradient}{Variation in \textbackslash lambda Across a Gradient}}\label{variation-in-lambda-across-a-gradient}}
+
+\hypertarget{hypothetical-relationshiop-of--0.5}{%
+\subsubsection{Hypothetical relationshiop of
+-0.5}\label{hypothetical-relationshiop-of--0.5}}
+
+Previous work has investigated the biases in \(\lambda\) estimates when
+using different methods White et al. (2007). However, the focus of the
+present work is to investigate biases when estimating the relationship
+parameters across an environmental gradient.
+
+To do this, we again sampled body sizes from distributions described by
+five different \(\lambda\)'s. However, because the estimates from the
+binning methods varied depending on the value of \(\lambda\) (see
+results), we chose three different ``windows'' of lambda values. The
+three windows were steep
+(\(\lambda = [-1.50,~ -1.75,~ -2.00,~ -2.25, -2.50]\)), medium
+(\(\lambda = [-1.00,~ -1.25,~ -1.50,~ -1.75,~ -2.00]\)), and shallow
+((\(\lambda = [-0.50,~ -0.75,~ -1.00,~ -1.25,~ -1.50]\)).
+
+\textbf{NOTE} This MS seems to be getting long. Possibly present
+``medium'' window in main results, and include other windows in SI
+
+Each sample of body sizes from different \(\lambda\) values was assumed
+to come from a community across a hypothetical environmental gradient
+\(X\). Values of \(X\) for each community are uniformly distributed from
+-1 to 1 (site-specific x-values: \((-1.0, -0.5, 0, 0.5, 1.0)\). For each
+\(\lambda\) ``window,'' the relative change in \(\lambda\) values
+(\(\beta_{env}\)) across the gradient is equal to -0.5, such that our
+model is:
+
+\textbf{JEFF} Is this correct? for some reason this seems wrong to me.
+
+\[\beta_{env} = Norm(\alpha -0.5X,~\sigma)\]
+
+We repeated the data simulation 1000 times (reps). The main results
+presented here were not dependent on the range of \emph{x}-values or the
+number of sites (supplemental information).
+
+\hypertarget{varying-hypothetical-relationship}{%
+\subsubsection{Varying Hypothetical
+Relationship}\label{varying-hypothetical-relationship}}
+
+Finally, we also varied the relationship (\(\beta_{env}\)) across the
+hypothetical gradient. We tested \(\beta_{env}\) values of 0, -0.25, and
+-0.5 by sampling 5 body sizes communities uniformly distributed across
+the following sets of \(\lambda\) values: {[}-1.5{]}, {[}-1.25,
+-1.75{]}, {[}-1, -2{]}.
+
+\hypertarget{estimation-of-relationship-across-gradient-beta_env}{%
 \subsection{\texorpdfstring{Estimation of relationship across gradient,
-\(\beta_1\)}{Estimation of relationship across gradient, \textbackslash beta\_1}}\label{estimation-of-relationship-across-gradient-beta_1}}
-
-For each simulation replicate, we estimated how the size spectra
-parameters varied across the hypothetical gradient. Simple OLS
-regression were conducted in the form
-\(\lambda_{estimate} ~ \beta_0 + \beta_1 * X\). The distribution of the
-relationship coefficient, \(\beta_1\), were plotted compared to the
-known relationship.
+\(\beta_{env}\)}{Estimation of relationship across gradient, \textbackslash beta\_\{env\}}}\label{estimation-of-relationship-across-gradient-beta_env}}
+
+For each simulation replicate and size spectra method, we estimated how
+the size spectra parameters varied across the hypothetical gradient. We
+did this by performing simple OLS regression of the
+\(\lambda_{estimate}\) across the hypothetical gradient (\(X\)):
+
+\[\lambda_{estimate} = \beta_0 + \beta_{env} * X\]
+
+The distribution of the relationship coefficient, \(\beta_{env}\), were
+plotted compared to the known relationship.
 
@@
 across an acid mine drainage (AMD) stress gradient. Details of the
-sample collection and processing can be found in (Pomeranz et al. 2019).
-Briefly, all individuals from each sample were identified to the lowest
-practical taxonomic unit and body lengths were measured using image
-processing software from photos taken with a camera mounted to a
+sample collection and processing can be found in (Pomeranz et al.
+2019a). Briefly, all individuals from each sample were identified to the
+lowest practical taxonomic unit and body lengths were measured using
+image processing software from photos taken with a camera mounted to a
 dissecting microscope. Body mass was estimated using taxon-specific
@@
 
+\hypertarget{data-availabity}{%
+\subsection{Data Availabity}\label{data-availabity}}
+
+The data used in this manuscript is already publicly available (Pomeranz
+et al. 2019b) data dryad DOI
+\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s},
+(National Ecological Observatory Network (NEON) 2022). R Scripts to
+reproduce the full simulation and analysis are available at:
+({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).
+
 \hypertarget{results}{%
 \section{Results}\label{results}}
+
+\hypertarget{lambda-estimates}{%
+\section{\texorpdfstring{\(\lambda\)
+estimates}{\textbackslash lambda estimates}}\label{lambda-estimates}}
+
+There was considerable variation in the \(\lambda\) estimate across
+methodologies (Figure 1). The distribution of estimates from the MLE
+method was almost always symmetrical and centered at the known value of
+\(\lambda\) (Figure 1, blue distribution and dashed line). The
+distribution of estimates from the binning methods were generally wider,
+and sometimes showed assymetrical distributions.
+
+\begin{figure}
+\centering
+\includegraphics{figures/est_lambda_est_b_density.png}
+\caption{Figure 1. Distribution of Lambda estimates by method (color)
+from random samples of body sizes from bounded power law distributions
+with varying exponents (-2.5, -0.5). The figure is facetted by the known
+lambda parameter (facet title) and is also shown as the dashed line in
+each facet. Note that the x-axis varies in each facet.}
+\end{figure}
+
+Furthermore, the binning methods systematically over estimated steep
+size spectra relationships (\(\lambda = ~\sim-2.5~to-1.5\), Figure 2),
+and slightly underestimated shallow size spectra relationships
+(\(\lambda = ~\sim-0.5\)). This finding was more pronounced in the NAS
+method compared with the ELBn method.
+
+\begin{figure}
+\centering
+\includegraphics{figures/known_est_b_line.png}
+\caption{Figure 2. Plot of the known and estimated lambda from different
+methods (color). The dashed line is the 1:1 line, and almost perfectly
+matches the MLE (Blue) estimate. The two binning method cross the 1:1
+line, meaning they systematically underestimate shallow lambdas, and
+over estimate steeper lambdas.}
+\end{figure}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_sim_main.png}
-\caption{Relationship estimates across the hypothetical gradient for
-each replicate. Each panel is a different method for estimating the size
-spectra parameter.}
-\end{figure}
-
-Alternate plot version
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_main_v2.png}
-\caption{alternate plot version, thoughts on which you prefer??}
-\end{figure}
-
-All methods performed reasonably well in detecting the known
-relationship across the hypothetical environmental gradient (mean
-\(\beta_1\) estimates ranged from -0.419 to -0.498). However, the SD for
-the NAS and ELBn methods were 2-3x larger that that of the MLE method
-(Figure \textsubscript{XX}). This result was consistent across
-simulation parameters, including sample size, number of sites, and
-\(\lambda\) values.
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_sim_relationship_density.png}
-\caption{Halfeye plots for the estimated relationship coefficient
-estimates across 1000 replicates and three different methods. The point
-below each distribution plot is the median estimate, the thick and thin
-bars represent the 66th and 95th percentile, respectively. Vertical line
-is the known relationship. The median estimate for the MLE method was
-closest to the known relationship, and had the narrowest distribution
-across replicates.}
-\end{figure}
-
-\newpage
-
-\begin{verbatim}
-##   Model   p50   p25  p975
-## 1   MLE -0.49 -0.50 -0.48
-## 2  ELBn -0.49 -0.52 -0.29
-## 3   NAS -0.44 -0.46 -0.30
-\end{verbatim}
-
-\begin{figure}
-\centering
-\includegraphics{figures/n_vary_main.png}
-\caption{Individual regression estimates across the hypothetical
-gradient based on sample size (rows) and methodology used (columns).}
-\end{figure}
-
-\newpage
-
-\hypertarget{no-relationship-1}{%
-\subsection{No relationship}\label{no-relationship-1}}
-
-All of the methods performed similarly when there was no relationship
-across the hypothetical gradient. The mean \(\beta_1\) estimates for all
-three methods ranged from -0.0005 to 0.0002. However, the distribution
-of coefficient estimates was much wider for the two binning methods; the
-95th percentile of the MLE distributions is narrower thatn the 66th
-percentiles for the NAS and ELBn methods.
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_static_b_relationship_density.png}
-\caption{Individual regression estimates when no relationship exists
-across the hypothetical environmental gradient. The methods are
-separated by rows. The median for all methocs is approximately zero, but
-the distribution for the MLE method is narrower than the two binning
-methods.}
-\end{figure}
-
-\begin{verbatim}
-##   Model p50   p25 p975
-## 1   MLE   0 -0.01 0.04
-## 2   NAS   0 -0.04 0.12
-## 3  ELBn   0 -0.05 0.13
-\end{verbatim}
-
-\hypertarget{small-variation-in-lambda}{%
-\subsection{Small variation in lambda}\label{small-variation-in-lambda}}
-
-The performance of all methods to detect small variation in the
-\(\lambda\) parameter (between -1.9 to -2.1) across the environmental
-gradient was also investigated. The median \(\beta_1\) coefficient
-estimates for the MLE method was -0.1, which exactly matched the value
-for the hypothetical relationship, while the other two methods both had
-median estimates of 0.9. The upper 97.5 percentile of the distribution
-was below 0 for the MLE method, meaning that most of the estimates had
-the correct sign. In contrast, the upper 97.5 percentile for both the
-NAS and ELBn methods were greater than 0 at 0.027 and 0.052,
-respectively. (Figure 4).
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_sim_small_relationship_density.png}
-\caption{Distribution of the estimated relationship for each method.
-Here, \(\lambda\) varied from -1.9 to -2.1, with a known relationaship
-of -0.1 across the gradient indicated by the dashed vertical line. The
-distribution of relationship estimates was approximately centered at the
-known relationship for all methods, with a narrower distribution
-observed for the MLE estimates.}
-\end{figure}
-
-\begin{verbatim}
-##   Model   p50   p25  p975
-## 1   MLE -0.10 -0.11 -0.06
-## 2   NAS -0.09 -0.13  0.03
-## 3  ELBn -0.09 -0.13  0.05
-\end{verbatim}
+\hypertarget{lambda-windows}{%
+\subsubsection{\texorpdfstring{\(\lambda\)
+``windows''}{\textbackslash lambda ``windows''}}\label{lambda-windows}}
+
+Because of the different performance of the two binning methods at steep
+and shallow values of lambdas, we first investigated how well the
+methods capture a known relationship of \(\beta_{env} = -0.5\) across a
+hypothetical environmental gradient. The lambda ``windows'' are steep
+(-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).
+
+\begin{figure}
+\centering
+\includegraphics{figures/lambda_angle_plot.png}
+\caption{Figure 3. Distribution of relationship estimates
+(\(beta_{env}\)) in three different ``windows'' of lambda values. The
+dashed vertical line is the known relationship value of -0.5.}
+\end{figure}
+
+The distribution of relationship estimates from the MLE method (Figure
+3, blue) is approximately centered at the known relationship value in
+each of the ``windows.'' The distribution for the MLE estimate is widest
+in the ``steep'' window, but is still markedly narrower than the binning
+methods. In addition to having much wider distributions, the binning
+methods systematically overestimate the known relationship.
+
+\hypertarget{varying-the-known-relationship}{%
+\subsubsection{Varying the Known
+Relationship}\label{varying-the-known-relationship}}
+
+Finally, we were interested in finding if the three methods performed
+differently with varying values of the known relationship. Here, we
+present estimates from each method when the known relationship
+(\(\beta_{env}\)) across the hypothetical gradient was -0.5, -0.25, and
+0, respectively (Figure 4).
+
+\begin{figure}
+\centering
+\includegraphics{figures/vary_beta_plot.png}
+\caption{Figure 4. Individual regressions for each rep (N = 1000) and
+method (columns) for three different known relationship values (rows):
+0, 0.25, and 0.5}
+\end{figure}
+
+Likewise, we also present the distribution of the \(\beta_{env}\)
+relationship estimates (Figure 5). All the methods perform similarly and
+reasonably well when no relationship is present. The performance of the
+binning methods declines (center of distributions), with relationships
+being systematically over estimated. Likewise, the width of the
+distributions increases with stronger relationships across a
+hypothetical gradient. However, the distribution of estimates from the
+MLE method is nearly always centered at the known value, and has
+relatively narrow distributions.
+
+\begin{figure}
+\centering
+\includegraphics{figures/vary_beta_density_plot.png}
+\caption{Figure 5. Distribution of relationship estimates (beta\_1) when
+estimating from different known relationships}
+\end{figure}
 
@@
 parameters consistently increase (become flatter) in the AMD data (Fig.
-XX A). Likewise, the size spectra parameters consistently increase
-(become steeper) with increasing temperature across the NEON sites (Fig
-XX D)
-
-\textbf{NOTE} I need to re-plot these in ggplot and and stitich
-together. Currently stitching pre-made png files, I think that's why it
-looks so funky.
+6A). However, the magnitude of the change across the AMD gradient is
+weakest using the MLE method, and the width of the estimates (\(\pm\) 1
+SD) is larger in the binning methods, particularly the ELBn method (Fig.
+6B). Likewise, the size spectra parameters consistently increase (become
+steeper) with increasing temperature across the NEON sites (Fig. 6C).
+The magnitude of the relationships is weakest in the MLE method,
+although the uncertainty around the estimates are similar (Fig. 6D).
 
@@
 \includegraphics{figures/empirical_combined.png}
-\caption{Figure of empirical data estimates. All of the methods estimate
-the same sign of the relationship, but the estimates from the binning
-methods are generally greater than the MLE estimates.}
+\caption{Figure 6. of empirical data estimates. All of the methods
+estimate the same sign of the relationship, but the estimates from the
+binning methods are generally greater than the MLE estimates.
+\textbf{NOTE} I need to re-plot these in ggplot and and stitch together.
+Currently stitching pre-made png files, I think that's why it looks so
+funky.}
 \end{figure}
@@
 data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
-change in \(\lambda\) estimates ranges from 0.06 to 0.17, depending on
-the method used.
-
-\hypertarget{simulating-shallower-lambdas}{%
-\subsection{Simulating shallower
-lambdas}\label{simulating-shallower-lambdas}}
-
-Estimates of the relationship between lambda and a hypothetical gradient
-varied depending on the method used. However, interpretations of
-empirical data were broadly consistent across methodologies. Upon closer
-inspection, the values of size spectra parameters in the empirical data
-were considerably shallower than -2. Given that we found the performance
-of all methods increased with shallower \(\lambda\)'s we wanted to
-investigate how the methods performed with simulated values closer to
-the empirical estimates of parameters describing size spectra
-relationships. Therefore, we repeated the simulation process as in the
-main analysis, but used \(\lambda\) values ranging from -1.1 to -1.5,
-with a known relationship of \(\beta_1 = -0.2\)
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/shallow_lambda_main.png}
-\caption{Relationship estimates across the hypothetical gradient for
-each replicate. Each panel is a different method for estimating the size
-spectra parameter.}
-\end{figure}
-
-We found generally less variability in the relationship estimates
-(\(\beta_1\)) across a gradient of distributions described with
-shallower \(\lambda\) parameters. However, the median MLE estimate was
-once again closest to the hypothetical relationship and had the
-narrowest distribution. The median ELBn estimate was close to the
-hypothetical relationship but had a slightly wider distribution.
-Finally, 95\% of the NAS estimates were greater than the hypothetical
-relationship. In general, the distributions from the different
-methodologies followed the same overall pattern, with the MLE estimates
-being the smallest, and the binning methods being larger (i.e., more
-positive).
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_shallow_lambda_relationship_density.png}
-\caption{Distribution of relationship coefficient estimates. Vertical
-line is the known relationship. All methods under estimate the value,
-but the mean magnitude and distribution of values is greater for the
-ELBn and NAS methods.}
-\end{figure}
-
-\begin{verbatim}
-##   Model   p50   p25  p975
-## 1   MLE -0.20 -0.21 -0.18
-## 2  ELBn -0.19 -0.21 -0.16
-## 3   NAS -0.17 -0.18 -0.15
-\end{verbatim}
-
-\newpage
+change in \(\lambda\) estimates varies by 0.11 units depending on the
+methods used (range of absolute change across gradient: 0.06 to 0.17).
 
@@
 
-Measuring parameters describing the decline in abundance with increasing
-body size in communities is being done with increasing frequency across
-ecology. Previous work has investigated the accuracy and inherent biases
-associated with different estimation methods. However, how these
-inaccuracies and biases compound across environmental gradients remains
-uncertain, making it difficult to detect variation in size spectra
-parameters across environmental gradients with confidence. Here, we
-sampled body sizes from known distributions with varying parameters
-(\(\lambda\)) and estimated the coefficient of the relationship
-(\(\beta_1\)) across a hypothetical gradient in order to assess how the
-results would vary depnding on the method used. Likewise, we compared
-how the interpretation of previously published results could change
-depending on the methodology used to estimate the relationship.
-
-Generally, the estimate from the MLE method was always closer to the
-known relationship in simulated data, and always had the smallest
-variation around the estimate.
+The relationship between body size and abundance has been extensively
+studied in a wide range of taxa inhabiting both terrestrial and aquatic
+ecosystems (reviewed by White et al. (2007)). Empirical patterns are
+remarkably consistent, and can be explained by the metabolic theory of
+Ecology (Brown et al. 2004). Measuring parameters describing the decline
+in abundance with increasing body size in communities is being done with
+increasing frequency across ecology. Previous work has investigated the
+accuracy and inherent biases associated with different estimation
+methods. However, how these inaccuracies and biases compound across
+environmental gradients remains uncertain, making it difficult to detect
+variation in size spectra parameters across environmental gradients with
+confidence. Here, we sampled body sizes from known distributions with
+varying parameters (\(\lambda\)) and estimated the coefficient of the
+relationship (\(\beta_{env}\)) across a hypothetical gradient in order
+to assess how the results would vary depending on the method used.
+Likewise, we compared how the interpretation of previously published
+results could change depending on the methodology used to estimate the
+relationship.
+
+Generally, the estimate of the size spectra parameter from the MLE
+method was always closer to the known \(\lambda\) value, regardless of
+how steep or shallow the \(\lambda\) exponent was. This is in agreement
+with Edwards et at. (2017) which found that the MLE estimate more
+consistently recovered known \(\lambda\) values ranging from -2.5 to
+-0.5. We further this work by discovering that the magnitude of the
+error associated with the two binning methods increases as the
+\(\lambda\) value moves away from \(\sim\) -1.
+
+It is unsurprising that all methods performed less well in estimating
+steep values of \(\lambda\) (i.e., -2.5). The rapid decline of large
+body sizes makes it unlikely to sample any individuals with a large body
+size. Furthermore, the binning methods inherently ``lose'' data by
+aggregating individual observations into a single bin.
+
+Likewise, the MLE method also resulted in the most accurate
+\(\beta_{env}\) estimates for the hypothetical relationship in simulated
+data regardless of the \(\lambda\) ``window'' investigated. Furthermore,
+the variation around the estimate was always smallest with the MLE
+method. However, the variation in the MLE estimate did increase with
+steeper \(\lambda\) ``windows.'' This is not surprising, however because
+steep relationships like this make it very unlikely to sample the
+largest body sizes, and capturing the ``true'' relationships without the
+full range of body sizes present is more difficult (Edwards et
+al.~2017). Despite this, the MLE method is nearly always the best option
+to use, likely because all the data are used in the analysis.
 
@@
 variation within the data by collapsing body size variation into a
-single values within bins. For example, all individuals placed into a
+single value within each bin. For example, all individuals placed into a
 bin that ranges from 2-4 grams of mass are all treated as having a mass
@@
 is taken for this bin, despite that fact that there is almost certainly
-variation in the abundance of individuals that weigh \textasciitilde2
-grams or \textasciitilde4 grams. By homogenizing the data in this way,
-it seems that binning methods will always produce noisier results than
-MLE from the binning process alone, since binning is akin to deleting
-information that the model could otherwise use.
-
-One of the benefits of using MLE is that all data points are retained
-for use within the model.
-
-The performance of all methods improved as sample size increases, and as
-the exponent of the size spectra relationship gets larger (more
-shallow). As either or both of these variables change, the difference in
-estimated relationship coefficients declines across the methods. This is
-particularly interesting given the fact that both empirical data sets of
-stream communities examined here are shallower than the expected
-relationship of \(\lambda = -2\), and the conclusions of the change in
-size spectra relationships across the gradients are not dependent on
-method used. If communities under study are in fact described by
-shallower exponents, the method used may not be critical in the
-conclusions reached. However, the MLE method performs as well or better
-than the two binning methods examined here under all contexts and should
-be the preferred method in future studies. At a minimum, future studies
-should report MLE estimates to ensure that the results are not dependent
-on the methodology used.
-
-Despite the drop in performance with reduced variation in \(\lambda\)
-values, when a significant relationship was observed, it was generally
-in the correct direction and of a similar magnitude. This suggests that
-previously reported significant changes in size spectra parameters
-across environmental gradients and in experimental manipulations are
-plausible, and the magnitude of the relative change is a reasonable
-estimate. Given that all of the data within a study is treated
-identically, the the over all change in size spectra parameters is
-likely reasonable. However, the biases and inconsistencies in
-relationship estimates presented here suggest that it would be difficult
-if not impossible to compare the relative changes across different
-published studies which use different methods. Publication of individual
-body size data with future studies of size spectra relationships would
-greatly aid in our ability to generalize changes to this fundamental
-aspect of community organization across spatiotemporal scales and in
-response to environmental stressors and perturbations.
+variation in the abundance of individuals that weigh \textasciitilde2,
+\textasciitilde3, or \textasciitilde4 grams. By homogenizing the data in
+this way, it seems that binning methods will always produce noisier
+results than MLE from the binning process alone, since binning is akin
+to deleting information that the model could otherwise use. One of the
+benefits of using MLE is that all data points are retained for use
+within the model.
+
+The variation in estimated relationships across the empirical datasets
+varied by \textasciitilde0.1 units depending on the method used. At
+first glance this may not appear to be a concerning magnitude of
+variation. However, when one considers the range of variation reported
+in many seminal works on variation in size spectra relationships is
+\textasciitilde0.1-0.2 (SI Table \emph{xx}), it becomes clear how
+important an error in estimates of this magnitude is.
+
+Despite the differences in magnitude of the empirical relationship
+coefficients, they were generally in the correct direction and of a
+similar magnitude. This suggests that previously reported significant
+changes in size spectra parameters across environmental gradients and in
+experimental manipulations are plausible. Given that all of the data
+within a study is treated identically, the over all change in size
+spectra parameters is likely reasonable. However, the biases and
+inconsistencies in relationship estimates presented here suggest that it
+would be difficult if not impossible to directly compare the relative
+changes across different published studies which use different methods.
+However, the relatively systematic error of \(\lambda\) estimates across
+known exponents suggests that it may be possible to correct previous
+estimates onto the same scale (Figure 2).
+
+Despite this possibility, publication of individual body size data with
+future studies of size spectra relationships would greatly aid in our
+ability to generalize changes to this fundamental aspect of community
+organization across spatiotemporal scales and in response to
+environmental conditions.
 
@@
 
-Both of the binning methods can perform as well as MLE under certain
-circumstances (large sample size, shallow \(\lambda\)). However, it
-would be extremely difficult to know if this was the case beforehand.
-Furthermore, in order to confirm this result, MLE methods would likely
-need to be employed. With the publication of the sizespectra package
-(2017), producing MLE estimates of size spectra parameters is a
-relatively easy task. Therefore, we recommend using this in all future
-studies of size spectra relationships.
-
-We reiterate the recommendations of White et al.~(2007)(2007), Sprules
-and Barth (2016)(2016) and Edwards et al.~(2017)(2017) to estimate size
-spectra relationships using MLE methods due to their superior
-performance in nearly every context. Furthermore, we strongly encourage
-authors to publish individual size data whenever possible. This will
-allow for the consistent re-analysis of existing data sets as
-methodologies develop and improve. This will aid in the ability for size
-spectra work to be synthesized between research groups and across
-scales.
+The MLE method outperformed both of the binning methods under nearly any
+measure. With the publication of the \emph{sizespectra} package (2017),
+producing MLE estimates of size spectra parameters is a relatively easy
+task. Therefore, we recommend using this in all future studies of size
+spectra relationships.
+
+We reiterate the recommendations of (2017) to estimate size spectra
+relationships using MLE methods due to their superior performance in
+nearly every context. Furthermore, we strongly encourage authors to
+publish individual size data whenever possible. This will allow for the
+consistent re-analysis of existing data sets as methodologies develop
+and improve. This will aid in the ability for size spectra work to be
+synthesized between research groups and across scales.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-edwards2020}{}}%
-Edwards, A. M., J. P. W. Robinson, J. L. Blanchard, J. K. Baum, and M.
-J. Plank. 2020. \href{https://doi.org/10.3354/meps13230}{Accounting for
-the bin structure of data removes bias when fitting size spectra}.
-Marine Ecology Progress Series 636:19--33.
-
-\leavevmode\vadjust pre{\hypertarget{ref-edwards2017}{}}%
-Edwards, A. M., J. P. W. Robinson, M. J. Plank, J. K. Baum, and J. L.
-Blanchard. 2017. \href{https://doi.org/10.1111/2041-210X.12641}{Testing
-and recommending methods for fitting size spectra to data}. Methods in
-Ecology and Evolution 8:57--67.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcgarvey2019}{}}%
-McGarvey, D. J., T. E. Woods, and A. J. Kirk. 2019.
-\href{https://doi.org/10.3791/59945}{Modeling the {Size Spectrum} for
-{Macroinvertebrates} and {Fishes} in {Stream Ecosystems}}. Journal of
-Visualized Experiments: JoVE.
-
-\leavevmode\vadjust pre{\hypertarget{ref-NEON_Inverts2022}{}}%
-National Ecological Observatory Network (NEON). 2022.
-\href{https://doi.org/10.48443/GN8X-K322}{Macroinvertebrate collection
-({DP1}.20120.001)}. {National Ecological Observatory Network (NEON)}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-petchey2010}{}}%
-Petchey, O. L., and A. Belgrano. 2010.
-\href{https://doi.org/10.1098/rsbl.2010.0240}{Body-size distributions
-and size-spectra: Universal indicators of ecological status?} Biology
-Letters 6:434--437.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pomeranz2022}{}}%
-Pomeranz, J. P. F., J. R. Junker, and J. S. Wesner. 2022.
-\href{https://doi.org/10.1111/gcb.15862}{Individual size distributions
-across {North American} streams vary with local temperature}. Global
-Change Biology 28:848--858.
-
-\leavevmode\vadjust pre{\hypertarget{ref-pomeranz2019}{}}%
-Pomeranz, J. P. F., H. J. Warburton, and J. S. Harding. 2019.
-\href{https://doi.org/10.1111/fwb.13196}{Anthropogenic mining alters
-macroinvertebrate size spectra in streams}. Freshwater Biology
-64:81--92.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sprules2016}{}}%
-Sprules, W. G., and L. E. Barth. 2016.
-\href{https://doi.org/10.1139/cjfas-2015-0115}{Surfing the biomass size
-spectrum: Some remarks on history, theory, and application}. Canadian
-Journal of Fisheries and Aquatic Sciences 73:477--495.
-
-\leavevmode\vadjust pre{\hypertarget{ref-white2007}{}}%
+\leavevmode\hypertarget{ref-Brown1995}{}%
+Brown, J. H. 1995. Macroecology. {University of Chicago Press},
+{Chicago, IL}.
+
+\leavevmode\hypertarget{ref-Brown2004}{}%
+Brown, J. H., J. F. Gillooly, A. P. Allen, V. M. Savage, and G. B. West.
+2004. Toward a metabolic theory of ecology. Ecology 85:1771--1789.
+
+\leavevmode\hypertarget{ref-dossenaWarmingAltersCommunity2012}{}%
+Dossena, M., G. Yvon-Durocher, J. Grey, J. M. Montoya, D. M. Perkins, M.
+Trimmer, and G. Woodward. 2012. Warming alters community size structure
+and ecosystem functioning. Proceedings of the Royal Society B
+279:3011--3019.
+
+\leavevmode\hypertarget{ref-edwards2020}{}%
+Edwards, A. M., J. Robinson, J. Blanchard, J. Baum, and M. Plank. 2020.
+Accounting for the bin structure of data removes bias when fitting size
+spectra. Marine Ecology Progress Series 636:19--33.
+
+\leavevmode\hypertarget{ref-Edwards2017}{}%
+Edwards, A. M., J. Robinson, M. Plank, J. Baum, and J. Blanchard. 2017.
+Testing and recommending methods for fitting size spectra to data.
+Methods in Ecology and Evolution 8:57--67.
+
+\leavevmode\hypertarget{ref-jennings2004}{}%
+Jennings, S., and J. L. Blanchard. 2004. Fish abundance with no fishing:
+Predictions based on macroecological theory. Journal of Animal Ecology
+73:632--642.
+
+\leavevmode\hypertarget{ref-Jennings2002}{}%
+Jennings, S., K. J. Warr, and S. Mackinson. 2002. Use of size-based
+production and stable isotope analyses to predict trophic transfer
+efficiencies and predator-prey body mass ratios in food webs. Marine
+Ecology Progress Series 240:11--20.
+
+\leavevmode\hypertarget{ref-martinez2016}{}%
+Martínez, A., A. Larrañaga, A. Miguélez, G. Yvon-Durocher, and J. Pozo.
+2016. Land use change affects macroinvertebrate community size spectrum
+in streams: The case of {Pinus} radiata plantations. Freshwater Biology
+61:69--79.
+
+\leavevmode\hypertarget{ref-mcgarveySeasonalComparisonCommunitylevel2018}{}%
+McGarvey, D. J., and A. J. Kirk. 2018. Seasonal comparison of
+community-level size-spectra in southern coalfield streams of {West
+Virginia} ({USA}). Hydrobiologia 809:65--77.
+
+\leavevmode\hypertarget{ref-NEON_Inverts2022}{}%
+National Ecological Observatory Network (NEON). 2022. Macroinvertebrate
+collection ({DP1}.20120.001). {National Ecological Observatory Network
+(NEON)}.
+
+\leavevmode\hypertarget{ref-Perkins2018}{}%
+Perkins, D. M., I. Durance, F. K. Edwards, J. Grey, A. G. Hildrew, M.
+Jackson, J. I. Jones, R. B. Lauridsen, K. Layer-Dobra, M. S. A.
+Thompson, and G. Woodward. 2018. Bending the rules: Exploitation of
+allochthonous resources by a top-predator modifies size-abundance
+scaling in stream food webs. Ecology Letters.
+
+\leavevmode\hypertarget{ref-Petchey2010}{}%
+Petchey, O. L., and A. Belgrano. 2010. Body-size distributions and
+size-spectra: Universal indicators of ecological status? Biology Letters
+6:434--437.
+
+\leavevmode\hypertarget{ref-Pomeranz2022}{}%
+Pomeranz, J. P. F., J. R. Junker, and J. S. Wesner. 2022. Individual
+size distributions across {North American} streams vary with local
+temperature. Global Change Biology 28:848--858.
+
+\leavevmode\hypertarget{ref-pomeranzAnthropogenicMiningAlters2019}{}%
+Pomeranz, J. P. F., H. J. Warburton, and J. S. Harding. 2019a.
+Anthropogenic mining alters macroinvertebrate size spectra in streams.
+Freshwater Biology 64:81--92.
+
+\leavevmode\hypertarget{ref-Pomeranz2019}{}%
+Pomeranz, J. P. F., H. J. Warburton, and J. S. Harding. 2019b. Data
+from: {Anthropogenic} mining alters macroinvertebrate size spectra in
+streams. {Dryad}.
+
+\leavevmode\hypertarget{ref-sprulesSurfingBiomassSize2015}{}%
+Sprules, and Barth. 2015. Surfing the biomass size spectrum: Some
+remarks on history, theory, and application. Canadian Journal of
+Fisheries and Aquatic Sciences 73:477--495.
+
+\leavevmode\hypertarget{ref-white2008}{}%
+White, E. P., B. J. Enquist, and J. L. Green. 2008. On estimating the
+exponent of power-law frequency distributions. Ecology 89:905--912.
+
+\leavevmode\hypertarget{ref-White2007}{}%
 White, E. P., S. K. M. Ernest, A. J. Kerkhoff, and B. J. Enquist. 2007.
-\href{https://doi.org/10.1016/j.tree.2007.03.007}{Relationships between
-body size and abundance in ecology}. Trends in Ecology \& Evolution
-22:323--330.
+Relationships between body size and abundance in ecology. Trends in
+Ecology \& Evolution 22:323--330.
 
@@
 
-\hypertarget{range-of-body-sizes-m}{%
-\subsection{\texorpdfstring{Range of body sizes,
-\(M\)}{Range of body sizes, M}}\label{range-of-body-sizes-m}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_small_m_main.png}
-\caption{Individual regressions for five sites across a hypothetical
-gradient with a known relationship of 0.5. Range of body sizes is
-reduced and is from 1, to 100.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_small_m_relationship_density.png}
-\caption{Distribution of estimated relationship (\(\beta_1\))
-coefficient's for five sites across a hypothetical gradient with known
-value of 0.5. Range of body sizes is reduced and is from 1, to 100.}
-\end{figure}
-
-\newpage
-
-\hypertarget{large-environmental-gradient}{%
-\subsection{Large environmental
-gradient}\label{large-environmental-gradient}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_large_x_main.png}
-\caption{Individual regressions for five sites across a hypothetical
-gradient with a known relationship of 0.5. Range of environmental values
-(\emph{x}-axis) increased to be -1000, to 1000.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_large_x_est_b_density.png}
-\caption{Distribution of estimated \(\lambda\) coefficient for five
-sites across a hypothetical gradient with known values. Range of
-environmental values (\emph{x}-axis) increased to be -1000, to 1000.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_large_x_relationship_density.png}
-\caption{Distribution of estimated relationship (\(\beta_1\))
-coefficient's for five sites across a hypothetical gradient with known
-value of 0.5. Range of environmental values (\emph{x}-axis) increased to
-be -1000, to 1000.}
-\end{figure}
-
-\newpage
-
-\hypertarget{varying-number-of-sites}{%
-\subsection{Varying number of sites}\label{varying-number-of-sites}}
-
-\hypertarget{sites}{%
-\subsubsection{10 sites}\label{sites}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_10_sites_main.png}
-\caption{Individual regressions for ten sites across a hypothetical
-gradient with a known relationship of 0.5. All other parameters are the
-same as in the main analysis}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_10_sites_est_b_density.png}
-\caption{Distribution of estimated \(\lambda\) coefficient for ten sites
-across a hypothetical gradient with known values. All other parameters
-are the same as in the main analysis}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_10_sites_relationship_density.png}
-\caption{Distribution of estimated relationship (\(\beta_1\))
-coefficient's for ten sites across a hypothetical gradient with known
-value of 0.5. All other parameters are the same as in the main analysis}
-\end{figure}
-
-\newpage
-
-\hypertarget{three-sites}{%
-\subsubsection{Three sites}\label{three-sites}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_3_sites_main.png}
-\caption{Individual regressions for three sites across a hypothetical
-gradient with a known relationship of 0.5. All other parameters are the
-same as in the main analysis.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_3_sites_est_b_density.png}
-\caption{Distribution of estimated \(\lambda\) coefficient for three
-sites across a hypothetical gradient with known values. All other
-parameters are the same as in the main analysis.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_3_sites_relationship_density.png}
-\caption{Distribution of estimated relationship (\(\beta_1\))
-coefficient's for three sites across a hypothetical gradient with known
-value of 0.5. All other parameters are the same as in the main analysis}
-\end{figure}
-
-\newpage
-
-\hypertarget{sample-size}{%
-\subsection{Sample size}\label{sample-size}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/n_vary_est_b.png}
-\caption{Distribution of size spectra parameter estimates. Vertical line
-is the known parameter (dashed line) wich describes the bounded power
-law distribution from which the body size estimates were sampled. As n
-increases (top to bottom) and \(\lambda\) increases (left to right), the
-accuracy of the estimate improves across all methods.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/n_vary_main.png}
-\caption{Individual regressions when varying sample size}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/n_vary_relationship_density.png}
-\caption{Distribtuion of the relationship coefficients with varying
-sample size}
-\end{figure}
-
-\newpage
-
 \hypertarget{lambda-and-relationship-estimates}{%
-\subsection{\texorpdfstring{\(\lambda\) and relationship
+\section{\texorpdfstring{\(\lambda\) and relationship
 estimates}{\textbackslash lambda and relationship estimates}}\label{lambda-and-relationship-estimates}}
-
-\textbf{NOTE} I'm considering dropping the \(\lambda\) estimate plots
-from the manuscript. I think it's too much info, but if it adds to the
-story I'm happy to leave it in. \textbf{END}
 
@@
 
-\newpage
-
-\hypertarget{number-of-sites-3}{%
-\subsubsection{Number of sites = 3}\label{number-of-sites-3}}
+\hypertarget{varying-number-of-sites}{%
+\subsection{Varying number of sites}\label{varying-number-of-sites}}
+
+\hypertarget{sites}{%
+\subsubsection{10 sites}\label{sites}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_10_sites_main.png}
+\caption{Individual regressions for ten sites across a hypothetical
+gradient with a known relationship of 0.5. All other parameters are the
+same as in the main analysis}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_10_sites_est_b_density.png}
+\caption{Distribution of estimated \(\lambda\) coefficient for ten sites
+across a hypothetical gradient with known values. All other parameters
+are the same as in the main analysis}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_10_sites_relationship_density.png}
+\caption{Distribution of estimated relationship (\(\beta_1\))
+coefficient's for ten sites across a hypothetical gradient with known
+value of 0.5. All other parameters are the same as in the main analysis}
+\end{figure}
+
+\newpage
+
+\hypertarget{three-sites}{%
+\subsubsection{Three sites}\label{three-sites}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_3_sites_main.png}
+\caption{Individual regressions for three sites across a hypothetical
+gradient with a known relationship of 0.5. All other parameters are the
+same as in the main analysis.}
+\end{figure}
+
+\newpage
 
@@
 \includegraphics{figures/PLB_3_sites_est_b_density.png}
-\caption{Distribution of estimated \(\lambda\) coefficient across a
-hypothetical gradient with known values (dashed line). The number of
-sites is reduced to 3, compared with 5 sites presented in the main
-anaysis.}
+\caption{Distribution of estimated \(\lambda\) coefficient for three
+sites across a hypothetical gradient with known values. All other
+parameters are the same as in the main analysis.}
 \end{figure}
@@
 \includegraphics{figures/PLB_3_sites_relationship_density.png}
-\caption{Relationship estimate}
-\end{figure}
-
-\newpage
-
-\hypertarget{number-of-sites-10}{%
-\subsubsection{Number of sites = 10}\label{number-of-sites-10}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_10_sites_est_b_density.png}
-\caption{Distribution of estimated \(\lambda\) coefficient across a
-hypothetical gradient with known values(dashed line). The number of
-sites is increased to 10, compared with 5 sites presented in the main
-anaysis.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_10_sites_relationship_density.png}
-\caption{Relationship estimate}
-\end{figure}
-
-\newpage
-
-\hypertarget{environmental-gradient}{%
-\subsubsection{Environmental Gradient}\label{environmental-gradient}}
+\caption{Distribution of estimated relationship (\(\beta_1\))
+coefficient's for three sites across a hypothetical gradient with known
+value of 0.5. All other parameters are the same as in the main analysis}
+\end{figure}
+
+\newpage
+
+\hypertarget{large-environmental-gradient}{%
+\subsection{Large environmental
+gradient}\label{large-environmental-gradient}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_large_x_main.png}
+\caption{Individual regressions for five sites across a hypothetical
+gradient with a known relationship of 0.5. Range of environmental values
+(\emph{x}-axis) increased to be -1000, to 1000.}
+\end{figure}
+
+\newpage
 
@@
 \caption{Distribution of estimated \(\lambda\) coefficient for five
-sites across a hypothetical gradient with known values (dashed line).
-The range of the environmental gradient was increased (-100 to 100).}
+sites across a hypothetical gradient with known values. Range of
+environmental values (\emph{x}-axis) increased to be -1000, to 1000.}
 \end{figure}
@@
 \includegraphics{figures/PLB_large_x_relationship_density.png}
-\caption{Relationship estimate}
+\caption{Distribution of estimated relationship (\(\beta_1\))
+coefficient's for five sites across a hypothetical gradient with known
+value of 0.5. Range of environmental values (\emph{x}-axis) increased to
+be -1000, to 1000.}
+\end{figure}
+
+\newpage
+
+\hypertarget{range-of-body-sizes-m}{%
+\subsection{\texorpdfstring{Range of body sizes,
+\(M\)}{Range of body sizes, M}}\label{range-of-body-sizes-m}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_small_m_main.png}
+\caption{Individual regressions for five sites across a hypothetical
+gradient with a known relationship of 0.5. Range of body sizes is
+reduced and is from 1, to 100.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/PLB_small_m_relationship_density.png}
+\caption{Distribution of estimated relationship (\(\beta_1\))
+coefficient's for five sites across a hypothetical gradient with known
+value of 0.5. Range of body sizes is reduced and is from 1, to 100.}
+\end{figure}
+
+\newpage
+
+\hypertarget{sample-size-n}{%
+\subsection{\texorpdfstring{Sample size,
+\(n\)}{Sample size, n}}\label{sample-size-n}}
+
+The number of observations in our simulations may bias the results.
+Therefore, we repeated the simulations described above, but varied the
+sample size \(n\). We tested values of
+\(n = 200, 500, 1000, 5000, 10 000\). Results of this analysis are
+presented in the Supplemental Information.
+
+\begin{figure}
+\centering
+\includegraphics{figures/n_vary_main.png}
+\caption{Individual regression estimates across the hypothetical
+gradient based on sample size (rows) and methodology used (columns).
+(match this figure to ``new'' style if we like that better)}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics{figures/n_vary_est_b.png}
+\caption{Distribution of size spectra parameter estimates. Vertical line
+is the known parameter (dashed line) wich describes the bounded power
+law distribution from which the body size estimates were sampled. As n
+increases (top to bottom) and \(\lambda\) increases (left to right), the
+accuracy of the estimate improves across all methods.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{figures/n_vary_relationship_density.png}
+\caption{Distribtuion of the relationship coefficients with varying
+sample size}
 \end{figure}
@@
 
+\hypertarget{comparison-with-other-published-estimates}{%
+\section{Comparison with other published
+estimates}\label{comparison-with-other-published-estimates}}
+
+SI Table. This table shows published estimates of the variation in size
+spectra slopes (or exponents) in empirical studies. It is unclear how to
+directly compare estimates of the slope with different methods.However,
+the published estimates here range from \textasciitilde0.1 to 0.2 across
+the gradients studied. For comparison, the 2.5-95\% quantiles around the
+relationship estimate for the MLE method were \textasciitilde0.1,
+whereas for the ELBn and NAS method they were \textasciitilde0.25 and
+\textasciitilde0.2, respectively. b\_diff is the change in estimate
+(b-low - b-min). System refers to stream communities or mesocosm
+experiments. Method: MLE = maximum likelihod estimate, ELB = equal
+logarthmic binning, the number before indicates the number of bins used.
+The normalization process shifts the estimates by an absolute value of
+1.0. Hence, direct comparison the relative change in normalized and
+non-normalized studies should not introduce any bias. The O'Gorman et
+al.~2017 study used average species size and abundance (Local Size
+Density Relationship, \emph{sensu} White et al.~2007) as opposed to
+individual size distribution. These methods are related, but it is
+unclear how to directly compare estimates from each method.
+
+\begin{longtable}[]{@{}lrrlrrlll@{}}
+\toprule
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Citation\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+\(\beta_1\) Difference\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+Error\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Error Type\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+\(\beta_{low}\)\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+\(\beta_{high}\)\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+System\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Driver\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+Method\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Pomeranz et al.~2021\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.12\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-1.37\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-1.25\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+Streams\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Temperature\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+MLE\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Yvon Durocher et al.~2011 (Community)\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.13\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.79\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.92\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+FW mesocosms\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Temperature\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+10 ELB, not normalized\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Dossena et al.~2012 (October)\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.14\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+0.070\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+se\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+FW mesocosms\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Temperature\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+6 ELB, not normalized\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+O'Gorman et al.~2017\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.15\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.85\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.70\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+Streams\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Temperature\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+ln (mean species abundance) \textasciitilde{} ln(mean species
+mass)\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Martinez et al.~2016\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.15\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-1.11\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.96\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+Streams\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Land Use\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+6 ELB, not normalized\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Yvon Durocher et al.~2011 (Phytoplankton)\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.19\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.46\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-0.65\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+FW mesocosms\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Temperature\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+10 ELB, not normalized\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+McGarvey and Kirk 2018\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.19\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-1.81\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+-1.62\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+Streams\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Seasonality\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+NAS, Log2 bins\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
+Dossena et al.~2012 (April)\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
+0.21\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
+0.064\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+se\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
+NA\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
+FW mesocosms\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
+Temperature\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
+6 ELB, not normalized\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
 \end{document}

### 2023-02-23 Manuscript.Rmd
pre wesner pull commit

@@
 
-1.  The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra relationships.
+1.
+The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra, which are described by a power law.
+The focus of size spectra analysis is to estimate the exponent ($\lambda$) of the power law.
 2.
-Many methods have been proposed for constructing size spectra relationships, but recent work has shown that varied methods return biased estimates of relationship parameters, and different methods in fact are not estimating the same parameter.
+Many methods have been proposed for estimating $\lambda$ most of which involve binning the data, summing abundance within bins, and then fitting a ordinary least squares (OLS) regression in log-log space.
+However, recent work has shown that binning procedures may return biased estimates of size spectra exponents compared to procedures that directly estimate $\lambda$ using maximum likelihood estimation (MLE).
 Despite this variability in estimates, it is unclear if the relative change across environmental gradients is consistent across methodologies.
-Here, we simulate data sets across an hypothetical environmental gradient and estimate the size spectra parameter (slope or exponent, depending on method) at each site, and, importantly, estimate the relationship of parameters across the hypothetical gradient.
-We also use two previously published body size datasets across an anthropogenic stress gradient and an environmental temperature gradient and assess how the conclusions of these studies would vary based on methods used.
+Here, we used simulation to compare the ability of two binning methods (NAS, ELBn) and MLE to 1) recapture known values of $\lambda$, and 2) recapture parameters in a linear regression measuring the change in $\lambda$ across a hypothetical environmental gradient.
+We also compared the methods using two previously published body size datasets across a pollution gradient and a temperature gradient
 3.
-We find that maximum likelihood methods always perform better than common binning methods.
+Maximum likelihood methods always performed better than common binning methods, which demonstrated consistent bias depending on the simulated values of $\lambda$.
+This bias carried over to the regressions, which were more accurate when $\lambda$ was estimated using MLE compared to the binning procedures.
 Additionally, the variance in estimates using MLE methods is markedly reduced when compared to binning methods.
@@
 The uncertainty and variation in estimates when using binning methods is often greater than or equal to the variation previously published in experimental and observational studies, bringing into question the effect size of previously published results.
-However, re-analysis of two previously published datasets does not markedly alter the conclusions reached.
-Further study is needed to identify how and when these estimates may be biased, and when the general pattern of results can be considered to be "true".
+However, while the methods produced different slope estimates from previously published datasets, they were in qualitative agreement on the sign of those slopes (i.e., all negative or all positive).
+Our results provide further support for the direct estimation of $\lambda$ using MLE (or similar procedures) over the more common methods of binning.
 
@@
 Body size distributions are a fundamental characteristic of communities.
-In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities (Damuth 1981, 1991, 1998).
-This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology (Nee 1991, Brown et al.
-2004).
+In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981, @Damuth1991, @Damuth1998].
+This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991, @Brown2004].
 The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010].
-Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], to human activities [@jennings2004].
+Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], and in response to human activities [@jennings2004].
 Likewise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized.
@@
 Individual size distributions (ISD *sensu* [@White2007], also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
-ISDs show the relationship between the abundances of individual body size regardless of taxonomy.
+ISDs represent frequency distributions of individual body sizes within a community, regardless of taxonomy.
 Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
 Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972, @Andersen2006a].
-Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions of log transformed data of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
+Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions in log-log space of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
 Myriad binning methods have been proposed, including linear and logarithmic bin widths.
-Likewise, methods can rely on the absolute count in the bins as well as normalization techniques where the count is divided by the bin width (especially common with logarithmic binning).
-Alternatively, $\lambda$ can be estimated directly using maximum likelihood techniques.
+Likewise, some methods rely on the absolute counts in the bins and others employ normalization techniques such as dividing the count by the bin width (especially common with logarithmic binning).
+As an alternative to binning methods, $\lambda$ can be estimated directly on un-binned data using maximum likelihood techniques.
 
-Previous work has shown that the estimates of $\lambda$ differ between MLE and OLS techniques.
-OLS methods are particularly sensitive to decisions made in the binning process.
+Previous work has shown that the estimates of $\lambda$ differ between MLE and size-binned OLS techniques.
+Size-binned OLS methods are particularly sensitive to decisions made in the binning process.
 Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ (White et al.
@@
 While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time.
-In other words, if the data within a study are all treated the same, does a relative change of OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?
+In other words, if the data within a study are all treated the same, does a relative change of size-binned OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?
 
-We had three primary objectives in this study: 1) how well do different methods estimate site-specific $\lambda$'s?
-2) Do the estimates of the hypothetical relationship obtained from different methods agree?
-and 3) Are the conclusions reached from previously published data sets dependent on the method used?
-
-In order to answer the first two questions, we simulate body size observations from bounded power law distributions with varied $\lambda$ exponents across a hypothetical environmental gradient and compare the results obtained using three different methodologies common in the literature.
-To answer the third question, we re-analyze two previously published data sets of community body sizes across a stress and environmental gradient, respectively [@pomeranz2019, @pomeranz2022].
-We find that the MLE method more accurately estimates the site-specific $\lambda$ exponents as well as the relative change in exponents across the hypothetical gradient, as well as having smaller variation around the estimates.
-The logarithmic binning methods generally perform well, but have larger variation around the site-specific estimates.
-In addition, the variation of the estimates of the hypothetical relationship obtained from the binning methods is dependent on the value of $\lambda$ used (under and over estimates steep and shallow $\lambda$, respectively).
+We had Two primary objectives in this study: 1) to compare how well different methods estimate site-specific $\lambda$'s 2) To compare how different methods of estimating $\lambda$ impact subsequent regressions that use $\lambda$'s as a response variable.
+We did this using both simulated and empirical data sets.
 
@@
 
-In order to investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@edwards2017].
+To investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@edwards2017].
 Let $M$ be a random variable of body sizes described by the probability density function:
@@
 $M_{min} = 0.0026$ and $M_{max} = 1.2 *10^3$.
-These values are based on empirical body sizes of stream benthic communities reported in [@pomeranz2022].
-Our results are not dependent on the range of body sizes, and we show the results of other body size ranges in the supplemental information.
+These values were based on empirical body sizes of stream benthic communities reported in [@pomeranz2022].
+Our results were not dependent on the range of body sizes (Supplemental Information).
 
@@
 
-We sampled $n = 1000$ body sizes from distributions described by five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5).
-The values of $\lambda$ describe how quickly the abundance of large body sizes decline within a community, e.g., a value of -1.5 means there would be a relatively high number of large body sizes (shallow) whereas a value of -2.5 means there would be relatively very few large body sizes (steep).
-We used three different methods (described below) to estimate the value of $\lambda$.
-We repeated this process 1000 times (reps) and plot the distribution of values obtained for each method.
+Using the procedure above, we simulated $n = 1000$ body sizes from five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5).
+The values of $\lambda$ describe how quickly the abundance of large body sizes decline within a community.
+For example, a value of -1.5 means there would be a relatively high number of large body sizes (shallow) whereas a value of -2.5 means there would be relatively very few large body sizes (steep).
 
@@
 
-For each sample of $M$ and each of the 1000 replicates, we estimated the exponent $\lambda$ directly using MLE methods modified from the the `sizeSpectra` package (Edwards et al.
-2020).
-Throughout the manuscript, these estimates are referred to as MLE.
+After simulating data, we used three different methods (described below) to estimate the value of $\lambda$ (maximum likelihood, equal logarithmic bins, and normalized abundance spectra).
+We repeated this process 1000 times (reps) and plotted the distribution of values obtained for each method.
+
+## Maximum likelihood estimation (MLE)
+
+We estimated the exponent $\lambda$ directly using MLE methods modified from the the `sizeSpectra` package [@edwards2017]. Throughout the manuscript, these estimates are referred to as MLE.
 
@@
 
-In addition, we use two common binning methods to estimate the OLS slope parameter in log-log space.
-It is important to note that the binning methods are estimating a slope ($\beta_{1}$) from an OLS regression analysis, and not $\lambda$ directly.
-However, in the interest of clarity, we will refer to all estimates as $\lambda$ regardless of the method used, and reserve the term $\beta_{env}$ to refer to the estimate of the size spectra relationships across the hypothetical environmental gradient.
+In addition, to estimate the OLS slope parameter in log-log space, we used two common binning methods.
+.
+It is important to note that the log-log regression is estimating $\lambda$ as the regressionslope ($\beta_{1}$), and not $\lambda$ directly.
+In the interest of clarity, we will refer to all estimates of lambda as $\lambda$ regardless of the method used, and reserve the terms “parameters”, “intercept”, and “slope” $\beta_{env}$ to refer to the parameters of linear regression that uses the lambdas as a response variable.
 
@@
 ### Log~2~ Bins: NAS
-The second method was similar to ELBn, but bins of Log~2~ widths are used.
+The second method was similar to ELBn, but bins of log~2~ widths are used.
 The number of bins is dependent on the range of body sizes present in the data.
@@
 The count in each bin is normalized in the same way as described above.
-This method has been  used [@jennings2004, @Jennings2002, @sprules2015, @pomeranz2019; @mcgarvey2018) and is referred to as the Normalized Abundance Spectrum (NAS).
+This method has been  used extensively in the literature [\e.g., @jennings2004, @Jennings2002, @sprules2015, @pomeranz2019; @mcgarvey2018) and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 
-Where $\beta_{1}$ is considered to be the $\lambda_{estimate}$.
+Where $N$ is the count in each bin, $M$ is the mid-point of each bin, and $\beta_{1}$ is considered to be the $\lambda_{estimate}$.
 
 ## Variation in $\lambda$ Across a Gradient
-
-### Hypothetical relationshiop of -0.5
 
@@
 
-To do this, we again sampled body sizes from distributions described by five different $\lambda$'s.
-However, because the estimates from the binning methods varied depending on the value of $\lambda$ (see results), we chose three different "windows" of lambda values.
-The three windows were steep ($\lambda = [-1.50,~ -1.75,~ -2.00,~ -2.25, -2.50]$), medium ($\lambda = [-1.00,~ -1.25,~ -1.50,~ -1.75,~ -2.00]$), and shallow (($\lambda = [-0.50,~ -0.75,~ -1.00,~ -1.25,~ -1.50]$).
+To do this, we first simulated  $\lambda$'s from a linear regression with a known slope $\beta_{1}$ of -0.5 and intercept of -1.5 across a generic predictor variable x with values ranging from -1 to 1:
 
-**NOTE** This MS seems to be getting long. Possibly present "medium" window in main results, and include other windows in SI
+$$\lambda_{ik} = \beta_{0} + \beta_{0}x_{ik}$$
 
-Each sample of body sizes from different $\lambda$ values was assumed to come from a community across a hypothetical environmental gradient $X$.
-Values of $X$ for each community are uniformly distributed from -1 to 1 (site-specific x-values: $(-1.0,~-0.5,~0,~0.5,~1.0)$.
-For each $\lambda$ "window", the relative change in $\lambda$ values ($\beta_{env}$) across the gradient is equal to -0.5, such that our model is:
+This produced $k$ = 1000  values from each value of $x$ (-1, -0.5, 0, 0.5, 1), resulting in 5000 total $\lambda$’s.
+From each $\lambda$ we simulated 1000 individual body sizes using the procedure described above.
 
-**JEFF** Is this correct? for some reason this seems wrong to me.
-
-$$\beta_{env} = Norm(\alpha -0.5X,~\sigma)$$
-
-We repeated the data simulation 1000 times (reps). The main results presented here were not dependent on the range of *x*-values or the number of sites (supplemental information).
-
-### Varying Hypothetical Relationship
-
-Finally, we also varied the relationship ($\beta_{env}$) across the hypothetical gradient.
-We tested $\beta_{env}$ values of 0, -0.25, and -0.5 by sampling 5 body sizes communities uniformly distributed across the following sets of $\lambda$ values: [-1.5], [-1.25, -1.75], [-1, -2].
-
-## Estimation of relationship across gradient, $\beta_{env}$
-
-For each simulation replicate and size spectra method, we estimated how the size spectra parameters varied across the hypothetical gradient.
-We did this by performing simple OLS regression of the $\lambda_{estimate}$ across the hypothetical gradient ($X$):
-
-$$\lambda_{estimate} = \beta_0 + \beta_{env} * X$$
-
-The distribution of the relationship coefficient, $\beta_{env}$, were plotted compared to the known relationship.
+Once the body size data were simulated, we then reversed the procedure.
+First, we used the three methods (NAS, ELBn, MLE) to estimate $\lambda$ for each of the 5000 body size datasets.
+Then we fit a linear regression to those data with the predictor variable $x$ and compared the resulting intercept ($\beta_{0}$) and slope ($\beta_{1}$) parameter estimates to the known value.
+The main results presented here were not dependent on the range of x-values or the number of sites (Supplemental Information).
 
@@
 
-We re-analyze two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients.
+We re-analyzed two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients.
 In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage (AMD) stress gradient.
@@
 This allows us to determine whether or not the main results would differ depending on the method used.
+
+## Performance Metrics
+
+We compared performance of each procedure (NAS, ELBn, MLE) by first plotting the distribution of lambda or regression parameter estimates ($\lambda$, $\beta_{0}$, $\beta_{1}$) against the known values.
+We calculated bias for the procedures overall as the median absolute difference (averaged across all simulations) between the known values and the modeled estimates.
+Finally, for each procedure we estimated the width of the 95% CI’s to compare uncertainty.
 

### 2023-02-23 Manuscript.Rmd
MS draft 23 feb 2023

@@
 output:
+   pdf_document:
+     toc: yes
+     keep_tex: true
+     pandoc_args:
+     - '--lua-filter=scholarly-metadata.lua'
+     - '--lua-filter=author-info-blocks.lua'
    word_document:
@@
       toc: no
-      pandoc_args:
-      - '--lua-filter=scholarly-metadata.lua'
-      - '--lua-filter=author-info-blocks.lua'
-   pdf_document:
-     toc: yes
-     keep_tex: true
       pandoc_args:
@@
 ```
-
-**NOTE** I'm still working on integrating with Zotero and importing all the necessary references (thanks Jim for the help and instructions!).
-Don't worry about the **??** references in the rendered version for now.
 
@@
 Body size distributions are a fundamental characteristic of communities.
-In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981, @Damuth1991, @Damuth1998].
-This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991, @Brown2004].
+In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981; @Damuth1991; @Damuth1998].
+This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991; @Brown2004].
 The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010].
@@
 Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
-Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972, @Andersen2006a].
+Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972; @Andersen2006a].
 Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions in log-log space of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
@@
 Size-binned OLS methods are particularly sensitive to decisions made in the binning process.
-Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ (White et al.
-2008, Edwards et al.
-2017), and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008 @edwards2020].
+Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ [@white2008; @edwards2017], and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008; @edwards2020].
 However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied based on method, the relative change across the environmental gradient was consistent across methods [@pomeranz2022].
@@
 This method has been used extensively in previous studies.
-For example see [@perkins2018, @martinez2016, @dossena2012].
+For example see [@perkins2018; @martinez2016; @dossena2012].
 In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes.
 The process of normalizing shifts the estimate by -1.
-In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015, @edwards2017, @pomeranz2022].
+In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015; @edwards2017; @pomeranz2022].
 Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
@@
 The count in each bin is normalized in the same way as described above.
-This method has been  used extensively in the literature [\e.g., @jennings2004, @Jennings2002, @sprules2015, @pomeranz2019; @mcgarvey2018) and is referred to as the Normalized Abundance Spectrum (NAS).
+This method has been  used extensively in the literature e.g., [@jennings2004; @Jennings2002; @sprules2015; @pomeranz2019; @mcgarvey2018] and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 There was considerable variation in the $\lambda$ estimate across methodologies (Figure 1).
-The distribution of estimates from the MLE method was almost always symmetrical and centered at the known value of $\lambda$ (Figure 1, blue distribution and dashed line).
-The distribution of estimates from the binning methods were generally wider, and sometimes showed assymetrical distributions.
+The distribution of estimates from the MLE method was always symmetrical and centered at the known value of $\lambda$ (Figure 1).
+The distribution of estimates from the binning methods were generally wider and occasionally asymmetrical or bimodal.
+On average, the 95% confidence intervals produced by the NAS and ELBn methods were ~2 times wider than those produced by MLE (Table 1), indicating greater consistency of estimates from MLE.
+Similarly, estimates of  $\lambda$ deviated from the true $\lambda$ by an average of 0.04 or 0.05 absolute units for the NAS and ELBn methods, more than 4 times higher than deviation (0.01) for the MLE (Table 1).
 
@@
 Note that the x-axis varies in each facet.](figures/est_lambda_est_b_density.png)
+
+```{r, echo=FALSE}
+t1 <- read.csv("tables/bias_table.csv")
+knitr::kable(t1,
+             col.names = c("Target",
+                           "Method",
+                           "n",
+                           "Median Range of 95% CI",
+                           "Median Absolute Deviation",
+                           "SD Absolute Deviation"),
+                           caption = "Table 1.
+Summary of three methods in recapturing the known values of lambda or the regression slopes simulated in this study.
+Performance is determined by comparinged by uncertainty (range of 95% CI's) and the absolute distance of the model estimates from the known estimates (median and sd of the difference).
+Values are summarized across all n = 9000 or 6000 simulated datasets.
+See figures for more specific comparisons.")
+
+```
 
@@
 
-The distribution of relationship estimates from the MLE method (Figure 3, blue) is approximately centered at the known relationship value in each of the "windows".
-The distribution for the MLE estimate is widest in the "steep" window, but is still markedly narrower than the binning methods.
-In addition to having much wider distributions, the binning methods systematically overestimate the known relationship.
+The MLE method (Figure 3, blue) recaptured the known slope value in each of the “windows.”, with median absolute deviation of ~0.01 units (Table 1).
+By contrast, the binning methods systematically overestimated the known slope (Figure 3), with median absolute deviations >4 times higher than the MLE.
+Similarly, uncertainty in the slope estimates derived from binning methods was twice that of the uncertainty in the MLE method (Table 1).
 
@@
 
-Finally, we were interested in finding if the three methods performed differently with varying values of the known relationship.
-Here, we present estimates from each method when the known relationship ($\beta_{env}$) across the hypothetical gradient was -0.5, -0.25, and 0, respectively (Figure 4).
+All methods recaptured the correct sign of the slopes, yielding qualitative consistency (Figure 4).
+However, the binning methods again systematically underestimated the true value of the slope by ~0.05 units (Figure 5).
+Likewise, uncertainty in the slope estimates was always greater in the binning methods, with the width of the distributions increases with stronger relationships across a hypothetical gradient.
+By comparison, the MLE showed no evidence of bias and was always centered at the known value with relatively narrow variation.
 
-![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows): 0, 0.25, and 0.5](figures/vary_beta_plot.png)
-
-Likewise, we also present the distribution of the $\beta_{env}$ relationship estimates (Figure 5).
-All the methods perform similarly and reasonably well when no relationship is present.
-The performance of the binning methods declines (center of distributions), with relationships being systematically over estimated.
-Likewise, the width of the distributions increases with stronger relationships across a hypothetical gradient.
-However, the distribution of estimates from the MLE method is nearly always centered at the known value, and has relatively narrow distributions.
+![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows from bottom to top): 0, 0.25, and 0.5](figures/vary_beta_plot.png)
 
@@
 
-For both empirical data sets, the direction and magnitude of change (i.e.
-$\beta_1$ coefficients) are generally in agreement.
-Size spectra parameters consistently increase (become flatter) in the AMD data (Fig.
-6A).
-However, the magnitude of the change across the AMD gradient is weakest using the MLE method, and the width of the estimates ($\pm$ 1 SD) is larger in the binning methods, particularly the ELBn method (Fig.
+Both empirical data sets yielded similar results to the simulated data; the direction of the coefficients (i.e., $\beta_1$ coefficients) were the same among methods, with positive slopes across the AMD gradient (Fig.
+6A) and negative slopes across the temperature gradient.
+However, the magnitude of change differed between methods.
+The AMD slopes ranged from ~0.62 with MLE to ~0.78 with NAS (Figure 6b), while temperature slopes ranged from -0.0058 with MLE to -0.0019 with ELBn (Figure 6d).
+As with simulated data, slope uncertainty ($\pm$ 1 SD) was larger in the binning methods, particularly the ELBn method (Fig.
 6B).
@@
 6C).
-The magnitude of the relationships is weakest in the MLE method, although the uncertainty around the estimates are similar (Fig.
-6D).
 
@@
 
-The $\beta_1$ coefficient estimates for the AMD data are all positive and of relatively similar magnitudes across methods.
-The MLE estimate has the smallest magnitude and SD (0.063 $\pm$ 0.011 SD), The ELBn method is slightly larger and has a wider SD : (0.072 $\pm$ 0.018).
-Finally, the NAS method has the largest estimate and an intermediate standard deviation (0.076 $\pm$ 0.013).
-The range of the gradient in the AMD data is relatively small (9.5), but the absolute change in the size spectra parameter ($\lambda$) across the AMD gradient vary by ~0.13 units depending on the method used (range of absolute change: 0.59 to 0.72).
-
-Similar to the AMD results, the $\beta_1$ coefficient estimates for the NEON data are all negative and have similar magnitudes across methods.
-The MLE estimate is once again the smallest (-0.006 $\pm$ 0.001 SD).
-The NAS and ELBn methods are larger at -0.004 $\pm$ 0.002, and -0.002 $\pm$ 0.002, respectively.
-Although the estimates are similar and have a small magnitude, the range of the gradient in this data is larger (\~30$\circ$C), meaning that the absolute change in $\lambda$ estimates varies by 0.11 units depending on the methods used (range of absolute change across gradient: 0.06 to 0.17).
-
 # Discussion
 
-The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@brown1995, GastonBlackburn2000, @White2007]).
-Empirical patterns are remarkably consistent, and can be explained by the metabolic theory of Ecology [@Brown2004].
+The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@brown1995; GastonBlackburn2000; @White2007]).
+Empirical data shows generally consistent patterns and can be explained by the metabolic theory of ecology [@Brown2004].
 Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology.
-Previous work has investigated the accuracy and inherent biases associated with different estimation methods.
+Previous work has investigated the accuracy and inherent biases associated with different estimation methods [@White2007; @edwards2017; @edwards2020].
 However, how these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
-Here, we sampled body sizes from known distributions with varying parameters ($\lambda$) and estimated the coefficient of the relationship ($\beta_{env}$) across a hypothetical gradient in order to assess how the results would vary depending on the method used.
-Likewise, we compared how the interpretation of previously published results could change depending on the methodology used to estimate the relationship.
-
-Generally, the estimate of the size spectra parameter from the MLE method was always closer to the known $\lambda$ value, regardless of how steep or shallow the $\lambda$ exponent was.
-This is in agreement with Edwards et at.
-(2017) which found that the MLE estimate more consistently recovered known $\lambda$ values ranging from -2.5 to -0.5.
-We further this work by discovering that the magnitude of the error associated with the two binning methods increases as the $\lambda$ value moves away from $\sim$ -1.
-
-It is unsurprising that all methods performed less well in estimating steep values of $\lambda$ (i.e., -2.5).
-The rapid decline of large body sizes makes it unlikely to sample any individuals with a large body size.
-Furthermore, the binning methods inherently "lose" data by aggregating individual observations into a single bin.
-
-Likewise, the MLE method also resulted in the most accurate $\beta_{env}$ estimates for the hypothetical relationship in simulated data regardless of the $\lambda$ "window" investigated.
-Furthermore, the variation around the estimate was always smallest with the MLE method.
-However, the variation in the MLE estimate did increase with steeper $\lambda$ "windows".
-This is not surprising, however because steep relationships like this make it very unlikely to sample the largest body sizes, and capturing the "true" relationships without the full range of body sizes present is more difficult (Edwards et al.
-2017).
-Despite this, the MLE method is nearly always the best option to use, likely because all the data are used in the analysis.
+The most important outcome of our work is that binning methods not only generate biased $\lambda$ values for individual data sets, but that bias carries over to affect the parameters of subsequent regressions that use those $\lambda$’s as response variables.
+This makes it challenging to understand how $\lambda$ varies in response to environmental gradients if binning is used to estimate size spectra exponents.
 
@@
 For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
-Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams.
-By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use.
-One of the benefits of using MLE is that all data points are retained for use within the model.
+Likewise, a single abundance value is taken for each bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams.
+Moreover, the number of bins that can be produced by any dataset is limited most often to less than six.
+That means a subsequent log-log regression used to estimate $\lambda$ would contains n = 6 data points, even though there likely 100’s or 1000’s of individual body sizes available in a typical body size data set.
+By homogenizing the data in this way, binning methods produce noisier results than MLE, since binning is akin to deleting information that the model could otherwise use.
+By contrast, the MLE uses all the individual body size data to directly estimate $\lambda$, meaning that it not only produces more accurate estimates, but does so with less certainty than binning, even when the underlying data sets are identical.
 
-The variation in estimated relationships across the empirical datasets varied by ~0.1 units depending on the method used.
-At first glance this may not appear to be a concerning magnitude of variation.
-However, when one considers the range of variation reported in many seminal works on variation in size spectra relationships is ~0.1-0.2 (SI Table *xx*), it becomes clear how important an error in estimates of this magnitude is.
+The error in estimating individual $\lambda$’s was compounded when estimating linear models with those $\lambda$’s as response variables.
+The MLE method also resulted in the most accurate $\beta_{env}$ estimates.
+Furthermore, variation around the estimate was always smallest with the MLE method.
 
-Despite the differences in magnitude of the empirical relationship coefficients, they were generally in the correct direction and of a similar magnitude.
+Although there were differences in magnitude of the empirical relationship coefficients, they were generally in the correct direction and of a similar magnitude.
 This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
@@
 However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.
-However, the relatively systematic error of $\lambda$ estimates across known exponents suggests that it may be possible to correct previous estimates onto the same scale (Figure 2).
 
-Despite this possibility, publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
+The publication of individual body size data with future studies of size spectra would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
 
@@
 With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
-Therefore, we recommend using this in all future studies of size spectra relationships.
+Therefore, we recommend using it in all future studies of size spectra relationships rather than binning.
 
@@
 
+![Distribution of estimated \$\\lambda\$ coefficient for five sites across a hypothetical gradient with known values(dashed line).
+Range of body sizes is smaller than main anaysis and ranges from 1, to 100.](figures/PLB_small_m_est_b_density.png)
+
 ![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
@@
 
-\newpage
-
-### Range of body sizes = 1 to 100
-
-![Distribution of estimated \$\\lambda\$ coefficient for five sites across a hypothetical gradient with known values(dashed line).
-Range of body sizes is smaller than main anaysis and ranges from 1, to 100.](figures/PLB_small_m_est_b_density.png)
-
-\newpage
-
-![Relationship estimate](figures/PLB_small_m_relationship_density.png)
-
 # Comparison with other published estimates
@@
 This table shows published estimates of the variation in size spectra slopes (or exponents) in empirical studies.
-It is unclear how to directly compare estimates of the slope with different methods.However, the published estimates here range from ~0.1 to 0.2 across the gradients studied.
+It is unclear how to directly compare estimates of the slope with different methods.
+However, the published estimates here range from ~0.1 to 0.2 across the gradients studied.
 For comparison, the 2.5-95% quantiles around the relationship estimate for the MLE method were ~0.1, whereas for the ELBn and NAS method they were ~0.25 and ~0.2, respectively.
@@
 
-```{r, echo=FALSE}
-df <- read.csv("lit comparison.csv")
-knitr::kable(df,
-             col.names = c("Citation",
-                           "$\\beta_1$ Difference",
-                           "Error", 
-                           "Error Type",
-                           "$\\beta_{low}$",
-                           "$\\beta_{high}$",
-                           "System",
-                           "Driver", 
-                           "Method"))
-
-```
-
-```{r, echo=TRUE}
-# kable() above looks awful
-# here is another bad option
-# if we decide to include this I'll make it look nice in excel. 
-read.csv("lit comparison.csv")
-```
+Supplementary table comparing with other published results will be submitted as a separate file for formatting.

### 2023-02-23 Manuscript.tex
MS draft 23 feb 2023

@@
 Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
-\date{07 January, 2023}
+\date{23 February, 2023}
 
@@
 
-\textbf{NOTE} I'm still working on integrating with Zotero and importing
-all the necessary references (thanks Jim for the help and
-instructions!). Don't worry about the \textbf{??} references in the
-rendered version for now.
-
 \hypertarget{abstract}{%
@@
   The distribution of body size in communities is remarkably consistent
-  across habitats and taxa and can be represented by size spectra
-  relationships.
+  across habitats and taxa and can be represented by size spectra, which
+  are described by a power law. The focus of size spectra analysis is to
+  estimate the exponent (\(\lambda\)) of the power law.
 \item
-  Many methods have been proposed for constructing size spectra
-  relationships, but recent work has shown that varied methods return
-  biased estimates of relationship parameters, and different methods in
-  fact are not estimating the same parameter. Despite this variability
-  in estimates, it is unclear if the relative change across
-  environmental gradients is consistent across methodologies. Here, we
-  simulate data sets across an hypothetical environmental gradient and
-  estimate the size spectra parameter (slope or exponent, depending on
-  method) at each site, and, importantly, estimate the relationship of
-  parameters across the hypothetical gradient. We also use two
-  previously published body size datasets across an anthropogenic stress
-  gradient and an environmental temperature gradient and assess how the
-  conclusions of these studies would vary based on methods used.
+  Many methods have been proposed for estimating \(\lambda\) most of
+  which involve binning the data, summing abundance within bins, and
+  then fitting a ordinary least squares (OLS) regression in log-log
+  space. However, recent work has shown that binning procedures may
+  return biased estimates of size spectra exponents compared to
+  procedures that directly estimate \(\lambda\) using maximum likelihood
+  estimation (MLE). Despite this variability in estimates, it is unclear
+  if the relative change across environmental gradients is consistent
+  across methodologies. Here, we used simulation to compare the ability
+  of two binning methods (NAS, ELBn) and MLE to 1) recapture known
+  values of \(\lambda\), and 2) recapture parameters in a linear
+  regression measuring the change in \(\lambda\) across a hypothetical
+  environmental gradient. We also compared the methods using two
+  previously published body size datasets across a pollution gradient
+  and a temperature gradient
 \item
-  We find that maximum likelihood methods always perform better than
-  common binning methods. Additionally, the variance in estimates using
-  MLE methods is markedly reduced when compared to binning methods.
+  Maximum likelihood methods always performed better than common binning
+  methods, which demonstrated consistent bias depending on the simulated
+  values of \(\lambda\). This bias carried over to the regressions,
+  which were more accurate when \(\lambda\) was estimated using MLE
+  compared to the binning procedures. Additionally, the variance in
+  estimates using MLE methods is markedly reduced when compared to
+  binning methods.
 \item
@@
   in experimental and observational studies, bringing into question the
-  effect size of previously published results. However, re-analysis of
-  two previously published datasets does not markedly alter the
-  conclusions reached. Further study is needed to identify how and when
-  these estimates may be biased, and when the general pattern of results
-  can be considered to be ``true.''
+  effect size of previously published results. However, while the
+  methods produced different slope estimates from previously published
+  datasets, they were in qualitative agreement on the sign of those
+  slopes (i.e., all negative or all positive). Our results provide
+  further support for the direct estimation of \(\lambda\) using MLE (or
+  similar procedures) over the more common methods of binning.
 \end{enumerate}
@@
 value of -0.75 across all biological communities (Damuth 1981, 1991,
-1998). This is thought to be a consequence of simple size-dependent
-metabolic constraints on organisms' energy use predicted by the
-metabolic theory of ecology (Nee 1991, Brown et al.~2004). The
-remarkable consistency of these relationships across spatiotemporal
+Damuth 1998). This is thought to be a consequence of simple
+size-dependent metabolic constraints on organisms' energy use predicted
+by the metabolic theory of ecology (Nee et al. 1991, Brown et al. 2004).
+The remarkable consistency of these relationships across spatiotemporal
 scales and ecosystems has led them to be recommended as a ``universal''
-indicator of ecological status (Petchey and Belgrano 2010)(Petchey and
-Belgrano 2010). Variation in size-abundance relationships have been
-documented through space (Pomeranz 2021), time (newish Sprules papers?),
-seasonality (McGarvey and Kirk), to human activities (fish abundance
-with no fishing paper). Likwise, variation in size spectra relationships
-have been used to explain fundamental differences in how communities are
+indicator of ecological status (Petchey and Belgrano 2010). Variation in
+size-abundance relationships have been documented through space
+(Pomeranz et al. 2022), time (Evans et al. 2022), seasonality (McGarvey
+and Kirk 2018), and in response to human activities (Jennings and
+Blanchard 2004). Likewise, variation in size spectra relationships have
+been used to explain fundamental differences in how communities are
 organized. For example, external resource subsidies ``bend the rules''
 and allow higher abundances of large body sizes than would be expected
-(Perkins). However, recent research has shown that these results may be
-an artifact of how the data were treated. Edwards et al.~(2020?)
-reanalyzed a time series of marine fisheries data and found that
-previously reported changes through time were actually dependent on the
-methodolugy used.
-
-Individual size distributions (ISD \emph{sensu} (White et al. 2007)),
+(Perkins et al. 2018). However, recent research has shown that these
+results may be an artifact of how the data were treated. Edwards et
+al.~(2020) reanalyzed a time series of marine fisheries data and found
+that previously reported changes through time were actually dependent on
+the methodology used.
+
+Individual size distributions (ISD \emph{sensu} (White et al. 2007),
 also referred to as abundance size spectra) are one of the
-size-abundance relationships commonly used. ISDs show the relationship
-between the abundances of individual body size regardless of taxonomy.
-Generally, there is a negative relationship between individual body size
-(M) on the \emph{x}-axis and abundance (N) on the \emph{y}-axis.
-Theoretical and empirical data support this relationship being described
-as a simple power law with exponent \(\lambda\) in the form of
-\(N \sim M^{\lambda}\), where \(\lambda = 2\). Commonly, \(N\) is the
-count of body sizes grouped into bins, and \(\lambda\) is estimated as
-the slope from OLS regressions of log transformed data of \(N\) and the
-mid point of the body size bin \(M_{bin}\):
-\(\log(N_{count}) = \lambda \log(M_{bin})\). Myriad binning methods have
-been proposed, including linear and logarithmic bin widths. Likewise,
-methods can rely on the absolute count in the bins as well as
-normalization techniques where the count is divided by the bin width
-(especially common with logarithmic binning). Alternatively, \(\lambda\)
-can be estimated directly using maximum likelihood techniques.
+size-abundance relationships commonly used. ISDs represent frequency
+distributions of individual body sizes within a community, regardless of
+taxonomy. Generally, there is a negative relationship between individual
+body size (M) on the \emph{x}-axis and abundance (N) on the
+\emph{y}-axis. Theoretical and empirical data support this relationship
+being described as a simple power law with exponent \(\lambda\) in the
+form of \(N \sim M^{\lambda}\), where \(\lambda = 2\) (Sheldon and Kerr
+1972, Andersen and Beyer 2006). Commonly, \(N\) is the count of body
+sizes grouped into bins, and \(\lambda\) is estimated as the slope from
+OLS regressions in log-log space of \(N\) and the mid point of the body
+size bin \(M_{bin}\): \(\log(N_{count}) = \lambda \log(M_{bin})\).
+Myriad binning methods have been proposed, including linear and
+logarithmic bin widths. Likewise, some methods rely on the absolute
+counts in the bins and others employ normalization techniques such as
+dividing the count by the bin width (especially common with logarithmic
+binning). As an alternative to binning methods, \(\lambda\) can be
+estimated directly on un-binned data using maximum likelihood
+techniques.
 
 Previous work has shown that the estimates of \(\lambda\) differ between
-MLE and OLS techniques. OLS methods are particularly sensitive to
-decisions made in the binning process. Simulation studies have shown
-that MLE offers consistently more accurate estimates of \(\lambda\)
-(White et al.~2008, Edwards et al.~2017), and reanalysis of empirical
-data also indicates that the conclusions are dependent on the
-methodology used Edwards et al. (2020). However, recent empirical
-analysis of stream macroinvertebrate communities across the National
-Ecological Observatory Network (NEON, USA) showed that while the
-estimates of \(\lambda\) varied, the relative change across the
-environmental gradient was consistent regardless of method used
-(Pomeranz et al. 2022). While there is a growing consensus that MLE
-methods offer more reliable estimates of \(\lambda\), and binning
-methods result in biased estimates, it remains unclear if these biases
-are consistent and systematic or stochastic, and whether or not the
-relative change in ISD parameters is consistent across space and time.
-In other words, if the data within a study are all treated the same,
-does a relative change of OLS slope parameters of 0.1 coincide with a
-relative change of MLE estimates of 0.1?
-
-We had three primary objectives in this study: 1) how well do different
-methods estimate site-specific \(\lambda\)'s? 2) Do the estimates of the
-hypothetical relationship obtained from different methods agree? and 3)
-Are the conclusions reached from previously published data sets
-dependent on the method used?
-
-In order to answer the first two questions, we simulate body size
-observations from bounded power law distributions with varied
-\(\lambda\) exponents across a hypothetical environmental gradient and
-compare the results obtained using three different methodologies common
-in the literature. To answer the third question, we re-analyze two
-previously published data sets of stream community body sizes across a
-stress and environmental gradient, respectively. We find that the MLE
-method more accurately estimates the site-specific \(\lambda\) exponents
-as well as the relative change in exponents across the hypothetical
-gradient, as well as having smaller variation around the estimates. The
-logarithmic binning methods generally perform well, but have larger
-variation around the site-specific estimates. In addition, the variation
-of the estimates of the hypothetical relationship obtained from the
-binning methods is dependent on the value of \(\lambda\) used (under and
-over estimates steep and shallow \(\lambda\), respectively).
+MLE and size-binned OLS techniques. Size-binned OLS methods are
+particularly sensitive to decisions made in the binning process.
+Simulation studies have shown that MLE offers consistently more accurate
+estimates of \(\lambda\) (White et al. 2008, Edwards et al. 2017), and
+reanalysis of empirical data also indicates that the conclusions are
+dependent on the methodology used (White et al. 2008, Edwards et al.
+2020). However, recent empirical analysis of stream macroinvertebrate
+communities across the National Ecological Observatory Network (NEON,
+USA) showed that while the estimates of \(\lambda\) varied based on
+method, the relative change across the environmental gradient was
+consistent across methods (Pomeranz et al. 2022). While there is a
+growing consensus that MLE methods offer more reliable estimates of
+\(\lambda\), and binning methods result in biased estimates, it remains
+unclear if these biases are consistent and systematic or stochastic, and
+whether or not the relative change in ISD parameters is consistent
+across space and time. In other words, if the data within a study are
+all treated the same, does a relative change of size-binned OLS slope
+parameters of 0.1 coincide with a relative change of MLE \(\lambda\)
+estimates of 0.1?
+
+We had Two primary objectives in this study: 1) to compare how well
+different methods estimate site-specific \(\lambda\)'s 2) To compare how
+different methods of estimating \(\lambda\) impact subsequent
+regressions that use \(\lambda\)'s as a response variable. We did this
+using both simulated and empirical data sets.
 
@@
 
-In order to investigate the performance of commonly used methods, we
-simulate body size observations from a bounded power law distribution
-using the inverse method, as described in (Edwards et al. 2017). Let
-\(M\) be a random variable of body sizes described by the probability
-density function:
+To investigate the performance of commonly used methods, we simulate
+body size observations from a bounded power law distribution using the
+inverse method, as described in (Edwards et al. 2017). Let \(M\) be a
+random variable of body sizes described by the probability density
+function:
 
@@
 
-\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values are based
-on empirical body sizes of stream benthic communities reported in
-(Pomeranz et al. 2022). Our results are not dependent on the range of
-body sizes, and we show the results of other body size ranges in the
-supplemental information.
+\(M_{min} = 0.0026\) and \(M_{max} = 1.2 *10^3\). These values were
+based on empirical body sizes of stream benthic communities reported in
+(Pomeranz et al. 2022). Our results were not dependent on the range of
+body sizes (Supplemental Information).
 
@@
 
-We sampled \(n = 1000\) body sizes from distributions described by five
-different \(\lambda\)'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The values
-of \(\lambda\) describe how quickly the abundance of large body sizes
-decline within a community, e.g., a value of -1.5 means there would be a
-relatively high number of large body sizes (shallow) whereas a value of
--2.5 means there would be relatively very few large body sizes (steep).
-We used three different methods (described below) to estimate the value
-of \(\lambda\). We repeated this process 1000 times (reps) and plot the
-distribution of values obtained for each method.
+Using the procedure above, we simulated \(n = 1000\) body sizes from
+five different \(\lambda\)'s: (-1.5, -1.75, -2.00, -2.25, -2.5). The
+values of \(\lambda\) describe how quickly the abundance of large body
+sizes decline within a community. For example, a value of -1.5 means
+there would be a relatively high number of large body sizes (shallow)
+whereas a value of -2.5 means there would be relatively very few large
+body sizes (steep).
 
@@
 
-For each sample of \(M\) and each of the 1000 replicates, we estimated
-the exponent \(\lambda\) directly using MLE methods modified from the
-the \texttt{sizeSpectra} package (Edwards et al.~2020). Throughout the
-manuscript, these estimates are referred to as MLE.
+After simulating data, we used three different methods (described below)
+to estimate the value of \(\lambda\) (maximum likelihood, equal
+logarithmic bins, and normalized abundance spectra). We repeated this
+process 1000 times (reps) and plotted the distribution of values
+obtained for each method.
+
+\hypertarget{maximum-likelihood-estimation-mle}{%
+\subsection{Maximum likelihood estimation
+(MLE)}\label{maximum-likelihood-estimation-mle}}
+
+We estimated the exponent \(\lambda\) directly using MLE methods
+modified from the the \texttt{sizeSpectra} package (Edwards et al.
+2017). Throughout the manuscript, these estimates are referred to as
+MLE.
 
@@
 
-In addition, we use two common binning methods to estimate the OLS slope
-parameter in log-log space. It is important to note that the binning
-methods are estimating a slope (\(\beta_{1}\)) from an OLS regression
-analysis, and not \(\lambda\) directly. However, in the interest of
-clarity, we will refer to all estimates as \(\lambda\) regardless of the
-method used, and reserve the term \(\beta_{env}\) to refer to the
-estimate of the size spectra relationships across the hypothetical
-environmental gradient.
+In addition, to estimate the OLS slope parameter in log-log space, we
+used two common binning methods. . It is important to note that the
+log-log regression is estimating \(\lambda\) as the regressionslope
+(\(\beta_{1}\)), and not \(\lambda\) directly. In the interest of
+clarity, we will refer to all estimates of lambda as \(\lambda\)
+regardless of the method used, and reserve the terms ``parameters,''
+``intercept,'' and ``slope'' \(\beta_{env}\) to refer to the parameters
+of linear regression that uses the lambdas as a response variable.
 
@@
 covering the range of body sizes. This method has been used extensively
-in previous studies. For example see Dossena et al. (2012). In the
-present study, the count in each bin was normalized by dividing by the
-bin width to account for the unequal bin sizes. The process of
-normalizing shifts the estimate by -1. In other words, an un-normalized
-estimate of -0.75 would result in an estimate of -1.75 when normalizing
-the data (Sprules and Barth 2015 pp. Edwards2018, Pomeranz2022).
-Throughout the manuscript, the normalized equal logarithmic binning
-method will be referred to as ELBn.
+in previous studies. For example see (Dossena et al. 2012, Martínez et
+al. 2016, Perkins et al. 2018). In the present study, the count in each
+bin was normalized by dividing by the bin width to account for the
+unequal bin sizes. The process of normalizing shifts the estimate by -1.
+In other words, an un-normalized estimate of -0.75 would result in an
+estimate of -1.75 when normalizing the data (Sprules and Barth 2015,
+Edwards et al. 2017, Pomeranz et al. 2022). Throughout the manuscript,
+the normalized equal logarithmic binning method will be referred to as
+ELBn.
 
@@
 
-The second method was similar to ELBn, but bins of Log\textsubscript{2}
-widths are used. The number of bins is dependent on the size ranges
-present in the data. When working with empirical data with different
-size ranges this can alter the number of bins per site. However, since
-the data here were simulated from a known size range, the number of bins
-for each site is identical. The count in each bin is normalized in the
-same way. This method has been used {[}Jennings and Blanchard (2004),
-Jennings et al. (2002), Sprules and Barth (2015), Pomeranz et al.
-(2019a); McGarvey and Kirk (2018)) and is referred to as the Normalized
-Abundance Spectrum (NAS).
+The second method was similar to ELBn, but bins of log\textsubscript{2}
+widths are used. The number of bins is dependent on the range of body
+sizes present in the data. When working with empirical data with
+different size ranges this can alter the number of bins per site, and it
+is recommended to construct the bins based on the global min and max
+body size present. However, since the data here were simulated from a
+known size range, the number of bins for each site is identical. The
+count in each bin is normalized in the same way as described above. This
+method has been used extensively in the literature e.g., (Jennings et
+al. 2002, Jennings and Blanchard 2004, Sprules and Barth 2015, McGarvey
+and Kirk 2018, Pomeranz et al. 2019a) and is referred to as the
+Normalized Abundance Spectrum (NAS).
 
 For the two binning methods, simple OLS regression were conducted in the
-form \(Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)\). Where
-\(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).
+form \[Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)\]
+
+Where \(N\) is the count in each bin, \(M\) is the mid-point of each
+bin, and \(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).
 
@@
 Gradient}{Variation in \textbackslash lambda Across a Gradient}}\label{variation-in-lambda-across-a-gradient}}
-
-\hypertarget{hypothetical-relationshiop-of--0.5}{%
-\subsubsection{Hypothetical relationshiop of
--0.5}\label{hypothetical-relationshiop-of--0.5}}
 
@@
 
-To do this, we again sampled body sizes from distributions described by
-five different \(\lambda\)'s. However, because the estimates from the
-binning methods varied depending on the value of \(\lambda\) (see
-results), we chose three different ``windows'' of lambda values. The
-three windows were steep
-(\(\lambda = [-1.50,~ -1.75,~ -2.00,~ -2.25, -2.50]\)), medium
-(\(\lambda = [-1.00,~ -1.25,~ -1.50,~ -1.75,~ -2.00]\)), and shallow
-((\(\lambda = [-0.50,~ -0.75,~ -1.00,~ -1.25,~ -1.50]\)).
-
-\textbf{NOTE} This MS seems to be getting long. Possibly present
-``medium'' window in main results, and include other windows in SI
-
-Each sample of body sizes from different \(\lambda\) values was assumed
-to come from a community across a hypothetical environmental gradient
-\(X\). Values of \(X\) for each community are uniformly distributed from
--1 to 1 (site-specific x-values: \((-1.0, -0.5, 0, 0.5, 1.0)\). For each
-\(\lambda\) ``window,'' the relative change in \(\lambda\) values
-(\(\beta_{env}\)) across the gradient is equal to -0.5, such that our
-model is:
-
-\textbf{JEFF} Is this correct? for some reason this seems wrong to me.
-
-\[\beta_{env} = Norm(\alpha -0.5X,~\sigma)\]
-
-We repeated the data simulation 1000 times (reps). The main results
-presented here were not dependent on the range of \emph{x}-values or the
-number of sites (supplemental information).
-
-\hypertarget{varying-hypothetical-relationship}{%
-\subsubsection{Varying Hypothetical
-Relationship}\label{varying-hypothetical-relationship}}
-
-Finally, we also varied the relationship (\(\beta_{env}\)) across the
-hypothetical gradient. We tested \(\beta_{env}\) values of 0, -0.25, and
--0.5 by sampling 5 body sizes communities uniformly distributed across
-the following sets of \(\lambda\) values: {[}-1.5{]}, {[}-1.25,
--1.75{]}, {[}-1, -2{]}.
-
-\hypertarget{estimation-of-relationship-across-gradient-beta_env}{%
-\subsection{\texorpdfstring{Estimation of relationship across gradient,
-\(\beta_{env}\)}{Estimation of relationship across gradient, \textbackslash beta\_\{env\}}}\label{estimation-of-relationship-across-gradient-beta_env}}
-
-For each simulation replicate and size spectra method, we estimated how
-the size spectra parameters varied across the hypothetical gradient. We
-did this by performing simple OLS regression of the
-\(\lambda_{estimate}\) across the hypothetical gradient (\(X\)):
-
-\[\lambda_{estimate} = \beta_0 + \beta_{env} * X\]
-
-The distribution of the relationship coefficient, \(\beta_{env}\), were
-plotted compared to the known relationship.
+To do this, we first simulated \(\lambda\)'s from a linear regression
+with a known slope \(\beta_{1}\) of -0.5 and intercept of -1.5 across a
+generic predictor variable x with values ranging from -1 to 1:
+
+\[\lambda_{ik} = \beta_{0} + \beta_{0}x_{ik}\]
+
+This produced \(k\) = 1000 values from each value of \(x\) (-1, -0.5, 0,
+0.5, 1), resulting in 5000 total \(\lambda\)'s. From each \(\lambda\) we
+simulated 1000 individual body sizes using the procedure described
+above.
+
+Once the body size data were simulated, we then reversed the procedure.
+First, we used the three methods (NAS, ELBn, MLE) to estimate
+\(\lambda\) for each of the 5000 body size datasets. Then we fit a
+linear regression to those data with the predictor variable \(x\) and
+compared the resulting intercept (\(\beta_{0}\)) and slope
+(\(\beta_{1}\)) parameter estimates to the known value. The main results
+presented here were not dependent on the range of x-values or the number
+of sites (Supplemental Information).
 
@@
 
-We re-analyze two data sets of benthic macroinvertebrate communities
+We re-analyzed two data sets of benthic macroinvertebrate communities
 from stream habitats across two different gradients. In the first,
@@
 
-The second data set was from the wadeable stream sites of the National
-Ecological Observatory Network (National Ecological Observatory Network
-(NEON) 2022). NEON stream sites are located across a wide temperature
-gradient in the United States, from Puerto Rico to Alaska. Quantitative
-macroinvertebrate samples were collected using the most appropriate
-method based on the local habitat. All individuals were identified and
-had their body lengths measured, and body mass was estimated using
-published length weight regressions. This data has been analyzed
-previously using size spectra methods as described in (Pomeranz et al.
-2022). Detailed methods of the sampling collection and data processing
-methods can be found on the NEON website
-({[}\url{https://www.neonscience.org/data-collection/macroinvertebrates}{]},
-or macroinvertebrate DPI pubs). Coefficient estimates (\(\pm\) 1
-standard deviation) are compared across methods. This allows us to
-determine whether or not the main results would differ depending on the
-method used.
+The second data set was from the wadeable stream sites of NEON (National
+Ecological Observatory Network (NEON) 2022). NEON stream sites are
+located across a wide temperature gradient in the United States, from
+Puerto Rico to Alaska. Quantitative macroinvertebrate samples were
+collected using the most appropriate method based on the local habitat.
+All individuals were identified and had their body lengths measured, and
+body mass was estimated using published length weight regressions. This
+data has been analyzed previously using size spectra methods as
+described in (Pomeranz et al. 2022). Detailed methods of the sampling
+collection and data processing methods can be in the macroinvertebrate
+DPI pubs found on the NEON website
+\url{https://www.neonscience.org/data-collection/macroinvertebrates}.
+Coefficient estimates (\(\pm\) 1 standard deviation) are compared across
+methods. This allows us to determine whether or not the main results
+would differ depending on the method used.
+
+\hypertarget{performance-metrics}{%
+\subsection{Performance Metrics}\label{performance-metrics}}
+
+We compared performance of each procedure (NAS, ELBn, MLE) by first
+plotting the distribution of lambda or regression parameter estimates
+(\(\lambda\), \(\beta_{0}\), \(\beta_{1}\)) against the known values. We
+calculated bias for the procedures overall as the median absolute
+difference (averaged across all simulations) between the known values
+and the modeled estimates. Finally, for each procedure we estimated the
+width of the 95\% CI's to compare uncertainty.
 
@@
 ({[}\url{https://github.com/Jpomz/detecting-spectra-differences}{]}).
+(To be archived upon acceptance.)
 
@@
 methodologies (Figure 1). The distribution of estimates from the MLE
-method was almost always symmetrical and centered at the known value of
-\(\lambda\) (Figure 1, blue distribution and dashed line). The
-distribution of estimates from the binning methods were generally wider,
-and sometimes showed assymetrical distributions.
+method was always symmetrical and centered at the known value of
+\(\lambda\) (Figure 1). The distribution of estimates from the binning
+methods were generally wider and occasionally asymmetrical or bimodal.
+On average, the 95\% confidence intervals produced by the NAS and ELBn
+methods were \textasciitilde2 times wider than those produced by MLE
+(Table 1), indicating greater consistency of estimates from MLE.
+Similarly, estimates of \(\lambda\) deviated from the true \(\lambda\)
+by an average of 0.04 or 0.05 absolute units for the NAS and ELBn
+methods, more than 4 times higher than deviation (0.01) for the MLE
+(Table 1).
 
@@
 
+\begin{longtable}[]{@{}llrrrr@{}}
+\caption{Table 1. Summary of three methods in recapturing the known
+values of lambda or the regression slopes simulated in this study.
+Performance is determined by comparinged by uncertainty (range of 95\%
+CI's) and the absolute distance of the model estimates from the known
+estimates (median and sd of the difference). Values are summarized
+across all n = 9000 or 6000 simulated datasets. See figures for more
+specific comparisons.}\tabularnewline
+\toprule
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Target\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+Method\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+n\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+Median Range of 95\% CI\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+Median Absolute Deviation\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+SD Absolute Deviation\strut
+\end{minipage}\tabularnewline
+\midrule
+\endfirsthead
+\toprule
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Target\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+Method\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+n\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+Median Range of 95\% CI\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+Median Absolute Deviation\strut
+\end{minipage} &
+\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+SD Absolute Deviation\strut
+\end{minipage}\tabularnewline
+\midrule
+\endhead
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Lambda\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+MLE\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+9000\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+0.0661400\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+0.0120454\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+0.0191500\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Lambda\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+NAS\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+9000\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+0.1335267\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+0.0460093\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+0.0929513\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Lambda\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+ELBn\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+8579\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+0.1672598\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+0.0350186\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+0.0675122\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Regression Slope\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+MLE\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+6000\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+0.0382185\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+0.0086982\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+0.0089766\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Regression Slope\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+NAS\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+6000\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+0.0799201\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+0.0513434\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+0.0442020\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
+Regression Slope\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
+ELBn\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
+6000\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
+0.0923495\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
+0.0300186\strut
+\end{minipage} &
+\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
+0.0353218\strut
+\end{minipage}\tabularnewline
+\bottomrule
+\end{longtable}
+
 Furthermore, the binning methods systematically over estimated steep
@@
 
-The distribution of relationship estimates from the MLE method (Figure
-3, blue) is approximately centered at the known relationship value in
-each of the ``windows.'' The distribution for the MLE estimate is widest
-in the ``steep'' window, but is still markedly narrower than the binning
-methods. In addition to having much wider distributions, the binning
-methods systematically overestimate the known relationship.
+The MLE method (Figure 3, blue) recaptured the known slope value in each
+of the ``windows.'' with median absolute deviation of
+\textasciitilde0.01 units (Table 1). By contrast, the binning methods
+systematically overestimated the known slope (Figure 3), with median
+absolute deviations \textgreater4 times higher than the MLE. Similarly,
+uncertainty in the slope estimates derived from binning methods was
+twice that of the uncertainty in the MLE method (Table 1).
 
@@
 
-Finally, we were interested in finding if the three methods performed
-differently with varying values of the known relationship. Here, we
-present estimates from each method when the known relationship
-(\(\beta_{env}\)) across the hypothetical gradient was -0.5, -0.25, and
-0, respectively (Figure 4).
+All methods recaptured the correct sign of the slopes, yielding
+qualitative consistency (Figure 4). However, the binning methods again
+systematically underestimated the true value of the slope by
+\textasciitilde0.05 units (Figure 5). Likewise, uncertainty in the slope
+estimates was always greater in the binning methods, with the width of
+the distributions increases with stronger relationships across a
+hypothetical gradient. By comparison, the MLE showed no evidence of bias
+and was always centered at the known value with relatively narrow
+variation.
 
@@
 \caption{Figure 4. Individual regressions for each rep (N = 1000) and
-method (columns) for three different known relationship values (rows):
-0, 0.25, and 0.5}
-\end{figure}
-
-Likewise, we also present the distribution of the \(\beta_{env}\)
-relationship estimates (Figure 5). All the methods perform similarly and
-reasonably well when no relationship is present. The performance of the
-binning methods declines (center of distributions), with relationships
-being systematically over estimated. Likewise, the width of the
-distributions increases with stronger relationships across a
-hypothetical gradient. However, the distribution of estimates from the
-MLE method is nearly always centered at the known value, and has
-relatively narrow distributions.
+method (columns) for three different known relationship values (rows
+from bottom to top): 0, 0.25, and 0.5}
+\end{figure}
 
@@
 
-For both empirical data sets, the direction and magnitude of change
-(i.e.~\(\beta_1\) coefficients) are generally in agreement. Size spectra
-parameters consistently increase (become flatter) in the AMD data (Fig.
-6A). However, the magnitude of the change across the AMD gradient is
-weakest using the MLE method, and the width of the estimates (\(\pm\) 1
-SD) is larger in the binning methods, particularly the ELBn method (Fig.
-6B). Likewise, the size spectra parameters consistently increase (become
-steeper) with increasing temperature across the NEON sites (Fig. 6C).
-The magnitude of the relationships is weakest in the MLE method,
-although the uncertainty around the estimates are similar (Fig. 6D).
+Both empirical data sets yielded similar results to the simulated data;
+the direction of the coefficients (i.e., \(\beta_1\) coefficients) were
+the same among methods, with positive slopes across the AMD gradient
+(Fig. 6A) and negative slopes across the temperature gradient. However,
+the magnitude of change differed between methods. The AMD slopes ranged
+from \textasciitilde0.62 with MLE to \textasciitilde0.78 with NAS
+(Figure 6b), while temperature slopes ranged from -0.0058 with MLE to
+-0.0019 with ELBn (Figure 6d). As with simulated data, slope uncertainty
+(\(\pm\) 1 SD) was larger in the binning methods, particularly the ELBn
+method (Fig. 6B). Likewise, the size spectra parameters consistently
+increase (become steeper) with increasing temperature across the NEON
+sites (Fig. 6C).
 
@@
 estimate the same sign of the relationship, but the estimates from the
-binning methods are generally greater than the MLE estimates.
-\textbf{NOTE} I need to re-plot these in ggplot and and stitch together.
-Currently stitching pre-made png files, I think that's why it looks so
-funky.}
-\end{figure}
-
-\newpage
-
-The \(\beta_1\) coefficient estimates for the AMD data are all positive
-and of relatively similar magnitudes across methods. The MLE estimate
-has the smallest magnitude and SD (0.063 \(\pm\) 0.011 SD), The ELBn
-method is slightly larger and has a wider SD : (0.072 \(\pm\) 0.018).
-Finally, the NAS method has the largest estimate and an intermediate
-standard deviation (0.076 \(\pm\) 0.013). The range of the gradient in
-the AMD data is relatively small (9.5), but the absolute change in the
-size spectra parameter (\(\lambda\)) across the AMD gradient vary by
-\textasciitilde0.13 units depending on the method used (range of
-absolute change: 0.59 to 0.72).
-
-Similar to the AMD results, the \(\beta_1\) coefficient estimates for
-the NEON data are all negative and have similar magnitudes across
-methods. The MLE estimate is once again the smallest (-0.006 \(\pm\)
-0.001 SD). The NAS and ELBn methods are larger at -0.004 \(\pm\) 0.002,
-and -0.002 \(\pm\) 0.002, respectively. Although the estimates are
-similar and have a small magnitude, the range of the gradient in this
-data is larger (\textasciitilde30\(\circ\)C), meaning that the absolute
-change in \(\lambda\) estimates varies by 0.11 units depending on the
-methods used (range of absolute change across gradient: 0.06 to 0.17).
+binning methods are generally greater than the MLE estimates.}
+\end{figure}
+
+\newpage
 
@@
 studied in a wide range of taxa inhabiting both terrestrial and aquatic
-ecosystems (reviewed by White et al. (2007)). Empirical patterns are
-remarkably consistent, and can be explained by the metabolic theory of
-Ecology (Brown et al. 2004). Measuring parameters describing the decline
-in abundance with increasing body size in communities is being done with
-increasing frequency across ecology. Previous work has investigated the
-accuracy and inherent biases associated with different estimation
-methods. However, how these inaccuracies and biases compound across
-environmental gradients remains uncertain, making it difficult to detect
-variation in size spectra parameters across environmental gradients with
-confidence. Here, we sampled body sizes from known distributions with
-varying parameters (\(\lambda\)) and estimated the coefficient of the
-relationship (\(\beta_{env}\)) across a hypothetical gradient in order
-to assess how the results would vary depending on the method used.
-Likewise, we compared how the interpretation of previously published
-results could change depending on the methodology used to estimate the
-relationship.
-
-Generally, the estimate of the size spectra parameter from the MLE
-method was always closer to the known \(\lambda\) value, regardless of
-how steep or shallow the \(\lambda\) exponent was. This is in agreement
-with Edwards et at. (2017) which found that the MLE estimate more
-consistently recovered known \(\lambda\) values ranging from -2.5 to
--0.5. We further this work by discovering that the magnitude of the
-error associated with the two binning methods increases as the
-\(\lambda\) value moves away from \(\sim\) -1.
-
-It is unsurprising that all methods performed less well in estimating
-steep values of \(\lambda\) (i.e., -2.5). The rapid decline of large
-body sizes makes it unlikely to sample any individuals with a large body
-size. Furthermore, the binning methods inherently ``lose'' data by
-aggregating individual observations into a single bin.
-
-Likewise, the MLE method also resulted in the most accurate
-\(\beta_{env}\) estimates for the hypothetical relationship in simulated
-data regardless of the \(\lambda\) ``window'' investigated. Furthermore,
-the variation around the estimate was always smallest with the MLE
-method. However, the variation in the MLE estimate did increase with
-steeper \(\lambda\) ``windows.'' This is not surprising, however because
-steep relationships like this make it very unlikely to sample the
-largest body sizes, and capturing the ``true'' relationships without the
-full range of body sizes present is more difficult (Edwards et
-al.~2017). Despite this, the MLE method is nearly always the best option
-to use, likely because all the data are used in the analysis.
+ecosystems (reviewed by {[}Brown (1995); GastonBlackburn2000; White et
+al. (2007){]}). Empirical data shows generally consistent patterns and
+can be explained by the metabolic theory of ecology (Brown et al. 2004).
+Measuring parameters describing the decline in abundance with increasing
+body size in communities is being done with increasing frequency across
+ecology. Previous work has investigated the accuracy and inherent biases
+associated with different estimation methods (White et al. 2007, Edwards
+et al. 2017, Edwards et al. 2020). However, how these inaccuracies and
+biases compound across environmental gradients remains uncertain, making
+it difficult to detect variation in size spectra parameters across
+environmental gradients with confidence. The most important outcome of
+our work is that binning methods not only generate biased \(\lambda\)
+values for individual data sets, but that bias carries over to affect
+the parameters of subsequent regressions that use those \(\lambda\)'s as
+response variables. This makes it challenging to understand how
+\(\lambda\) varies in response to environmental gradients if binning is
+used to estimate size spectra exponents.
 
@@
 of 3 grams, the midpoint of that bin. Likewise, a single abundance value
-is taken for this bin, despite that fact that there is almost certainly
+is taken for each bin, despite that fact that there is almost certainly
 variation in the abundance of individuals that weigh \textasciitilde2,
-\textasciitilde3, or \textasciitilde4 grams. By homogenizing the data in
-this way, it seems that binning methods will always produce noisier
-results than MLE from the binning process alone, since binning is akin
-to deleting information that the model could otherwise use. One of the
-benefits of using MLE is that all data points are retained for use
-within the model.
-
-The variation in estimated relationships across the empirical datasets
-varied by \textasciitilde0.1 units depending on the method used. At
-first glance this may not appear to be a concerning magnitude of
-variation. However, when one considers the range of variation reported
-in many seminal works on variation in size spectra relationships is
-\textasciitilde0.1-0.2 (SI Table \emph{xx}), it becomes clear how
-important an error in estimates of this magnitude is.
-
-Despite the differences in magnitude of the empirical relationship
-coefficients, they were generally in the correct direction and of a
-similar magnitude. This suggests that previously reported significant
-changes in size spectra parameters across environmental gradients and in
-experimental manipulations are plausible. Given that all of the data
-within a study is treated identically, the over all change in size
-spectra parameters is likely reasonable. However, the biases and
-inconsistencies in relationship estimates presented here suggest that it
-would be difficult if not impossible to directly compare the relative
-changes across different published studies which use different methods.
-However, the relatively systematic error of \(\lambda\) estimates across
-known exponents suggests that it may be possible to correct previous
-estimates onto the same scale (Figure 2).
-
-Despite this possibility, publication of individual body size data with
-future studies of size spectra relationships would greatly aid in our
-ability to generalize changes to this fundamental aspect of community
-organization across spatiotemporal scales and in response to
-environmental conditions.
+\textasciitilde3, or \textasciitilde4 grams. Moreover, the number of
+bins that can be produced by any dataset is limited most often to less
+than six. That means a subsequent log-log regression used to estimate
+\(\lambda\) would contains n = 6 data points, even though there likely
+100's or 1000's of individual body sizes available in a typical body
+size data set. By homogenizing the data in this way, binning methods
+produce noisier results than MLE, since binning is akin to deleting
+information that the model could otherwise use. By contrast, the MLE
+uses all the individual body size data to directly estimate \(\lambda\),
+meaning that it not only produces more accurate estimates, but does so
+with less certainty than binning, even when the underlying data sets are
+identical.
+
+The error in estimating individual \(\lambda\)'s was compounded when
+estimating linear models with those \(\lambda\)'s as response variables.
+The MLE method also resulted in the most accurate \(\beta_{env}\)
+estimates. Furthermore, variation around the estimate was always
+smallest with the MLE method.
+
+Although there were differences in magnitude of the empirical
+relationship coefficients, they were generally in the correct direction
+and of a similar magnitude. This suggests that previously reported
+significant changes in size spectra parameters across environmental
+gradients and in experimental manipulations are plausible. Given that
+all of the data within a study is treated identically, the over all
+change in size spectra parameters is likely reasonable. However, the
+biases and inconsistencies in relationship estimates presented here
+suggest that it would be difficult if not impossible to directly compare
+the relative changes across different published studies which use
+different methods.
+
+The publication of individual body size data with future studies of size
+spectra would greatly aid in our ability to generalize changes to this
+fundamental aspect of community organization across spatiotemporal
+scales and in response to environmental conditions.
 
@@
 producing MLE estimates of size spectra parameters is a relatively easy
-task. Therefore, we recommend using this in all future studies of size
-spectra relationships.
+task. Therefore, we recommend using it in all future studies of size
+spectra relationships rather than binning.
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Brown1995}{}%
+\leavevmode\hypertarget{ref-Andersen2006a}{}%
+Andersen, K. H., and J. E. Beyer. 2006. Asymptotic {Size Determines
+Species Abundance} in the {Marine Size Spectrum}. The American
+Naturalist 168:54--61.
+
+\leavevmode\hypertarget{ref-brown1995}{}%
 Brown, J. H. 1995. Macroecology. {University of Chicago Press},
@@
 
-\leavevmode\hypertarget{ref-dossenaWarmingAltersCommunity2012}{}%
+\leavevmode\hypertarget{ref-Damuth1981}{}%
+Damuth, J. 1981. Population density and body size in mammals. Nature
+290:699--700.
+
+\leavevmode\hypertarget{ref-Damuth1991}{}%
+Damuth, J. 1991. Of size and abundance. Nature 351:268--269.
+
+\leavevmode\hypertarget{ref-Damuth1998}{}%
+Damuth, J. D. 1998. Common rules for animals and plants. Nature
+395:115--116.
+
+\leavevmode\hypertarget{ref-dossena2012}{}%
 Dossena, M., G. Yvon-Durocher, J. Grey, J. M. Montoya, D. M. Perkins, M.
@@
 
-\leavevmode\hypertarget{ref-Edwards2017}{}%
+\leavevmode\hypertarget{ref-edwards2017}{}%
 Edwards, A. M., J. Robinson, M. Plank, J. Baum, and J. Blanchard. 2017.
@@
 Methods in Ecology and Evolution 8:57--67.
+
+\leavevmode\hypertarget{ref-evans2022}{}%
+Evans, T. M., Z. S. Feiner, L. G. Rudstam, D. M. Mason, J. M. Watkins,
+E. D. Reavie, A. E. Scofield, L. E. Burlakova, A. Y. Karatayev, and W.
+G. Sprules. 2022. Size spectra analysis of a decade of {Laurentian Great
+Lakes} data. Canadian Journal of Fisheries and Aquatic Sciences
+79:183--194.
 
@@
 
-\leavevmode\hypertarget{ref-mcgarveySeasonalComparisonCommunitylevel2018}{}%
+\leavevmode\hypertarget{ref-mcgarvey2018}{}%
 McGarvey, D. J., and A. J. Kirk. 2018. Seasonal comparison of
@@
 
-\leavevmode\hypertarget{ref-Perkins2018}{}%
+\leavevmode\hypertarget{ref-Nee1991}{}%
+Nee, S., A. F. Read, J. J. D. Greenwood, and P. H. Harvey. 1991. The
+relationship between abundance and body size in {British} birds. Nature
+351:312--313.
+
+\leavevmode\hypertarget{ref-perkins2018}{}%
 Perkins, D. M., I. Durance, F. K. Edwards, J. Grey, A. G. Hildrew, M.
@@
 
-\leavevmode\hypertarget{ref-Pomeranz2022}{}%
+\leavevmode\hypertarget{ref-pomeranz2022}{}%
 Pomeranz, J. P. F., J. R. Junker, and J. S. Wesner. 2022. Individual
@@
 
-\leavevmode\hypertarget{ref-pomeranzAnthropogenicMiningAlters2019}{}%
+\leavevmode\hypertarget{ref-pomeranz2019}{}%
 Pomeranz, J. P. F., H. J. Warburton, and J. S. Harding. 2019a.
@@
 
-\leavevmode\hypertarget{ref-sprulesSurfingBiomassSize2015}{}%
+\leavevmode\hypertarget{ref-Sheldon1972}{}%
+Sheldon, R. W., and S. R. Kerr. 1972. The {Population Density} of
+{Monsters} in {Loch Ness1}. Limnology and Oceanography 17:796--798.
+
+\leavevmode\hypertarget{ref-sprules2015}{}%
 Sprules, and Barth. 2015. Surfing the biomass size spectrum: Some
@@
 \centering
+\includegraphics{figures/PLB_small_m_est_b_density.png}
+\caption{Distribution of estimated \$\textbackslash lambda\$ coefficient
+for five sites across a hypothetical gradient with known values(dashed
+line). Range of body sizes is smaller than main anaysis and ranges from
+1, to 100.}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics{figures/PLB_small_m_relationship_density.png}
@@
 
-\newpage
-
-\hypertarget{range-of-body-sizes-1-to-100}{%
-\subsubsection{Range of body sizes = 1 to
-100}\label{range-of-body-sizes-1-to-100}}
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_small_m_est_b_density.png}
-\caption{Distribution of estimated \$\textbackslash lambda\$ coefficient
-for five sites across a hypothetical gradient with known values(dashed
-line). Range of body sizes is smaller than main anaysis and ranges from
-1, to 100.}
-\end{figure}
-
-\newpage
-
-\begin{figure}
-\centering
-\includegraphics{figures/PLB_small_m_relationship_density.png}
-\caption{Relationship estimate}
-\end{figure}
-
 \hypertarget{comparison-with-other-published-estimates}{%
@@
 spectra slopes (or exponents) in empirical studies. It is unclear how to
-directly compare estimates of the slope with different methods.However,
+directly compare estimates of the slope with different methods. However,
 the published estimates here range from \textasciitilde0.1 to 0.2 across
@@
 
-\begin{longtable}[]{@{}lrrlrrlll@{}}
-\toprule
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Citation\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-\(\beta_1\) Difference\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-Error\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Error Type\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
-\(\beta_{low}\)\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
-\(\beta_{high}\)\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-System\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Driver\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-Method\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Pomeranz et al.~2021\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.12\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--1.37\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--1.25\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-Streams\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Temperature\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-MLE\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Yvon Durocher et al.~2011 (Community)\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.13\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.79\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.92\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-FW mesocosms\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Temperature\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-10 ELB, not normalized\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Dossena et al.~2012 (October)\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.14\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-0.070\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-se\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-FW mesocosms\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Temperature\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-6 ELB, not normalized\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-O'Gorman et al.~2017\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.15\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.85\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.70\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-Streams\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Temperature\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-ln (mean species abundance) \textasciitilde{} ln(mean species
-mass)\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Martinez et al.~2016\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.15\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--1.11\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.96\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-Streams\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Land Use\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-6 ELB, not normalized\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Yvon Durocher et al.~2011 (Phytoplankton)\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.19\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.46\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--0.65\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-FW mesocosms\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Temperature\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-10 ELB, not normalized\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-McGarvey and Kirk 2018\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.19\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--1.81\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
--1.62\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-Streams\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Seasonality\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-NAS, Log2 bins\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.23}}\raggedright
-Dossena et al.~2012 (April)\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.11}}\raggedleft
-0.21\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.03}}\raggedleft
-0.064\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-se\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.08}}\raggedleft
-NA\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.07}}\raggedright
-FW mesocosms\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.06}}\raggedright
-Temperature\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 8\tabcolsep) * \real{0.28}}\raggedright
-6 ELB, not normalized\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
+Supplementary table comparing with other published results will be
+submitted as a separate file for formatting.
 

### 2023-02-23 1b_sim_sliding_windows.R

@@ -42,6 +42,7 @@ est_lambda <- sim_result(n = 1000,
 plot_sim(est_lambda)
 
 # match colors, jitter points, increase alpha, try GAM/loess, 
+# est_lambda <- readRDS("data_sim/est_lambda_dat.rds")
   
 (known_est_b_line <- est_lambda %>%
   mutate(Model = factor(name,
@@ -54,18 +55,22 @@ plot_sim(est_lambda)
            y = estimate, 
            color = name)) +
   geom_point(
-    position = position_jitter(w = 0.025, h = 0),
+    position = position_jitter(w = 0.005, h = 0),
     alpha = 0.25
   ) +
   stat_smooth(method = "lm",
               se = FALSE, 
               geom = "line",
-              size = 1.75) +
-  geom_abline(linetype = "dashed",
+              size = 1) +
+  geom_abline(linetype = "solid",
               size = 1) +
   scale_color_manual(
     values = c("#FF914A", "#019AFF",  "#FF1984" )) +
-  theme_bw())
+  theme_bw() +
+    facet_grid(rows = vars(Model)) +
+    labs(y = "\u03bb estimate",
+         x = "Known \u03bb") +
+    theme(legend.position = "none"))
 
 ggsave(plot = known_est_b_line,
        filename = "figures/known_est_b_line.png")
@@ -76,7 +81,7 @@ saveRDS(est_lambda, file = paste0("data_sim/",
 
 # 2 lambda windows ####
 
-# 2A ####
+# 2A steep ####
 # steep lambda [-1.5, -2.5]
 # for reproducibility
 set.seed(2806)
@@ -101,7 +106,7 @@ saveRDS(steep_lambda, file = paste0("data_sim/",
                                   substitute(steep_lambda),
                                   "_dat.rds"))
 
-# 2b ####
+# 2b medium ####
 # medium lambda [-1, -2]
 # for reproducibility
 set.seed(2806)
@@ -126,7 +131,7 @@ saveRDS(med_lambda, file = paste0("data_sim/",
                                     substitute(med_lambda),
                                     "_dat.rds"))
 
-# 2c ####
+# 2c shallow ####
 # shallow lambda [-0.5, -1.5]
 # for reproducibility
 set.seed(2806)
@@ -155,7 +160,7 @@ saveRDS(shallow_lambda, file = paste0("data_sim/",
[…]

### 2023-02-23 2b_integrated plots.R

@@ -165,9 +165,13 @@ est_lambda %>%
              linetype = "dashed") +
   labs(
     x = "Lambda estimate") +
-  facet_wrap(~known_b, scales = "free")
+  facet_wrap(~known_b, scales = "free") +
+  scale_y_discrete(breaks=c("MLE",
+                            "ELBn", 
+                            "NAS"),
+                   labels= NULL)
 
-ggsave(filename = "figures/est_lambda_est_b_density_fixed_x.png",
+ggsave(filename = "figures/est_lambda_est_b_density.png",
        units = "in", 
        height = 6,
        width = 6)

### 2023-02-23 old_scripts/Overview.Rmd
massive reorganization before submission. made "old_scripts/" to make sure didn't accidentally delete anything important.

--- 2022-05-24 Overview.Rmd
@@
   + b = c(-1.5, -2, -2.5); b = c(-1.9, -2, -2.1)
+  + hypothetical environmental gradient (x) = c(-1, 0, 1) paired with b parameters above
+  + Repeat simulation 1000 times (reps)
   + xmin and xmax based on macroinvertebrate sizes
   + Also use Edwards values to test sensitivity
-  + Simulate from other distributions? rlnorm(), rweibull() etc.?
+  + Simulate from other distributions? rlnorm(), rweibull() etc.? --> Not done yet. Maybe leave out untill reviewers ask for it??
 2. Analyze simulated data
   + MLE methods from Edwards et al. 2018
-  + log2 bins - Pomeranz, Blanchard, etc
-  + ELB - 6, 8, 10 bins. Woodward's group generally uses 6 equal bins
+  + NAS = log2 bins, normalized - Pomeranz, Blanchard, etc
+  + ELBn = 6 equal log bins, normalized. Woodward's group generally uses 6 equal bins. 
   + Others from Edwards' 8 methods?
@@
   + what is *relative* difference in estimates? i.e. do all methods result in a relative change of 0.5 between data "groups"?
+  + Do methods accurately detect no relationship across "sites"?
+  + Plot estimated CSS parameter for each rep
+  + OLS regression for each rep: est_css_param ~ env_gradient
+4. Sensitivity
+  + Change range of "x"?
+  + Change range of "m" body sizes
+  + More "sites"
+5. Empirical data - do conclusions of previous studies vary when using different methods?
+  + AMD data from Pomeranz et al. 2018
+  + NEON Data from Pomeranz et al. 2022
+
+## Results
+
+Figures are supposed to come after this, not before... Latex??
+
+![Fig 1.
+Distribution of CSS parameter.
+Vertical line is the known parameter wich describes the bounded power law distribution from which the body size estimates were sampled.](figures/PLB_sim_est_b_density.png)
+\newpage
+
+![Fig 2.
+OLS Regression lines for each rep across the hypothetical environmental gradient.
+Body sizes were sampled such that the CSS parameter should change by 0.5 units across each unit of x.](figures/PLB_sim_main.png)
+
+\newpage
+
+![Fig 3.
+Density distribution for the estimated relationship across the hypothetical environmental gradient.
+the MLE method is centered at 0.5, which is the "known" relationship simulated in the main analysis.
+Both of the binning methods underestimate the hypothetical relationship.](figures/PLB_sim_relationship_density.png)
+
+\newpage
+
+### No relationship
+
+![Fig 4.
+Distribution of CSS parameter.
+Vertical line is the known parameter wich describes the bounded power law distribution from which the body size estimates were sampled.](figures/PLB_static_b_est_b_density.png)
+\newpage
+
+![Fig 5.
+OLS Regression lines for each rep across the hypothetical environmental gradient.
+Body sizes were sampled such that the CSS parameter should not change across values of x.](figures/PLB_static_b_main.png)
+
+\newpage
+
+![Fig 6.
+Density distribution for the estimated relationship across the hypothetical environmental gradient.
+All methods are centered at 0, which is the "known" relationship simulated in this analysis.
+Both of the binning methods have a much wider distribution than the MLE method.](figures/PLB_static_b_relationship_density.png)
+
+### Small variation of CSS over environmental gradient
+
+\newpage
+
+![Fig 7.
+Distribution of CSS parameter.
+Vertical line is the known parameter wich describes the bounded power law distribution from which the body size estimates were sampled.](figures/PLB_sim_small_est_b_density.png)
+
+\newpage
+
+![Fig 8.
+OLS Regression lines for each rep across the hypothetical environmental gradient.
+Body sizes were sampled such that the CSS parameter should change by 0.1 units across values of x.](figures/PLB_sim_small_main.png)
+\newpage
+
+![Fig 9.
+Density distribution for the estimated relationship across the hypothetical environmental gradient.
+All methods are centered at -0.1, which is the "known" relationship simulated in this analysis.
+Both of the binning methods have a much wider distribution than the MLE method, and occasionally result in a positive relationship.](figures/PLB_sim_small_relationship_density.png)
+\newpage
+
+## Empirical data
+
+### Acid Mine Drainage (AMD) gradient
+
+The data of Pomeranz et al. 2018 was re-analyzed using the three methods presented here. The goal here was to see if the main conclusions reached would vary based on methodologies.
+
+![Fig 10. CSS parameter estimates are positively related to AMD stress. The conclusions are in agreement with Pomeranz et al. 2018, and are not dependent on methodology used.](figures/AMD_plot.png)
+
+\newpage
+
+![Fig 11. Mean (+- SD) estimated slope coefficient relating change in CSS parameter across gradient of AMD stress.](figures/amd_relationship.png)
+
+\newpage
+
+### Environmental Temperature Gradient
+
+![Fig 10.
+CSS parameter estimates are weakly negatively related to AMD stress.
+The conclusions are in agreement with Pomeranz et al.
+2022, and are not dependent on methodology used.](figures/NEON_plot.png)
+
+\newpage
+
+![Fig 11. Mean (+- SD) estimated slope coefficient relating change in CSS parameter across gradient of environmental temperature across the NEON sites.](figures/neon_relationship.png)
+
+## Conclusions
++ MLE offers the best estimates of individual CSS parameters, and the best estimate of their relative change across hypothetical gradients
++ Normalized binning methods generally overestimate CSS parameters (i.e., "shallower" slopes than expected)
++ Normalized Binning methods (ELBn, NAS) *generally* correctly estimate the relationship across hypothetical and empirical gradients. 
++ Need to determine if it's possible to correctly propagate error / systematic bias in estimates, but the present study indicates that comparing the *relative* changes across gradients in different studies should be reasonable.
+
+## SI
+
+Sensitivity analyses. Need to come back to these and add captions and think about implications
+
+### Large X (-100, 100)
+
+![Large x-range.](figures/PLB_large_x_est_b_density.png)
+
+![Large x-range.](figures/PLB_large_x_main.png)
+
+\newpage
+
+![Large x-range.](figures/PLB_large_x_relationship_density.png)
+
+\newpage
+
+### Small body size range (m) (1, 100)
+![Small (M) x-range.](figures/PLB_small_m_est_b_density.png)
+
+\newpage
+
+![Small (M) x-range.](figures/PLB_small_m_main.png)
+
+![Small (M) x-range.](figures/PLB_small_m_relationship_density.png)
+\newpage
+
+![10 sites.](figures/PLB_10_sites_est_b_density.png)
+
+\newpage
+
+![10 sites.](figures/PLB_10_sites_main.png)
+
+\newpage
+
+![10 sites.](figures/PLB_10_sites_relationship_density.png)

### 2023-02-23 1a_simulations_sensitivity.R

@@ -55,43 +55,9 @@ saveRDS(PLB_sim, file = paste0("data_sim/",
 plot_sim(PLB_sim)
 #rm(PLB_sim)
 
-# PLB no relationship
-# for reproducibility
-set.seed(2806)
-
-PLB_static_b <- sim_result(n = 1000,
-                           b = rep(-2, length(env_gradient)),
-                           env_gradient = env_gradient,
-                           rep = rep, 
-                           m_lower = m_lower,
-                           m_upper = m_upper,
-                           distribution = "PLB")
-saveRDS(PLB_static_b, file = paste0("data_sim/",
-                                    substitute(PLB_static_b),
-                                    "_dat.rds"))
-plot_sim(PLB_static_b)
-rm(PLB_static_b)
 
 
-# PLB "small" relationship
-# lambda varies from -1.9 to -2.1
-# for reproducibility
-set.seed(2806)
 
-PLB_sim_small <- sim_result(n = 1000,
-                            b = seq(from = -1.9, to = -2.1, length.out = 5),
-                            env_gradient = env_gradient,
-                            rep = rep, 
-                            m_lower = m_lower,
-                            m_upper = m_upper,
-                            distribution = "PLB")
-
-saveRDS(PLB_sim_small, file = paste0("data_sim/",
-                               substitute(PLB_sim_small),
-                               "_dat.rds"))
-
-plot_sim(PLB_sim_small)
-rm(PLB_sim_small)
 
 # sensitivity, supplementary information ----------------------------------
 
@@ -249,21 +215,6 @@ saveRDS(PLB_sim_n500, file = paste0("data_sim/",
 #rm(PLB_sim_n500)
 
 
-
-# re run ------------------------------------------------------------------
-
-
-# Re run this ####
-# threw an error in group 517
-# tomorrow, attempt to use purrr::possibly() as a wrapper
-# should hopefully throw an NA instead of stopping the run...
-# https://purrr.tidyverse.org/reference/safely.html 
-
-# seems to throw lots or errors or have too few bins
-# changing to n = 200 on 6/13
-
-## manuscript --> sample size not the main objective, but anecdotally seems that n= 100 is too few observations. 
-
 n = 200 
 # for reproducibility
 set.seed(2806)
@@ -275,51 +226,6 @@ PLB_sim_n200 <- sim_result(n = n,
[…]

### 2023-02-23 old_scripts/deleted_from_1_sim.R

@@ -1,0 +1,123 @@
+# old stuff removed from 1_simulations.R
+# PLB no relationship
+# for reproducibility
+set.seed(2806)
+
+PLB_static_b <- sim_result(n = 1000,
+                           b = rep(-2, length(env_gradient)),
+                           env_gradient = env_gradient,
+                           rep = rep, 
+                           m_lower = m_lower,
+                           m_upper = m_upper,
+                           distribution = "PLB")
+saveRDS(PLB_static_b, file = paste0("data_sim/",
+                                    substitute(PLB_static_b),
+                                    "_dat.rds"))
+plot_sim(PLB_static_b)
+rm(PLB_static_b)
+
+
+# PLB "small" relationship
+# lambda varies from -1.9 to -2.1
+# for reproducibility
+set.seed(2806)
+
+PLB_sim_small <- sim_result(n = 1000,
+                            b = seq(from = -1.9, to = -2.1, length.out = 5),
+                            env_gradient = env_gradient,
+                            rep = rep, 
+                            m_lower = m_lower,
+                            m_upper = m_upper,
+                            distribution = "PLB")
+
+saveRDS(PLB_sim_small, file = paste0("data_sim/",
+                                     substitute(PLB_sim_small),
+                                     "_dat.rds"))
+
+plot_sim(PLB_sim_small)
+rm(PLB_sim_small)
+
+# sim n200
+
+# re run ------------------------------------------------------------------
+
+
+# Re run this ####
+# threw an error in group 517
+# tomorrow, attempt to use purrr::possibly() as a wrapper
+# should hopefully throw an NA instead of stopping the run...
+# https://purrr.tidyverse.org/reference/safely.html 
+
+# seems to throw lots or errors or have too few bins
+# changing to n = 200 on 6/13
+
+## manuscript --> sample size not the main objective, but anecdotally seems that n= 100 is too few observations. 
+
+
+
+# addedd purrr::possibly() to custom functions
+# double check this simulation and make sure it looks right
+# also double check the number of iterations it thre out (only one window popped up while I was watching...)
+# make sure plot_sim() function still works with NA's... may return a list???
+# may need to modify code in subsequent calculations
[…]

### 2023-07-18 1a_sim_sliding_windows.R
update script names to match readme

@@ -31,7 +31,10 @@ rep = 1000
 m_lower = 0.0026
 m_upper = 1.2e3
 
-est_lambda <- sim_result(n = 1000,
+# number of body sizes r=to sample, n
+n = 999
+
+est_lambda <- sim_result(n = n,
                       b = lambda,
                       env_gradient = env_gradient,
                       rep = rep, 
@@ -92,7 +95,7 @@ env_gradient = seq(from = -1, to = 1, length.out = 5)
 lambda
 env_gradient
 
-steep_lambda <- sim_result(n = 1000,
+steep_lambda <- sim_result(n = n,
                          b = lambda,
                          env_gradient = env_gradient,
                          rep = rep, 
@@ -117,7 +120,7 @@ env_gradient = seq(from = -1, to = 1, length.out = 5)
 lambda
 env_gradient
 
-med_lambda <- sim_result(n = 1000,
+med_lambda <- sim_result(n = n,
                            b = lambda,
                            env_gradient = env_gradient,
                            rep = rep, 
@@ -141,7 +144,7 @@ lambda = seq(from = -0.5, to = -1.5, length.out = 5)
 lambda
 env_gradient
 
-shallow_lambda <- sim_result(n = 1000,
+shallow_lambda <- sim_result(n = n,
                            b = lambda,
                            env_gradient = env_gradient,
                            rep = rep, 
@@ -170,7 +173,7 @@ lambda = seq(from = -1.5, to = -1.5, length.out = 5)
 lambda
 env_gradient
 
-relationship_0 <- sim_result(n = 1000,
+relationship_0 <- sim_result(n = n,
                              b = lambda,
                              env_gradient = env_gradient,
                              rep = rep, 
@@ -196,7 +199,7 @@ lambda = seq(from = -1.25, to = -1.75, length.out = 5)
 lambda
 env_gradient
 
-relationship_025 <- sim_result(n = 1000,
+relationship_025 <- sim_result(n = n,
                              b = lambda,
                              env_gradient = env_gradient,
                              rep = rep, 
@@ -221,7 +224,7 @@ lambda = seq(from = -1, to = -2, length.out = 5)
 lambda
 env_gradient
 
-relationship_05 <- sim_result(n = 1000,
+relationship_05 <- sim_result(n = n,
                                b = lambda,
                                env_gradient = env_gradient,
                                rep = rep, 

### 2023-07-18 1b_simulations_sensitivity.R

@@ -35,13 +35,19 @@ env_gradient = seq(from = -1, to = 1, length.out = 5)
 lambda
 env_gradient
 
+# number of replications to perform
 rep = 1000
+
 # body sizes in NEON data set varies from 0.0026 to 1.2e4
 # setting sample range based on that minimum, but reducing max by order of magnitude
 m_lower = 0.0026
 m_upper = 1.2e3
 
-PLB_sim <- sim_result(n = 1000,
+# number of body sizes to sample, n
+# 999 so as not to confuse with 1000 reps
+n = 999
+
+PLB_sim <- sim_result(n = n,
                       b = lambda,
                       env_gradient = env_gradient,
                       rep = rep, 
@@ -65,7 +71,7 @@ plot_sim(PLB_sim)
 # for reproducibility
 set.seed(2806)
 
-PLB_large_x <- sim_result(n = 1000,
+PLB_large_x <- sim_result(n = n,
                           b = lambda,
                           env_gradient = seq(from = -100,
                                              to = 100,
@@ -86,7 +92,7 @@ rm(PLB_large_x)
 # for reproducibility
 set.seed(2806)
 
-PLB_small_m <- sim_result(n = 1000,
+PLB_small_m <- sim_result(n = n,
                           b = lambda,
                           env_gradient = env_gradient,
                           rep = rep, 
@@ -146,7 +152,11 @@ rm(PLB_10_sites)
 
 # sample size of body sizes -----------------------------------------------
 # varying the sample size of simulated data
-# 10,000, 5,000, 500, 100
+# 10,000, 5,000, 500, 200
+  # originally tried n = 100, but some of the 
+  # iterations could not fit size spectra methods
+  # Maybe an indication that n=200 is an 
+  # approximate minimial sample size?
 
 #rep = 1000
 n = 10000 

### 2023-07-18 2b_integrated_plots.R

@@ -116,7 +116,10 @@ ggsave(filename = "figures/vary_beta_plot.png",
        width = 6)
 
 
+
+
 rel_data_summary = rel_data %>% calc_relationship_estimate(.)
+
 saveRDS(rel_data_summary, file = "data_sim/rel_data_summary.rds")
 
 rel_data_summary %>%
@@ -146,6 +149,33 @@ ggsave(filename = "figures/vary_beta_density_plot.png",
        units = "in", 
        height = 6,
        width = 6)
+
+## export one plot as pdf for inkscape practice
+rel_data_summary %>%
+  filter(known_relationship == -0.5) %>%
+  mutate(Model = factor(name,
+                        levels = 
+                          c("MLE",
+                            "ELBn", 
+                            "NAS"))) %>%
+  ggplot(aes(x = estimate, 
+             y = Model,
+             fill = Model))+
+  stat_halfeye(.width = c(0.66, 0.95)) +
+  scale_fill_manual(
+    values = c("#019AFF",
+               "#FF914A",
+               "#FF1984" )) +
+  theme_bw() +
+  geom_vline(
+    aes(xintercept = known_relationship),
+    linetype = "dashed") +
+  labs(x = "Relationship estimate") +
+  facet_wrap(known_relationship~., 
+             ncol = 1) +
+  NULL 
+
+# export this figure using the plot window
 
 est_lambda %>%
   mutate(Model = factor(name,

### 2023-07-18 custom_functions.R

@@ -146,13 +146,9 @@ compare_slopes <- function(data, dw_range, rsp_var, ...){
   # AS_coefs<- coef(AS_lm)
   # AS_conf <- confint(AS_lm)
   
-  # equal logarithmic binning method
+  # equal logarithmic binning method, non-normalized
   
-  # breaks_log_6 <- exp(seq(floor(log(min(dw_range))),
-  #                         ceiling(log(max(dw_range))),
-  #                         length.out = 7))
-  # mid_bin_log_6 <- log10(breaks_log_6[floor(length(breaks_log_6)/2)])
-  # ELB6
+  # 6 equal logarthmic bins
    ELB <- bin_and_center(data, var = rsp_var, breaks = breaks_log_6)
   ELB$log_mids_center <- ELB$log_mids - mid_bin_log_6
   # ELB_lm <- lm(log_count~log_mids_center,
@@ -160,6 +156,8 @@ compare_slopes <- function(data, dw_range, rsp_var, ...){
   # # ELB_coef <- coef(ELB_lm)
   # # ELB_conf <- confint(ELB_lm)
   # ELB_out <- est_ci(ELB_lm)
+  
+  # ELB normalized
   ELBn_lm <- lm(log_count_corrected~log_mids_center,
               data = ELB)
   # ELBn_coef <- coef(ELBn_lm)
@@ -175,25 +173,9 @@ compare_slopes <- function(data, dw_range, rsp_var, ...){
                    #ELB_out,
                    ELBn_out,
                    mle_b))
-  slopes$name <- c("NAS", #"AS", 
+  slopes$name <- c("L2n", #"AS", 
                    #"ELB",
                    "ELBn", "MLE")
-  # slopes = data.frame(NAS = NAS_out,
-  #                     # NAS_25 = NAS_conf[2],
-  #                     # NAS_975 = NAS_conf[4],
-  #                     AS = AS_out,
-  #                     # AS_25 = AS_conf[2],
-  #                     # AS_975 = AS_conf[4],
-  #                     ELB6 = ELB_out,
-  #                     # ELB_25 = ELB_conf[2],
-  #                     # ELB_95 = ELB_conf[4],
-  #                     ELBn = ELBn_out,
-  #                     # ELBn_25 = ELBn_conf[2],
-  #                     # ELBn_975 = ELBn_conf[4],
-  #                     mle = mle_b
-  #                     # mle_25 = mle_b$minCI,
-  #                     # mle_975 = mle_b$maxCI
[…]

### 2023-07-18 Manuscript.Rmd
unstaged changes to manuscript before 1st submission

@@
 4.
-The uncertainty and variation in estimates when using binning methods is often greater than or equal to the variation previously published in experimental and observational studies, bringing into question the effect size of previously published results.
+The uncertainty and variation in estimates when using binning methods can be of similar magnitudes as the variation previously published in experimental and observational studies, bringing into question the effect size of previously published results.
 However, while the methods produced different slope estimates from previously published datasets, they were in qualitative agreement on the sign of those slopes (i.e., all negative or all positive).
@@
 
-We had Two primary objectives in this study: 1) to compare how well different methods estimate site-specific $\lambda$'s 2) To compare how different methods of estimating $\lambda$ impact subsequent regressions that use $\lambda$'s as a response variable.
+We had two primary objectives in this study: 1) to compare how well different methods estimate site-specific $\lambda$'s and 2) to compare how different methods of estimating $\lambda$ impact subsequent regressions that use $\lambda$'s as a response variable.
 We did this using both simulated and empirical data sets.
@@
 After simulating data, we used three different methods (described below) to estimate the value of $\lambda$ (maximum likelihood, equal logarithmic bins, and normalized abundance spectra).
-We repeated this process 1000 times (reps) and plotted the distribution of values obtained for each method.
+We repeated the process of data simulation and parameter estimation 1000 times (reps) and plotted the distribution of values obtained for each method.
 
@@
 
-### Equal Logarithmic Bins: ELBn
+## Equal Logarithmic Bins: ELBn
 
 In addition, to estimate the OLS slope parameter in log-log space, we used two common binning methods.
-.
 It is important to note that the log-log regression is estimating $\lambda$ as the regressionslope ($\beta_{1}$), and not $\lambda$ directly.
-In the interest of clarity, we will refer to all estimates of lambda as $\lambda$ regardless of the method used, and reserve the terms “parameters”, “intercept”, and “slope” $\beta_{env}$ to refer to the parameters of linear regression that uses the lambdas as a response variable.
+In the interest of clarity, we will refer to all estimates of lambda as $\lambda$ regardless of the method used, and reserve the terms “parameters”, “intercept”, and “slope” $\beta_{env}$ to refer to the parameters of linear regressions that use the lambdas as a response variable.
 
-For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes.
-This method has been used extensively in previous studies.
-For example see [@perkins2018; @martinez2016; @dossena2012].
+For the first binning method, we created six equal logarithmic bins covering the range of body sizes.
+This method has been used extensively in previous studies (e.g., see Perkins et al.
+[-@perkins2018], Martinez et al.
+[-@martinez2016], and Dossena et al.
+[-@dossena2012].
 In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes.
@@
 ### Log~2~ Bins: NAS
-The second method was similar to ELBn, but bins of log~2~ widths are used.
+The second binning method was similar to ELBn, but bins of log~2~ width are used.
 The number of bins is dependent on the range of body sizes present in the data.
-When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global min and max body size present.
+When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global minimum and maximum body size present.
 However, since the data here were simulated from a known size range, the number of bins for each site is identical.
-The count in each bin is normalized in the same way as described above.
-This method has been  used extensively in the literature e.g., [@jennings2004; @Jennings2002; @sprules2015; @pomeranz2019; @mcgarvey2018] and is referred to as the Normalized Abundance Spectrum (NAS).
+The count in each bin is normalized in the same way as described above for the ELBn approach.
+This method has been  used extensively in the literature (e.g., Jennings et al.
+[-@jennings2004; -@Jennings2002], Sprules and Barth [-@sprules2015], Pomeranz et al.
+[-@pomeranz2019], McGarvey and Kirk [-@mcgarvey2018]), and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 
-Where $N$ is the count in each bin, $M$ is the mid-point of each bin, and $\beta_{1}$ is considered to be the $\lambda_{estimate}$.
+Where $N$ is the count in each bin, $M$ is the mid-point of each bin, and $\beta_{1}$ is considered to be the $\lambda$ estimate.
 
@@
 
-Previous work has investigated the biases in $\lambda$ estimates when using different methods [-@edwards2017, @edwards2020, @White2007].
+Previous work has investigated the biases in $\lambda$ estimates when using different methods [@edwards2017; @edwards2020; @White2007].
 However, the focus of the present work is to investigate biases when estimating the relationship parameters across an environmental gradient.
@@
 
-$$\lambda_{ik} = \beta_{0} + \beta_{0}x_{ik}$$
+$$\lambda_{ik} = \beta_{0} + \beta_{1}x_{ik}$$
 
@@
 Once the body size data were simulated, we then reversed the procedure.
-First, we used the three methods (NAS, ELBn, MLE) to estimate $\lambda$ for each of the 5000 body size datasets.
-Then we fit a linear regression to those data with the predictor variable $x$ and compared the resulting intercept ($\beta_{0}$) and slope ($\beta_{1}$) parameter estimates to the known value.
+First, we used the three methods (NAS, ELBn, MLE) to estimate $\lambda$ for each of the 5000 body size data sets.
+Then, we fit a linear regression to those data with the predictor variable $x$ and compared the resulting intercept ($\beta_{0}$) and slope ($\beta_{1}$) parameter estimates to the known value.
 The main results presented here were not dependent on the range of x-values or the number of sites (Supplemental Information).
@@
 Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat.
-All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions.
-This data has been analyzed previously using size spectra methods as described in [@pomeranz2022].
-Detailed methods of the sampling collection and data processing methods can be in the macroinvertebrate DPI pubs found on the NEON website <https://www.neonscience.org/data-collection/macroinvertebrates>.
-Coefficient estimates ($\pm$ 1 standard deviation) are compared across methods.
+All individuals were identified and had their body lengths measured, and body mass was estimated using published length-weight regressions.
+This data has been analyzed previously using size spectra methods as described in Pomeranz et al.
+[-@pomeranz2022].
+Detailed methods of the sampling collection and data processing methods can be found in the macroinvertebrate data product information documents found on the NEON website <https://data.neonscience.org/data-products/DP1.20120.001>.
+Estimates of the slope coefficient ($\beta_1$ ) $\pm$ 1 SD, are compared across methods.
 This allows us to determine whether or not the main results would differ depending on the method used.
@@
 
-The data used in this manuscript is already publicly available [@Pomeranz2019] data dryad DOI <https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s>, [@NEON_Inverts2022].
+The empirical data used in this manuscript is already publicly available (Pomeranz et al.
+[-@Pomeranz2019]; data dryad DOI <https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s>, and  [@NEON_Inverts2022].
 R Scripts to reproduce the full simulation and analysis are available at: ([<https://github.com/Jpomz/detecting-spectra-differences>]).
@@
 On average, the 95% confidence intervals produced by the NAS and ELBn methods were ~2 times wider than those produced by MLE (Table 1), indicating greater consistency of estimates from MLE.
-Similarly, estimates of  $\lambda$ deviated from the true $\lambda$ by an average of 0.04 or 0.05 absolute units for the NAS and ELBn methods, more than 4 times higher than deviation (0.01) for the MLE (Table 1).
+Similarly, estimates of  $\lambda$ deviated from the true $\lambda$ by an average of 0.04 or 0.05 absolute units for the NAS and ELBn methods, more than four times higher than the deviation (0.01) observed for the MLE (Table 1).
 
 ![Figure 1.
-Distribution of Lambda estimates by method (color) from random samples of body sizes from bounded power law distributions with varying exponents (-2.5, -0.5).
+Distribution of Lambda estimates by method (color) from random samples of body sizes from bounded power law distributions with varying exponents (-2.5 to -0.5).
 The figure is facetted by the known lambda parameter (facet title) and is also shown as the dashed line in each facet.
 Note that the x-axis varies in each facet.](figures/est_lambda_est_b_density.png)
-
-```{r, echo=FALSE}
-t1 <- read.csv("tables/bias_table.csv")
-knitr::kable(t1,
-             col.names = c("Target",
-                           "Method",
-                           "n",
-                           "Median Range of 95% CI",
-                           "Median Absolute Deviation",
-                           "SD Absolute Deviation"),
-                           caption = "Table 1.
-Summary of three methods in recapturing the known values of lambda or the regression slopes simulated in this study.
-Performance is determined by comparinged by uncertainty (range of 95% CI's) and the absolute distance of the model estimates from the known estimates (median and sd of the difference).
-Values are summarized across all n = 9000 or 6000 simulated datasets.
-See figures for more specific comparisons.")
-
-```
 
@@
 The two binning method cross the 1:1 line, meaning they systematically underestimate shallow lambdas, and over estimate steeper lambdas.](figures/known_est_b_line.png)
+
+#### **Table 1**
+
+(*Submitted as additional file on biorxiv.*) Summary of three methods in recapturing the known values of lambda or the regression slopes simulated in this study.
+Performance is determined by comparinged by uncertainty (range of 95% CI's) and the absolute distance of the model estimates from the known estimates (median and SD of the difference).
+Values are summarized across all n = 9000 or 6000 simulated data sets.
+See figures for more specific comparisons.
 
@@
 However, the binning methods again systematically underestimated the true value of the slope by ~0.05 units (Figure 5).
-Likewise, uncertainty in the slope estimates was always greater in the binning methods, with the width of the distributions increases with stronger relationships across a hypothetical gradient.
+Likewise, uncertainty in the slope estimates was always greater in the binning methods, with the width of the distributions increasing with stronger relationships across a hypothetical gradient.
 By comparison, the MLE showed no evidence of bias and was always centered at the known value with relatively narrow variation.
@@
 
-The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by [@brown1995; GastonBlackburn2000; @White2007]).
+The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by Brown [-@brown1995], and White et al.
+[-@White2007]).
 Empirical data shows generally consistent patterns and can be explained by the metabolic theory of ecology [@Brown2004].
@@
 Previous work has investigated the accuracy and inherent biases associated with different estimation methods [@White2007; @edwards2017; @edwards2020].
-However, how these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
+However, the extent to which these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
 The most important outcome of our work is that binning methods not only generate biased $\lambda$ values for individual data sets, but that bias carries over to affect the parameters of subsequent regressions that use those $\lambda$’s as response variables.
@@
 Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies.
-However, aggregating individuals into logarithmic bins removes a large source of the variation within the data by collapsing body size variation into a single value within each bin.
+However, aggregating individuals into logarithmic bins removes a large amount of information within the data by collapsing body size variation into a single value within each bin.
 For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
@@
 By homogenizing the data in this way, binning methods produce noisier results than MLE, since binning is akin to deleting information that the model could otherwise use.
-By contrast, the MLE uses all the individual body size data to directly estimate $\lambda$, meaning that it not only produces more accurate estimates, but does so with less certainty than binning, even when the underlying data sets are identical.
+By contrast, the MLE uses all the individual body size data to directly estimate $\lambda$, meaning that it not only produces more accurate estimates, but does so with less uncertainty than binning, even when the underlying data sets are identical.
 
-The error in estimating individual $\lambda$’s was compounded when estimating linear models with those $\lambda$’s as response variables.
-The MLE method also resulted in the most accurate $\beta_{env}$ estimates.
-Furthermore, variation around the estimate was always smallest with the MLE method.
-
-Although there were differences in magnitude of the empirical relationship coefficients, they were generally in the correct direction and of a similar magnitude.
-This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
-Given that all of the data within a study is treated identically, the over all change in size spectra parameters is likely reasonable.
-However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.
+Although there were differences in the value of the empirical relationship parameters, they were generally in the correct direction and of a similar magnitude.
+This suggests that previously reported changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
+However, the biases and inconsistencies in the estimates of both lambda and environmental response parameters presented here suggest that it may be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.
 
@@
 The MLE method outperformed both of the binning methods under nearly any measure.
-With the publication of the *sizespectra* package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
+With the publication of the *sizeSpectra* package [@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
 Therefore, we recommend using it in all future studies of size spectra relationships rather than binning.
 
-We reiterate the recommendations of [-@White2007, -@sprules2015, -@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
+We reiterate the recommendations of White et al.
+[-@White2007], Sprules and Barth [-@sprules2015], and Edwards et al.
+[-@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
 Furthermore, we strongly encourage authors to publish individual size data whenever possible.

### 2023-07-18 Manuscript.tex
unstaged changes to manuscript before 1st submission

@@
 \usepackage[margin=1in]{geometry}
-\usepackage{longtable,booktabs}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 Gjoni\textsuperscript{4} \and Jeff S. Wesner\textsuperscript{4}}
-\date{23 February, 2023}
+\date{20 March, 2023}
 
@@
   The uncertainty and variation in estimates when using binning methods
-  is often greater than or equal to the variation previously published
-  in experimental and observational studies, bringing into question the
+  can be of similar magnitudes as the variation previously published in
+  experimental and observational studies, bringing into question the
   effect size of previously published results. However, while the
@@
 
-We had Two primary objectives in this study: 1) to compare how well
-different methods estimate site-specific \(\lambda\)'s 2) To compare how
-different methods of estimating \(\lambda\) impact subsequent
+We had two primary objectives in this study: 1) to compare how well
+different methods estimate site-specific \(\lambda\)'s and 2) to compare
+how different methods of estimating \(\lambda\) impact subsequent
 regressions that use \(\lambda\)'s as a response variable. We did this
@@
 to estimate the value of \(\lambda\) (maximum likelihood, equal
-logarithmic bins, and normalized abundance spectra). We repeated this
-process 1000 times (reps) and plotted the distribution of values
-obtained for each method.
+logarithmic bins, and normalized abundance spectra). We repeated the
+process of data simulation and parameter estimation 1000 times (reps)
+and plotted the distribution of values obtained for each method.
 
@@
 \hypertarget{equal-logarithmic-bins-elbn}{%
-\subsubsection{Equal Logarithmic Bins:
+\subsection{Equal Logarithmic Bins:
 ELBn}\label{equal-logarithmic-bins-elbn}}
@@
 In addition, to estimate the OLS slope parameter in log-log space, we
-used two common binning methods. . It is important to note that the
+used two common binning methods. It is important to note that the
 log-log regression is estimating \(\lambda\) as the regressionslope
@@
 ``intercept,'' and ``slope'' \(\beta_{env}\) to refer to the parameters
-of linear regression that uses the lambdas as a response variable.
-
-For the first binning method, we created 6 equal logarithmic bins
+of linear regressions that use the lambdas as a response variable.
+
+For the first binning method, we created six equal logarithmic bins
 covering the range of body sizes. This method has been used extensively
-in previous studies. For example see (Dossena et al. 2012, Martínez et
-al. 2016, Perkins et al. 2018). In the present study, the count in each
-bin was normalized by dividing by the bin width to account for the
+in previous studies (e.g., see Perkins et al. (2018), Martinez et al.
+(2016), and Dossena et al. (2012). In the present study, the count in
+each bin was normalized by dividing by the bin width to account for the
 unequal bin sizes. The process of normalizing shifts the estimate by -1.
@@
 
-The second method was similar to ELBn, but bins of log\textsubscript{2}
-widths are used. The number of bins is dependent on the range of body
-sizes present in the data. When working with empirical data with
-different size ranges this can alter the number of bins per site, and it
-is recommended to construct the bins based on the global min and max
-body size present. However, since the data here were simulated from a
-known size range, the number of bins for each site is identical. The
-count in each bin is normalized in the same way as described above. This
-method has been used extensively in the literature e.g., (Jennings et
-al. 2002, Jennings and Blanchard 2004, Sprules and Barth 2015, McGarvey
-and Kirk 2018, Pomeranz et al. 2019a) and is referred to as the
-Normalized Abundance Spectrum (NAS).
+The second binning method was similar to ELBn, but bins of
+log\textsubscript{2} width are used. The number of bins is dependent on
+the range of body sizes present in the data. When working with empirical
+data with different size ranges this can alter the number of bins per
+site, and it is recommended to construct the bins based on the global
+minimum and maximum body size present. However, since the data here were
+simulated from a known size range, the number of bins for each site is
+identical. The count in each bin is normalized in the same way as
+described above for the ELBn approach. This method has been used
+extensively in the literature (e.g., Jennings et al. (2002, 2004),
+Sprules and Barth (2015), Pomeranz et al. (2019a), McGarvey and Kirk
+(2018)), and is referred to as the Normalized Abundance Spectrum (NAS).
 
@@
 Where \(N\) is the count in each bin, \(M\) is the mid-point of each
-bin, and \(\beta_{1}\) is considered to be the \(\lambda_{estimate}\).
+bin, and \(\beta_{1}\) is considered to be the \(\lambda\) estimate.
 
@@
 Previous work has investigated the biases in \(\lambda\) estimates when
-using different methods White et al. (2007). However, the focus of the
-present work is to investigate biases when estimating the relationship
-parameters across an environmental gradient.
+using different methods (White et al. 2007, Edwards et al. 2017, Edwards
+et al. 2020). However, the focus of the present work is to investigate
+biases when estimating the relationship parameters across an
+environmental gradient.
 
@@
 
-\[\lambda_{ik} = \beta_{0} + \beta_{0}x_{ik}\]
+\[\lambda_{ik} = \beta_{0} + \beta_{1}x_{ik}\]
 
@@
 First, we used the three methods (NAS, ELBn, MLE) to estimate
-\(\lambda\) for each of the 5000 body size datasets. Then we fit a
+\(\lambda\) for each of the 5000 body size data sets. Then, we fit a
 linear regression to those data with the predictor variable \(x\) and
@@
 All individuals were identified and had their body lengths measured, and
-body mass was estimated using published length weight regressions. This
+body mass was estimated using published length-weight regressions. This
 data has been analyzed previously using size spectra methods as
-described in (Pomeranz et al. 2022). Detailed methods of the sampling
-collection and data processing methods can be in the macroinvertebrate
-DPI pubs found on the NEON website
-\url{https://www.neonscience.org/data-collection/macroinvertebrates}.
-Coefficient estimates (\(\pm\) 1 standard deviation) are compared across
-methods. This allows us to determine whether or not the main results
-would differ depending on the method used.
+described in Pomeranz et al. (2022). Detailed methods of the sampling
+collection and data processing methods can be found in the
+macroinvertebrate data product information documents found on the NEON
+website \url{https://data.neonscience.org/data-products/DP1.20120.001}.
+Estimates of the slope coefficient (\(\beta_1\) ) \(\pm\) 1 SD, are
+compared across methods. This allows us to determine whether or not the
+main results would differ depending on the method used.
 
@@
 
-The data used in this manuscript is already publicly available (Pomeranz
-et al. 2019b) data dryad DOI
-\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s},
+The empirical data used in this manuscript is already publicly available
+(Pomeranz et al. (2019b); data dryad DOI
+\url{https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s}, and
 (National Ecological Observatory Network (NEON) 2022). R Scripts to
@@
 by an average of 0.04 or 0.05 absolute units for the NAS and ELBn
-methods, more than 4 times higher than deviation (0.01) for the MLE
-(Table 1).
+methods, more than four times higher than the deviation (0.01) observed
+for the MLE (Table 1).
 
@@
 from random samples of body sizes from bounded power law distributions
-with varying exponents (-2.5, -0.5). The figure is facetted by the known
-lambda parameter (facet title) and is also shown as the dashed line in
-each facet. Note that the x-axis varies in each facet.}
-\end{figure}
-
-\begin{longtable}[]{@{}llrrrr@{}}
-\caption{Table 1. Summary of three methods in recapturing the known
-values of lambda or the regression slopes simulated in this study.
-Performance is determined by comparinged by uncertainty (range of 95\%
-CI's) and the absolute distance of the model estimates from the known
-estimates (median and sd of the difference). Values are summarized
-across all n = 9000 or 6000 simulated datasets. See figures for more
-specific comparisons.}\tabularnewline
-\toprule
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Target\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-Method\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-n\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-Median Range of 95\% CI\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-Median Absolute Deviation\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-SD Absolute Deviation\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Target\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-Method\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-n\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-Median Range of 95\% CI\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-Median Absolute Deviation\strut
-\end{minipage} &
-\begin{minipage}[b]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-SD Absolute Deviation\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Lambda\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-MLE\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-9000\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-0.0661400\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-0.0120454\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-0.0191500\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Lambda\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-NAS\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-9000\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-0.1335267\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-0.0460093\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-0.0929513\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Lambda\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-ELBn\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-8579\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-0.1672598\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-0.0350186\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-0.0675122\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Regression Slope\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-MLE\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-6000\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-0.0382185\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-0.0086982\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-0.0089766\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Regression Slope\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-NAS\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-6000\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-0.0799201\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-0.0513434\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-0.0442020\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.17}}\raggedright
-Regression Slope\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.07}}\raggedright
-ELBn\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.05}}\raggedleft
-6000\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.23}}\raggedleft
-0.0923495\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.26}}\raggedleft
-0.0300186\strut
-\end{minipage} &
-\begin{minipage}[t]{(\columnwidth - 5\tabcolsep) * \real{0.22}}\raggedleft
-0.0353218\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
+with varying exponents (-2.5 to -0.5). The figure is facetted by the
+known lambda parameter (facet title) and is also shown as the dashed
+line in each facet. Note that the x-axis varies in each facet.}
+\end{figure}
 
@@
 \end{figure}
+
+\hypertarget{table-1}{%
+\paragraph{\texorpdfstring{\textbf{Table 1}}{Table 1}}\label{table-1}}
+
+(\emph{Submitted as additional file on biorxiv.}) Summary of three
+methods in recapturing the known values of lambda or the regression
+slopes simulated in this study. Performance is determined by comparinged
+by uncertainty (range of 95\% CI's) and the absolute distance of the
+model estimates from the known estimates (median and SD of the
+difference). Values are summarized across all n = 9000 or 6000 simulated
+data sets. See figures for more specific comparisons.
 
@@
 estimates was always greater in the binning methods, with the width of
-the distributions increases with stronger relationships across a
+the distributions increasing with stronger relationships across a
 hypothetical gradient. By comparison, the MLE showed no evidence of bias
@@
 studied in a wide range of taxa inhabiting both terrestrial and aquatic
-ecosystems (reviewed by {[}Brown (1995); GastonBlackburn2000; White et
-al. (2007){]}). Empirical data shows generally consistent patterns and
-can be explained by the metabolic theory of ecology (Brown et al. 2004).
-Measuring parameters describing the decline in abundance with increasing
-body size in communities is being done with increasing frequency across
-ecology. Previous work has investigated the accuracy and inherent biases
+ecosystems (reviewed by Brown (1995), and White et al. (2007)).
+Empirical data shows generally consistent patterns and can be explained
+by the metabolic theory of ecology (Brown et al. 2004). Measuring
+parameters describing the decline in abundance with increasing body size
+in communities is being done with increasing frequency across ecology.
+Previous work has investigated the accuracy and inherent biases
 associated with different estimation methods (White et al. 2007, Edwards
-et al. 2017, Edwards et al. 2020). However, how these inaccuracies and
-biases compound across environmental gradients remains uncertain, making
-it difficult to detect variation in size spectra parameters across
-environmental gradients with confidence. The most important outcome of
-our work is that binning methods not only generate biased \(\lambda\)
-values for individual data sets, but that bias carries over to affect
-the parameters of subsequent regressions that use those \(\lambda\)'s as
-response variables. This makes it challenging to understand how
-\(\lambda\) varies in response to environmental gradients if binning is
-used to estimate size spectra exponents.
+et al. 2017, Edwards et al. 2020). However, the extent to which these
+inaccuracies and biases compound across environmental gradients remains
+uncertain, making it difficult to detect variation in size spectra
+parameters across environmental gradients with confidence. The most
+important outcome of our work is that binning methods not only generate
+biased \(\lambda\) values for individual data sets, but that bias
+carries over to affect the parameters of subsequent regressions that use
+those \(\lambda\)'s as response variables. This makes it challenging to
+understand how \(\lambda\) varies in response to environmental gradients
+if binning is used to estimate size spectra exponents.
 
@@
 accounts for their wide use in ecological studies. However, aggregating
-individuals into logarithmic bins removes a large source of the
-variation within the data by collapsing body size variation into a
-single value within each bin. For example, all individuals placed into a
-bin that ranges from 2-4 grams of mass are all treated as having a mass
-of 3 grams, the midpoint of that bin. Likewise, a single abundance value
-is taken for each bin, despite that fact that there is almost certainly
+individuals into logarithmic bins removes a large amount of information
+within the data by collapsing body size variation into a single value
+within each bin. For example, all individuals placed into a bin that
+ranges from 2-4 grams of mass are all treated as having a mass of 3
+grams, the midpoint of that bin. Likewise, a single abundance value is
+taken for each bin, despite that fact that there is almost certainly
 variation in the abundance of individuals that weigh \textasciitilde2,
@@
 meaning that it not only produces more accurate estimates, but does so
-with less certainty than binning, even when the underlying data sets are
-identical.
-
-The error in estimating individual \(\lambda\)'s was compounded when
-estimating linear models with those \(\lambda\)'s as response variables.
-The MLE method also resulted in the most accurate \(\beta_{env}\)
-estimates. Furthermore, variation around the estimate was always
-smallest with the MLE method.
-
-Although there were differences in magnitude of the empirical
-relationship coefficients, they were generally in the correct direction
+with less uncertainty than binning, even when the underlying data sets
+are identical.
+
+Although there were differences in the value of the empirical
+relationship parameters, they were generally in the correct direction
 and of a similar magnitude. This suggests that previously reported
-significant changes in size spectra parameters across environmental
-gradients and in experimental manipulations are plausible. Given that
-all of the data within a study is treated identically, the over all
-change in size spectra parameters is likely reasonable. However, the
-biases and inconsistencies in relationship estimates presented here
-suggest that it would be difficult if not impossible to directly compare
-the relative changes across different published studies which use
-different methods.
+changes in size spectra parameters across environmental gradients and in
+experimental manipulations are plausible. However, the biases and
+inconsistencies in the estimates of both lambda and environmental
+response parameters presented here suggest that it may be difficult if
+not impossible to directly compare the relative changes across different
+published studies which use different methods.
 
@@
 The MLE method outperformed both of the binning methods under nearly any
-measure. With the publication of the \emph{sizespectra} package (2017),
-producing MLE estimates of size spectra parameters is a relatively easy
-task. Therefore, we recommend using it in all future studies of size
-spectra relationships rather than binning.
-
-We reiterate the recommendations of (2017) to estimate size spectra
+measure. With the publication of the \emph{sizeSpectra} package (Edwards
+et al. 2017), producing MLE estimates of size spectra parameters is a
+relatively easy task. Therefore, we recommend using it in all future
+studies of size spectra relationships rather than binning.
+
+We reiterate the recommendations of White et al. (2007), Sprules and
+Barth (2015), and Edwards et al. (2017) to estimate size spectra
 relationships using MLE methods due to their superior performance in

### 2023-07-24 2b_integrated_plots.R
change order of facet rows vary known relationship plot

@@ -36,12 +36,17 @@ angles <- bind_rows(steep = angle_A,
 
 saveRDS(angles, file = "data_sim/angles.rds")
 
-lambda_window <- angles %>%
-  mutate(Model = factor(name,
+##### Need to chnage order of legend to be L2n, ELBn, MLE ####
+
+
+(lambda_window <- angles %>%
+  mutate(
+    Model = case_when(name == "NAS" ~ "L2n", .default = name),
+    Model = factor(Model,
                         levels = 
                           c("MLE",
                             "ELBn", 
-                            "NAS")),
+                            "L2n")),
          sim = factor(id, 
                       levels = c("steep", 
                       "medium", 
@@ -61,7 +66,7 @@ lambda_window <- angles %>%
   labs(x = "Relationship estimate") +
   facet_wrap(sim~., 
              ncol = 1) +
-  NULL
+  NULL)
 
 ggsave(plot = lambda_window,
        filename = "figures/lambda_angle_plot.png",
@@ -76,12 +81,20 @@ rel_data <- bind_rows(beta_0 = rel_0,
 
 saveRDS(rel_data, file = "data_sim/rel_data.rds")
 
+
+###### Flip facet-wrap rows, top = 0, bottom = -0.5
 rel_data %>%
-  mutate(Model = factor(name,
-                        levels = 
-                          c("MLE",
-                            "ELBn", 
-                            "NAS"))) %>%
+  mutate(
+    Model = case_when(name == "NAS" ~ "L2n", .default = name),
+    Model = factor(Model,
+                   levels = 
+                     c("MLE",
+                       "ELBn", 
+                       "L2n")),
+    # set levels of known relaitonship for facet_wrap() plotting order below
+    known_relationship = factor(known_relationship,
+                                levels = c("0", "-0.25", "-0.5"))) %>%
+  filter(rep <= 100) %>%
   ggplot(aes(x = env_gradient,
              y = estimate, 
              group = rep,
@@ -90,24 +103,14 @@ rel_data %>%
     values = c("#019AFF", "#FF914A", "#FF1984" )) +
   stat_smooth(geom = "line",
               method = "lm",
-              alpha = 0.15,
+              alpha = 0.05,
               se = FALSE, 
               color = "black") +
-  geom_point(alpha = 0.5) +
+  geom_point(alpha = 0.1) +
   facet_wrap(known_relationship~Model)+
-  # labs(title = sim_data$distribution,
-  #      subtitle = 
-  #        paste0("m_range =(",
[…]

### 2023-07-24 2b_integrated_plots.R
change order of facet rows vary known relationship plot

@@ -36,12 +36,17 @@ angles <- bind_rows(steep = angle_A,
 
 saveRDS(angles, file = "data_sim/angles.rds")
 
-lambda_window <- angles %>%
-  mutate(Model = factor(name,
+##### Need to chnage order of legend to be L2n, ELBn, MLE ####
+
+
+(lambda_window <- angles %>%
+  mutate(
+    Model = case_when(name == "NAS" ~ "L2n", .default = name),
+    Model = factor(Model,
                         levels = 
                           c("MLE",
                             "ELBn", 
-                            "NAS")),
+                            "L2n")),
          sim = factor(id, 
                       levels = c("steep", 
                       "medium", 
@@ -61,7 +66,7 @@ lambda_window <- angles %>%
   labs(x = "Relationship estimate") +
   facet_wrap(sim~., 
              ncol = 1) +
-  NULL
+  NULL)
 
 ggsave(plot = lambda_window,
        filename = "figures/lambda_angle_plot.png",
@@ -76,12 +81,19 @@ rel_data <- bind_rows(beta_0 = rel_0,
 
 saveRDS(rel_data, file = "data_sim/rel_data.rds")
 
+
+###### Flip facet-wrap rows, top = 0, bottom = -0.5
 rel_data %>%
-  mutate(Model = factor(name,
-                        levels = 
-                          c("MLE",
-                            "ELBn", 
-                            "NAS"))) %>%
+  mutate(
+    Model = case_when(name == "NAS" ~ "L2n", .default = name),
+    Model = factor(Model,
+                   levels = 
+                     c("MLE",
+                       "ELBn", 
+                       "L2n")),
+    known_relationship = factor(known_relationship,
+                                levels = c("0", "-0.25", "-0.5"))) %>%
+  filter(rep <= 100) %>%
   ggplot(aes(x = env_gradient,
              y = estimate, 
              group = rep,
@@ -90,24 +102,14 @@ rel_data %>%
     values = c("#019AFF", "#FF914A", "#FF1984" )) +
   stat_smooth(geom = "line",
               method = "lm",
-              alpha = 0.15,
+              alpha = 0.05,
               se = FALSE, 
               color = "black") +
-  geom_point(alpha = 0.5) +
+  geom_point(alpha = 0.1) +
   facet_wrap(known_relationship~Model)+
-  # labs(title = sim_data$distribution,
-  #      subtitle = 
-  #        paste0("m_range =(",
-  #               sim_data$m_lower,
-  #               ",",
[…]

### 2023-07-25 3_bias.R
fix formatting of script 3

@@ -4,12 +4,15 @@ library(tidybayes)
 source("custom_functions.R")
 
 # load data
-est_lambda <- readRDS("data_sim/est_lambda_dat.rds") %>% mutate(target = known_b,
-                                                                target_name = "Lambda") #figure 1 sims
-angles = readRDS("data_sim/angles.rds") %>% mutate(target = known_relationship,
-                                                   minCI = estimate - 1.96*std.error,
-                                                   maxCI = estimate + 1.96*std.error,
-                                                   target_name = "Regression Slope") #figure 3 sims
+est_lambda <- readRDS("data_sim/est_lambda_dat.rds") %>% 
+  mutate(target = known_b,
+         target_name = "Lambda") #figure 1 sims
+
+angles = readRDS("data_sim/angles.rds") %>% 
+  mutate(target = known_relationship,
+         minCI = estimate - 1.96*std.error,
+         maxCI = estimate + 1.96*std.error,
+         target_name = "Regression Slope") #figure 3 sims
 
 # rel_data = readRDS("data_sim/rel_data.rds") # figure 4 sims
 

### 2023-07-25 1a_sim_sliding_windows.R
fix a small typo in one of the comments

@@ -31,7 +31,7 @@ rep = 1000
 m_lower = 0.0026
 m_upper = 1.2e3
 
-# number of body sizes r=to sample, n
+# number of body sizes to sample, n
 n = 999
 
 est_lambda <- sim_result(n = n,

### 2023-07-25 2b_integrated_plots.R
change order of legends to match what's plotted in figs

@@ -56,9 +56,12 @@ saveRDS(angles, file = "data_sim/angles.rds")
              fill = Model))+
   stat_halfeye(.width = c(0.66, 0.95)) +
   scale_fill_manual(
-    values = c("#019AFF",
+    values = c("#FF1984",
                "#FF914A",
-               "#FF1984" )) +
+               "#019AFF"),
+    breaks = c("L2n",
+               "ELBn", 
+               "MLE")) +
   theme_bw() +
   geom_vline(
     aes(xintercept = known_relationship),
@@ -139,9 +142,12 @@ rel_data_summary %>%
              fill = Model))+
   stat_halfeye(.width = c(0.66, 0.95)) +
   scale_fill_manual(
-    values = c("#019AFF",
+    values = c("#FF1984",
                "#FF914A",
-               "#FF1984" )) +
+               "#019AFF"),
+    breaks = c("L2n",
+               "ELBn", 
+               "MLE")) +
   theme_bw() +
   geom_vline(
     aes(xintercept = known_relationship),
@@ -169,9 +175,12 @@ rel_data_summary %>%
              fill = Model))+
   stat_halfeye(.width = c(0.66, 0.95)) +
   scale_fill_manual(
-    values = c("#019AFF",
+    values = c("#FF1984",
                "#FF914A",
-               "#FF1984" )) +
+               "#019AFF"),
+    breaks = c("L2n",
+               "ELBn", 
+               "MLE")) +
   theme_bw() +
   geom_vline(
     aes(xintercept = known_relationship),
@@ -197,7 +206,12 @@ est_lambda %>%
         fill = Model)) +
   stat_halfeye(.width = c(0.66, 0.95)) +
   scale_fill_manual(
-    values = c("#019AFF", "#FF914A", "#FF1984" )) +
+    values = c("#FF1984",
+               "#FF914A",
+               "#019AFF"),
+    breaks = c("L2n",
+               "ELBn", 
+               "MLE")) +
   theme_bw() +
   geom_vline(aes(xintercept = known_b),
              linetype = "dashed") +

### 2023-07-25 3_bias.R
correct name for summary stat of all bias table

@@ -88,7 +88,7 @@ bias_table = all_dat %>%
   add_count %>% 
   group_by(target_name, name, n) %>%
   summarize(median_ci_range = median(conf_width),
-            mean_abs_bias = median(abs_bias),
+            median_abs_bias = median(abs_bias),
             sd_abs_bias = sd(abs_bias)) %>% 
   arrange(target_name, median_ci_range)
 

### 2023-07-25 SI_draft.Rmd
make separate rmd file for SI

--- 2023-07-18 Manuscript.Rmd
@@
 ---
-title: "Manuscript: Detecting differences in Size Spectra"
+title: "Supplementary Information for 'Detecting' Manuscript"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 ```
-
-# Abstract
-
-1.
-The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra, which are described by a power law.
-The focus of size spectra analysis is to estimate the exponent ($\lambda$) of the power law.
-2.
-Many methods have been proposed for estimating $\lambda$ most of which involve binning the data, summing abundance within bins, and then fitting a ordinary least squares (OLS) regression in log-log space.
-However, recent work has shown that binning procedures may return biased estimates of size spectra exponents compared to procedures that directly estimate $\lambda$ using maximum likelihood estimation (MLE).
-Despite this variability in estimates, it is unclear if the relative change across environmental gradients is consistent across methodologies.
-Here, we used simulation to compare the ability of two binning methods (NAS, ELBn) and MLE to 1) recapture known values of $\lambda$, and 2) recapture parameters in a linear regression measuring the change in $\lambda$ across a hypothetical environmental gradient.
-We also compared the methods using two previously published body size datasets across a pollution gradient and a temperature gradient
-3.
-Maximum likelihood methods always performed better than common binning methods, which demonstrated consistent bias depending on the simulated values of $\lambda$.
-This bias carried over to the regressions, which were more accurate when $\lambda$ was estimated using MLE compared to the binning procedures.
-Additionally, the variance in estimates using MLE methods is markedly reduced when compared to binning methods.
-4.
-The uncertainty and variation in estimates when using binning methods can be of similar magnitudes as the variation previously published in experimental and observational studies, bringing into question the effect size of previously published results.
-However, while the methods produced different slope estimates from previously published datasets, they were in qualitative agreement on the sign of those slopes (i.e., all negative or all positive).
-Our results provide further support for the direct estimation of $\lambda$ using MLE (or similar procedures) over the more common methods of binning.
-
-# Introduction
-
-Body size distributions are a fundamental characteristic of communities.
-In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981; @Damuth1991; @Damuth1998].
-This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991; @Brown2004].
-The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010].
-Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], and in response to human activities [@jennings2004].
-Likewise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized.
-For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected [@perkins2018].
-However, recent research has shown that these results may be an artifact of how the data were treated.
-Edwards et al.
-(2020) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodology used.
-
-Individual size distributions (ISD *sensu* [@White2007], also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
-ISDs represent frequency distributions of individual body sizes within a community, regardless of taxonomy.
-Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
-Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972; @Andersen2006a].
-Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions in log-log space of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
-Myriad binning methods have been proposed, including linear and logarithmic bin widths.
-Likewise, some methods rely on the absolute counts in the bins and others employ normalization techniques such as dividing the count by the bin width (especially common with logarithmic binning).
-As an alternative to binning methods, $\lambda$ can be estimated directly on un-binned data using maximum likelihood techniques.
-
-Previous work has shown that the estimates of $\lambda$ differ between MLE and size-binned OLS techniques.
-Size-binned OLS methods are particularly sensitive to decisions made in the binning process.
-Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ [@white2008; @edwards2017], and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008; @edwards2020].
-However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied based on method, the relative change across the environmental gradient was consistent across methods [@pomeranz2022].
-While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time.
-In other words, if the data within a study are all treated the same, does a relative change of size-binned OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?
-
-We had two primary objectives in this study: 1) to compare how well different methods estimate site-specific $\lambda$'s and 2) to compare how different methods of estimating $\lambda$ impact subsequent regressions that use $\lambda$'s as a response variable.
-We did this using both simulated and empirical data sets.
-
-# Methods
-
-## Data Simulation
-
-To investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@edwards2017].
-Let $M$ be a random variable of body sizes described by the probability density function:
-
-$$f(M) = CM^\lambda, M_{min} \le M \le M_{max}$$
-
-$M_{min} = 0.0026$ and $M_{max} = 1.2 *10^3$.
-These values were based on empirical body sizes of stream benthic communities reported in [@pomeranz2022].
-Our results were not dependent on the range of body sizes (Supplemental Information).
-
-## Site-specific $\lambda$ estimates
-
-Using the procedure above, we simulated $n = 1000$ body sizes from five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5).
-The values of $\lambda$ describe how quickly the abundance of large body sizes decline within a community.
-For example, a value of -1.5 means there would be a relatively high number of large body sizes (shallow) whereas a value of -2.5 means there would be relatively very few large body sizes (steep).
-
-## Estimation of Size spectra parameter $\lambda$
-
-After simulating data, we used three different methods (described below) to estimate the value of $\lambda$ (maximum likelihood, equal logarithmic bins, and normalized abundance spectra).
-We repeated the process of data simulation and parameter estimation 1000 times (reps) and plotted the distribution of values obtained for each method.
-
-## Maximum likelihood estimation (MLE)
-
-We estimated the exponent $\lambda$ directly using MLE methods modified from the the `sizeSpectra` package [@edwards2017]. Throughout the manuscript, these estimates are referred to as MLE.
-
-## Equal Logarithmic Bins: ELBn
-
-In addition, to estimate the OLS slope parameter in log-log space, we used two common binning methods.
-It is important to note that the log-log regression is estimating $\lambda$ as the regressionslope ($\beta_{1}$), and not $\lambda$ directly.
-In the interest of clarity, we will refer to all estimates of lambda as $\lambda$ regardless of the method used, and reserve the terms “parameters”, “intercept”, and “slope” $\beta_{env}$ to refer to the parameters of linear regressions that use the lambdas as a response variable.
-
-For the first binning method, we created six equal logarithmic bins covering the range of body sizes.
-This method has been used extensively in previous studies (e.g., see Perkins et al.
-[-@perkins2018], Martinez et al.
-[-@martinez2016], and Dossena et al.
-[-@dossena2012].
-In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes.
-The process of normalizing shifts the estimate by -1.
-In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015; @edwards2017; @pomeranz2022].
-Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
-
-### Log~2~ Bins: NAS
-The second binning method was similar to ELBn, but bins of log~2~ width are used.
-The number of bins is dependent on the range of body sizes present in the data.
-When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global minimum and maximum body size present.
-However, since the data here were simulated from a known size range, the number of bins for each site is identical.
-The count in each bin is normalized in the same way as described above for the ELBn approach.
-This method has been  used extensively in the literature (e.g., Jennings et al.
-[-@jennings2004; -@Jennings2002], Sprules and Barth [-@sprules2015], Pomeranz et al.
-[-@pomeranz2019], McGarvey and Kirk [-@mcgarvey2018]), and is referred to as the Normalized Abundance Spectrum (NAS).
-
-For the two binning methods, simple OLS regression were conducted in the form 
-$$Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)$$
-
-Where $N$ is the count in each bin, $M$ is the mid-point of each bin, and $\beta_{1}$ is considered to be the $\lambda$ estimate.
-
-## Variation in $\lambda$ Across a Gradient
-
-Previous work has investigated the biases in $\lambda$ estimates when using different methods [@edwards2017; @edwards2020; @White2007].
-However, the focus of the present work is to investigate biases when estimating the relationship parameters across an environmental gradient.
-
-To do this, we first simulated  $\lambda$'s from a linear regression with a known slope $\beta_{1}$ of -0.5 and intercept of -1.5 across a generic predictor variable x with values ranging from -1 to 1:
-
-$$\lambda_{ik} = \beta_{0} + \beta_{1}x_{ik}$$
-
-This produced $k$ = 1000  values from each value of $x$ (-1, -0.5, 0, 0.5, 1), resulting in 5000 total $\lambda$’s.
-From each $\lambda$ we simulated 1000 individual body sizes using the procedure described above.
-
-Once the body size data were simulated, we then reversed the procedure.
-First, we used the three methods (NAS, ELBn, MLE) to estimate $\lambda$ for each of the 5000 body size data sets.
-Then, we fit a linear regression to those data with the predictor variable $x$ and compared the resulting intercept ($\beta_{0}$) and slope ($\beta_{1}$) parameter estimates to the known value.
-The main results presented here were not dependent on the range of x-values or the number of sites (Supplemental Information).
-
-## Empirical Data
-
-We re-analyzed two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients.
-In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage (AMD) stress gradient.
-Details of the sample collection and processing can be found in [@pomeranz2019].
-Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
-Body mass was estimated using taxon-specific published length weight regressions.
-
-The second data set was from the wadeable stream sites of NEON [@NEON_Inverts2022].
-NEON stream sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska.
-Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat.
-All individuals were identified and had their body lengths measured, and body mass was estimated using published length-weight regressions.
-This data has been analyzed previously using size spectra methods as described in Pomeranz et al.
-[-@pomeranz2022].
-Detailed methods of the sampling collection and data processing methods can be found in the macroinvertebrate data product information documents found on the NEON website <https://data.neonscience.org/data-products/DP1.20120.001>.
-Estimates of the slope coefficient ($\beta_1$ ) $\pm$ 1 SD, are compared across methods.
-This allows us to determine whether or not the main results would differ depending on the method used.
-
-## Performance Metrics
-
-We compared performance of each procedure (NAS, ELBn, MLE) by first plotting the distribution of lambda or regression parameter estimates ($\lambda$, $\beta_{0}$, $\beta_{1}$) against the known values.
-We calculated bias for the procedures overall as the median absolute difference (averaged across all simulations) between the known values and the modeled estimates.
-Finally, for each procedure we estimated the width of the 95% CI’s to compare uncertainty.
-
-## Data Availabity
-
-The empirical data used in this manuscript is already publicly available (Pomeranz et al.
-[-@Pomeranz2019]; data dryad DOI <https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s>, and  [@NEON_Inverts2022].
-R Scripts to reproduce the full simulation and analysis are available at: ([<https://github.com/Jpomz/detecting-spectra-differences>]).
-(To be archived upon acceptance.)
-
-# Results
-
-# $\lambda$ estimates
-
-There was considerable variation in the $\lambda$ estimate across methodologies (Figure 1).
-The distribution of estimates from the MLE method was always symmetrical and centered at the known value of $\lambda$ (Figure 1).
-The distribution of estimates from the binning methods were generally wider and occasionally asymmetrical or bimodal.
-On average, the 95% confidence intervals produced by the NAS and ELBn methods were ~2 times wider than those produced by MLE (Table 1), indicating greater consistency of estimates from MLE.
-Similarly, estimates of  $\lambda$ deviated from the true $\lambda$ by an average of 0.04 or 0.05 absolute units for the NAS and ELBn methods, more than four times higher than the deviation (0.01) observed for the MLE (Table 1).
-
-![Figure 1.
-Distribution of Lambda estimates by method (color) from random samples of body sizes from bounded power law distributions with varying exponents (-2.5 to -0.5).
-The figure is facetted by the known lambda parameter (facet title) and is also shown as the dashed line in each facet.
-Note that the x-axis varies in each facet.](figures/est_lambda_est_b_density.png)
-
-Furthermore, the binning methods systematically over estimated steep size spectra relationships ($\lambda = ~\sim-2.5~to-1.5$, Figure 2), and slightly underestimated shallow size spectra relationships ($\lambda = ~\sim-0.5$).
-This finding was more pronounced in the NAS method compared with the ELBn method.
-
-![Figure 2.
-Plot of the known and estimated lambda from different methods (color).
-The dashed line is the 1:1 line, and almost perfectly matches the MLE (Blue) estimate.
-The two binning method cross the 1:1 line, meaning they systematically underestimate shallow lambdas, and over estimate steeper lambdas.](figures/known_est_b_line.png)
-
-#### **Table 1**
-
-(*Submitted as additional file on biorxiv.*) Summary of three methods in recapturing the known values of lambda or the regression slopes simulated in this study.
-Performance is determined by comparinged by uncertainty (range of 95% CI's) and the absolute distance of the model estimates from the known estimates (median and SD of the difference).
-Values are summarized across all n = 9000 or 6000 simulated data sets.
-See figures for more specific comparisons.
-
-## Relationship across the hypothetical environmental gradient
-
-### $\lambda$ "windows"
-
-Because of the different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of $\beta_{env} = -0.5$ across a hypothetical environmental gradient.
-The lambda "windows" are steep (-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).
-
-![Figure 3.
-Distribution of relationship estimates ($beta_{env}$) in three different "windows" of lambda values.
-The dashed vertical line is the known relationship value of -0.5.](figures/lambda_angle_plot.png)
-
-The MLE method (Figure 3, blue) recaptured the known slope value in each of the “windows.”, with median absolute deviation of ~0.01 units (Table 1).
-By contrast, the binning methods systematically overestimated the known slope (Figure 3), with median absolute deviations >4 times higher than the MLE.
-Similarly, uncertainty in the slope estimates derived from binning methods was twice that of the uncertainty in the MLE method (Table 1).
-
-### Varying the Known Relationship
-
-All methods recaptured the correct sign of the slopes, yielding qualitative consistency (Figure 4).
-However, the binning methods again systematically underestimated the true value of the slope by ~0.05 units (Figure 5).
-Likewise, uncertainty in the slope estimates was always greater in the binning methods, with the width of the distributions increasing with stronger relationships across a hypothetical gradient.
-By comparison, the MLE showed no evidence of bias and was always centered at the known value with relatively narrow variation.
-
-![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows from bottom to top): 0, 0.25, and 0.5](figures/vary_beta_plot.png)
-
-![Figure 5. Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
-
-## Empirical data
-
-Both empirical data sets yielded similar results to the simulated data; the direction of the coefficients (i.e., $\beta_1$ coefficients) were the same among methods, with positive slopes across the AMD gradient (Fig.
-6A) and negative slopes across the temperature gradient.
-However, the magnitude of change differed between methods.
-The AMD slopes ranged from ~0.62 with MLE to ~0.78 with NAS (Figure 6b), while temperature slopes ranged from -0.0058 with MLE to -0.0019 with ELBn (Figure 6d).
-As with simulated data, slope uncertainty ($\pm$ 1 SD) was larger in the binning methods, particularly the ELBn method (Fig.
-6B).
-Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig.
-6C).
-
-\newpage
-
-![Figure 6.
-of empirical data estimates.
-All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.](figures/empirical_combined.png)
-
-\newpage
-
-# Discussion
-
-The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by Brown [-@brown1995], and White et al.
-[-@White2007]).
-Empirical data shows generally consistent patterns and can be explained by the metabolic theory of ecology [@Brown2004].
-Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology.
-Previous work has investigated the accuracy and inherent biases associated with different estimation methods [@White2007; @edwards2017; @edwards2020].
-However, the extent to which these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
-The most important outcome of our work is that binning methods not only generate biased $\lambda$ values for individual data sets, but that bias carries over to affect the parameters of subsequent regressions that use those $\lambda$’s as response variables.
-This makes it challenging to understand how $\lambda$ varies in response to environmental gradients if binning is used to estimate size spectra exponents.
-
-Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies.
-However, aggregating individuals into logarithmic bins removes a large amount of information within the data by collapsing body size variation into a single value within each bin.
-For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
-Likewise, a single abundance value is taken for each bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams.
-Moreover, the number of bins that can be produced by any dataset is limited most often to less than six.
-That means a subsequent log-log regression used to estimate $\lambda$ would contains n = 6 data points, even though there likely 100’s or 1000’s of individual body sizes available in a typical body size data set.
-By homogenizing the data in this way, binning methods produce noisier results than MLE, since binning is akin to deleting information that the model could otherwise use.
-By contrast, the MLE uses all the individual body size data to directly estimate $\lambda$, meaning that it not only produces more accurate estimates, but does so with less uncertainty than binning, even when the underlying data sets are identical.
-
-Although there were differences in the value of the empirical relationship parameters, they were generally in the correct direction and of a similar magnitude.
-This suggests that previously reported changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
-However, the biases and inconsistencies in the estimates of both lambda and environmental response parameters presented here suggest that it may be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.
-
-The publication of individual body size data with future studies of size spectra would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
-
-## Concluding Remarks
-
-The MLE method outperformed both of the binning methods under nearly any measure.
-With the publication of the *sizeSpectra* package [@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
-Therefore, we recommend using it in all future studies of size spectra relationships rather than binning.
-
-We reiterate the recommendations of White et al.
-[-@White2007], Sprules and Barth [-@sprules2015], and Edwards et al.
-[-@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
-Furthermore, we strongly encourage authors to publish individual size data whenever possible.
-This will allow for the consistent re-analysis of existing data sets as methodologies develop and improve.
-This will aid in the ability for size spectra work to be synthesized between research groups and across scales.
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-\newpage
 

### 2023-07-25 figures_draft.Rmd
make separate rmd file for figs and knit for submission

--- 2023-07-18 Manuscript.Rmd
@@
 ---
-title: "Manuscript: Detecting differences in Size Spectra"
+title: "Figures for 'Detecting' Manuscript"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-# Abstract
-
-1.
-The distribution of body size in communities  is remarkably consistent across habitats and taxa and can be represented by size spectra, which are described by a power law.
-The focus of size spectra analysis is to estimate the exponent ($\lambda$) of the power law.
-2.
-Many methods have been proposed for estimating $\lambda$ most of which involve binning the data, summing abundance within bins, and then fitting a ordinary least squares (OLS) regression in log-log space.
-However, recent work has shown that binning procedures may return biased estimates of size spectra exponents compared to procedures that directly estimate $\lambda$ using maximum likelihood estimation (MLE).
-Despite this variability in estimates, it is unclear if the relative change across environmental gradients is consistent across methodologies.
-Here, we used simulation to compare the ability of two binning methods (NAS, ELBn) and MLE to 1) recapture known values of $\lambda$, and 2) recapture parameters in a linear regression measuring the change in $\lambda$ across a hypothetical environmental gradient.
-We also compared the methods using two previously published body size datasets across a pollution gradient and a temperature gradient
-3.
-Maximum likelihood methods always performed better than common binning methods, which demonstrated consistent bias depending on the simulated values of $\lambda$.
-This bias carried over to the regressions, which were more accurate when $\lambda$ was estimated using MLE compared to the binning procedures.
-Additionally, the variance in estimates using MLE methods is markedly reduced when compared to binning methods.
-4.
-The uncertainty and variation in estimates when using binning methods can be of similar magnitudes as the variation previously published in experimental and observational studies, bringing into question the effect size of previously published results.
-However, while the methods produced different slope estimates from previously published datasets, they were in qualitative agreement on the sign of those slopes (i.e., all negative or all positive).
-Our results provide further support for the direct estimation of $\lambda$ using MLE (or similar procedures) over the more common methods of binning.
-
-# Introduction
-
-Body size distributions are a fundamental characteristic of communities.
-In general, size-abundance relationships are expected to scale with the value of -0.75 across all biological communities [@Damuth1981; @Damuth1991; @Damuth1998].
-This is thought to be a consequence of simple size-dependent metabolic constraints on organisms’ energy use predicted by the metabolic theory of ecology [@Nee1991; @Brown2004].
-The remarkable consistency of these relationships across spatiotemporal scales and ecosystems has led them to be recommended as a "universal" indicator of ecological status [@Petchey2010].
-Variation in size-abundance relationships have been documented through space [@pomeranz2022], time [@evans2022], seasonality [@mcgarvey2018], and in response to human activities [@jennings2004].
-Likewise, variation in size spectra relationships have been used to explain fundamental differences in how communities are organized.
-For example, external resource subsidies "bend the rules" and allow higher abundances of large body sizes than would be expected [@perkins2018].
-However, recent research has shown that these results may be an artifact of how the data were treated.
-Edwards et al.
-(2020) reanalyzed a time series of marine fisheries data and found that previously reported changes through time were actually dependent on the methodology used.
-
-Individual size distributions (ISD *sensu* [@White2007], also referred to as abundance size spectra) are one of the size-abundance relationships commonly used.
-ISDs represent frequency distributions of individual body sizes within a community, regardless of taxonomy.
-Generally, there is a negative relationship between individual body size (M) on the *x*-axis  and abundance (N) on the *y*-axis.
-Theoretical and empirical data support this relationship being described as a simple power law with exponent $\lambda$ in the form of $N \sim M^{\lambda}$, where $\lambda = 2$ [@Sheldon1972; @Andersen2006a].
-Commonly, $N$ is the count of body sizes grouped into bins, and $\lambda$ is estimated as the slope from OLS regressions in log-log space of $N$ and the mid point of the body size bin $M_{bin}$: $\log(N_{count}) = \lambda \log(M_{bin})$.
-Myriad binning methods have been proposed, including linear and logarithmic bin widths.
-Likewise, some methods rely on the absolute counts in the bins and others employ normalization techniques such as dividing the count by the bin width (especially common with logarithmic binning).
-As an alternative to binning methods, $\lambda$ can be estimated directly on un-binned data using maximum likelihood techniques.
-
-Previous work has shown that the estimates of $\lambda$ differ between MLE and size-binned OLS techniques.
-Size-binned OLS methods are particularly sensitive to decisions made in the binning process.
-Simulation studies have shown that MLE offers consistently more accurate estimates of $\lambda$ [@white2008; @edwards2017], and reanalysis of empirical data also indicates that the conclusions are dependent on the methodology used [@white2008; @edwards2020].
-However, recent empirical analysis of stream macroinvertebrate communities across the National Ecological Observatory Network (NEON, USA) showed that while the estimates of $\lambda$ varied based on method, the relative change across the environmental gradient was consistent across methods [@pomeranz2022].
-While there is a growing consensus that MLE methods offer more reliable estimates of $\lambda$, and binning methods result in biased estimates, it remains unclear if these biases are consistent and systematic or stochastic, and whether or not the relative change in ISD parameters is consistent across space and time.
-In other words, if the data within a study are all treated the same, does a relative change of size-binned OLS slope parameters of 0.1 coincide with a relative change of MLE $\lambda$ estimates of 0.1?
-
-We had two primary objectives in this study: 1) to compare how well different methods estimate site-specific $\lambda$'s and 2) to compare how different methods of estimating $\lambda$ impact subsequent regressions that use $\lambda$'s as a response variable.
-We did this using both simulated and empirical data sets.
-
-# Methods
-
-## Data Simulation
-
-To investigate the performance of commonly used methods, we simulate body size observations from a bounded power law distribution using the inverse method, as described in [@edwards2017].
-Let $M$ be a random variable of body sizes described by the probability density function:
-
-$$f(M) = CM^\lambda, M_{min} \le M \le M_{max}$$
-
-$M_{min} = 0.0026$ and $M_{max} = 1.2 *10^3$.
-These values were based on empirical body sizes of stream benthic communities reported in [@pomeranz2022].
-Our results were not dependent on the range of body sizes (Supplemental Information).
-
-## Site-specific $\lambda$ estimates
-
-Using the procedure above, we simulated $n = 1000$ body sizes from five different $\lambda$'s: (-1.5, -1.75, -2.00, -2.25, -2.5).
-The values of $\lambda$ describe how quickly the abundance of large body sizes decline within a community.
-For example, a value of -1.5 means there would be a relatively high number of large body sizes (shallow) whereas a value of -2.5 means there would be relatively very few large body sizes (steep).
-
-## Estimation of Size spectra parameter $\lambda$
-
-After simulating data, we used three different methods (described below) to estimate the value of $\lambda$ (maximum likelihood, equal logarithmic bins, and normalized abundance spectra).
-We repeated the process of data simulation and parameter estimation 1000 times (reps) and plotted the distribution of values obtained for each method.
-
-## Maximum likelihood estimation (MLE)
-
-We estimated the exponent $\lambda$ directly using MLE methods modified from the the `sizeSpectra` package [@edwards2017]. Throughout the manuscript, these estimates are referred to as MLE.
-
-## Equal Logarithmic Bins: ELBn
-
-In addition, to estimate the OLS slope parameter in log-log space, we used two common binning methods.
-It is important to note that the log-log regression is estimating $\lambda$ as the regressionslope ($\beta_{1}$), and not $\lambda$ directly.
-In the interest of clarity, we will refer to all estimates of lambda as $\lambda$ regardless of the method used, and reserve the terms “parameters”, “intercept”, and “slope” $\beta_{env}$ to refer to the parameters of linear regressions that use the lambdas as a response variable.
-
-For the first binning method, we created six equal logarithmic bins covering the range of body sizes.
-This method has been used extensively in previous studies (e.g., see Perkins et al.
-[-@perkins2018], Martinez et al.
-[-@martinez2016], and Dossena et al.
-[-@dossena2012].
-In the present study, the count in each bin was normalized by dividing by the bin width to account for the unequal bin sizes.
-The process of normalizing shifts the estimate by -1.
-In other words, an un-normalized estimate of -0.75 would result in an estimate of -1.75 when normalizing the data [@sprules2015; @edwards2017; @pomeranz2022].
-Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn.
-
-### Log~2~ Bins: NAS
-The second binning method was similar to ELBn, but bins of log~2~ width are used.
-The number of bins is dependent on the range of body sizes present in the data.
-When working with empirical data with different size ranges this can alter the number of bins per site, and it is recommended to construct the bins based on the global minimum and maximum body size present.
-However, since the data here were simulated from a known size range, the number of bins for each site is identical.
-The count in each bin is normalized in the same way as described above for the ELBn approach.
-This method has been  used extensively in the literature (e.g., Jennings et al.
-[-@jennings2004; -@Jennings2002], Sprules and Barth [-@sprules2015], Pomeranz et al.
-[-@pomeranz2019], McGarvey and Kirk [-@mcgarvey2018]), and is referred to as the Normalized Abundance Spectrum (NAS).
-
-For the two binning methods, simple OLS regression were conducted in the form 
-$$Log_{10}(N) = \beta_0 + \beta_1 * Log_{10}(M)$$
-
-Where $N$ is the count in each bin, $M$ is the mid-point of each bin, and $\beta_{1}$ is considered to be the $\lambda$ estimate.
-
-## Variation in $\lambda$ Across a Gradient
-
-Previous work has investigated the biases in $\lambda$ estimates when using different methods [@edwards2017; @edwards2020; @White2007].
-However, the focus of the present work is to investigate biases when estimating the relationship parameters across an environmental gradient.
-
-To do this, we first simulated  $\lambda$'s from a linear regression with a known slope $\beta_{1}$ of -0.5 and intercept of -1.5 across a generic predictor variable x with values ranging from -1 to 1:
-
-$$\lambda_{ik} = \beta_{0} + \beta_{1}x_{ik}$$
-
-This produced $k$ = 1000  values from each value of $x$ (-1, -0.5, 0, 0.5, 1), resulting in 5000 total $\lambda$’s.
-From each $\lambda$ we simulated 1000 individual body sizes using the procedure described above.
-
-Once the body size data were simulated, we then reversed the procedure.
-First, we used the three methods (NAS, ELBn, MLE) to estimate $\lambda$ for each of the 5000 body size data sets.
-Then, we fit a linear regression to those data with the predictor variable $x$ and compared the resulting intercept ($\beta_{0}$) and slope ($\beta_{1}$) parameter estimates to the known value.
-The main results presented here were not dependent on the range of x-values or the number of sites (Supplemental Information).
-
-## Empirical Data
-
-We re-analyzed two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients.
-In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage (AMD) stress gradient.
-Details of the sample collection and processing can be found in [@pomeranz2019].
-Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope.
-Body mass was estimated using taxon-specific published length weight regressions.
-
-The second data set was from the wadeable stream sites of NEON [@NEON_Inverts2022].
-NEON stream sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska.
-Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat.
-All individuals were identified and had their body lengths measured, and body mass was estimated using published length-weight regressions.
-This data has been analyzed previously using size spectra methods as described in Pomeranz et al.
-[-@pomeranz2022].
-Detailed methods of the sampling collection and data processing methods can be found in the macroinvertebrate data product information documents found on the NEON website <https://data.neonscience.org/data-products/DP1.20120.001>.
-Estimates of the slope coefficient ($\beta_1$ ) $\pm$ 1 SD, are compared across methods.
-This allows us to determine whether or not the main results would differ depending on the method used.
-
-## Performance Metrics
-
-We compared performance of each procedure (NAS, ELBn, MLE) by first plotting the distribution of lambda or regression parameter estimates ($\lambda$, $\beta_{0}$, $\beta_{1}$) against the known values.
-We calculated bias for the procedures overall as the median absolute difference (averaged across all simulations) between the known values and the modeled estimates.
-Finally, for each procedure we estimated the width of the 95% CI’s to compare uncertainty.
-
-## Data Availabity
-
-The empirical data used in this manuscript is already publicly available (Pomeranz et al.
-[-@Pomeranz2019]; data dryad DOI <https://datadryad.org/stash/dataset/doi:10.5061/dryad.v6g985s>, and  [@NEON_Inverts2022].
-R Scripts to reproduce the full simulation and analysis are available at: ([<https://github.com/Jpomz/detecting-spectra-differences>]).
-(To be archived upon acceptance.)
-
-# Results
-
-# $\lambda$ estimates
-
-There was considerable variation in the $\lambda$ estimate across methodologies (Figure 1).
-The distribution of estimates from the MLE method was always symmetrical and centered at the known value of $\lambda$ (Figure 1).
-The distribution of estimates from the binning methods were generally wider and occasionally asymmetrical or bimodal.
-On average, the 95% confidence intervals produced by the NAS and ELBn methods were ~2 times wider than those produced by MLE (Table 1), indicating greater consistency of estimates from MLE.
-Similarly, estimates of  $\lambda$ deviated from the true $\lambda$ by an average of 0.04 or 0.05 absolute units for the NAS and ELBn methods, more than four times higher than the deviation (0.01) observed for the MLE (Table 1).
-
 ![Figure 1.
@@
 
-Furthermore, the binning methods systematically over estimated steep size spectra relationships ($\lambda = ~\sim-2.5~to-1.5$, Figure 2), and slightly underestimated shallow size spectra relationships ($\lambda = ~\sim-0.5$).
-This finding was more pronounced in the NAS method compared with the ELBn method.
-
 ![Figure 2.
-Plot of the known and estimated lambda from different methods (color).
-The dashed line is the 1:1 line, and almost perfectly matches the MLE (Blue) estimate.
-The two binning method cross the 1:1 line, meaning they systematically underestimate shallow lambdas, and over estimate steeper lambdas.](figures/known_est_b_line.png)
-
-#### **Table 1**
-
-(*Submitted as additional file on biorxiv.*) Summary of three methods in recapturing the known values of lambda or the regression slopes simulated in this study.
-Performance is determined by comparinged by uncertainty (range of 95% CI's) and the absolute distance of the model estimates from the known estimates (median and SD of the difference).
-Values are summarized across all n = 9000 or 6000 simulated data sets.
-See figures for more specific comparisons.
-
-## Relationship across the hypothetical environmental gradient
-
-### $\lambda$ "windows"
-
-Because of the different performance of the two binning methods at steep and shallow values of lambdas, we first investigated how well the methods capture a known relationship of $\beta_{env} = -0.5$ across a hypothetical environmental gradient.
-The lambda "windows" are steep (-2.5, -1.5), medium (-2, -1) and shallow (-1.5, -0.5).
-
-![Figure 3.
 Distribution of relationship estimates ($beta_{env}$) in three different "windows" of lambda values.
@@
 
-The MLE method (Figure 3, blue) recaptured the known slope value in each of the “windows.”, with median absolute deviation of ~0.01 units (Table 1).
-By contrast, the binning methods systematically overestimated the known slope (Figure 3), with median absolute deviations >4 times higher than the MLE.
-Similarly, uncertainty in the slope estimates derived from binning methods was twice that of the uncertainty in the MLE method (Table 1).
+![Figure 3.
+Individual regressions by method (columns) for three different known relationship values (rows from bottom to top): 0, 0.25, and 0.5.
+There were a total of 1000 replicates simulated, but only 100 replicates are plotted here with `alpha = 0.05` to illustrate the variability in the regression lines.](figures/vary_beta_plot.png)
 
-### Varying the Known Relationship
+![Figure 4. Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
 
-All methods recaptured the correct sign of the slopes, yielding qualitative consistency (Figure 4).
-However, the binning methods again systematically underestimated the true value of the slope by ~0.05 units (Figure 5).
-Likewise, uncertainty in the slope estimates was always greater in the binning methods, with the width of the distributions increasing with stronger relationships across a hypothetical gradient.
-By comparison, the MLE showed no evidence of bias and was always centered at the known value with relatively narrow variation.
-
-![Figure 4. Individual regressions for each rep (N = 1000) and method (columns) for three different known relationship values (rows from bottom to top): 0, 0.25, and 0.5](figures/vary_beta_plot.png)
-
-![Figure 5. Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
-
-## Empirical data
-
-Both empirical data sets yielded similar results to the simulated data; the direction of the coefficients (i.e., $\beta_1$ coefficients) were the same among methods, with positive slopes across the AMD gradient (Fig.
-6A) and negative slopes across the temperature gradient.
-However, the magnitude of change differed between methods.
-The AMD slopes ranged from ~0.62 with MLE to ~0.78 with NAS (Figure 6b), while temperature slopes ranged from -0.0058 with MLE to -0.0019 with ELBn (Figure 6d).
-As with simulated data, slope uncertainty ($\pm$ 1 SD) was larger in the binning methods, particularly the ELBn method (Fig.
-6B).
-Likewise, the size spectra parameters consistently increase (become steeper) with increasing temperature across the NEON sites (Fig.
-6C).
-
-\newpage
-
-![Figure 6.
+![Figure 5.
 of empirical data estimates.
 All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.](figures/empirical_combined.png)
-
-\newpage
-
-# Discussion
-
-The relationship between body size and abundance has been extensively studied in a wide range of taxa inhabiting both terrestrial and aquatic ecosystems (reviewed by Brown [-@brown1995], and White et al.
-[-@White2007]).
-Empirical data shows generally consistent patterns and can be explained by the metabolic theory of ecology [@Brown2004].
-Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology.
-Previous work has investigated the accuracy and inherent biases associated with different estimation methods [@White2007; @edwards2017; @edwards2020].
-However, the extent to which these inaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence.
-The most important outcome of our work is that binning methods not only generate biased $\lambda$ values for individual data sets, but that bias carries over to affect the parameters of subsequent regressions that use those $\lambda$’s as response variables.
-This makes it challenging to understand how $\lambda$ varies in response to environmental gradients if binning is used to estimate size spectra exponents.
-
-Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies.
-However, aggregating individuals into logarithmic bins removes a large amount of information within the data by collapsing body size variation into a single value within each bin.
-For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin.
-Likewise, a single abundance value is taken for each bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh \~2, \~3, or \~4 grams.
-Moreover, the number of bins that can be produced by any dataset is limited most often to less than six.
-That means a subsequent log-log regression used to estimate $\lambda$ would contains n = 6 data points, even though there likely 100’s or 1000’s of individual body sizes available in a typical body size data set.
-By homogenizing the data in this way, binning methods produce noisier results than MLE, since binning is akin to deleting information that the model could otherwise use.
-By contrast, the MLE uses all the individual body size data to directly estimate $\lambda$, meaning that it not only produces more accurate estimates, but does so with less uncertainty than binning, even when the underlying data sets are identical.
-
-Although there were differences in the value of the empirical relationship parameters, they were generally in the correct direction and of a similar magnitude.
-This suggests that previously reported changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible.
-However, the biases and inconsistencies in the estimates of both lambda and environmental response parameters presented here suggest that it may be difficult if not impossible to directly compare the relative changes across different published studies which use different methods.
-
-The publication of individual body size data with future studies of size spectra would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental conditions.
-
-## Concluding Remarks
-
-The MLE method outperformed both of the binning methods under nearly any measure.
-With the publication of the *sizeSpectra* package [@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task.
-Therefore, we recommend using it in all future studies of size spectra relationships rather than binning.
-
-We reiterate the recommendations of White et al.
-[-@White2007], Sprules and Barth [-@sprules2015], and Edwards et al.
-[-@edwards2017] to estimate size spectra relationships using MLE methods due to their superior performance in nearly every context.
-Furthermore, we strongly encourage authors to publish individual size data whenever possible.
-This will allow for the consistent re-analysis of existing data sets as methodologies develop and improve.
-This will aid in the ability for size spectra work to be synthesized between research groups and across scales.
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
-
-\newpage
-
-# Supplementary material {.unnumbered}
-
-```{=tex}
-\beginsupplement
-```
-# $\lambda$ and relationship estimates
-
-In the main analysis, we simulate body size data from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.
-For the results presented in the main text, we held the number of sites at 5, scaled the environmental gradient from -1 to 1, and set the minimum and maximum body sizes to $0.0026$ and $1.2 *10^3$ respectively.
-Here, we plot the results of varying the number of sites (3, 10), increasing the scale of the environmental gradient (-100 to 100) and decreasing the range of body sizes (min = 1, max = 100).
-Generally, the results reported in the main manuscript are robust to changing these parameters: the MLE estimate is nearly always closer to the known parameters, and the variation in these estimates is usually smaller than the binning methods.
-
-## Varying number of sites
-
-### 10 sites
-
-![Individual regressions for ten sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis](figures/PLB_10_sites_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for ten sites across a hypothetical gradient with known values.
-All other parameters are the same as in the main analysis](figures/PLB_10_sites_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for ten sites across a hypothetical gradient with known value of 0.5.
-All other parameters are the same as in the main analysis](figures/PLB_10_sites_relationship_density.png)
-
-\newpage
-
-### Three sites
-
-![Individual regressions for three sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis.](figures/PLB_3_sites_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for three sites across a hypothetical gradient with known values.
-All other parameters are the same as in the main analysis.](figures/PLB_3_sites_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for three sites across a hypothetical gradient with known value of 0.5.
-All other parameters are the same as in the main analysis](figures/PLB_3_sites_relationship_density.png)
-
-\newpage
-
-## Large environmental gradient
-
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_relationship_density.png)
-
-\newpage
-
-## Range of body sizes, $M$
-
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_main.png)
-
-\newpage
-
-![Distribution of estimated \$\\lambda\$ coefficient for five sites across a hypothetical gradient with known values(dashed line).
-Range of body sizes is smaller than main anaysis and ranges from 1, to 100.](figures/PLB_small_m_est_b_density.png)
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
-Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_relationship_density.png)
-
-\newpage
-
-## Sample size, $n$
-
-The number of observations in our simulations may bias the results.
-Therefore, we repeated the simulations described above, but varied the sample size $n$.
-We tested values of $n = 200, 500, 1000, 5000, 10 000$.
-Results of this analysis are presented in the Supplemental Information.
-
-![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).
-(match this figure to "new" style if we like that better)](figures/n_vary_main.png)
-
-![Distribution of size spectra parameter estimates.
-Vertical line is the known parameter (dashed line) wich describes the bounded power law distribution from which the body size estimates were sampled.
-As n increases (top to bottom) and $\lambda$ increases (left to right), the accuracy of the estimate improves across all methods.](figures/n_vary_est_b.png)
-
-\newpage
-
-![Distribtuion of the relationship coefficients with varying sample size](figures/n_vary_relationship_density.png)
-
-# Comparison with other published estimates
-
-SI Table.
-This table shows published estimates of the variation in size spectra slopes (or exponents) in empirical studies.
-It is unclear how to directly compare estimates of the slope with different methods.
-However, the published estimates here range from ~0.1 to 0.2 across the gradients studied.
-For comparison, the 2.5-95% quantiles around the relationship estimate for the MLE method were ~0.1, whereas for the ELBn and NAS method they were ~0.25 and ~0.2, respectively.
-b_diff is the change in estimate (b-low - b-min).
-System refers to stream communities or mesocosm experiments.
-Method: MLE = maximum likelihod estimate, ELB = equal logarthmic binning, the number before indicates the number of bins used.
-The normalization process shifts the estimates by an absolute value of 1.0.
-Hence, direct comparison the relative change in normalized and non-normalized studies should not introduce any bias.
-The O'Gorman et al.
-2017 study used average species size and abundance (Local Size Density Relationship, *sensu* White et al.
-2007) as opposed to individual size distribution.
-These methods are related, but it is unclear how to directly compare estimates from each method.
-
-Supplementary table comparing with other published results will be submitted as a separate file for formatting.

### 2023-07-25 5a_SI_figs.R
update script for maiing SI figs and tables

@@ -1,9 +1,11 @@
-# sample size analysis
+# figures and tables for SI
 
 library(tidyverse)
 library(sizeSpectra) #bounded power law and MLE functions
 source("custom_functions.R")
 library(tidybayes)
+
+# body values sample size ####
 
 # read in datasets with variable sample size, n
 # All other params are as in main analysis
@@ -11,201 +13,101 @@ n200 <- readRDS("data_sim/PLB_sim_n200_dat.rds")
 n500 <- readRDS("data_sim/PLB_sim_n500_dat.rds")
 n5000 <- readRDS("data_sim/PLB_sim_n5000_dat.rds")
 n10000 <- readRDS("data_sim/PLB_sim_n10000_dat.rds")
+# n1000 are the results from the main analysis
+# the actual number of body sizes sampled is n = 999
 n1000 <- readRDS("data_sim/PLB_sim_dat.rds")
-n1000$n <- 1000
+n1000$n <- 999
 
 
-### ELBn struggles to estimate b when sampled from -2 and -2.5 PLB ###
-# Filter out all reps which have an estimate == NA
-
-# make vector of reps which have an NA value for the estimate
-#rep_na <- n100$rep[which(is.na(n100$estimate))]
-
-# filter out all the reps with NA values
-# the !rep means (reps which do not match values in rep_na)
-#PLB_sim_n100 <- n100 %>%
-#  filter(!rep %in% rep_na)
-
-# combine all data sets
+# combine all sample size data sets data sets ####
 dat <- bind_rows(n200, n500, n5000, n10000, n1000)
 
-# labda ~n line graph old? ####
-# # How does the lambda estimate vary by sample size?
-# ggplot(dat, aes(x = n, y = estimate, color = name)) +
-#   geom_point(alpha = 0.1) +
-#   #stat_smooth(method = "loess") +
-#   stat_summary(geom = "line",  fun = mean, size = 1.5) +
-#   geom_hline(aes(yintercept = known_b), linetype = "dashed") +
-#   facet_wrap(~known_b,
-#              scales = "free") +
-#   theme_bw() +
-#   labs(title = "Lambda ~ n",
-#        caption = "dashed line is known lambda value") +
-#   NULL
-# 
-# ggsave("figures/lambda_n_5_sites.png",
-#        width = 8,
-#        height = 8)
 
-# lambda ~ n density old? ####
-# how does estimate of lambda vary with sample size?
-# ggplot(dat,
-#        aes(x = estimate,
-#            y = ..scaled..,
[…]

### 2023-07-25 SI_draft.Rmd
Separate SI rmd and knit files

@@
 
-In the main analysis, we simulate body size data from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.
-For the results presented in the main text, we held the number of sites at 5, scaled the environmental gradient from -1 to 1, and set the minimum and maximum body sizes to $0.0026$ and $1.2 *10^3$ respectively.
-Here, we plot the results of varying the number of sites (3, 10), increasing the scale of the environmental gradient (-100 to 100) and decreasing the range of body sizes (min = 1, max = 100).
+In the main analysis, we sampled body size values from bounded power law distributions while varying the $\lambda$ exponent which describes the distribution.
+For the results presented in the main text, we held the number of body size values constant at n=999, held the number of sites at 5, scaled the environmental gradient from -1 to 1, and set the minimum and maximum body sizes to $0.0026$ and $1.2 *10^3$ respectively.
+Here, we plot the mean and standard deviation of the absolute value of deviations of the estimates from the known $\lambda$ values for the main results, and when varying the number of body sizes sampled (n = 200, 500, 5000, 10000), varying the number of sites (3, 10), increasing the scale of the environmental gradient (-100 to 100) and decreasing the range of body sizes (min = 1, max = 100).
 Generally, the results reported in the main manuscript are robust to changing these parameters: the MLE estimate is nearly always closer to the known parameters, and the variation in these estimates is usually smaller than the binning methods.
-
-## Varying number of sites
-
-### 10 sites
-
-![Individual regressions for ten sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis](figures/PLB_10_sites_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for ten sites across a hypothetical gradient with known values.
-All other parameters are the same as in the main analysis](figures/PLB_10_sites_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for ten sites across a hypothetical gradient with known value of 0.5.
-All other parameters are the same as in the main analysis](figures/PLB_10_sites_relationship_density.png)
-
-\newpage
-
-### Three sites
-
-![Individual regressions for three sites across a hypothetical gradient with a known relationship of 0.5. All other parameters are the same as in the main analysis.](figures/PLB_3_sites_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for three sites across a hypothetical gradient with known values.
-All other parameters are the same as in the main analysis.](figures/PLB_3_sites_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for three sites across a hypothetical gradient with known value of 0.5.
-All other parameters are the same as in the main analysis](figures/PLB_3_sites_relationship_density.png)
-
-\newpage
-
-## Large environmental gradient
-
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_main.png)
-
-\newpage
-
-![Distribution of estimated $\lambda$ coefficient for five sites across a hypothetical gradient with known values.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_est_b_density.png)
-
-\newpage
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
-Range of environmental values (*x*-axis) increased to be -1000, to 1000.](figures/PLB_large_x_relationship_density.png)
-
-\newpage
-
-## Range of body sizes, $M$
-
-![Individual regressions for five sites across a hypothetical gradient with a known relationship of 0.5. Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_main.png)
-
-\newpage
-
-![Distribution of estimated \$\\lambda\$ coefficient for five sites across a hypothetical gradient with known values(dashed line).
-Range of body sizes is smaller than main anaysis and ranges from 1, to 100.](figures/PLB_small_m_est_b_density.png)
-
-![Distribution of estimated relationship ($\beta_1$) coefficient's for five sites across a hypothetical gradient with known value of 0.5.
-Range of body sizes is reduced and is from 1, to 100.](figures/PLB_small_m_relationship_density.png)
-
-\newpage
 
@@
 The number of observations in our simulations may bias the results.
-Therefore, we repeated the simulations described above, but varied the sample size $n$.
-We tested values of $n = 200, 500, 1000, 5000, 10 000$.
-Results of this analysis are presented in the Supplemental Information.
+Therefore, we repeated the simulations described in the main, but varied the number of body size values sampled.
+We tested values of $n = 200, 500, 999, 5000, 10 000$.
+All other parameters of our simulations were kept the same as the simulations presented in the main text.
 
-![Individual regression estimates across the hypothetical gradient based on sample size (rows) and methodology used (columns).
-(match this figure to "new" style if we like that better)](figures/n_vary_main.png)
+All methods improved with increasing sample size (Fig.
+S1).
+However, the accuracy of the MLE method was far superior.
+For example, the average absolute deviation for the MLE method with only 200 body sizes values was similar to the ELBn method with 5,000 body size values and smaller than the L2n method with 10,000 body size values (Table S1).
 
-![Distribution of size spectra parameter estimates.
-Vertical line is the known parameter (dashed line) wich describes the bounded power law distribution from which the body size estimates were sampled.
-As n increases (top to bottom) and $\lambda$ increases (left to right), the accuracy of the estimate improves across all methods.](figures/n_vary_est_b.png)
+![Figure S1.
+Absolute deviation in estimate from known value of $\lambda$.
+Points are jittered horizontally for visualization, and are deviations for each simulated replicate (n = 1000) across varying sample sizes of body size values.
+Lines are the average absolute deviation from the estimate.
+Colors indicate which method was used to estimate $\lambda$.](figures/abs_dev_by_body_n.png)
+```{r, echo=FALSE, results='asis', message=FALSE}
+library(knitr)
+table1 <- read_csv("tables/abs_dev_n_vary.csv") %>%
+  mutate(diff_abs_mean = round(diff_abs_mean, 3),
+         diff_abs_sd = round(diff_abs_sd, 3)) %>%
+  rename(method = name,
+         "mean abs diff"=diff_abs_mean,
+         "SD abs diff"=diff_abs_sd)
+kable(table1,
+         caption = "Table S1.
+Deviation of estimates across methods for different sample sizes of body values (body_n).
+The mean and standard deviation for each row (mean abs diff and sd abs diff, respectively) in the table are calculated across all replicates for each value of lambda (1000 reps * 5 lambda values = 5000 values.",
+      padding = 10)
+```
 
-\newpage
+## Range of values for hypothetical environmental gradient (x) and body sizes (m)
 
-![Distribtuion of the relationship coefficients with varying sample size](figures/n_vary_relationship_density.png)
+We also wanted to ensure that our results were not sensitive to the values for the hypothetical gradient or the range of body sizes sampled.
+For this, we varied the values of the hypothetical gradient from -100 to 100, instead of from -1 to 1 as in the main text while keeping all other parameters of the simulation as described in the main text.
+We found no differences in the deviation of estimates when the hypothetical gradient was changed across any of the methods (Table S2).
 
-# Comparison with other published estimates
+When the range of body sizes sampled was reduced to be from 1 to 100, the ELBn method had slightly greater absolute deviations (0.02) from the known values of $\lambda$, although the variation in the estimates was reduced by ~0.016 (Table S2).
+In contrast, the L2n method had generally lower absolute deviations (0.05) and lower variability (0.04) with smaller body size ranges.
+Finally, the MLE method had slightly higher absolute deviations (0.004) and variability (0.001) when the range of body sizes was reduced, but was more similar to estimated values presented in the main results (Table S2).
 
-SI Table.
-This table shows published estimates of the variation in size spectra slopes (or exponents) in empirical studies.
-It is unclear how to directly compare estimates of the slope with different methods.
-However, the published estimates here range from ~0.1 to 0.2 across the gradients studied.
-For comparison, the 2.5-95% quantiles around the relationship estimate for the MLE method were ~0.1, whereas for the ELBn and NAS method they were ~0.25 and ~0.2, respectively.
-b_diff is the change in estimate (b-low - b-min).
-System refers to stream communities or mesocosm experiments.
-Method: MLE = maximum likelihod estimate, ELB = equal logarthmic binning, the number before indicates the number of bins used.
-The normalization process shifts the estimates by an absolute value of 1.0.
-Hence, direct comparison the relative change in normalized and non-normalized studies should not introduce any bias.
-The O'Gorman et al.
-2017 study used average species size and abundance (Local Size Density Relationship, *sensu* White et al.
-2007) as opposed to individual size distribution.
-These methods are related, but it is unclear how to directly compare estimates from each method.
+```{r, echo=FALSE, results='asis', message=FALSE}
+library(knitr)
+table2 <- read_csv("tables/si_table2.csv") %>%
+  mutate(diff_abs_mean = round(diff_abs_mean, 3),
+         diff_abs_sd = round(diff_abs_sd, 3)) %>%
+  rename(method = name,
+         "mean abs diff"=diff_abs_mean,
+         "SD abs diff"=diff_abs_sd)
+kable(table2,
+         caption = "Table S2. Deviation of estimates across methods when the values of the hypothetical gradient and the range of body sizes are changed.")
+```
 
-Supplementary table comparing with other published results will be submitted as a separate file for formatting.
+## Varying number of sites
+
+The results presented in the main text used five sites equally spaced across a hypothetical environmental gradient.
+We also ran the simulations using 3 and 10 sites, while keeping all other parameters as described in the main text.
+Generally, the average absolute deviation in estimates increased with the number of sites, whereas the variation in the deviations decreased, although the magnitude of these changes was slight (Table S3).
+The change in both of these metrics was the smallest for the MLE method (both = 0.001) and was approximately 0.007 to 0.012 for the absolute deviations and variation in the binning methods.
+
+```{r, echo=FALSE, results='asis', message=FALSE}
+library(knitr)
+table3 <- read_csv("tables/si_table3.csv") %>%
+  mutate(diff_abs_mean = round(diff_abs_mean, 3),
+         diff_abs_sd = round(diff_abs_sd, 3)) %>%
+  rename(method = name,
+         "mean abs diff"=diff_abs_mean,
+         "SD abs diff"=diff_abs_sd)
+kable(table3,
+         caption = "Table S3. Deviation of estimates across methods when the number of sites was changed. The results presented in the main text use 5 sites.")
+```

### 2023-07-25 figures_draft.Rmd
tweak some figures

@@
 ![Figure 2.
-Distribution of relationship estimates ($beta_{env}$) in three different "windows" of lambda values.
+Distribution of relationship estimates ($\beta_{env}$) in three different "windows" of lambda values.
 The dashed vertical line is the known relationship value of -0.5.](figures/lambda_angle_plot.png)
@@
 
-![Figure 4. Distribution of relationship estimates (beta_1) when estimating from different known relationships](figures/vary_beta_density_plot.png)
+![Figure 4. Distribution of relationship estimates ($\beta_{env}$) when estimating from different known relationships](figures/vary_beta_density_plot.png)
 
 ![Figure 5.
-of empirical data estimates.
-All of the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally greater than the MLE estimates.](figures/empirical_combined.png)
+Estimates of change in exponent for size-abundance relationships across gradients from empirical data estimates.
+Panels A and C show the individual λ estimates for each site and the the line shows the estimated relationship based on method (color) for the natural pollution and temperature gradients, respectively.
+Panels B and D show the mean estimated relationship coefficient ($\beta$, point) ± 1 standard deviation (error bars) from the OLS model for both empirical data sets.
+All the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally larger than the MLE estimates.](figures/empirical_combined.png)

### 2023-07-25 2b_integrated_plots.R

@@ -87,6 +87,7 @@ saveRDS(rel_data, file = "data_sim/rel_data.rds")
 
 ###### Flip facet-wrap rows, top = 0, bottom = -0.5
 rel_data %>%
+  filter(rep<=100) %>%
   mutate(
     Model = case_when(name == "NAS" ~ "L2n", .default = name),
     Model = factor(Model,
@@ -112,8 +113,8 @@ rel_data %>%
   geom_point(alpha = 0.1) +
   facet_wrap(known_relationship~Model)+
   theme_bw() +
-  theme(legend.position="none")+
-  labs(x = "Hypothetical environemntal gradient")
+  theme(legend.position="none") +
+  labs(x = "Hypothetical environemntal gradient") +
   NULL
 
 ggsave(filename = "figures/vary_beta_plot.png",
@@ -130,7 +131,6 @@ saveRDS(rel_data_summary, file = "data_sim/rel_data_summary.rds")
 
 rel_data_summary %>%
   mutate(
-    
     Model = case_when(name == "NAS" ~ "L2n", .default = name),
     Model = factor(Model,
                    levels = 

### 2023-07-26 SI_draft.Rmd
minor chnages to SI and re-knit

@@
 Therefore, we repeated the simulations described in the main, but varied the number of body size values sampled.
-We tested values of $n = 200, 500, 999, 5000, 10 000$.
+We tested values of $n = (200,~500,~999,~5000,~10~000)$.
 All other parameters of our simulations were kept the same as the simulations presented in the main text.
@@
   rename(method = name,
+         "Body sample size" = body_n,
          "mean abs diff"=diff_abs_mean,
@@
          caption = "Table S1.
-Deviation of estimates across methods for different sample sizes of body values (body_n).
+Deviation of estimates across methods for different sample sizes of body size values.
 The mean and standard deviation for each row (mean abs diff and sd abs diff, respectively) in the table are calculated across all replicates for each value of lambda (1000 reps * 5 lambda values = 5000 values.",
@@
 kable(table2,
-         caption = "Table S2. Deviation of estimates across methods when the values of the hypothetical gradient and the range of body sizes are changed.")
+         caption = "Table S2. Deviation of estimates across methods when the values of the hypothetical gradient (large x) and the range of body sizes (small m) are changed.")
 ```

### 2023-11-14 3_bias.R
added 95% ci analysis to 3_bias.R

@@ -92,7 +92,7 @@ bias_table = all_dat %>%
             sd_abs_bias = sd(abs_bias)) %>% 
   arrange(target_name, median_ci_range)
 
-write_csv(bias_table, file = "tables/bias_table.csv")
+#write_csv(bias_table, file = "tables/bias_table.csv")
 
 
 all_dat %>% 
@@ -110,3 +110,108 @@ all_dat %>%
   NULL
   
 
+# proportion of 95% CI with target ####
+lambda_ci_prop <- est_lambda %>%
+  select(known_b, rep, estimate, minCI, maxCI, name) %>%
+  mutate(in_ci = known_b > minCI & known_b < maxCI) %>%
+  na.omit() %>%
+  group_by(name) %>%
+  summarize(count = n(),
+            proportion = sum(in_ci, na.rm = TRUE) / count) %>%
+  mutate(target_name = "Lambda")
+
+beta_ci_prop <- angles %>%
+  select(known_relationship, rep, estimate, minCI, maxCI, name, id) %>%
+  mutate(in_ci = known_relationship > minCI & known_relationship< maxCI) %>%
+  na.omit() %>%
+  group_by(name) %>%
+  summarize(count = n(),
+            proportion = sum(in_ci, na.rm = TRUE) / count) %>%
+  mutate(target_name = "lambda scenarios")
+
+rel_beta_ci_prop <- rel_data_summary %>%
+  select(known_relationship, rep, estimate, minCI, maxCI, name) %>%
+  mutate(in_ci = known_relationship > minCI & known_relationship< maxCI) %>%
+  na.omit() %>%
+  group_by(name) %>%
+  summarize(count = n(),
+            proportion = sum(in_ci, na.rm = TRUE) / count) %>%
+  mutate(target_name = "Varying beta")
+
+beta_props <- bind_rows(beta_ci_prop, rel_beta_ci_prop) %>%
+  group_by(name) %>%
+  summarise(count = sum(count), 
+            proportion = mean(proportion)) %>%
+  mutate(target_name = "Regression Slope")
+
+
+
+ci_prop <- bind_rows(lambda_ci_prop, beta_props)
+
+table <- left_join(bias_table, ci_prop)
+write_csv(table, file = "tables/bias_table.csv")
+
+# plot of 95% CI's ####
+# estimated lambda
+est_lambda %>%
+  filter(rep < 500) %>%
+  select(known_b, rep, estimate, minCI, maxCI, name) %>%
+  group_by(name, known_b) %>%
[…]

### 2023-11-14 2b_integrated_plots.R
updated order of facet plots, set scales to one decimal, added "true" line to plots, top of plot 2 for SI

@@ -36,9 +36,7 @@ angles <- bind_rows(steep = angle_A,
 
 saveRDS(angles, file = "data_sim/angles.rds")
 
-##### Need to chnage order of legend to be L2n, ELBn, MLE ####
-
-
+# figure 3 ####
 (lambda_window <- angles %>%
   mutate(
     Model = case_when(name == "NAS" ~ "L2n", .default = name),
@@ -85,29 +83,41 @@ rel_data <- bind_rows(beta_0 = rel_0,
 saveRDS(rel_data, file = "data_sim/rel_data.rds")
 
 
+# figure 4 ####
 ###### Flip facet-wrap rows, top = 0, bottom = -0.5
+
+ab_data <- data.frame(
+  slope = c(0, -0.25, -0.5),
+  intercept = -1.5,
+  known_relationship = c(0, -0.25, -0.5)
+)
+
+ab_data <- ab_data %>%
+  expand_grid(method = c("MLE", "ELBn", "L2N"))
+
 rel_data %>%
   filter(rep<=100) %>%
   mutate(
     Model = case_when(name == "NAS" ~ "L2n", .default = name),
     Model = factor(Model,
                    levels = 
-                     c("MLE",
-                       "ELBn", 
-                       "L2n")),
+                     c("L2n",
+                       "ELBn",
+                       "MLE"
+                       )),
     # set levels of known relaitonship for facet_wrap() plotting order below
     known_relationship = factor(known_relationship,
                                 levels = c("0", "-0.25", "-0.5"))) %>%
-  filter(rep <= 100) %>%
+  filter(rep <= 500) %>%
   ggplot(aes(x = env_gradient,
              y = estimate, 
              group = rep,
              color = Model)) +
   scale_color_manual(
-    values = c("#019AFF", "#FF914A", "#FF1984" )) +
+    values = c("#FF1984", "#FF914A", "#019AFF")) +
   stat_smooth(geom = "line",
               method = "lm",
-              alpha = 0.05,
+              alpha = 0.15,
               se = FALSE, 
               color = "black") +
   geom_point(alpha = 0.1) +
@@ -115,6 +125,12 @@ rel_data %>%
   theme_bw() +
   theme(legend.position="none") +
   labs(x = "Hypothetical environemntal gradient") +
+  geom_abline(data = ab_data,
+              aes(intercept = intercept, slope = known_relationship),
+              color = "red",
+              linetype = "dashed",
[…]

### 2023-11-14 SI_draft.Rmd
SI add bin widths, top panel fig 2, update captions, number of replicates

@@
 ```
+
+# ELBn bin widths
+
+The ELBn method uses six bins of equal width on a logarithmic scale.
+To illustrate how the widths of these bins varies, we provide the following example when the body sizes range from 1 to 100 g vs 1 to 100g.
+
+```{r, echo = FALSE, message=FALSE}
+dw_range_100 <- c(1, 100)
+
+elbn_100 <- exp(seq(floor(log(min(dw_range_100))),
+                          ceiling(log(max(dw_range_100))),
+                          length.out = 7))
+
+dw_range_1000 <- c(1, 1000)
+
+elbn_1000<- exp(seq(floor(log(min(dw_range_1000))),
+                          ceiling(log(max(dw_range_1000))),
+                          length.out = 7))
+
+elb_df <- data.frame(x = c(elbn_100[1:6], elbn_1000[1:6]),
+           xend = c(elbn_100[2:7], elbn_1000[2:7]),
+          y  = rep(c("range = 100", "range = 1000"), each= 6),
+          yend = rep(c("range = 100", "range = 1000"), each= 6),
+          color = rep(c("black", "red"), 6))
+
+l2n_100 <- 2^(floor(log2(min(dw_range_100))):
+                  ceiling(log2(max(dw_range_100))) )
+
+l2n_1000 <- 2^(floor(log2(min(dw_range_1000))):
+                  ceiling(log2(max(dw_range_1000))) )
+
+# length(l2n_100)
+# length(l2n_1000)
+
+l2n_df <- data.frame(x = c(l2n_100[1:7], l2n_1000[1:10]),
+           xend = c(l2n_100[2:8], l2n_1000[2:11]),
+          y  = c(rep("range = 100", 7),rep("range = 1000", 10)),
+          yend = c(rep("range = 100", 7),rep("range = 1000", 10)),
+          color = c(rep(c("black", "red"), length.out = 7), rep(c("black", "red"), length.out = 10))) 
+```
+
+```{r, echo=FALSE}
+elb_df %>%
+  slice(1:5, 7:10) %>%
+  ggplot(aes(x = x,
+             xend = xend,
+             y = y,
+             yend = yend,
+             color = color)) +
+  geom_segment(linewidth = 3,
+               linejoin = 'bevel',
+               lineend = 'butt') +
+  geom_point(shape = "|", size = 7, color = "black") +
+  theme_bw() +
+  coord_cartesian(xlim = (c(0, 50))) +
+  labs(x = "body size",
+       y = "") +
+  theme(legend.position = "none")
+
+```
+Figure S1.
+The edges of the first 4 bins are shown when the body sizes range from 0-100 (bottom line) and 0-1000 (top line).
+The bins have alternating colors, and black segments are used to emphasize bin edges.
+As you can see, the range of the smallest bins are similar, but the edges quickly diverge in the large bin sizes.
+Only the first 4 bins are shown, and the image is cropped for visualization (i.e., the edges of the last bins are not shown).
+
+# L2n bin widths
+
+To contrast the variable bin widths, we show that the widths of the bins in the L2n method are always the same, but the number of bins changes to accomodate the range of body sizes present.
+
+```{r, echo=FALSE}
+l2n_df %>%
+  ggplot(aes(x = x,
+             xend = xend,
+             y = y,
+             yend = yend,
+             color = color)) +
+  geom_segment(linewidth = 3,
+               linejoin = 'bevel',
+               lineend = 'butt') +
+  geom_point(shape = "|", size = 7, color = "black") +
+  theme_bw() +
+  #coord_cartesian(xlim = (c(0, 50))) +
+  labs(x = "body size",
+       y = "") +
+  theme(legend.position = "none") +
+  scale_x_log10()
+```
+
+Figure S2.
+The edges of all bins for the L2n method are shown when body sizes range from 0-100 (bottom line) and 0-1000 (top line).
+The bins have alternating colors, and black segments are used to emphasize bin edges.
+As you can see, the bin widths are identical, but the number of bins increases with the range of body sizes in the data.
+The plot is presented on a log~10~ scale for visualization.
+
+# Top panel of Figure 2, main text
+
+![](figures/SI_fig2_top_panel.png)
+Figure S3.
+The variation in the distributions of $\lambda$ estimates for "steep" values in Figure 2 in the main text makes it difficult to compare them with the "shallower" distirbutions.
+Here, we repeat just the top panel of Figure 2 from the main text to show the variation in estimates across methods.
+As in the other values of $\lambda$, MLE is the most accurate and precise.
+The L2N method performs slightly worse, and the ELBn method is much more variable, likely due to the fact that the presence/absence of even a single large individual could drastically alter the widths of the bins present within an individual simulation replicate (See main text).
+
 # $\lambda$ and relationship estimates
@@
 
-![Figure S1.
+![Figure S4.
 Absolute deviation in estimate from known value of $\lambda$.
@@
 Colors indicate which method was used to estimate $\lambda$.](figures/abs_dev_by_body_n.png)
+
 ```{r, echo=FALSE, results='asis', message=FALSE}
@@
 kable(table2,
-         caption = "Table S2. Deviation of estimates across methods when the values of the hypothetical gradient (large x) and the range of body sizes (small m) are changed.")
+         caption = "Table S2.
+Deviation of estimates across methods when the values of the hypothetical gradient (large x) and the range of body sizes (small m) are changed compared with the results presented in the manuscript (main).
+")
 ```
@@
 ```
+
+# Altering the number of replicates
+
+The results presented in the main text for experiment 2.1 use 1000 replicates for each scenario.
+This number was chosen arbitrarily - we wanted a large enough number to capture the variability, but a small enough number to make repeated simulations tractable.
+Here, we alter the number of replicates to show how the results presented in the main text were largely invariant to this decision.
+
+![Figure S5.
+We varied the number of replicates (n = 10, 50, 100, 200, 250, 500, 750, 1000) and plotted the distribution of $\beta$ estimates across the three $\lambda$ scenarios.
+The point represents the median of the distribution, and the bold and narrow bars represent the 66th and 95th percentile of the distributions, respectively.
+There was some variation in estimates when n < ~250, but generally the results were surprisingly robust to the number of replicates chosen.
+Note that the y-axis varies across rows.](figures/SI_vary_rep_n.png)
+
+# Confidence intervals
+
+We examined the proportion of 95% Confidence Intervals which had the true value for site specific $\lambda$'s (S6) as well as the CI's for the $\beta_{env}$ across $\lambda$ scenarios (S7) as well as when the effect size of $\beta_{env}$ varies.
+
+![Figure S6.
+CI's and $\lambda$ estimates.
+CI's are colored if they contain the true value (blue) or if they do not (red).
+Only the first 500 replicates are displayed for visualization purposes.
+They are ordered by the rank of the lower 95% CI bound](figures/lamba_ci_true.png)
+
+![Figure S7.
+CI's and $\beta$ estimates when varying the known value of $\beta$.
+CI's are colored if they contain the true value (blue) or if they do not (red).
+Only the first 500 replicates are displayed for visualization purposes.
+They are ordered by the rank of the lower 95% CI bound](figures/rel_data_ci_true.png)
+
+![Figure S8.
+CI's and $\beta$ estimates when varying the scenario of $\lambda$ parameter space.
+CI's are colored if they contain the true value (blue) or if they do not (red).
+Only the first 500 replicates are displayed for visualization purposes.
+They are ordered by the rank of the lower 95% CI bound](figures/angles_ci_true.png)

### 2023-11-14 5a_SI_figs.R

@@ -5,6 +5,8 @@ library(sizeSpectra) #bounded power law and MLE functions
 source("custom_functions.R")
 library(tidybayes)
 
+
+
 # body values sample size ####
 
 # read in datasets with variable sample size, n

### 2023-11-14 figures_draft.Rmd
update fig files and rmd

@@
 ![Figure 1.
+Conceptual figure of the simulation procedure used in experiment 2.
+A) we set $\lambda$ (points) to vary at a known relationship ($\beta$, dashed line).
+B) Using known values of $\lambda$, we sampled 999 body sizes for each site.
+C) We estimated site-specific $\lambda$ using the three methods (ELbn, L2n, MLE, see main text) and plotted them across the hypothetical environmental gradient (points, colored by method).
+D) We then estimated $\beta$ using OLS regressions for each method separately (colored lines).
+E) We repeated A-D 1,000 times to get a distribution of $\beta$ estimates.
+F) We compared the distribution of $\beta$ estimates (colored density plots) with the known values of $\beta$ (dashed vertical line).](conceptual.png)
+
+![Figure 2.
 Distribution of Lambda estimates by method (color) from random samples of body sizes from bounded power law distributions with varying exponents (-2.5 to -0.5).
@@
 
-![Figure 2.
-Distribution of relationship estimates ($\beta_{env}$) in three different "windows" of lambda values.
+![Figure 3.
+Distribution of relationship estimates ($\beta_{env}$) in three different "scenarios" of lambda values; steep: $\lamba$ = -2.5 to -1.5; medium: $\lamba$ = -2.0 to -1.0; steep: $\lamba$ = -1.5 to -0.5.
 The dashed vertical line is the known relationship value of -0.5.](figures/lambda_angle_plot.png)
 
-![Figure 3.
-Individual regressions by method (columns) for three different known relationship values (rows from bottom to top): 0, 0.25, and 0.5.
-There were a total of 1000 replicates simulated, but only 100 replicates are plotted here with `alpha = 0.05` to illustrate the variability in the regression lines.](figures/vary_beta_plot.png)
+![Figure 4.
+Individual regressions by method (columns) for three different known relationship values (rows from bottom to top, red dashed line).
+There were a total of 1000 replicates simulated for each combination of method and known relationship, but only 500 are plotted here to illustrate the variability in the regression lines.](figures/vary_beta_plot.png)
 
-![Figure 4. Distribution of relationship estimates ($\beta_{env}$) when estimating from different known relationships](figures/vary_beta_density_plot.png)
+![Figure 5. Distribution of relationship estimates ($\beta_{env}$) when estimating from different known relationships](figures/vary_beta_density_plot.png)
 
-![Figure 5.
+![Figure 6.
 Estimates of change in exponent for size-abundance relationships across gradients from empirical data estimates.
-Panels A and C show the individual λ estimates for each site and the the line shows the estimated relationship based on method (color) for the natural pollution and temperature gradients, respectively.
+Panels A and C show the individual λ estimates for each site and the line shows the estimated relationship based on method (color) for the natural pollution and temperature gradients, respectively.
 Panels B and D show the mean estimated relationship coefficient ($\beta$, point) ± 1 standard deviation (error bars) from the OLS model for both empirical data sets.
 All the methods estimate the same sign of the relationship, but the estimates from the binning methods are generally larger than the MLE estimates.](figures/empirical_combined.png)
+
+![Figure 7. Decision tree to aid researchers in choosing an analytical framework for estimating Individual Size Distributions. Major pros and cons of each framework are included. ](decision_tree.png)
